### 2011-02-25 workflow.Rnw
Adding more files.

\documentclass[11pt]{article}
\usepackage{fancyvrb,natbib,url,array,dcolumn}

\title{Collaboration}

\author{Jake Bowers\thanks{Many thanks to Kevin Quinn for direct help on this document, Mika LaVaque-Manty, Ben Hansen, and Mark Fredrickson for many useful discussions on this topic, and Kieran Healy for inspiration from afar.}}

\date{\today}

\newcolumntype{.}{D{.}{.}{2.1}} 
\begin{document}
\maketitle

\begin{verse}
Do I contradict myself? \\
Very well then I contradict myself, \\
(I am large, I contain multitudes.) \\
\citep{whitman1855song}
\end{verse}

An idea is born in a coffee shop, a seminar, a quiet walk in 2011. This idea inspires a seminar paper in 2011. And a conference paper arises from the seminar paper in 2012. A dissertation chapter descends from the conference paper in 2013. Other dissertation chapters take up 2014. A submission to a journal happens in 2015. Revision and resubmission wait until 2017 while harried editors, reviewers and author strive to balance teaching, service, and life. In 2018 a child is born and a paper is published. In 2020 a graduate student in a coffee shop reads the paper and has an idea that challenges the analysis in the paper. What would happen if only the author had thought to control for X? Or run one more MCMC chain? Or chosen a different likelihood function? Will the United Nations (now eager to act based the paper) make a wrong move? Time to go back to the analyses run sometime between 2011 and 2020 to check. Easy. Right? Collaboration between the past self and the present self should be simple --- after all you are the same person in 2011 and 2020, right? And the way Microsoft Word/Stata/SPSS/R/LISREL stores files and the way that your Windows/Mac machine reads and writes them is the same, right? And you know exactly which bit of code produced which table and which figure, right? And you wrote your code following Nagler's Maxims \citet{nagler1995coding} and being following King's Replication Standard \citet{king1995replication} right?

Just in case one has some doubt about answering these questions, this piece aims to amplify some of what we already ought to know \citet{nagler1995coding}, and to update some of those ideas given current practices, platforms, and propensities.

\vspace{4em}

When we call some piece of work ``scientific'' we are talking about how it was done. By calling it ``scientific'' we are also using a short-hand to suggest that (1) the findings of the work are not a matter of opinion and (2) other people could reproduce the findings. That is, what the work represents is not a matter of opinion but is a shared experience --- and an experience shared without respect to the identities of others (although requiring some shared technical training and research resources). This short-hand means that ``scientific'' work should make us change how we act --- or at least, it ought to stand on stronger epistemological ground than other claims about experiences which ought to be shared or shareable.

Assume we want other scientists to believe us when we say something. More narrowly, assume we want other people to believe us when we say something about data: ``data'' here can be words, numbers, musical notes, images, ideas, etc \ldots The point is that we are making some claims about patterns that we see in some collection of stuff. Now, it might be easy to convince other people that ``this collection of stuff is different from this collection of stuff'' if those people were looking over our shoulders the whole time that we made decisions about collecting the stuff and analyzed it [where ``analyze'' means something like ``break it up into understandable parts'']. Unfortunately, we can't assume that people are willing to shadow a researcher throughout her career. Rather, we do our work alone or in small groups and want to convince other distant people about our analyses.

Now, say your collections of stuff are large or complex and where your chosen tools of analyses are computer programs. How can we convince people that what we did with some data with some program is credible: not a matter of whim or opinion, and reproducible by others who didn't shadow us as we wrote our papers? Here are a few concrete actions that people can take to enhance the believeability of their work. In addition, these actions make collaboration within research groups more effective (since believeability comes in part from reproducibility and research groups often need to be able to reproduce in part or in whole what different people in the group have done).

Here is some advice that is not currently emphasized but which might be profitably added to previous advice.

\section{Show me the data.}
All files containing commands operating on data must refer to a data file. A reference to a data file is a line of code the analysis program will use to operate on (``load''/ ``open'' / ``get'' / ``use'') the data file. One should not have to edit this line on different computers or platforms in order to execute this command.  Using R, for example, all analysis files should have load('thedata.rda') or \Verb+read.csv('http://www.mywebsite.org/Data/thedata.csv')+ or some equivalent line in them, and 'thedata.csv' should be stored in some place easy to find (like in the same directory as the file or perhaps in 'Data/thedata.rda').

Where should you store the data files? A very obvious solution is to always make sure that the data file used by a command file is in the same directory as the command file. More elegant solutions (pointed to below) require all co-authors to have the same directory structure so that load('Data/thedata.rda') means the same thing on all computers used to work on said project.

The principle of modularity \cite{nagler1995coding} suggests that you separate data cleaning, processing, recoding, and merging from analysis in different files. So, perhaps your analysis oriented files will load('cleandata.rda') or something but with a comment in the code telling the future you (among others) that cleandata.rda was created from create-cleandata.R which in turn begins with read.csv('dirtydata.csv') which, from a comment in the file, was downloaded from \url{http://www.greatfreedata.gov/thedata?dirtydata.csv} on April 1, 2011. Such a data processing file will typically end with something like \Verb+save('cleandata.rda')+ so that we are doubly certain about the provenance of the data.

Now, if in 5 years we wonder where 'cleandata.rda' came from, we might \Verb+ grep cleandata *.R+ to search for occurrences of this file on our Unix/OS X system. However, if such searching among files is a burden, and even nicer solution is to maintain a file for each project called ``MANIFEST.txt'' or ``INDEX.txt'' or ``README.txt'' which lists the data and command files with brief descriptions of their functions and relations.

The principle regarding data is to know where the data came from and what operations were performed on which set of data. I have seen command files which do not begin with a call to load data and I have despaired. What is the meaning of \Verb+please-fit var1 with var2+ if I do not know where var1 var2 come from.\footnote{The command \Verb+please-fit+ comes from the MayIPleaseDoStatistics package for R which emphasizes politeness in data analysis. It is highly recommended for use with toddlers.}

\section{Code not Clicking}
All tables and graphs should be as much as possible generated from code and not clicking (let alone copying and pasting). If I wanted to re-create the figure you created but with red lines, and only for people in the South, I should be able to do so with just a few edits to the code.

Using R, for example: I would write

\begin{verbatim}
pdf('fig1.pdf')
plot(stuff)
dev.off()
\end{verbatim}

Now, in the future if I wonder how ``that plot on page 10'' was created, if I suitably commented my manuscript or inserted text into my MANIFEST.txt file I will know (1) ``that plot'' is from a file called ``fig1.pdf'' and (2) fig1.pdf was created in \Verb+figs1and2.R+. In the future, if I wanted to quickly change fig1.pdf, I could edit the commands quickly and easily to do so. [Another good practice is to name the files descriptively, so that we don't in fact have ``fig1.pdf'' but ``political participation by education in the south.pdf'' although such a name may prove burdensome to operating systems which do not care for spaces in file names or to folks who are used to typing full filenames rather than using command line completion facilities in modern shells.]

An even nicer way to ensure that figures and tables are reproducible or easily modifiable by your older and wiser but more harried self is to use literate programming. For example, one may insert the following chunk of code into either a LaTeX or an OpenOffice document, ask R to process the document, and then produce a pdf file in which the code has disappeared to be replaced with an actual figure:

\begin{Verbatim}
 <<fig1,fig=TRUE>>==
 plot(stuff)
 @
\end{Verbatim}

The principle here is to know where tables and figures came from --- to know what commands created which figures and tables.

\section{Fishing rods and not fish}
The source code of any pdf document exchanged by the group must be available and executable. Say you run some command and then make a table or a graph. Then you produce a nice little report on your work for use among the working group. Eventually we want to use pieces of that report (tables, graphs, paragraphs) in a publishable paper. Thus, that report must not have been created using copying and pasting. It must be created using LaTeX (or HTML or OpenOffice or something else that the working group standardizes on). And, in fact, it must be created using the reproducible markup system like Sweave (if you are using LaTeX and R as your working group tools). Alternatively, you can put comments into your report detailing where the figures and tables came from (i.e. which command file produced them). Or include details about which code produced which figures in your MANIFEST.txt.

In order to know that ``that plot on page 10'' is from a file called ``fig1.pdf'' you need to document it somehow. I can imagine a system with very disciplined use of Word/OpenOffice --- perhaps using special auxiliary ``listoffigures.txt'' files or something. In my experience, it is much easier to just use LaTeX (which can be used in a very user friendly way via LyX).

The Sweave system is even nicer because you can have a LaTeX file with R code right inside of it! I showed Sweave with plotting above so here is an example in which a regression table is produced automagically:

<<maketable, echo=FALSE, results=tex>>=
##source("http://artsci.wustl.edu/~jgill/Models/graph.summary.R")
dem.nations.df<-read.csv("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv")

lm1<-lm(protac2000~Gini2004+meanpr,data=dem.nations.df)

themat<-cbind(summary(lm1)$coef[,c("Estimate","Std. Error")],confint(lm1))
n<-length(lm1$fitted.values)
est.sigma<-round(summary(lm1)$sigma,3)
library(xtable)

colnames(themat)[3:4]<-c("~{}","~{}")
thextab<-xtable(themat,align=c("l",rep(".",4)),
                caption="People living in countries with unequal
  income distributions report less protest activity to World Values
  Survey interviewers than people living in countries with relatively
  more equal income distributions, controlling for mean political
  rights as measured by Freedom House 1980--2000",label="tab:protest")

print(thextab,include.colnames=FALSE,
      add.to.row=list(pos=list(0,3),
                command=c("& \\multicolumn{1}{c}{Coef} & \\multicolumn{1}{c}{Std. Err.} 
& \\multicolumn{2}{c}{95\\% CI} \\\\", 
                  paste("\\multicolumn{4}{l}{n: ",n,",   $\\hat{\\sigma}$:",est.sigma,"} \\\\ "))
                )
      )
@

Right now it is hard to beat Sweave as a system for enhancing the collaboration of research groups doing quantitative work in political science. Something will make our lives easier than Sweave (or odfWeave) in the future. Then we'll change.

\section{Enabling revision. Preventing clobbering.}

Group work requires version control.  If you are using some kind of weak version control system (which is much much better than none at all), like Dropbox, then you must communicate with other folks in your group before you edit existing files. That is, only one of you can edit a given file at a time. This prevents your work (or your colleagues work) from getting lost when you both try to save the same file on top of each other. If you find that you are needing to work on the same files at the same time, then you should work on establishing your own shared version control system. Free options include launchpad, github, sourceforge for open source projects (i.e. papers you are writing which you are happy to share with others). Each of those services include paid versions too. One may also use Dropbox as a kind of server for version control: checking out files from the Dropbox directory into a local working directory.

% (the bazaar system appears particularly well suited for use with DropBox http://marcioandreyoliveira.blogspot.com/2009/05/get-your-own-bazaar-hosting-service-for.html)

We use subversion with our own research group, and I use it for all of my own projects. Subversion and bazaar and git are  all great. They mainly differ in the extent to which you need to run a server. Subversion requires a server and we are lucky that the National Center for Supercomputing Applications provides such a  server for us. However, the price of such hosting may not be that much using one of the many webhosting services out there.

An excellent, simple, and robust version control system is to merely rename your files with the date and time of saving them:  thedoc.txt becomes thedoc25092009-22:40.txt. And then when you reach certain milestones you can rename the file accordingly: thedocAPSA2009.txt [for the one sent to discussants at APSA] or thedocAPSR2015.txt [for the version eventually sent to the APSR six years after you presented it at APSA]. The systems I mentioned above all allow this kind of thing and are more elegant, but you can do it by hand too as long as you don't mind taking up a lot of disk space and having many many ``thedoc...'' files around.

\section{Obey Nagler's Maxims}
Read \citet{nagler1995coding}. Comment your code. Luckily, if you are
using a literate programming practice (Sweave, odfWeave, using R2HTML,
etc..), you can write paragraphs to surround your code. If not, then
use whatever commenting system exists in your analysis system.

A given analysis of a given collection is a long series of
decisions. You need to be able to justify them. Often you justify them
in the paper you are writing (in the text surrounding the R code
chunks in your Sweave document). Some decisions will be too small and
technical for inclusion in the paper itself. These need to be
documented in the code itself.

Using R within Sweave, for example, the comment is marked using the pound sign. Notice that I've left in a regression not run for the paper directly, but run as a part of the due diligence process of writing the paper.

\begin{verbatim}
<<reg1>>
##Fitting the regression with x/1000 for ease of interpretation
thelm<-lm(y~I(x/1000))
##Fits using a quadratic term as suggested by Blah(1965) did not enhance the fit of the model
##thelmsq<-lm(y~I(x/1000)^2)

@
\end{verbatim}

If you need to sleuth down something more than a small detail, but yet the sleuthing is really a footnote to the article, then you can create another file to explore and report on those questions for the research group. The writing in such a memo can be more informal, but the norms of reproducibility ought to be the same as for the main paper. After all you will be releasing both documents (probably) publicly upon publication of the article if not before.

(and read Jonathan Nagler's piece on this too)

\section{File Naming}

(6) Name your files with evocative and descriptive names. Your collaborators are less likely to call you at midnight asking for help if your files are named ``regressions.R'' than if your files are called ``temp'' (note the use of the extension .R to tell us that the file contains R commands. use extensions like this as a standard practice to help you and your computer get along.)

(other stuff I've forgotten?)

[examples: Examples of reprodible research using R vignette system [jake's Manifest Effects paper?], and a more complex system that required more time instensive simulation using a Makefile [jake and ben and marks's JASA archive]?]

\section{Constructive Collaboration}

We all always collaborate. Many of us collaborate with groups of people at one moment in time as we race to produce a paper in time for some deadline. All of us collaborate with ourselves over time. An idealized version of self-collaboration goes something like this: We begin a project with a few notes. We write a conference paper. We revise this paper and send it to a journal. The journal asks for revisions. We revise and resubmit. Our paper is accepted. We revise a bit more. The paper is published. Later another scholar reacts to our paper. We write another paper based on said reactions (draft, conference paper, submission, revision, acceptance, revision). These processes occur over years of time. So, I collaborate with a self 5 or 10 or even more years hence when I write a paper. [Ask anyone who has turned their dissertation into a book how long it took from them from dissertation prospectus to physical published book for an example of a common time span in which self-collaboration is required.]

The time-frames over which collaboration are required --- whether among a group of people working together or within a single scholar's productive life --- are much longer than any given version of any given software will easily exist. All except for plain text. Thus, even as we extol version control systems, one must have a way to ensure future access to them in a form that will still be around when the cockroaches take over political science departments (and thereby finally make C-x M-x C-c  a reasonable way to instruct Emacs to do something [insert url citing this argument --- that Emacs was really made for cockroaches or other six legged creatures.]

In the end, following these practices allows your analyses of your collections to be credible --- and, in the case of criticism, you can always tell antagonists to go redo your analysis if they really don't believe you.

"[I]f the empiricalbasis for an articleor book cannotbe reproduced,of what use to the discipline are its conclusions? What purposedoes an articlelike this serve? \cite[445]{king1995replication}

\bibliographystyle{asa} 
\bibliography{/Users/jwbowers/Documents/BIB/trunk/big}

\end{document}

### 2011-02-25 regsummary.R

@@ -1,0 +1,53 @@
+##a fork from http://artsci.wustl.edu/~jgill/Models/graph.summary.R
+
+
+graph.summary <- function (in.object, alpha = 0.05, digits=3) # DESIGNED BY NEAL, CODED BY JEFF, FORKED BY JAKE
+{
+    lo <- in.object$coefficient - qnorm(1-alpha/2) * summary(in.object)$coef[,2]
+    hi <- in.object$coefficient + qnorm(1-alpha/2) * summary(in.object)$coef[,2]
+    out.mat <- round(cbind(in.object$coefficient, summary(in.object)$coef[,2], lo, hi),digits)
+    blanks <- "                                                                          "
+    dashes <- "--------------------------------------------------------------------------"
+    bar.plot <- NULL
+    scale.min <- floor(min(out.mat[,3])); scale.max <- ceiling(max(out.mat[,4]))
+    for (i in 1:nrow(out.mat))  {
+        ci.half.length <- abs(out.mat[i,1]-out.mat[i,3])
+        ci.start <- out.mat[i,1] - ci.half.length
+        ci.stop <- out.mat[i,1] + ci.half.length
+        bar <- paste("|",substr(dashes,1,ci.half.length), "o", substr(dashes,1,ci.half.length), "|", sep="", collapse="")
+        start.buf <- substr(blanks,1,round(abs(scale.min - ci.start)))
+        stop.buf <- substr(blanks,1,round(abs(scale.max - ci.stop)))
+        bar.plot <- rbind( bar.plot, paste(start.buf,bar, stop.buf, sep="", collapse="") )
+    }
+    out.df <- data.frame( matrix(NA,nrow=nrow(out.mat),ncol=ncol(out.mat)), bar.plot[1:length(bar.plot)] )
+    out.df[1:nrow(out.mat),1:ncol(out.mat)] <- out.mat
+    CI.label <- paste( "CIs:", substr(blanks,1,abs(scale.min)-2-4),"ZE+RO",
+        substr(blanks,1,abs(scale.max)-2), sep="", collapse="" )
+    dimnames(out.df)[[1]] <- dimnames(summary(in.object)$coef)[[1]]
+    dimnames(out.df)[[2]] <- c("Coef","Std.Err.", paste(1-alpha,"Lower"),paste(1-alpha,"Upper"),CI.label)
+    if (substr(in.object$call[1],1,2) == "gl")  print(in.object$family)
[…]

### 2011-02-25 workflow.Rnw
Changed main document to Sweave.

@@
-\documentclass[11pt]{article}
-\usepackage{fancyvrb,natbib,url,array,dcolumn}
+\documentclass[12pt]{article}
+\usepackage{fancyvrb,natbib,url,array,dcolumn,float,booktabs}
 
@@
 
-\newcolumntype{.}{D{.}{.}{2.1}} 
+\newcolumntype{.}{D{.}{.}{1.2}}
+
+\usepackage[noae,nogin]{Sweave} %Declaring this explicitly so that I can modify the Sweave environment here rather in Sweave.sty which will be c
+\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=.5em,fontsize=\footnotesize} 
+\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=.5em,fontsize=\footnotesize} 
+\DefineVerbatimEnvironment{Scode}{Verbatim}{xleftmargin=.5em,fontsize=\footnotesize} 
+\fvset{listparameters={\setlength{\topsep}{0pt}}} 
+\renewenvironment{Schunk}{\vspace{\topsep}}{\vspace{\topsep}} 
+\SweaveOpts{keep.source=TRUE}  %Show comments
+\fvset{fontsize=\normalsize} %slightly smaller verbatim font
+
+%% These next lines tell latex that it is ok to have a single graphic
+%% taking up most of a page, and they also decrease the space arou
+%% figures and tables.
+\renewcommand\floatpagefraction{.9}
+\renewcommand\topfraction{.9}
+\renewcommand\bottomfraction{.9}
+\renewcommand\textfraction{.1}   
+\setcounter{totalnumber}{50}
+\setcounter{topnumber}{50}
+\setcounter{bottomnumber}{50}
+\setlength{\intextsep}{2ex}
+\setlength{\floatsep}{2ex}
+\setlength{\textfloatsep}{2ex}
+
 \begin{document}
 \maketitle
+
+<<initialize,echo=FALSE,results=hide>>= 
+##First, just setup the \R environment:
+
+options(SweaveHooks=list(
+          fig=function(){
+            par(mar=c(3.5, 3, 1.1, 0),
+                pty="s",
+                mgp=c(1.5,0.5,0),
+                oma=c(0,0,0,0))},#,
+          echo=function(){options(continue=" ") ##Don't show "+" prompts,
+                         options(prompt=" ")}
+          ),
+        digits=3,
+        scipen=8,
+        width=100
+        )
+
+@
 
@@
 
-An idea is born in a coffee shop, a seminar, a quiet walk in 2011.
-This idea inspires a seminar paper in 2011.
+An idea is born in a coffee shop, a seminar, a quiet walk.
+It is winter 2011.
+This idea inspires a seminar paper in the spring.
 And a conference paper arises from the seminar paper in 2012.
@@
 I have seen command files which do not begin with a call to load data and I have despaired.
-What is the meaning of \Verb+please-fit var1 with var2+ if I do not know where var1 var2 come from.\footnote{The command \Verb+please-fit+ comes from the MayIPleaseDoStatistics package for R which emphasizes politeness in data analysis.
+What is the meaning of \Verb+please-fit(var1 with var2)+ if I do not know where var1 var2 come from.\footnote{The command \Verb+please-fit+ comes from the MayIPleaseDoStatistics package for R which emphasizes politeness in data analysis.
 It is highly recommended for use with toddlers.}
@@
 
-Using R, for example: I would write
+Using R, for example, I might specify that the file \Verb+fig1.pdf+ was produced by the command, where the \Verb+please-plot+ function was loaded elsewhere.
 
@@
 pdf('fig1.pdf')
-plot(stuff)
+please-plot(outcome by explanatory)
 dev.off()
@@
 
-\begin{Verbatim}
- <<fig1,fig=TRUE>>==
- plot(stuff)
+<<read.data>>=
+
+##This read.csv line is slightly more complex than needed because of
+##the particular encoding used for characters in the source file.
+
+dem.nations.df<-read.csv(url("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv",encoding="ISO_8859-1"),as.is=TRUE)
+
+row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful
+
+good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
 @
-\end{Verbatim}
+
+<<fig1code,echo=TRUE,fig=FALSE,width=3,height=3>>=
+##The code to produce the following figure
+par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
+with(good.df,plot(Gini2004/100,protac2000,
+     xlab='Gini Coefficient 2004 (UNDP)',
+     ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
+     cex=.8))
+
+with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
+     text(Gini2004/100,protac2000,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
+@
+
+\begin{figure}[ht]
+  \begin{center}
+<<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>= 
+<<fig1code>>
+@
+    \caption{Protest activity by income inequality from \citep{norris2009data}.}\label{fig:giniprot}
+  \end{center}
+\end{figure}
 
@@
 ##source("http://artsci.wustl.edu/~jgill/Models/graph.summary.R")
-dem.nations.df<-read.csv("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv")
 
-lm1<-lm(protac2000~Gini2004+meanpr,data=dem.nations.df)
+lm1<-lm(protac2000~I(Gini2004/100)+meanpr,data=good.df)
 
@@
 colnames(themat)[3:4]<-c("~{}","~{}")
+row.names(themat)<-c("Intercept","Income Inequality (lower=more equal)","Mean Political Rights (lower=more rights)")
 thextab<-xtable(themat,align=c("l",rep(".",4)),
@@
   Survey interviewers than people living in countries with relatively
-  more equal income distributions, controlling for mean political
-  rights as measured by Freedom House 1980--2000",label="tab:protest")
+  more equal income distributions, adjusting for average political
+  rights as measured by Freedom House 1980--2000. Data from \\citep{norris2009data}.",label="tab:protest")
 
 print(thextab,include.colnames=FALSE,
+      hline.after=NULL,
       add.to.row=list(pos=list(0,3),
-                command=c("& \\multicolumn{1}{c}{Coef} & \\multicolumn{1}{c}{Std. Err.} 
-& \\multicolumn{2}{c}{95\\% CI} \\\\", 
-                  paste("\\multicolumn{4}{l}{n: ",n,",   $\\hat{\\sigma}$:",est.sigma,"} \\\\ "))
+                command=c("\\toprule \n & \\multicolumn{1}{c}{Coef} & \\multicolumn{1}{c}{Std. Err.} 
+& \\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ", 
+                  paste("& \\multicolumn{4}{c}{n: ",n,",   $\\hat{\\sigma}$:",est.sigma,"} \\\\ \\bottomrule"))
                 )
@@
 
-Group work requires version control.
-If you are using some kind of weak version control system (which is much much better than none at all), like Dropbox, then you must communicate with other folks in your group before you edit existing files.
-That is, only one of you can edit a given file at a time.
-This prevents your work (or your colleagues work) from getting lost when you both try to save the same file on top of each other.
-If you find that you are needing to work on the same files at the same time, then you should work on establishing your own shared version control system.
-Free options include launchpad, github, sourceforge for open source projects (i.e.
-papers you are writing which you are happy to share with others).
-Each of those services include paid versions too.
-One may also use Dropbox as a kind of server for version control: checking out files from the Dropbox directory into a local working directory.
+Group work requires version control.\footnote{See X for more
+  discussion of what version control is.} Many people are familiar
+with ``Track Changes'' feature in OpenOffice/Word/Pages or the fact that
+Dropbox allows one to recover previous versions of files. These are
+both kinds of version control. When collaborating with yourself or
+others, it is useful to see what has changed, to feel free to
+experiment and then to dump the experiment in favor of previous work,
+to have multiple ``releases'' of the same document (one to MPSA, one
+to APSR, one to your parents) without requiring that spawn many copies
+of the same document and risk confusion and clobbering. Clobbering is
+what happens when your future self or your other collaborators saves
+an old version of a file over a new version --- erasing good work by
+accident.
+
+Of course if you rely on Dropbox or ``track changes'' for version
+control, you must communicate with other folks in your group before
+you edit existing files. Only one of you can edit and save a given
+file at a time. This prevents your work (or your colleagues work) from
+getting lost when you both try to save the same file on top of each
+other. If you find that you are needing to work on the same files at
+the same time, then you should work on establishing your own shared
+version control system. Free options include launchpad, github,
+sourceforge for open source projects (i.e. papers you are writing
+which you are happy to share with others). Each of those services
+include paid versions too. One may also use Dropbox as a kind of
+server for version control: checking out files from the Dropbox
+directory into a local working directory. (Notice that this is
+different from directly working on files within your Dropbox managed
+directories.)
 
@@
 
-We use subversion with our own research group, and I use it for all of my own projects.
-Subversion and bazaar and git are  all great.
-They mainly differ in the extent to which you need to run a server.
-Subversion requires a server and we are lucky that the National Center for Supercomputing Applications provides such a  server for us.
-However, the price of such hosting may not be that much using one of the many webhosting services out there.
+We use subversion with our own research group, and I use it for all of
+my own projects. Subversion and bazaar and git are all great. They
+mainly differ in the extent to which you need to run a
+server. Subversion requires a server and we are lucky that the
+National Center for Supercomputing Applications provides such a server
+for us. However, the price of such hosting may not be that much using
+one of the many webhosting services out there.
 
-An excellent, simple, and robust version control system is to merely rename your files with the date and time of saving them:  thedoc.txt becomes thedoc25092009-22:40.txt.
-And then when you reach certain milestones you can rename the file accordingly: thedocAPSA2009.txt [for the one sent to discussants at APSA] or thedocAPSR2015.txt [for the version eventually sent to the APSR six years after you presented it at APSA].
-The systems I mentioned above all allow this kind of thing and are more elegant, but you can do it by hand too as long as you don't mind taking up a lot of disk space and having many many ``thedoc...'' files around.
+Fancy version control systems are not required, however. I suspect
+that Google Docs allows a kind of version tracking and collaboration
+as well.  An excellent, simple, and robust version control system is
+to merely rename your files with the date and time of saving them:
+thedoc.tex becomes thedoc25092009-22:40.tex.
+If you are wise enough to have saved your documents as plain text (Such as LaTeX source (with or without R/Stata code chunks) \footnote{A quick Google search of ``Sweave for Stata'' turned up lots of resources for literate programming with Stata.} then you can easily compare documents using the many utilities available for comparing text files [FileMerge on OS X is pretty, but ediff and diff for Unix are very useful, I am sure that Windows has many other options].
+Adobe Acrobat allows one to compare differences in pdf files.
+OpenOffice supports a ``Compare Documents'' option.
+
+When you reach certain milestones you can rename the file accordingly:
+thedocAPSA2009.tex [for the one sent to discussants at APSA] or
+thedocAPSR2015.tex [for the version eventually sent to the APSR six
+years after you presented it at APSA]. The systems I mentioned above
+all allow this kind of thing and are much more elegant and capable,
+but you can do it by hand too as long as you don't mind taking up a
+lot of disk space and having many many ``thedoc...''  files around.
 
@@
 using a literate programming practice (Sweave, odfWeave, using R2HTML,
-etc..), you can write paragraphs to surround your code. If not, then
-use whatever commenting system exists in your analysis system.
+etc..), you can write paragraphs to surround your code as well a
+technical comments in the code itself. If you are not using a strictly
+literate programming practice then use whatever commenting protocol
+exists in your analysis language.
 
@@
 
-\begin{verbatim}
-<<reg1>>
+\begin{Verbatim}
+<<reg1>> =
 ##Fitting the regression with x/1000 for ease of interpretation
-thelm<-lm(y~I(x/1000))
-##Fits using a quadratic term as suggested by Blah(1965) did not enhance the fit of the model
-##thelmsq<-lm(y~I(x/1000)^2)
-
+thelm<-lm(y~I(x/1000)+z,data=fabulous.df)
+##Fits using a quadratic term as suggested by Blah(1865) did not enhance the fit of the model
+##thelmsq<-lm(y~I(x/1000)^2+z,data=fabulous.df)
 @
-\end{verbatim}
+\end{Verbatim}
 
@@
 
-(and read Jonathan Nagler's piece on this too)
-
 \section{File Naming}
 
-(6) Name your files with evocative and descriptive names.
-Your collaborators are less likely to call you at midnight asking for help if your files are named ``regressions.R'' than if your files are called ``temp'' (note the use of the extension .R to tell us that the file contains R commands.
-use extensions like this as a standard practice to help you and your computer get along.)
+Name your files with evocative and descriptive names. Your
+collaborators are less likely to call you at midnight asking for help
+if your files are named ``regressions.R'' than if your files are
+called ``temp9'' or ``supercalifragilisticexpialidocious'' (note the
+use of the extension .R to tell us that the file contains R
+commands. use extensions like this as a standard practice to help you
+and your computer get along.)
 
 (other stuff I've forgotten?)
+
+\section{Exemplars}
+
+Lots of people are thinking about ``reproducible research'' and ``literate programming'' these days.
+Google those terms.
+I have experimented with three systems so far: (1) for one paper I simply included my Sweave document and data files into a compressed archive; (2) for another more computing intensive paper...
 

### 2011-02-25 workflow.Rnw
Next

@@
 
-\title{Collaboration}
+\title{A paper is to write}
 
-\author{Jake Bowers\thanks{Many thanks to Kevin Quinn for direct help on this document, Mika LaVaque-Manty, Ben Hansen, and Mark Fredrickson for many useful discussions on this topic, and Kieran Healy for inspiration from afar.}}
+\author{Jake Bowers\thanks{Many thanks to Kevin Quinn and Cara Wong for direct help on this document, Mika LaVaque-Manty, Ben Hansen, and Mark Fredrickson for many useful discussions on this topic, and Kieran Healy for inspiration from afar.}}
 
@@
 
-An idea is born in a coffee shop, a seminar, a quiet walk.
-It is winter 2011.
+An idea is born in a coffee shop, a seminar, a quiet walk. It is a cold winter in 2011, but the idea energizes the student. The idea is why he went to graduate school.
+
 This idea inspires a seminar paper in the spring.
-And a conference paper arises from the seminar paper in 2012.
+And a conference paper arises from the seminar paper in collaboration with another student in 2012.
 A dissertation chapter descends from the conference paper in 2013.
 Other dissertation chapters take up 2014.
-A submission to a journal happens in 2015.
-Revision and resubmission wait until 2017 while harried editors, reviewers and author strive to balance teaching, service, and life.
+A submission to a journal with the original co-author and a new collaborator happens in 2015.
+Revision and resubmission wait until 2017 while harried editors, reviewers and authors strive to balance teaching, service, and life.
+The three collaborators have luckily found work in three different universities around the world.
 In 2018 a child is born and a paper is published.
 In 2020 a graduate student in a coffee shop reads the paper and has an idea that challenges the analysis in the paper.
-What would happen if only the author had thought to control for X?
+What would happen if only the first author had thought to control for X?
 Or run one more MCMC chain?
@@
 Will the United Nations (now eager to act based the paper) make a wrong move?
-Time to go back to the analyses run sometime between 2011 and 2020 to check.
-Easy.
+
+The first author convenes a three way video conference with the other collaborators while commuting home in his flying car.
+The group must go back to the analyses.
+Which ones?
+The ones from 2011?
+Or 2012?
+Or 2018?
+Where are the files?
+The next day, one member of the group finds some of the files.
+Now the checks and re-analyses should be easy.
 Right?
-Collaboration between the past self and the present self should be simple --- after all you are the same person in 2011 and 2020, right?
-And the way Microsoft Word/Stata/SPSS/R/LISREL stores files and the way that your Windows/Mac machine reads and writes them is the same, right?
-And you know exactly which bit of code produced which table and which figure, right?
-And you wrote your code following Nagler's Maxims \citet{nagler1995coding} and being following King's Replication Standard \citet{king1995replication} right?
+The student-now-faculty member is the same person and should remember the reason for those bits of code (or at least should remember which series of mouse clicks were used to produce the numbers for that crucial table).
+And, of course, the way Microsoft Word/Stata/SPSS/R/LISREL stores files and the way that your Windows/Mac machine reads and writes them is the same, right?
+(And Windows and Mac are still relevant operating systems.) And the group knows exactly which bit of code produced which table and which figure, right?
+And they wrote their code following Nagler's Maxims \citet{nagler1995coding} and being following King's Replication Standard \citet{king1995replication} right?
 
-Just in case one has some doubt about answering these questions, this piece aims to amplify some of what we already ought to know \citet{nagler1995coding}, and to update some of those ideas given current practices, platforms, and propensities.
+Just in case one has some doubt about answering these questions, this piece aims to amplify some of what we already ought to know from \citet{nagler1995coding} and \citet{king1995replication}, and to update some of those ideas given current practices, platforms, and propensities.
 
-\vspace{4em}
+\section{Why should we believe each other?}
 
@@
 In addition, these actions make collaboration within research groups more effective (since believeability comes in part from reproducibility and research groups often need to be able to reproduce in part or in whole what different people in the group have done).
-
-Here is some advice that is not currently emphasized but which might be profitably added to previous advice.
 
@@
 An even nicer way to ensure that figures and tables are reproducible or easily modifiable by your older and wiser but more harried self is to use literate programming.
-For example, one may insert the following chunk of code into either a LaTeX or an OpenOffice document, ask R to process the document, and then produce a pdf file in which the code has disappeared to be replaced with an actual figure:
+For example, to produce the Figure~\ref{fig:giniprot}, I insert the following chunk of code into the  LaTeX document, and then I asked R to process the document, and then produce a pdf file in which the code has disappeared to be replaced with an actual figure.\footnote{One can insert the same code chunk in an OpenOffice document and use the odfWeave package to produce OpenOffice documents in which figures and tables replace code.
+One may also use the LyX interface to LaTeX to write LaTeX documents within a WYSIWYG environment.}
 
-<<read.data>>=
+<<read.data,echo=FALSE,results=hide>>=
 
@@
 It must be created using LaTeX (or HTML or OpenOffice or something else that the working group standardizes on).
-And, in fact, it must be created using the reproducible markup system like Sweave (if you are using LaTeX and R as your working group tools).
+And, in fact, it must be created using the reproducible markup system like Sweave (if you are using LaTeX and R as your working group tools) \citep{Leis:2005}.
 Alternatively, you can put comments into your report detailing where the figures and tables came from (i.e.
@@
 In order to know that ``that plot on page 10'' is from a file called ``fig1.pdf'' you need to document it somehow.
-I can imagine a system with very disciplined use of Word/OpenOffice --- perhaps using special auxiliary ``listoffigures.txt'' files or something.
+I can imagine a system with very disciplined use of a word processor (like Open Office or Word) --- perhaps using special auxiliary ``listoffigures.txt'' files or something.
 In my experience, it is much easier to just use LaTeX (which can be used in a very user friendly way via LyX).
@@
   discussion of what version control is.} Many people are familiar
-with ``Track Changes'' feature in OpenOffice/Word/Pages or the fact that
+with ``Track Changes'' feature in modern WYSIWYG word processors or the fact that
 Dropbox allows one to recover previous versions of files. These are
@@
 Google those terms.
-I have experimented with three systems so far: (1) for one paper I simply included my Sweave document and data files into a compressed archive; (2) for another more computing intensive paper...
+Of course the devil is the details: Here I list a few of my own attempts at enabling reproducible research.
+You'll find many other inspiring examples on the web.
 
-[examples: Examples of reprodible research using R vignette system [jake's Manifest Effects paper?], and a more complex system that required more time instensive simulation using a Makefile [jake and ben and marks's JASA archive]?]
+I have experimented with three systems so far: (1) for one paper we simply included a Sweave document and data files into a compressed archive \cite{bowers2005dataverse}; (2) for another more computing intensive paper we assembled a set of files which enabled reproduction of our results using the ``make'' system \citep{bowers2008dataverse}; and (3) recently I have tried the ``compendium'' approach \citep{gentleman2005reproducible,gentleman2007statistical} which embeds an academic paper with the R package system \cite{bowers2011dataverse}.
+The benefit of this last system is that one is not required to have access to unix: one may download, run, and view the paper from within R --- which is platform independent.
 
@@
 
-The time-frames over which collaboration are required --- whether among a group of people working together or within a single scholar's productive life --- are much longer than any given version of any given software will easily exist.
-All except for plain text.
-Thus, even as we extol version control systems, one must have a way to ensure future access to them in a form that will still be around when the cockroaches take over political science departments (and thereby finally make C-x M-x C-c  a reasonable way to instruct Emacs to do something [insert url citing this argument --- that Emacs was really made for cockroaches or other six legged creatures.]
+The time-frames over which collaboration are required --- whether
+among a group of people working together or within a single scholar's
+productive life --- are much longer than any given version of any
+given software will easily exist. All except for plain text. Thus,
+even as we extol version control systems, one must have a way to
+ensure future access to them in a form that will still be around when
+the cockroaches take over political science departments.\footnote{
+  (and thereby finally make Ctrl-a Ctrl-x Esc-x Ctrl-c a reasonable
+  way to instruct Emacs to do something
+  (\url{http://kieran.healy.usesthis.com/} )}
 
 In the end, following these practices allows your analyses of your collections to be credible --- and, in the case of criticism, you can always tell antagonists to go redo your analysis if they really don't believe you.
+Let me end with the most pithy statement in favor of the idea that collaboration is at the heart of our endeavor that I have found:
 
-"[I]f the empiricalbasis for an articleor book cannotbe reproduced,of what use to the discipline are its conclusions? What purposedoes an articlelike this serve? \cite[445]{king1995replication}
+"[I]f the empirical basis for an article or book cannot be reproduced, of what use to the discipline are its conclusions?
+What purpose does an article like this serve?'' \cite[445]{king1995replication}
 
@@
 \end{document}
+<<>>=
+options(prompt="> ",continue="+ ")
+@

### 2011-02-26 workflow.Rnw
Began section on unit testing

@@
 \usepackage{fancyvrb,natbib,url,array,dcolumn,float,booktabs}
+\usepackage{hyperref}
 
@@
 
-\section{Obey Nagler's Maxims}
+\section{Write code for your audience.}
 Read \citet{nagler1995coding}. Comment your code. Luckily, if you are
@@
 After all you will be releasing both documents (probably) publicly upon publication of the article if not before.
+
+\section{Unit Testing}
+
+Imagine that you reported  bootstrap confidence intervals in your famous article.
+Now the statisticians at the UN are worrying about your bootstrap confidence interval.
+You'd like to evaluate your bootstrap procedure.
+Although nice code exists for bootstrapping linear models, no nice code exists to bootstrap the bootstrap.
+Of course, the code required is not complex, but since you are writing your own code you worry about getting it right.
+Now, 9 years after the idea, you've had lots of time to appreciate problems arising from bugs and errors in data analysis code.
+
+Now, if you had a moment to think in between teaching that new class, reading books for the awards committee, reading application files for the admissions committee, staying home with a sick child, and your own current research, you might say to yourself, ``Before I write new code, I should write a test of the code.
+I should write a little bit of code that let's me know that my double-bootstrap procedure actually does what it is supposed to do.''
+
+Of course, this idea, like most others is not new.
+When large groups of programmers write code for multi-million dollar programs the question about avoiding error looms large.
+The idea of \href{http://en.wikipedia.org/wiki/Test-driven_development}{test driven development} and the idea that one ought to create tests of \href{http://en.wikipedia.org/wiki/Unit_testing}{small parts of one's code} arose to address such concerns.
+For the social scientist collaborating with her future self and/or a small group of collaborators.
+
+Here is an example of this idea in a very simple form.
+I want to write a function to multiply a number by 2.
+If my function works, when I give it the number 4, I should see it return the number 8 and when I give it -4, I should get -8.
+
+<<unit.test>>=
+times.2.fn<-function(thenumber){
+  thenumber+2
+}
+
+test.times.2.fn<-function(){
+  if(times.2.fn(thenumber=4) == 8 &
+     times.2.fn(thenumber=-4) == -8){
+    print("It works!")
+  }else{print("It does not work!")}
+}
+
+test.times.2.fn()
+@
+
+Ack! I mistyped ``+'' for ``*''. Good thing I wrote the test!
 

### 2011-02-27 workflow.Rnw
Added vc files. A few more edits. Still not a first draft.

@@
 
-\title{A paper is to write}
+\title{A Few Easy Pieces: \\ How to enhance your relationship with your future self.}
 
-\author{Jake Bowers\thanks{Many thanks to Kevin Quinn and Cara Wong for direct help on this document, Mika LaVaque-Manty, Ben Hansen, and Mark Fredrickson for many useful discussions on this topic, and Kieran Healy for inspiration from afar.}}
+\author{Jake Bowers\thanks{Many thanks to Kevin Quinn and Cara Wong for direct help on this document, Mika LaVaque-Manty, Ben Hansen, and Mark Fredrickson for many useful discussions on this topic, and Kieran Healy for inspiration from afar.
+The source code for this document may be freely downloaded from \url{https://github.com/jwbowers/workflow}.}}
 
@@
         scipen=8,
-        width=100
+        width=100,
+       error=function(){options(prompt="> ",continue="+ ");NULL}
         )
@@
 Right?
-The student-now-faculty member is the same person and should remember the reason for those bits of code (or at least should remember which series of mouse clicks were used to produce the numbers for that crucial table).
-And, of course, the way Microsoft Word/Stata/SPSS/R/LISREL stores files and the way that your Windows/Mac machine reads and writes them is the same, right?
-(And Windows and Mac are still relevant operating systems.) And the group knows exactly which bit of code produced which table and which figure, right?
+The student-now-faculty member is the same person and should remember the reason for those bits of code (or at least should remember which series of mouse clicks were used to produce the numbers for that crucial table --- and those mouse clicks should not be time consuming to redo exactly as they were done in 2011, or was it 2015?).
+And, of course, the way Microsoft Word/Stata/SPSS/R/LISREL stores files and the way that your machine reads and writes them is the same, right --- since Windows and Mac OS X have always existed and will always continue to exist more or less as they current exist.
+And the group knows exactly which bit of code produced which table and which figure, right?
 And they wrote their code following Nagler's Maxims \citet{nagler1995coding} and being following King's Replication Standard \citet{king1995replication} right?
@@
 
-\section{Why should we believe each other?}
-
-When we call some piece of work ``scientific'' we are talking about how it was done.
-By calling it ``scientific'' we are also using a short-hand to suggest that (1) the findings of the work are not a matter of opinion and (2) other people could reproduce the findings.
-That is, what the work represents is not a matter of opinion but is a shared experience --- and an experience shared without respect to the identities of others (although requiring some shared technical training and research resources).
-This short-hand means that ``scientific'' work should make us change how we act --- or at least, it ought to stand on stronger epistemological ground than other claims about experiences which ought to be shared or shareable.
-
-Assume we want other scientists to believe us when we say something.
-More narrowly, assume we want other people to believe us when we say something about data: ``data'' here can be words, numbers, musical notes, images, ideas, etc \ldots The point is that we are making some claims about patterns that we see in some collection of stuff.
-Now, it might be easy to convince other people that ``this collection of stuff is different from this collection of stuff'' if those people were looking over our shoulders the whole time that we made decisions about collecting the stuff and analyzed it [where ``analyze'' means something like ``break it up into understandable parts''].
-Unfortunately, we can't assume that people are willing to shadow a researcher throughout her career.
-Rather, we do our work alone or in small groups and want to convince other distant people about our analyses.
-
-Now, say your collections of stuff are large or complex and where your chosen tools of analyses are computer programs.
-How can we convince people that what we did with some data with some program is credible: not a matter of whim or opinion, and reproducible by others who didn't shadow us as we wrote our papers?
-Here are a few concrete actions that people can take to enhance the believeability of their work.
-In addition, these actions make collaboration within research groups more effective (since believeability comes in part from reproducibility and research groups often need to be able to reproduce in part or in whole what different people in the group have done).
-
-\section{Show me the data.}
+\section{Meaningful code requires data.}
 All files containing commands operating on data must refer to a data file.
@@
 I have seen command files which do not begin with a call to load data and I have despaired.
-What is the meaning of \Verb+please-fit(var1 with var2)+ if I do not know where var1 var2 come from.\footnote{The command \Verb+please-fit+ comes from the MayIPleaseDoStatistics package for R which emphasizes politeness in data analysis.
+What is the meaning of \Verb+please-fit(var1 with var2)+ if I do not know where var1 var2 come from.\footnote{The command \Verb+please-fit+ and some other R functions used in this essay come from the MayIPleaseDoStatistics package for R which emphasizes politeness in data analysis.
 It is highly recommended for use with toddlers.}
 
-\section{Code not Clicking}
-All tables and graphs should be as much as possible generated from code and not clicking (let alone copying and pasting).
+\section{Type, don't click.}
+All tables and graphs should be generated from code and not clicking (let alone copying and pasting).
 If I wanted to re-create the figure you created but with red lines, and only for people in the South, I should be able to do so with just a few edits to the code.
@@
 
-\begin{verbatim}
+\begin{Verbatim}[fontsize=\small]
 pdf('fig1.pdf')
@@
 dev.off()
-\end{verbatim}
+\end{Verbatim}
 
@@
 In the future, if I wanted to quickly change fig1.pdf, I could edit the commands quickly and easily to do so.
-[Another good practice is to name the files descriptively, so that we don't in fact have ``fig1.pdf'' but ``political participation by education in the south.pdf'' although such a name may prove burdensome to operating systems which do not care for spaces in file names or to folks who are used to typing full filenames rather than using command line completion facilities in modern shells.]
+Another good practice is to name the files descriptively, so that we don't in fact have ``fig1.pdf'' but ``political participation by education in the south.pdf'' although such a name may prove burdensome to operating systems which do not care for spaces in file names or to folks who are used to typing full filenames rather than using command line completion facilities in modern shells.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
   \begin{center}
@@
 @
-    \caption{Protest activity by income inequality from \citep{norris2009data}.}\label{fig:giniprot}
+    \caption{Protest activity by income inequality \citep[from]{norris2009data}.}\label{fig:giniprot}
   \end{center}
@@
 
-\section{Fishing rods and not fish}
+\section{Enjoy literate data analysis.}
 The source code of any pdf document exchanged by the group must be available and executable.
@@
 The Sweave system is even nicer because you can have a LaTeX file with R code right inside of it!
-I showed Sweave with plotting above so here is an example in which a regression table is produced automagically:
+I showed Sweave with plotting in Figure~\ref{fig:giniplot}, those who view the code for this essay will see how the Table~\ref{tab:protest} was also generated directly from a regression object.
 
-<<maketable, echo=FALSE, results=tex>>=
-##source("http://artsci.wustl.edu/~jgill/Models/graph.summary.R")
+<<maketable, echo=FALSE, results=hide>>=
+lm1<-lm(protac2000~I(Gini2004/100)+meanpr,data=good.df)  #  Run the regression
 
-lm1<-lm(protac2000~I(Gini2004/100)+meanpr,data=good.df)
-
+##  The next lines extract relevant and useful output from the
+##  regression object, following Neal Beck's TPM article 2010.
 themat<-cbind(summary(lm1)$coef[,c("Estimate","Std. Error")],confint(lm1))
 n<-length(lm1$fitted.values)
-est.sigma<-round(summary(lm1)$sigma,3)
-library(xtable)
-
+r2<-signif(summary(lm1)$r.squared,2)
+est.sigma<-signif(summary(lm1)$sigma,2)
 colnames(themat)[3:4]<-c("~{}","~{}")
 row.names(themat)<-c("Intercept","Income Inequality (lower=more equal)","Mean Political Rights (lower=more rights)")
+
+library(xtable)
 thextab<-xtable(themat,align=c("l",rep(".",4)),
@@
   rights as measured by Freedom House 1980--2000. Data from \\citep{norris2009data}.",label="tab:protest") 
+@
 
+<<printtable,echo=FALSE,results=tex>>= 
 print(thextab,include.colnames=FALSE,
       hline.after=NULL,
-      add.to.row=list(pos=list(0,3),
-                command=c("\\toprule \n & \\multicolumn{1}{c}{Coef} & \\multicolumn{1}{c}{Std. Err.} 
-& \\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ", 
-                  paste("& \\multicolumn{4}{c}{n: ",n,",   $\\hat{\\sigma}$:",est.sigma,"} \\\\ \\bottomrule"))
+      table.placement="H",
+      add.to.row=list(
+        pos=list(0,3),
+        command=c(
+                  "\\toprule \n &
+ \\multicolumn{1}{c}{Coef} &
+ \\multicolumn{1}{c}{Std. Err.} &
+ \\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ", 
+                  paste("& \\multicolumn{4}{c}{n: ",n,",","   resid. sd:",signif(est.sigma),", R$^2$:",r2,"} \\\\ \\bottomrule",sep="")
                   )
-      )
+        ))
 @
@@
 
-\section{Enabling revision. Preventing clobbering.}
+\section{Enabling revision to prevent clobbering.}
 
@@
   discussion of what version control is.} Many people are familiar
-with ``Track Changes'' feature in modern WYSIWYG word processors or the fact that
+with the ``Track Changes'' feature in modern WYSIWYG word processors or the fact that
 Dropbox allows one to recover previous versions of files. These are
@@
 to merely rename your files with the date and time of saving them:
-thedoc.tex becomes thedoc25092009-22:40.tex.
+thedoc.tex becomes thedoc25-12-2011-23:50.tex.  Be sure to include
+year in the file names --- remember, the life of an idea is measured
+in years. Also, if you use this method, spend a little extra time to
+ensure that you do not clobber files when you make typos in the
+file-names.
+
 If you are wise enough to have saved your documents as plain text (Such as LaTeX source (with or without R/Stata code chunks) \footnote{A quick Google search of ``Sweave for Stata'' turned up lots of resources for literate programming with Stata.} then you can easily compare documents using the many utilities available for comparing text files [FileMerge on OS X is pretty, but ediff and diff for Unix are very useful, I am sure that Windows has many other options].
@@
 
-\begin{Verbatim}
-<<reg1>> =
-##Fitting the regression with x/1000 for ease of interpretation
-thelm<-lm(y~I(x/1000)+z,data=fabulous.df)
-##Fits using a quadratic term as suggested by Blah(1865) did not enhance the fit of the model
-##thelmsq<-lm(y~I(x/1000)^2+z,data=fabulous.df)
+<<reg1,echo=TRUE,results=verb>>=
+##  Repeat the regression run for Table~\ref{tab:protest} [use a label instead
+##  of a number to make moving sections and tables easier]
+##  Do political rights predict protest reporting?
+lm1<-lm(protac2000 ~ I(Gini2004/100) + meanpr, good.df)
+##  Q: Was it worthwhile controlling for political rights?
+lm2<-lm(protac2000~I(Gini2004/100),data=good.df)
+coef(lm2)[2]-coef(lm1)[2]
+##  A: Yes (around 3/4 an act of protest difference)
 @
-\end{Verbatim}
 
-If you need to sleuth down something more than a small detail, but yet the sleuthing is really a footnote to the article, then you can create another file to explore and report on those questions for the research group.
+If you need to track down something more than a small detail, but yet the sleuthing is really a footnote to the article, then you can create another file to explore and report on those questions for the research group.
 The writing in such a memo can be more informal, but the norms of reproducibility ought to be the same as for the main paper.
@@
 collaborators are less likely to call you at midnight asking for help
-if your files are named ``regressions.R'' than if your files are
-called ``temp9'' or ``supercalifragilisticexpialidocious'' (note the
+if your files are named ``inequality-and-protest-analyses.R'' than if your files are
+called ``temp9'' or ``supercalifragilisticexpialidocious'' (Note the
 use of the extension .R to tell us that the file contains R
-commands. use extensions like this as a standard practice to help you
+commands. Use extensions like this as a standard practice to help you
 and your computer get along.)
-
-(other stuff I've forgotten?)
 
@@
 I have experimented with three systems so far: (1) for one paper we simply included a Sweave document and data files into a compressed archive \cite{bowers2005dataverse}; (2) for another more computing intensive paper we assembled a set of files which enabled reproduction of our results using the ``make'' system \citep{bowers2008dataverse}; and (3) recently I have tried the ``compendium'' approach \citep{gentleman2005reproducible,gentleman2007statistical} which embeds an academic paper with the R package system \cite{bowers2011dataverse}.
-The benefit of this last system is that one is not required to have access to unix: one may download, run, and view the paper from within R --- which is platform independent.
+The benefit of this last system is that one is not required to have access to unix: the compendium is downloadable from within R using \Verb+install.packages()+ and it viewable using the \Verb+vignette()+ function.
 
@@
 We all always collaborate.
-Many of us collaborate with groups of people at one moment in time as we race to produce a paper in time for some deadline.
+Many of us collaborate with groups of people at one moment in time as we race against a deadline.
 All of us collaborate with ourselves over time.
@@
 So, I collaborate with a self 5 or 10 or even more years hence when I write a paper.
-[Ask anyone who has turned their dissertation into a book how long it took from them from dissertation prospectus to physical published book for an example of a common time span in which self-collaboration is required.]
+[What is a reasonable time-span for which to plan for self-collaboration on a single idea?
+Ask your advisers how long it took them from idea to dissertation to book (articles).]
 
@@
 productive life --- are much longer than any given version of any
-given software will easily exist. All except for plain text. Thus,
+given software will easily exist. Plain text is the exception. Thus,
 even as we extol version control systems, one must have a way to
@@
 
+Now, imagine that the UN never get's wind of your work, or even that your article does not spawn debate.
+Why then would you spend time to communicate with your future self and others?
+My own answer to this question is because I want my work to be credible and useful.
+What I report in my data analyses should have two main characteristics: (1) the findings of the work should not be a matter of opinion and (2) other people could reproduce the findings.
+That is, what the work represents is not a matter of opinion but is a shared experience --- and an experience shared without respect to the identities of others (although requiring some shared technical training and research resources).
+Such work should make us change how we act --- or at least, it ought to stand on stronger epistemological ground than other claims about experiences which ought to be shared or shareable.
+
+Assume we want others to believe us when we say something.
+More narrowly, assume we want other people to believe us when we say something about data: ``data'' here can be words, numbers, musical notes, images, ideas, etc \ldots The point is that we are making some claims about patterns that we see in some collection of stuff.
+Now, it might be easy to convince other people that ``this collection of stuff is different from this collection of stuff'' if those people were looking over our shoulders the whole time that we made decisions about collecting the stuff and analyzed it [where ``analyze'' means something like ``break it up into understandable parts''].
+Unfortunately, we can't assume that people are willing to shadow a researcher throughout her career.
+Rather, we do our work alone or in small groups and want to convince other distant people about our analyses.
+
+Now, say your collections of stuff are large or complex and where your chosen tools of analyses are computer programs.
+How can we convince people that what we did with some data with some program is credible: not a matter of whim or opinion, and reproducible by others who didn't shadow us as we wrote our papers?
+Here are a few concrete actions that people can take to enhance the believeability of their work.
+In addition, these actions make collaboration within research groups more effective (since believeability comes in part from reproducibility and research groups often need to be able to reproduce in part or in whole what different people in the group have done).
+
 In the end, following these practices allows your analyses of your collections to be credible --- and, in the case of criticism, you can always tell antagonists to go redo your analysis if they really don't believe you.
-Let me end with the most pithy statement in favor of the idea that collaboration is at the heart of our endeavor that I have found:
+Let me end with a pithy statement in favor of the idea that collaboration is at the heart of our endeavor:
 

### 2011-02-27 workflow.Rnw
Version sent to Wendy and Brian and others for comments.

@@
 \documentclass[12pt]{article}
-\usepackage{fancyvrb,natbib,url,array,dcolumn,float,booktabs}
+\usepackage{fancyvrb,natbib,url,array,dcolumn,float,booktabs,listings}
 \usepackage{hyperref}
@@
 
-An idea is born in a coffee shop, a seminar, a quiet walk. It is a cold winter in 2011, but the idea energizes the student. The idea is why he went to graduate school.
+An idea is born in a coffee shop, a seminar, a quiet walk. It is a
+cold winter in 2011, but the idea energizes the student. The idea is
+why he went to graduate school.
 
-This idea inspires a seminar paper in the spring.
-And a conference paper arises from the seminar paper in collaboration with another student in 2012.
-A dissertation chapter descends from the conference paper in 2013.
-Other dissertation chapters take up 2014.
-A submission to a journal with the original co-author and a new collaborator happens in 2015.
-Revision and resubmission wait until 2017 while harried editors, reviewers and authors strive to balance teaching, service, and life.
-The three collaborators have luckily found work in three different universities around the world.
-In 2018 a child is born and a paper is published.
-In 2020 a graduate student in a coffee shop reads the paper and has an idea that challenges the analysis in the paper.
-What would happen if only the first author had thought to control for X?
-Or run one more MCMC chain?
-Or chosen a different likelihood function?
-Will the United Nations (now eager to act based the paper) make a wrong move?
+This idea inspires a seminar paper in the spring. And a conference
+paper arises from the seminar paper in collaboration with another
+student in 2012. A dissertation chapter descends from the conference
+paper in 2013. Other dissertation chapters take up 2014. A submission
+to a journal with the original co-author and a new collaborator
+happens in 2015. Revision and resubmission wait until 2017 while
+harried editors, reviewers and authors strive to balance teaching,
+service, and life. The three collaborators have luckily found work in
+three different universities around the world. In 2018 a child is born
+and a paper is published. In 2020 a graduate student in a coffee shop
+reads the paper and has an idea that challenges the analysis in the
+paper. What would happen if only the first author had thought to
+control for X? Or run one more MCMC chain? Or chosen a different
+likelihood function? Will the United Nations (now eager to act based
+                                              the paper) make a wrong move?
 
-The first author convenes a three way video conference with the other collaborators while commuting home in his flying car.
-The group must go back to the analyses.
-Which ones?
-The ones from 2011?
-Or 2012?
-Or 2018?
-Where are the files?
-The next day, one member of the group finds some of the files.
-Now the checks and re-analyses should be easy.
-Right?
-The student-now-faculty member is the same person and should remember the reason for those bits of code (or at least should remember which series of mouse clicks were used to produce the numbers for that crucial table --- and those mouse clicks should not be time consuming to redo exactly as they were done in 2011, or was it 2015?).
-And, of course, the way Microsoft Word/Stata/SPSS/R/LISREL stores files and the way that your machine reads and writes them is the same, right --- since Windows and Mac OS X have always existed and will always continue to exist more or less as they current exist.
-And the group knows exactly which bit of code produced which table and which figure, right?
-And they wrote their code following Nagler's Maxims \citet{nagler1995coding} and being following King's Replication Standard \citet{king1995replication} right?
+The first author convenes a three way video conference with the other
+collaborators while commuting home in his flying car. The group must
+go back to the analyses. Which ones? The ones from 2011? Or 2012? Or
+2018? Where are the files? The next day, one member of the group finds
+some of the files. Now the checks and re-analyses should be
+easy. Right? The student-now-faculty member is the same person and
+should remember the reason for those bits of code (or at least should
+                                                   remember which series of mouse clicks were used to produce the numbers
+                                                   for that crucial table --- and those mouse clicks should not be time
+                                                   consuming to redo exactly as they were done in 2011, or was it
+                                                   2015?). And, of course, the way Microsoft Word/Stata/SPSS/R/LISREL
+stores files and the way that your machine reads and writes them is
+the same, right --- since Windows and Mac OS X have always existed and
+will always continue to exist more or less as they current exist. And
+the group knows exactly which bit of code produced which table and
+which figure, right? And they wrote their code following Nagler's
+Maxims \citet{nagler1995coding} and being following King's Replication
+Standard \citet{king1995replication} right?
 
-Just in case one has some doubt about answering these questions, this piece aims to amplify some of what we already ought to know from \citet{nagler1995coding} and \citet{king1995replication}, and to update some of those ideas given current practices, platforms, and propensities.
+Just in case one has some doubt about answering these questions, this
+piece aims to amplify some of what we already ought to know from
+\citet{nagler1995coding} and \citet{king1995replication}, and to
+update some of those ideas given current practices, platforms, and
+propensities.
 
 \section{Meaningful code requires data.}
-All files containing commands operating on data must refer to a data file.
-A reference to a data file is a line of code the analysis program will use to operate on (``load''/ ``open'' / ``get'' / ``use'') the data file.
-One should not have to edit this line on different computers or platforms in order to execute this command.
-Using R, for example, all analysis files should have load('thedata.rda') or \Verb+read.csv('http://www.mywebsite.org/Data/thedata.csv')+ or some equivalent line in them, and 'thedata.csv' should be stored in some place easy to find (like in the same directory as the file or perhaps in 'Data/thedata.rda').
 
-Where should you store the data files?
-A very obvious solution is to always make sure that the data file used by a command file is in the same directory as the command file.
-More elegant solutions (pointed to below) require all co-authors to have the same directory structure so that load('Data/thedata.rda') means the same thing on all computers used to work on said project.
+All files containing commands operating on data must refer to a data
+file. A reference to a data file is a line of code the analysis
+program will use to operate on (``load''/ ``open'' / ``get'' /
+                                ``use'') the data file. One should not have to edit this line on
+different computers or platforms in order to execute this command.
+Using R, for example, all analysis files should have
+load('thedata.rda') or
+\Verb+read.csv('http://www.mywebsite.org/Data/thedata.csv')+ or some
+equivalent line in them, and 'thedata.csv' should be stored in some
+place easy to find (like in the same directory as the file or perhaps
+                    in 'Data/thedata.rda').
 
-The principle of modularity \cite{nagler1995coding} suggests that you separate data cleaning, processing, recoding, and merging from analysis in different files.
-So, perhaps your analysis oriented files will load('cleandata.rda') or something but with a comment in the code telling the future you (among others) that cleandata.rda was created from create-cleandata.R which in turn begins with read.csv('dirtydata.csv') which, from a comment in the file, was downloaded from \url{http://www.greatfreedata.gov/thedata?dirtydata.csv} on April 1, 2011.
-Such a data processing file will typically end with something like \Verb+save('cleandata.rda')+ so that we are doubly certain about the provenance of the data.
+Where should you store the data files? A very obvious solution is to
+always make sure that the data file used by a command file is in the
+same directory as the command file. More elegant solutions (pointed to
+                                                            below) require all co-authors to have the same directory structure so
+that load('Data/thedata.rda') means the same thing on all computers
+used to work on said project.
 
-Now, if in 5 years we wonder where 'cleandata.rda' came from, we might \Verb+ grep cleandata *.R+ to search for occurrences of this file on our Unix/OS X system.
-However, if such searching among files is a burden, and even nicer solution is to maintain a file for each project called ``MANIFEST.txt'' or ``INDEX.txt'' or ``README.txt'' which lists the data and command files with brief descriptions of their functions and relations.
+The principle of modularity \cite{nagler1995coding} suggests that you
+separate data cleaning, processing, recoding, and merging from
+analysis in different files. So, perhaps your analysis oriented files
+will load('cleandata.rda') or something but with a comment in the code
+telling the future you (among others) that cleandata.rda was created
+from create-cleandata.R which in turn begins with
+read.csv('dirtydata.csv') which, from a comment in the file, was
+downloaded from
+\Verb+http://www.greatfreedata.gov/dirtydata.csv+ on April 1,
+2011. Such a data processing file will typically end with something
+like \Verb+save('cleandata.rda')+ so that we are doubly certain about
+the provenance of the data.
 
-The principle regarding data is to know where the data came from and what operations were performed on which set of data.
-I have seen command files which do not begin with a call to load data and I have despaired.
-What is the meaning of \Verb+please-fit(var1 with var2)+ if I do not know where var1 var2 come from.\footnote{The command \Verb+please-fit+ and some other R functions used in this essay come from the MayIPleaseDoStatistics package for R which emphasizes politeness in data analysis.
-It is highly recommended for use with toddlers.}
+Now, if in 5 years we wonder where 'cleandata.rda' came from, we might
+\Verb+ grep cleandata *.R+ to search for occurrences of this file on
+our Unix/OS X system. However, if such searching among files is a
+burden, and even nicer solution is to maintain a file for each project
+called ``MANIFEST.txt'' or ``INDEX.txt'' or ``README.txt'' which lists
+the data and command files with brief descriptions of their functions
+and relations.
+
+The principle regarding data is to know where the data came from and
+what operations were performed on which set of data. I have seen
+command files which do not begin with a call to load data and I have
+despaired. What is the meaning of \Verb+please-fit(var1 with var2)+ if
+I do not know where var1 var2 come from.\footnote{The command
+                                                  \Verb+please-fit+ and some other R functions used in this essay come
+                                                  from the \Verb+MayIPleaseDoStatistics+ package which emphasizes
+                                                  politeness in data analysis. The commands \Verb+please-fit+ and
+                                                  \Verb+please-plot+ can be blocked and \Verb+would-you-please-fit+ or
+                                                  \Verb+may-I-please-have-a-plot+ can be required using
+                                                  \Verb+options(politeness=99)+ }
 
 \section{Type, don't click.}
-All tables and graphs should be generated from code and not clicking (let alone copying and pasting).
-If I wanted to re-create the figure you created but with red lines, and only for people in the South, I should be able to do so with just a few edits to the code.
 
-Using R, for example, I might specify that the file \Verb+fig1.pdf+ was produced by the command, where the \Verb+please-plot+ function was loaded elsewhere.
+All tables and graphs should be generated from code and not clicking
+(let alone copying and pasting). If I wanted to re-create the figure
+you created but with red lines, and only for people in the South, I
+should be able to do so with just a few edits to the code.
+
+Using R, for example, I might specify that the file \Verb+fig1.pdf+
+  was produced by the command, where the \Verb+please-plot+ function was
+loaded elsewhere.
 
@@
 
-Now, in the future if I wonder how ``that plot on page 10'' was created, if I suitably commented my manuscript or inserted text into my MANIFEST.txt file I will know (1) ``that plot'' is from a file called ``fig1.pdf'' and (2) fig1.pdf was created in \Verb+figs1and2.R+.
-In the future, if I wanted to quickly change fig1.pdf, I could edit the commands quickly and easily to do so.
-Another good practice is to name the files descriptively, so that we don't in fact have ``fig1.pdf'' but ``political participation by education in the south.pdf'' although such a name may prove burdensome to operating systems which do not care for spaces in file names or to folks who are used to typing full filenames rather than using command line completion facilities in modern shells.
+Now, in the future if I wonder how ``that plot on page 10'' was
+created, if I suitably commented my manuscript or inserted text into
+my MANIFEST.txt file I will know (1) ``that plot'' is from a file
+called ``fig1.pdf'' and (2) fig1.pdf was created in
+\Verb+figs1and2.R+. In the future, if I wanted to quickly change
+fig1.pdf, I could edit the commands quickly and easily to do
+so. Another good practice is to name the files descriptively, so that
+we don't in fact have ``fig1.pdf'' but ``political participation by
+education in the south.pdf'' although such a name may prove burdensome
+to operating systems which do not care for spaces in file names or to
+folks who are used to typing full filenames rather than using command
+line completion facilities in modern shells.
 
-An even nicer way to ensure that figures and tables are reproducible or easily modifiable by your older and wiser but more harried self is to use literate programming.
-For example, to produce the Figure~\ref{fig:giniprot}, I insert the following chunk of code into the  LaTeX document, and then I asked R to process the document, and then produce a pdf file in which the code has disappeared to be replaced with an actual figure.\footnote{One can insert the same code chunk in an OpenOffice document and use the odfWeave package to produce OpenOffice documents in which figures and tables replace code.
-One may also use the LyX interface to LaTeX to write LaTeX documents within a WYSIWYG environment.}
+An even nicer way to ensure that figures and tables are reproducible
+or easily modifiable by your older and wiser but more harried self is
+to use literate programming. For example, to produce the
+Figure~\ref{fig:giniprot}, I insert the following chunk of code into
+the LaTeX document, and then I asked R to process the document, and
+then produce a pdf file in which the code has disappeared to be
+replaced with an actual figure.\footnote{One can insert the same code
+  chunk in an OpenOffice document and use the odfWeave package to
+  produce OpenOffice documents in which figures and tables replace
+  code. One may also use the LyX interface to LaTeX to write LaTeX
+  documents within a WYSIWYG environment.}
 
 <<read.data,echo=FALSE,results=hide>>=
-
-##This read.csv line is slightly more complex than needed because of
-##the particular encoding used for characters in the source file.
-
+##  This read.csv line is slightly more complex than needed because of
+##  the particular encoding used for characters in the source file.
 dem.nations.df<-read.csv(url("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv",encoding="ISO_8859-1"),as.is=TRUE)
@@
 
-<<fig1code,echo=TRUE,fig=FALSE,width=3,height=3>>=
-##The code to produce the following figure
-par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
+<<fig1code,echo=TRUE,fig=FALSE>>=
+##  Make a scatterplot of Protest by Inequality
 with(good.df,plot(Gini2004/100,protac2000,
@@
 
+##  Label a few interesting points
 with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
@@
 <<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>=
+par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
 <<fig1code>>
@@
 
-The principle here is to know where tables and figures came from --- to know what commands created which figures and tables.
+The principle here is to know where tables and figures came from ---
+to know what commands created which figures and tables.
 
 \section{Enjoy literate data analysis.}
-The source code of any pdf document exchanged by the group must be available and executable.
-Say you run some command and then make a table or a graph.
-Then you produce a nice little report on your work for use among the working group.
-Eventually we want to use pieces of that report (tables, graphs, paragraphs) in a publishable paper.
-Thus, that report must not have been created using copying and pasting.
-It must be created using LaTeX (or HTML or OpenOffice or something else that the working group standardizes on).
-And, in fact, it must be created using the reproducible markup system like Sweave (if you are using LaTeX and R as your working group tools) \citep{Leis:2005}.
-Alternatively, you can put comments into your report detailing where the figures and tables came from (i.e.
-which command file produced them).
-Or include details about which code produced which figures in your MANIFEST.txt.
 
-In order to know that ``that plot on page 10'' is from a file called ``fig1.pdf'' you need to document it somehow.
-I can imagine a system with very disciplined use of a word processor (like Open Office or Word) --- perhaps using special auxiliary ``listoffigures.txt'' files or something.
-In my experience, it is much easier to just use LaTeX (which can be used in a very user friendly way via LyX).
+The source code of any pdf document exchanged by the group must be
+available and executable. Say you run some command and then make a
+table or a graph. Then you produce a nice little report on your work
+for use among the working group. Eventually we want to use pieces of
+that report (tables, graphs, paragraphs) in a publishable paper. Thus,
+that report must not have been created using copying and pasting. It
+must be created using LaTeX (or HTML or OpenOffice or something else
+that the working group standardizes on). And, in fact, it must be
+created using the reproducible markup system like Sweave (if you are
+using LaTeX and R as your working group tools)
+\citep{Leis:2005}. Alternatively, you can put comments into your
+report detailing where the figures and tables came from (i.e. which
+command file produced them). Or include details about which code
+produced which figures in your MANIFEST.txt.
 
-The Sweave system is even nicer because you can have a LaTeX file with R code right inside of it!
-I showed Sweave with plotting in Figure~\ref{fig:giniplot}, those who view the code for this essay will see how the Table~\ref{tab:protest} was also generated directly from a regression object.
+In order to know that ``that plot on page 10'' is from a file called
+``fig1.pdf'' you need to document it somehow. I can imagine a system
+with very disciplined use of a word processor (like Open Office or
+Word) --- perhaps using special auxiliary ``listoffigures.txt'' files
+or something. In my experience, it is much easier to just use LaTeX
+(which can be used in a very user friendly way via LyX).
+
+The Sweave system is even nicer because you can have a LaTeX file with
+R code right inside of it! I showed Sweave with plotting in
+Figure~\ref{fig:giniprot}, those who view the code for this essay will
+see how the Table~\ref{tab:protest} was also generated directly from a
+regression object.
 
@@
              \\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ",
-                  paste("& \\multicolumn{4}{c}{n: ",n,",","   resid. sd:",signif(est.sigma),", R$^2$:",r2,"} \\\\ \\bottomrule",sep="")
+            paste("& \\multicolumn{4}{c}{n: ",n,",",
+                  "   resid.sd:",signif(est.sigma),
+                  ", R$^2$:",r2,"} \\\\ \\bottomrule",sep="")
             )
@@
 
-Right now it is hard to beat Sweave as a system for enhancing the collaboration of research groups doing quantitative work in political science.
-Something will make our lives easier than Sweave (or odfWeave) in the future.
-Then we'll change.
+Right now it is hard to beat Sweave as a system for enhancing the
+collaboration of research groups doing quantitative work in political
+science. Something will make our lives easier than Sweave (or
+                                                           odfWeave) in the future. Then we'll change.
 
@@
   discussion of what version control is.} Many people are familiar
-with the ``Track Changes'' feature in modern WYSIWYG word processors or the fact that
-Dropbox allows one to recover previous versions of files. These are
-both kinds of version control. When collaborating with yourself or
-others, it is useful to see what has changed, to feel free to
-experiment and then to dump the experiment in favor of previous work,
-to have multiple ``releases'' of the same document (one to MPSA, one
-to APSR, one to your parents) without requiring that spawn many copies
-of the same document and risk confusion and clobbering. Clobbering is
-what happens when your future self or your other collaborators saves
-an old version of a file over a new version --- erasing good work by
-accident.
+with the ``Track Changes'' feature in modern WYSIWYG word processors
+or the fact that Dropbox allows one to recover previous versions of
+files. These are both kinds of version control. When collaborating
+with yourself or others, it is useful to see what has changed, to feel
+free to experiment and then to dump the experiment in favor of
+previous work, to have multiple ``releases'' of the same document (one
+to MPSA, one to APSR, one to your parents) without requiring that
+spawn many copies of the same document and risk confusion and
+clobbering. Clobbering is what happens when your future self or your
+other collaborators saves an old version of a file over a new version
+--- erasing good work by accident.
 
@@
 
-% (the bazaar system appears particularly well suited for use with DropBox http://marcioandreyoliveira.blogspot.com/2009/05/get-your-own-bazaar-hosting-service-for.html)
-
 We use subversion with our own research group, and I use it for all of
@@
 for us. However, the price of such hosting may not be that much using
-one of the many webhosting services out there.
+one of the many webhosting services out there. [See Y in this issue
+for much more about version control and collaboration.]
 
@@
 
-If you are wise enough to have saved your documents as plain text (Such as LaTeX source (with or without R/Stata code chunks) \footnote{A quick Google search of ``Sweave for Stata'' turned up lots of resources for literate programming with Stata.} then you can easily compare documents using the many utilities available for comparing text files [FileMerge on OS X is pretty, but ediff and diff for Unix are very useful, I am sure that Windows has many other options].
-Adobe Acrobat allows one to compare differences in pdf files.
-OpenOffice supports a ``Compare Documents'' option.
+If you are wise enough to have saved your documents as plain text
+(Such as LaTeX source (with or without R/Stata code chunks)
+ \footnote{A quick Google search of ``Sweave for Stata'' turned up lots
+           of resources for literate programming with Stata.} then you can
+ easily compare documents using the many utilities available for
+ comparing text files [FileMerge on OS X is pretty, but ediff and diff
+                       for Unix are very useful, I am sure that Windows has many other
+                       options]. Adobe Acrobat allows one to compare differences in pdf
+ files. OpenOffice supports a ``Compare Documents'' option.
 
  When you reach certain milestones you can rename the file accordingly:
-thedocAPSA2009.tex [for the one sent to discussants at APSA] or
-thedocAPSR2015.tex [for the version eventually sent to the APSR six
-years after you presented it at APSA]. The systems I mentioned above
+ thedocAPSA2009.tex --- for the one sent to discussants at APSA --- or
+ thedocAPSR2015.tex --- for the version eventually sent to the APSR six
+ years after you presented it at APSA. The systems I mentioned above
  all allow this kind of thing and are much more elegant and capable,
@@
 
-\section{Write code for your audience.}
+\section{No data analyst is an island.}
+
 Read \citet{nagler1995coding}. Comment your code. Luckily, if you are
@@
 
-Using R within Sweave, for example, the comment is marked using the pound sign.
-Notice that I've left in a regression not run for the paper directly, but run as a part of the due diligence process of writing the paper.
+Using R within Sweave, for example, the comment is marked using the
+pound sign. Notice that I've left in a regression not run for the
+ paper directly, but run as a part of the due diligence process of
+ writing the paper.
 
 <<reg1,echo=TRUE,results=verb>>=
-##  Repeat the regression run for Table~\ref{tab:protest} [use a label instead
-##  of a number to make moving sections and tables easier]
+
+ ##  Repeat the regression run for Table~\ref{tab:protest}
  ##  Do political rights predict protest reporting?
-lm1<-lm(protac2000 ~ I(Gini2004/100) + meanpr, good.df)
+ lm1<-lm(protac2000 ~ I(Gini2004/100) + meanpr, data=good.df)
  ##  Q: Was it worthwhile controlling for political rights?
@@
 
-If you need to track down something more than a small detail, but yet the sleuthing is really a footnote to the article, then you can create another file to explore and report on those questions for the research group.
-The writing in such a memo can be more informal, but the norms of reproducibility ought to be the same as for the main paper.
-After all you will be releasing both documents (probably) publicly upon publication of the article if not before.
+ If you need to track down something more than a small detail, but yet
+ the sleuthing is really a footnote to the article, then you can create
+ another file to explore and report on those questions for the research
+ group. The writing in such a memo can be more informal, but the norms
+ of reproducibility ought to be the same as for the main paper. After
+ all you will be releasing both documents (probably) publicly upon
+ publication of the article if not before.
 
-\section{Unit Testing}
+ Notice one another benefit of coding for an audience: we learn by
+ teaching. By assuming that others will look at your code, you will be
+ more likely to write clearer code, or perhaps even to learn more about
+ what you are doing as you do it.
+
+ \section{Minimize error by testing.}
 
@@
 <<unit.test>>=
-times.2.fn<-function(thenumber){
-  thenumber+2
-}
 
+  ##  The test function:
   test.times.2.fn<-function(){
+    ##  This function tests times.2.fn
     if(times.2.fn(thenumber=4) == 8 &
@@
 
+##  The function:
+times.2.fn<-function(thenumber){
+  ##  This function multiplies a scalar number by 2
+  ##  thenumber is a scalar number
+  thenumber+2
+}
+
+##Use the test function
 test.times.2.fn()
@@
 
-\section{File Naming}
+\section{Take advantage of file names to convey information to your future self.}
 
@@
 
-\section{Exemplars}
+\section{Copy and improve on others' examples.}
 
-Lots of people are thinking about ``reproducible research'' and ``literate programming'' these days.
-Google those terms.
-Of course the devil is the details: Here I list a few of my own attempts at enabling reproducible research.
-You'll find many other inspiring examples on the web.
+Lots of people are thinking about ``reproducible research'' and
+``literate programming'' these days. Google those terms. Of course the
+devil is the details: Here I list a few of my own attempts at enabling
+reproducible research. You'll find many other inspiring examples on
+         the web. Luckily, the open source ethos aligns nicely with academic
+         incentives, so we are beginning to find more and more people offering
+         their files online for copying and improvement. By the way, if you do
+         copy and improve, it is polite to alert the person from whom you made
+         the copy about your work.
 
-I have experimented with three systems so far: (1) for one paper we simply included a Sweave document and data files into a compressed archive \cite{bowers2005dataverse}; (2) for another more computing intensive paper we assembled a set of files which enabled reproduction of our results using the ``make'' system \citep{bowers2008dataverse}; and (3) recently I have tried the ``compendium'' approach \citep{gentleman2005reproducible,gentleman2007statistical} which embeds an academic paper with the R package system \cite{bowers2011dataverse}.
-The benefit of this last system is that one is not required to have access to unix: the compendium is downloadable from within R using \Verb+install.packages()+ and it viewable using the \Verb+vignette()+ function.
+         I have experimented with three systems so far: (1) for one paper we
+         simply included a Sweave document and data files into a compressed
+         archive \cite{bowers2005dataverse}; (2) for another more computing
+         intensive paper we assembled a set of files which enabled reproduction
+         of our results using the ``make'' system \citep{bowers2008dataverse};
+         and (3) recently I have tried the ``compendium'' approach
+         \citep{gentleman2005reproducible,gentleman2007statistical} which
+         embeds an academic paper with the R package system
+         \cite{bowers2011dataverse}. The benefit of this last system is that
+         one is not required to have access to unix: the compendium is
+         downloadable from within R using \Verb+install.packages()+ and it
+         viewable using the \Verb+vignette()+ function. The idea that one ought
+         to be able to install and run and use an academic paper just as one
+         installs and uses statistical software packages is very attractive and
+         I anticipate that it will become ever easier to turn papers into R
+         packages as creative and energetic folks turn their attention to the
+         question of reproducible research.
 
-\section{Constructive Collaboration}
+         \section{Remember that research ought to be credibile communication.}
+         \begin{quote}
+         [I]f the empirical basis for an article or book cannot
+         be reproduced, of what use to the discipline are its conclusions?
+         What purpose does an article like this serve?
+         \cite[445]{king1995replication}
+         \end{quote}
 
-We all always collaborate.
-Many of us collaborate with groups of people at one moment in time as we race against a deadline.
-All of us collaborate with ourselves over time.
-An idealized version of self-collaboration goes something like this: We begin a project with a few notes.
-We write a conference paper.
-We revise this paper and send it to a journal.
-The journal asks for revisions.
-We revise and resubmit.
-Our paper is accepted.
-We revise a bit more.
-The paper is published.
-Later another scholar reacts to our paper.
-We write another paper based on said reactions (draft, conference paper, submission, revision, acceptance, revision).
-These processes occur over years of time.
-So, I collaborate with a self 5 or 10 or even more years hence when I write a paper.
-[What is a reasonable time-span for which to plan for self-collaboration on a single idea?
-Ask your advisers how long it took them from idea to dissertation to book (articles).]
-
+         We all always collaborate. Many of us collaborate with groups of
+         people at one moment in time as we race against a deadline. All of us
+         collaborate with ourselves over time.\footnote{What is a reasonable
+                                                        time-span for which to plan for self-collaboration on a single idea?
+                                                        Ask your advisers how long it took them from idea to dissertation to
+                                                        book (articles).}
          The time-frames over which collaboration are required --- whether
          among a group of people working together or within a single scholar's
-productive life --- are much longer than any given version of any
-given software will easily exist. Plain text is the exception. Thus,
-even as we extol version control systems, one must have a way to
-ensure future access to them in a form that will still be around when
-the cockroaches take over political science departments.\footnote{
-  (and thereby finally make Ctrl-a Ctrl-x Esc-x Ctrl-c a reasonable
-  way to instruct Emacs to do something
-  (\url{http://kieran.healy.usesthis.com/} )}
+productive life or probably both --- are much longer than any given
+version of any given software will easily exist. \href{http://en.wikipedia.org/wiki/Plain_text#Usage}{Plain text} is the
+exception. Thus, even as we extol version control systems, one must
+have a way to ensure future access to them in a form that will still
+be around when sentient cockroaches finally join political science
+departments (by then dominated by cetaceans after humans are mostly
+uploads).\footnote{The arrival of the six-legged social scientists
+  revives Emacs and finally makes Ctrl-a Ctrl-x
+  Esc-x Ctrl-c a \href{http://kieran.healy.usesthis.com/}{reasonable
+    key combination}.}
 
-Now, imagine that the UN never get's wind of your work, or even that your article does not spawn debate.
-Why then would you spend time to communicate with your future self and others?
-My own answer to this question is because I want my work to be credible and useful.
-What I report in my data analyses should have two main characteristics: (1) the findings of the work should not be a matter of opinion and (2) other people could reproduce the findings.
-That is, what the work represents is not a matter of opinion but is a shared experience --- and an experience shared without respect to the identities of others (although requiring some shared technical training and research resources).
-Such work should make us change how we act --- or at least, it ought to stand on stronger epistemological ground than other claims about experiences which ought to be shared or shareable.
+Now, imagine that the UN never hears of your work, or even that your
+article does not spawn debate. Why then would you spend time to
+communicate with your future self and others? My own answer to this
+question is because I want my work to be credible and useful to myself
+and other scholars even if each article does not immediately change
+the world.  What I report in my data analyses should have two main
+characteristics: (1) the findings of the work should not be a matter
+of opinion and (2) other people could reproduce the findings. That is,
+what the work represents is not a matter of opinion but is a shared
+experience --- and an experience shared without respect to the
+identities of others (although requiring some shared technical
+training and research resources).
+% Such work should make us change how
+% we act --- or at least, it ought to stand on stronger epistemological
+% ground than other claims about experiences which ought to be shared or
+% shareable.
 
-Assume we want others to believe us when we say something.
-More narrowly, assume we want other people to believe us when we say something about data: ``data'' here can be words, numbers, musical notes, images, ideas, etc \ldots The point is that we are making some claims about patterns that we see in some collection of stuff.
-Now, it might be easy to convince other people that ``this collection of stuff is different from this collection of stuff'' if those people were looking over our shoulders the whole time that we made decisions about collecting the stuff and analyzed it [where ``analyze'' means something like ``break it up into understandable parts''].
-Unfortunately, we can't assume that people are willing to shadow a researcher throughout her career.
-Rather, we do our work alone or in small groups and want to convince other distant people about our analyses.
+Assume we want others to believe us when we say something. More
+narrowly, assume we want other people to believe us when we say
+something about data: ``data'' here can be words, numbers, musical
+notes, images, ideas, etc \ldots The point is that we are making some
+claims about patterns in some collection of stuff. Now, it
+might be easy to convince others that ``this collection of stuff
+is different from this collection of stuff'' if those people were
+looking over our shoulders the whole time that we made decisions about
+collecting the stuff and broke it up into understandable
+parts and reorganized and summarized it. Unfortunately, we can't assume that people are willing to
+         shadow a researcher throughout her career. Rather, we do our work
+         alone or in small groups and want to convince other distant and future people
+         about our analyses.
 
-Now, say your collections of stuff are large or complex and where your chosen tools of analyses are computer programs.
-How can we convince people that what we did with some data with some program is credible: not a matter of whim or opinion, and reproducible by others who didn't shadow us as we wrote our papers?
-Here are a few concrete actions that people can take to enhance the believeability of their work.
-In addition, these actions make collaboration within research groups more effective (since believeability comes in part from reproducibility and research groups often need to be able to reproduce in part or in whole what different people in the group have done).
+         Now, say your collections of stuff are large or complex and your
+         chosen tools of analyses are computer programs. How can we convince
+         people that what we did with some data with some program is credible:
+           not a matter of whim or opinion, and reproducible by others who didn't
+shadow us as we wrote our papers? This essay has suggested a few
+concrete ways to enhance the believeability of such scholarly work. In
+addition, these actions (as summarized in the section headings of this
+essay) make collaboration within research groups more effective.
+Believeability comes in part from reproducibility and research groups
+often need to be able to reproduce in part or in whole what different
+people in the group have done.
 
-In the end, following these practices allows your analyses of your collections to be credible --- and, in the case of criticism, you can always tell antagonists to go redo your analysis if they really don't believe you.
-Let me end with a pithy statement in favor of the idea that collaboration is at the heart of our endeavor:
-
-"[I]f the empirical basis for an article or book cannot be reproduced, of what use to the discipline are its conclusions?
-What purpose does an article like this serve?'' \cite[445]{king1995replication}
+In the end, following these practices and those recommended by X and Y
+in this issue allows your computerized analyses of your collections
+stuff of to be credible.  Finally, if the UN quibbles with your
+analyses, your future self can shoot the archive required to reproduce
+your work (in still intelligible plain text, analyzed using commented
+code so that folks can translate to whatever system succeeds R, or
+since you used R, you can include a copy of R and all of the R
+packages you used in your final analyses in 2018 in the archive
+itself.) You can say, "Here is everything you need to reproduce my
+work." To be extra helpful you can add "Read the README file for
+futher instructions." And then you can get on with your life: maybe
+the next great idea will occur when your 4 year old asks a wacky
+question after stripping and painting her overly cooperative 1 year
+old brother purple, or teaching a class, or in a coffee shop, or on a
+quiet walk.
 

### 2011-03-14 workflow.Rnw
Colored links
Moving comments earlier in the document.
Subsequent sections refer to commenting code, so I think it makes
sense to move it up in the outline. This change also suggests moving
around other sections.

@@
 \usepackage{fancyvrb,natbib,url,array,dcolumn,float,booktabs,listings}
-\usepackage{hyperref}
+\usepackage[colorlinks=TRUE]{hyperref}
 
@@
 update some of those ideas given current practices, platforms, and
-propensities.
+propensities
+
+%%% This data frame will be used throughout the analyses.
+%%% Putting it up top, so we can use it in the comments section.
+
+<<read.data,echo=FALSE,results=hide>>=
+##  This read.csv line is slightly more complex than needed because of
+##  the particular encoding used for characters in the source file.
+dem.nations.df<-read.csv(url("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv",encoding="ISO_8859-1"),as.is=TRUE)
+
+row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful
+
+good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
+@.
+
+\section{No data analyst is an island.}
+% MMF: My original title before I saw you had you own comment section
+% \section{Comment on your own work. }
+
+Comments, or text inside of a file that is not executed, are a message to your
+future self and other consumers of your work. 
+Read \citet{nagler1995coding}. Comment your code. Luckily, if you are
+using a literate programming practice (Sweave, odfWeave, using R2HTML,
+etc..), you can write paragraphs to surround your code as well a
+technical comments in the code itself. If you are not using a strictly
+literate programming practice then use whatever commenting protocol
+exists in your analysis language. Here are some examples:
+\begin{verbatim}
+# This is a comment in R.
+# With a a second line.
+
+* This is a comment in a Stata .do file
+* You can also embed comments at the ends of lines using /* ... */
+
+/* SAS uses a similar convention to Stata
+   and allows multi-line comments like this.
+*/
+
+% Finally, the percent symbol is a comment in TeX.
+\end{verbatim}
+
+A given analysis of a given collection is a long series of
+decisions. You need to be able to justify them. Often you justify them
+in the paper you are writing (in the text surrounding the R code
+chunks in your Sweave document). Some decisions will be too small and
+technical for inclusion in the paper itself. These need to be
+documented in the code itself.
+
+Using R within Sweave, for example, the comment is marked using the
+pound sign. Notice that I've left in a regression not run for the
+paper directly, but run as a part of the due diligence process of
+writing the paper.
+
+<<reg1,echo=TRUE,results=verb>>=
+
+ ##  Repeat the regression run for Table~\ref{tab:protest}
+ ##  Do political rights predict protest reporting?
+ lm1<-lm(protac2000 ~ I(Gini2004/100) + meanpr, data=good.df)
+ ##  Q: Was it worthwhile controlling for political rights?
+ lm2<-lm(protac2000~I(Gini2004/100),data=good.df)
+ coef(lm2)[2]-coef(lm1)[2]
+ ##  A: Yes (around 3/4 an act of protest difference)
+@
+
+If you need to track down something more than a small detail, but yet
+the sleuthing is really a footnote to the article, then you can create
+another file to explore and report on those questions for the research
+group. The writing in such a memo can be more informal, but the norms
+of reproducibility ought to be the same as for the main paper. After
+all you will be releasing both documents (probably) publicly upon
+publication of the article if not before.
+
+Notice one another benefit of coding for an audience: we learn by
+teaching. By assuming that others will look at your code, you will be
+more likely to write clearer code, or perhaps even to learn more about
+what you are doing as you do it.
+
+Comment liberally. Comments are discarded when R runs analysis or LaTeX turns
+your poignant writings into a PDF, so only those who dig into your work will
+see them. They are footnotes on the process of producing your research.
 
@@
                     in 'Data/thedata.rda').
+It never hurts to drop in a comment about where the data is, if in doubt.
 
@@
   documents within a WYSIWYG environment.}
-
-<<read.data,echo=FALSE,results=hide>>=
-##  This read.csv line is slightly more complex than needed because of
-##  the particular encoding used for characters in the source file.
-dem.nations.df<-read.csv(url("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv",encoding="ISO_8859-1"),as.is=TRUE)
-
-row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful
-
-good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
-@
 
@@
 lot of disk space and having many many ``thedoc...''  files around.
-
-\section{No data analyst is an island.}
-
-Read \citet{nagler1995coding}. Comment your code. Luckily, if you are
-using a literate programming practice (Sweave, odfWeave, using R2HTML,
-etc..), you can write paragraphs to surround your code as well a
-technical comments in the code itself. If you are not using a strictly
-literate programming practice then use whatever commenting protocol
-exists in your analysis language.
-
-A given analysis of a given collection is a long series of
-decisions. You need to be able to justify them. Often you justify them
-in the paper you are writing (in the text surrounding the R code
-chunks in your Sweave document). Some decisions will be too small and
-technical for inclusion in the paper itself. These need to be
-documented in the code itself.
-
-Using R within Sweave, for example, the comment is marked using the
-pound sign. Notice that I've left in a regression not run for the
- paper directly, but run as a part of the due diligence process of
- writing the paper.
-
-<<reg1,echo=TRUE,results=verb>>=
-
- ##  Repeat the regression run for Table~\ref{tab:protest}
- ##  Do political rights predict protest reporting?
- lm1<-lm(protac2000 ~ I(Gini2004/100) + meanpr, data=good.df)
- ##  Q: Was it worthwhile controlling for political rights?
- lm2<-lm(protac2000~I(Gini2004/100),data=good.df)
- coef(lm2)[2]-coef(lm1)[2]
- ##  A: Yes (around 3/4 an act of protest difference)
-@
-
- If you need to track down something more than a small detail, but yet
- the sleuthing is really a footnote to the article, then you can create
- another file to explore and report on those questions for the research
- group. The writing in such a memo can be more informal, but the norms
- of reproducibility ought to be the same as for the main paper. After
- all you will be releasing both documents (probably) publicly upon
- publication of the article if not before.
-
- Notice one another benefit of coding for an audience: we learn by
- teaching. By assuming that others will look at your code, you will be
- more likely to write clearer code, or perhaps even to learn more about
- what you are doing as you do it.
 

### 2011-03-14 workflow.Rnw
moving Type Don't Click up
Order is now type, comment, sweave, data, rest.

@@
 @.
+\section{Type, don't click.}
+
+All tables and graphs should be generated from code and not clicking
+(let alone copying and pasting). If I wanted to re-create the figure
+you created but with red lines, and only for people in the South, I
+should be able to do so with just a few edits to the code.
+
+Using R, for example, I might specify that the file \Verb+fig1.pdf+
+  was produced by the command, where the \Verb+please-plot+ function was
+loaded elsewhere.
+
+\begin{Verbatim}[fontsize=\small]
+pdf('fig1.pdf')
+please-plot(outcome by explanatory)
+dev.off()
+\end{Verbatim}
+
+Now, in the future if I wonder how ``that plot on page 10'' was
+created, if I suitably commented my manuscript or inserted text into
+my MANIFEST.txt file I will know (1) ``that plot'' is from a file
+called ``fig1.pdf'' and (2) fig1.pdf was created in
+\Verb+figs1and2.R+. In the future, if I wanted to quickly change
+fig1.pdf, I could edit the commands quickly and easily to do
+so. Another good practice is to name the files descriptively, so that
+we don't in fact have ``fig1.pdf'' but
+``political-participation-by-education-in-the-south.pdf''. Now, if Figure is
+moved to an appendix, the file name is still meaningful.
+
+% An even nicer way to ensure that figures and tables are reproducible
+% or easily modifiable by your older and wiser but more harried self is
+% to use literate programming. For example, to produce the
+% Figure~\ref{fig:giniprot}, I insert the following chunk of code into
+% the LaTeX document, and then I asked R to process the document, and
+% then produce a pdf file in which the code has disappeared to be
+% replaced with an actual figure.\footnote{One can insert the same code
+%   chunk in an OpenOffice document and use the odfWeave package to
+%   produce OpenOffice documents in which figures and tables replace
+%   code. One may also use the LyX interface to LaTeX to write LaTeX
+%   documents within a WYSIWYG environment.}
+
+<<fig1code,echo=TRUE,fig=FALSE>>=
+##  Make a scatterplot of Protest by Inequality
+with(good.df,plot(Gini2004/100,protac2000,
+                  xlab='Gini Coefficient 2004 (UNDP)',
+                  ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
+                  cex=.8))
+
+##  Label a few interesting points
+with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
+     text(Gini2004/100,protac2000,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
+@
+
+\begin{figure}[H]
+  \begin{center}
+<<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>=
+par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
+<<fig1code>>
+@
+    \caption{Protest activity by income inequality \citep[from]{norris2009data}.}\label{fig:giniprot}
+  \end{center}
+\end{figure}
+
+The principle here is to know where tables and figures came from ---
+to know what commands created which figures and tables.
 
@@
 decisions. You need to be able to justify them. Often you justify them
-in the paper you are writing (in the text surrounding the R code
-chunks in your Sweave document). Some decisions will be too small and
+in the body of the paper you are writing. Some decisions will be too small and
 technical for inclusion in the paper itself. These need to be
@@
 
-Using R within Sweave, for example, the comment is marked using the
+Using R, for example, the comment is marked using the
 pound sign. Notice that I've left in a regression not run for the
@@
 see them. They are footnotes on the process of producing your research.
-
-\section{Meaningful code requires data.}
-
-All files containing commands operating on data must refer to a data
-file. A reference to a data file is a line of code the analysis
-program will use to operate on (``load''/ ``open'' / ``get'' /
-                                ``use'') the data file. One should not have to edit this line on
-different computers or platforms in order to execute this command.
-Using R, for example, all analysis files should have
-load('thedata.rda') or
-\Verb+read.csv('http://www.mywebsite.org/Data/thedata.csv')+ or some
-equivalent line in them, and 'thedata.csv' should be stored in some
-place easy to find (like in the same directory as the file or perhaps
-                    in 'Data/thedata.rda').
-It never hurts to drop in a comment about where the data is, if in doubt.
-
-Where should you store the data files? A very obvious solution is to
-always make sure that the data file used by a command file is in the
-same directory as the command file. More elegant solutions (pointed to
-                                                            below) require all co-authors to have the same directory structure so
-that load('Data/thedata.rda') means the same thing on all computers
-used to work on said project.
-
-The principle of modularity \cite{nagler1995coding} suggests that you
-separate data cleaning, processing, recoding, and merging from
-analysis in different files. So, perhaps your analysis oriented files
-will load('cleandata.rda') or something but with a comment in the code
-telling the future you (among others) that cleandata.rda was created
-from create-cleandata.R which in turn begins with
-read.csv('dirtydata.csv') which, from a comment in the file, was
-downloaded from
-\Verb+http://www.greatfreedata.gov/dirtydata.csv+ on April 1,
-2011. Such a data processing file will typically end with something
-like \Verb+save('cleandata.rda')+ so that we are doubly certain about
-the provenance of the data.
-
-Now, if in 5 years we wonder where 'cleandata.rda' came from, we might
-\Verb+ grep cleandata *.R+ to search for occurrences of this file on
-our Unix/OS X system. However, if such searching among files is a
-burden, and even nicer solution is to maintain a file for each project
-called ``MANIFEST.txt'' or ``INDEX.txt'' or ``README.txt'' which lists
-the data and command files with brief descriptions of their functions
-and relations.
-
-The principle regarding data is to know where the data came from and
-what operations were performed on which set of data. I have seen
-command files which do not begin with a call to load data and I have
-despaired. What is the meaning of \Verb+please-fit(var1 with var2)+ if
-I do not know where var1 var2 come from.\footnote{The command
-                                                  \Verb+please-fit+ and some other R functions used in this essay come
-                                                  from the \Verb+MayIPleaseDoStatistics+ package which emphasizes
-                                                  politeness in data analysis. The commands \Verb+please-fit+ and
-                                                  \Verb+please-plot+ can be blocked and \Verb+would-you-please-fit+ or
-                                                  \Verb+may-I-please-have-a-plot+ can be required using
-                                                  \Verb+options(politeness=99)+ }
-
-\section{Type, don't click.}
-
-All tables and graphs should be generated from code and not clicking
-(let alone copying and pasting). If I wanted to re-create the figure
-you created but with red lines, and only for people in the South, I
-should be able to do so with just a few edits to the code.
-
-Using R, for example, I might specify that the file \Verb+fig1.pdf+
-  was produced by the command, where the \Verb+please-plot+ function was
-loaded elsewhere.
-
-\begin{Verbatim}[fontsize=\small]
-pdf('fig1.pdf')
-please-plot(outcome by explanatory)
-dev.off()
-\end{Verbatim}
-
-Now, in the future if I wonder how ``that plot on page 10'' was
-created, if I suitably commented my manuscript or inserted text into
-my MANIFEST.txt file I will know (1) ``that plot'' is from a file
-called ``fig1.pdf'' and (2) fig1.pdf was created in
-\Verb+figs1and2.R+. In the future, if I wanted to quickly change
-fig1.pdf, I could edit the commands quickly and easily to do
-so. Another good practice is to name the files descriptively, so that
-we don't in fact have ``fig1.pdf'' but ``political participation by
-education in the south.pdf'' although such a name may prove burdensome
-to operating systems which do not care for spaces in file names or to
-folks who are used to typing full filenames rather than using command
-line completion facilities in modern shells.
-
-An even nicer way to ensure that figures and tables are reproducible
-or easily modifiable by your older and wiser but more harried self is
-to use literate programming. For example, to produce the
-Figure~\ref{fig:giniprot}, I insert the following chunk of code into
-the LaTeX document, and then I asked R to process the document, and
-then produce a pdf file in which the code has disappeared to be
-replaced with an actual figure.\footnote{One can insert the same code
-  chunk in an OpenOffice document and use the odfWeave package to
-  produce OpenOffice documents in which figures and tables replace
-  code. One may also use the LyX interface to LaTeX to write LaTeX
-  documents within a WYSIWYG environment.}
-
-<<fig1code,echo=TRUE,fig=FALSE>>=
-##  Make a scatterplot of Protest by Inequality
-with(good.df,plot(Gini2004/100,protac2000,
-                  xlab='Gini Coefficient 2004 (UNDP)',
-                  ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
-                  cex=.8))
-
-##  Label a few interesting points
-with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
-     text(Gini2004/100,protac2000,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
-@
-
-\begin{figure}[H]
-  \begin{center}
-<<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>=
-par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
-<<fig1code>>
-@
-    \caption{Protest activity by income inequality \citep[from]{norris2009data}.}\label{fig:giniprot}
-  \end{center}
-\end{figure}
-
-The principle here is to know where tables and figures came from ---
-to know what commands created which figures and tables.
 
@@
                                                            odfWeave) in the future. Then we'll change.
+
+\section{Meaningful code requires data.}
+
+All files containing commands operating on data must refer to a data
+file. A reference to a data file is a line of code the analysis
+program will use to operate on (``load''/ ``open'' / ``get'' /
+                                ``use'') the data file. One should not have to edit this line on
+different computers or platforms in order to execute this command.
+Using R, for example, all analysis files should have
+load('thedata.rda') or
+\Verb+read.csv('http://www.mywebsite.org/Data/thedata.csv')+ or some
+equivalent line in them, and 'thedata.csv' should be stored in some
+place easy to find (like in the same directory as the file or perhaps
+                    in 'Data/thedata.rda').
+It never hurts to drop in a comment about where the data is, if in doubt.
+
+Where should you store the data files? A very obvious solution is to
+always make sure that the data file used by a command file is in the
+same directory as the command file. More elegant solutions (pointed to
+                                                            below) require all co-authors to have the same directory structure so
+that load('Data/thedata.rda') means the same thing on all computers
+used to work on said project.
+
+The principle of modularity \cite{nagler1995coding} suggests that you
+separate data cleaning, processing, recoding, and merging from
+analysis in different files. So, perhaps your analysis oriented files
+will load('cleandata.rda') or something but with a comment in the code
+telling the future you (among others) that cleandata.rda was created
+from create-cleandata.R which in turn begins with
+read.csv('dirtydata.csv') which, from a comment in the file, was
+downloaded from
+\Verb+http://www.greatfreedata.gov/dirtydata.csv+ on April 1,
+2011. Such a data processing file will typically end with something
+like \Verb+save('cleandata.rda')+ so that we are doubly certain about
+the provenance of the data.
+
+Now, if in 5 years we wonder where 'cleandata.rda' came from, we might
+\Verb+ grep cleandata *.R+ to search for occurrences of this file on
+our Unix/OS X system. However, if such searching among files is a
+burden, and even nicer solution is to maintain a file for each project
+called ``MANIFEST.txt'' or ``INDEX.txt'' or ``README.txt'' which lists
+the data and command files with brief descriptions of their functions
+and relations.
+
+The principle regarding data is to know where the data came from and
+what operations were performed on which set of data. I have seen
+command files which do not begin with a call to load data and I have
+despaired. What is the meaning of \Verb+please-fit(var1 with var2)+ if
+I do not know where var1 var2 come from.\footnote{The command
+                                                  \Verb+please-fit+ and some other R functions used in this essay come
+                                                  from the \Verb+MayIPleaseDoStatistics+ package which emphasizes
+                                                  politeness in data analysis. The commands \Verb+please-fit+ and
+                                                  \Verb+please-plot+ can be blocked and \Verb+would-you-please-fit+ or
+                                                  \Verb+may-I-please-have-a-plot+ can be required using
+                                                  \Verb+options(politeness=99)+ }
 

### 2011-03-15 workflow.Rnw
Took into account Brian Gs comments

@@
 
-\author{Jake Bowers\thanks{Many thanks to Kevin Quinn and Cara Wong for direct help on this document, Mika LaVaque-Manty, Ben Hansen, and Mark Fredrickson for many useful discussions on this topic, and Kieran Healy for inspiration from afar.
-The source code for this document may be freely downloaded from \url{https://github.com/jwbowers/workflow}.}}
+\author{Jake Bowers\thanks{Many thanks to Mark Fredrickson, Brian
+    Gaines, Kieran Healy, Kevin Quinn and Cara Wong for direct help on
+    this document, Mika LaVaque-Manty, Ben Hansen, and Mark
+    Fredrickson for many useful discussions on this topic. The source
+    code for this document may be freely downloaded from
+    \url{https://github.com/jwbowers/workflow}.}}
 
@@
 the same, right --- since Windows and Mac OS X have always existed and
-will always continue to exist more or less as they current exist. And
+will always continue to exist more or less as they currently exist? And
 the group knows exactly which bit of code produced which table and
@@
 
-All tables and graphs should be generated from code and not clicking
+All tables and graphs should be generated from code, not clicking
 (let alone copying and pasting). If I wanted to re-create the figure
@@
 @
-    \caption{Protest activity by income inequality \citep[from]{norris2009data}.}\label{fig:giniprot}
+    \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}\label{fig:giniprot}
   \end{center}
@@
 of reproducibility ought to be the same as for the main paper. After
-all you will be releasing both documents (probably) publicly upon
+all, you will be releasing both documents (probably) publicly upon
 publication of the article if not before.
@@
 Word) --- perhaps using special auxiliary ``listoffigures.txt'' files
-or something. In my experience, it is much easier to just use LaTeX
+or something. In my experience, it is much easier just to use LaTeX
 (which can be used in a very user friendly way via LyX).
@@
 Using R, for example, all analysis files should have
-load('thedata.rda') or
+\Verb+load('thedata.rda')+ or
 \Verb+read.csv('http://www.mywebsite.org/Data/thedata.csv')+ or some
-equivalent line in them, and 'thedata.csv' should be stored in some
+equivalent line in them, and `thedata.csv' should be stored in some
 place easy to find (like in the same directory as the file or perhaps
-                    in 'Data/thedata.rda').
+                    in \Verb+'Data/thedata.rda'+).
 It never hurts to drop in a comment about where the data is, if in doubt.
 
-Where should you store the data files? A very obvious solution is to
-always make sure that the data file used by a command file is in the
+Where should one store data files? A very obvious solution is
+always to make sure that the data file used by a command file is in the
 same directory as the command file. More elegant solutions (pointed to
                                                             below) require all co-authors to have the same directory structure so
-that load('Data/thedata.rda') means the same thing on all computers
+that \Verb+load('Data/thedata.rda')+ means the same thing on all computers
 used to work on said project.
@@
 analysis in different files. So, perhaps your analysis oriented files
-will load('cleandata.rda') or something but with a comment in the code
+will \Verb+load('cleandata.rda')+ or something but with a comment in the code
 telling the future you (among others) that cleandata.rda was created
 from create-cleandata.R which in turn begins with
-read.csv('dirtydata.csv') which, from a comment in the file, was
+\Verb+read.csv('dirtydata.csv')+ which, from a comment in the file, was
 downloaded from
@@
 
-Now, if in 5 years we wonder where 'cleandata.rda' came from, we might
+Now, if in 5 years we wonder where `cleandata.rda' came from, we might
 \Verb+ grep cleandata *.R+ to search for occurrences of this file on
 our Unix/OS X system. However, if such searching among files is a
-burden, and even nicer solution is to maintain a file for each project
+burden, an even nicer solution is to maintain a file for each project
 called ``MANIFEST.txt'' or ``INDEX.txt'' or ``README.txt'' which lists
@@
 what operations were performed on which set of data. I have seen
-command files which do not begin with a call to load data and I have
+command files which do not begin with a call to load data, and I have
 despaired. What is the meaning of \Verb+please-fit(var1 with var2)+ if
@@
 
- Now, if you had a moment to think in between teaching that new class, reading books for the awards committee, reading application files for the admissions committee, staying home with a sick child, and your own current research, you might say to yourself, ``Before I write new code, I should write a test of the code.
+ Now, if you had a moment to think in between teaching that new class, reading books for the awards committee, reading application files for the admissions committee, staying home with a sick child, and undertaking the odd bit of your own current research, you might say to yourself, ``Before I write new code, I should write a test of the code.
 I should write a little bit of code that let's me know that my double-bootstrap procedure actually does what it is supposed to do.''
@@
 if your files are named ``inequality-and-protest-analyses.R'' than if your files are
-called ``temp9'' or ``supercalifragilisticexpialidocious'' (Note the
+called ``temp9'' or ``supercalifragilisticexpialidocious.'' (Note the
                                                             use of the extension .R to tell us that the file contains R
@@
 
-Now, imagine that the UN never hears of your work, or even that your
-article does not spawn debate. Why then would you spend time to
+But what if the UN never hears of your work, or, by some cruel fate, that your
+article does not spawn debate? Why then would you spend time to
 communicate with your future self and others? My own answer to this
-question is because I want my work to be credible and useful to myself
+question is that I want my work to be credible and useful to myself
 and other scholars even if each article does not immediately change
@@
 characteristics: (1) the findings of the work should not be a matter
-of opinion and (2) other people could reproduce the findings. That is,
+of opinion; and, (2) other people should be able to reproduce the findings. That is,
 what the work represents is not a matter of opinion but is a shared
@@
 packages you used in your final analyses in 2018 in the archive
-itself.) You can say, "Here is everything you need to reproduce my
-work." To be extra helpful you can add "Read the README file for
+itself.) You can say, ``Here is everything you need to reproduce my
+work." To be extra helpful you can add ``Read the README file for
 futher instructions." And then you can get on with your life: maybe
-the next great idea will occur when your 4 year old asks a wacky
-question after stripping and painting her overly cooperative 1 year
-old brother purple, or teaching a class, or in a coffee shop, or on a
-quiet walk.
+the next great idea will occur when your 4-year-old asks a wacky
+question after stripping and painting her overly cooperative
+1-year-old brother purple, or teaching a class, or in a coffee shop,
+or on a quiet walk.
 
-\bibliographystyle{asa}
+\bibliographystyle{apsr}
 \bibliography{/Users/jwbowers/Documents/BIB/trunk/big}

### 2011-03-16 workflow.Rnw
Took into account Kieran's comments

@@
 some of the files. Now the checks and re-analyses should be
-easy. Right? The student-now-faculty member is the same person and
+easy. Right? The student, now professor, is the same person and
 should remember the reason for those bits of code (or at least should
@@
 which figure, right? And they wrote their code following Nagler's
-Maxims \citet{nagler1995coding} and being following King's Replication
-Standard \citet{king1995replication} right?
+Maxims \citep{nagler1995coding} and  King's Replication
+Standard \citep{king1995replication} right?
 
@@
 (let alone copying and pasting). If I wanted to re-create the figure
-you created but with red lines, and only for people in the South, I
+you created but with red lines, including a new variable or specification, I
 should be able to do so with just a few edits to the code.
@@
 R code right inside of it! I showed Sweave with plotting in
-Figure~\ref{fig:giniprot}, those who view the code for this essay will
-see how the Table~\ref{tab:protest} was also generated directly from a
+Figure~\ref{fig:giniprot}. Those who view the code for this essay will
+see how Table~\ref{tab:protest} was also generated directly from a
 regression object.
@@
 collaboration of research groups doing quantitative work in political
-science. Something will make our lives easier than Sweave (or
-                                                           odfWeave) in the future. Then we'll change.
+science. Something other than than Sweave will make our lives easier
+in the future. Then we'll change.
 
@@
 
-\section{Enabling revision to prevent clobbering.}
+\section{Enabling revision to prevent clobbering and record history.}
 
-Group work requires version control.\footnote{See X for more
-  discussion of what version control is.} Many people are familiar
-with the ``Track Changes'' feature in modern WYSIWYG word processors
-or the fact that Dropbox allows one to recover previous versions of
-files. These are both kinds of version control. When collaborating
-with yourself or others, it is useful to see what has changed, to feel
-free to experiment and then to dump the experiment in favor of
-previous work, to have multiple ``releases'' of the same document (one
-to MPSA, one to APSR, one to your parents) without requiring that
-spawn many copies of the same document and risk confusion and
-clobbering. Clobbering is what happens when your future self or your
-other collaborators saves an old version of a file over a new version
---- erasing good work by accident.
+Group work requires version control.\footnote{See X and Y in this
+  issue for more discussion of what version control is.} Many people
+are familiar with the ``Track Changes'' feature in modern WYSIWYG word
+processors or the fact that Dropbox allows one to recover previous
+versions of files. These are both kinds of version control. When
+collaborating with yourself or others, it is useful to see what has
+changed, to feel free to experiment and then to dump the experiment in
+favor of previous work, to have multiple ``releases'' of the same
+document (one to MPSA, one to APSR, one to your parents) without
+requiring that spawn many copies of the same document and risk
+confusion and clobbering. Clobbering is what happens when your future
+self or your other collaborators saves an old version of a file over a
+new version --- erasing good work by accident.
 
@@
 getting lost when you both try to save the same file on top of each
-other. If you find that you are needing to work on the same files at
+other. If you find that you need to work on the same files at
 the same time, then you should work on establishing your own shared
@@
 ensure that you do not clobber files when you make typos in the
-file-names.
+file-names. And, you will find yourself spending extra time
+reconciling changes made by different collaborators by hand that
+modern version control systems (particularly the distributed kind
+these days) take care of quickly and easily.
 
@@
 
- Imagine that you reported  bootstrap confidence intervals in your famous article.
-Now the statisticians at the UN are worrying about your bootstrap confidence interval.
-You'd like to evaluate your bootstrap procedure.
-Although nice code exists for bootstrapping linear models, no nice code exists to bootstrap the bootstrap.
-Of course, the code required is not complex, but since you are writing your own code you worry about getting it right.
-Now, 9 years after the idea, you've had lots of time to appreciate problems arising from bugs and errors in data analysis code.
+ Imagine that you reported bootstrap confidence intervals in your
+ famous article. Now the statisticians at the UN are worrying about
+ your bootstrap confidence interval. You'd like to evaluate your
+ bootstrap procedure. Although nice code exists for bootstrapping
+ linear models, no nice code exists to bootstrap the bootstrap. Of
+ course, the code required is not complex, but since you are writing
+ your own code you worry about getting it right. Now, 9 years after
+ the idea, you've had lots of time to appreciate problems arising from
+ bugs and errors in data analysis and code.
 
- Now, if you had a moment to think in between teaching that new class, reading books for the awards committee, reading application files for the admissions committee, staying home with a sick child, and undertaking the odd bit of your own current research, you might say to yourself, ``Before I write new code, I should write a test of the code.
-I should write a little bit of code that let's me know that my double-bootstrap procedure actually does what it is supposed to do.''
+ Now, if you had a moment to think in between teaching that new class,
+ reading books for the awards committee, reading application files for
+ the admissions committee, staying home with a sick child, and
+ undertaking the odd bit of your own current research, you might say
+ to yourself, ``Before I write new code, I should write a test of the
+ code. I should write a little bit of code that let's me know that my
+ double-bootstrap procedure actually does what it is supposed to do.''
 
-         Of course, this idea, like most others is not new.
-When large groups of programmers write code for multi-million dollar programs the question about avoiding error looms large.
-The idea of \href{http://en.wikipedia.org/wiki/Test-driven_development}{test driven development} and the idea that one ought to create tests of \href{http://en.wikipedia.org/wiki/Unit_testing}{small parts of one's code} arose to address such concerns.
-For the social scientist collaborating with her future self and/or a small group of collaborators.
+ Of course, this idea, like most others, is not new. When large groups
+ of programmers write code for multi-million dollar programs the
+ question about avoiding error looms large. The idea of
+ \href{http://en.wikipedia.org/wiki/Test-driven_development}{test
+   driven development} and the idea that one ought to create tests of
+ \href{http://en.wikipedia.org/wiki/Unit_testing}{small parts of one's
+   code} arose to address such concerns. For the social scientist
+ collaborating with her future self and/or a small group of
+ collaborators.
 

### 2011-03-17 workflow.Rnw
Editing the document

@@
 
-\author{Jake Bowers\thanks{Many thanks to Mark Fredrickson, Brian
-    Gaines, Kieran Healy, Kevin Quinn and Cara Wong for direct help on
-    this document, Mika LaVaque-Manty, Ben Hansen, and Mark
-    Fredrickson for many useful discussions on this topic. The source
-    code for this document may be freely downloaded from
-    \url{https://github.com/jwbowers/workflow}.}}
+\author{Jake Bowers\thanks{I owe many thanks to Mark Fredrickson,
+    Brian Gaines, Mark Fredrickson, Kieran Healy, Kevin Quinn and Cara
+    Wong for direct help on this document and to Mika LaVaque-Manty
+    and Ben Hansen for many useful discussions on this topic. The
+    source code for this document may be freely downloaded and
+    modified from \url{https://github.com/jwbowers/workflow}.}}
 
@@
 \usepackage[noae,nogin]{Sweave} %Declaring this explicitly so that I can modify the Sweave environment here rather in Sweave.sty which will be c
-\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=.5em,fontsize=\footnotesize}
-\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=.5em,fontsize=\footnotesize}
-\DefineVerbatimEnvironment{Scode}{Verbatim}{xleftmargin=.5em,fontsize=\footnotesize}
+\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=.5em,fontsize=\small}
+\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=.5em,fontsize=\small}
+\DefineVerbatimEnvironment{Scode}{Verbatim}  {xleftmargin=.5em,fontsize=\small}
 \fvset{listparameters={\setlength{\topsep}{0pt}}}
@@
 \SweaveOpts{keep.source=TRUE}  %Show comments
-\fvset{fontsize=\normalsize} %slightly smaller verbatim font
+\fvset{fontsize=\small} %slightly smaller verbatim font
 
@@
   ##First, just setup the \R environment:
-
   options(SweaveHooks=list(
@@
 An idea is born in a coffee shop, a seminar, a quiet walk. It is a
-cold winter in 2011, but the idea energizes the student. The idea is
-why he went to graduate school.
+gray winter in 2011, but the idea dispells February's doldrums and
+energizes the student.% Now he knows why he went to graduate school.
 
@@
 harried editors, reviewers and authors strive to balance teaching,
-service, and life. The three collaborators have luckily found work in
-three different universities around the world. In 2018 a child is born
-and a paper is published. In 2020 a graduate student in a coffee shop
-reads the paper and has an idea that challenges the analysis in the
-paper. What would happen if only the first author had thought to
-control for X? Or run one more MCMC chain? Or chosen a different
-likelihood function? Will the United Nations (now eager to act based
-                                              the paper) make a wrong move?
+service, and life. By now, the three lucky collaborators work as
+professors in three different universities. In 2018 a child is born
+and a paper is published. The United Nations takes an interest in the
+paper in 2019 and hosts a conference to discuss implications of the
+research. In 2020 a first year graduate student in a coffee shop has
+an idea that challenges the analysis in the now famous paper. What
+would happen if they had controlled for X? Or run one more MCMC chain?
+Or chosen a different likelihood function?  Will the United Nations
+(now eager to act based the paper) make a wrong move?
 
 The first author convenes a three way video conference with the other
-collaborators while commuting home in his flying car. The group must
-go back to the analyses. Which ones? The ones from 2011? Or 2012? Or
-2018? Where are the files? The next day, one member of the group finds
-some of the files. Now the checks and re-analyses should be
-easy. Right? The student, now professor, is the same person and
-should remember the reason for those bits of code (or at least should
-                                                   remember which series of mouse clicks were used to produce the numbers
-                                                   for that crucial table --- and those mouse clicks should not be time
-                                                   consuming to redo exactly as they were done in 2011, or was it
-                                                   2015?). And, of course, the way Microsoft Word/Stata/SPSS/R/LISREL
-stores files and the way that your machine reads and writes them is
-the same, right --- since Windows and Mac OS X have always existed and
-will always continue to exist more or less as they currently exist? And
-the group knows exactly which bit of code produced which table and
-which figure, right? And they wrote their code following Nagler's
-Maxims \citep{nagler1995coding} and  King's Replication
-Standard \citep{king1995replication} right?
+collaborators during his homeward commute while his flying car is in
+auto-drive mode (video chatting during manual driving having been
+outlawed in his state in 2017). The group must go back to the
+analyses. Which ones?  The ones from 2011? Or 2012? Or 2018? Where are
+the files? The next day, one member of the group who has kept some
+hard-drives around out of nostalgia (the same guy who still owns
+cassette tapes) finds some of the files. Now the checks and
+re-analyses should be easy. Right? The student, now professor, is the
+same person and should remember the reason for those bits of code (or
+at least should remember which series of mouse clicks were used to
+produce the numbers for that crucial table --- and those mouse clicks
+should not be time consuming to redo exactly as they were done in
+2011, or was it 2015?). Right? And, of course, the way Microsoft
+Word/Stata/SPSS/R/LISREL stores files and the way that current machines in 2020
+read and write them is the same --- since Windows and Mac OS
+X have always existed and will always continue to exist more or less
+as they currently exist? Right? And the group knows exactly which bit of code
+produced which table and which figure, right? And they wrote their
+code following Nagler's Maxims \citep{nagler1995coding} and King's
+Replication Standard \citep{king1995replication} right?
+
+Right? Wrong? If the collaborators find themselves answering ``wrong'' to many of those questions
+
+If the original analyses were not done via typing in commands in plain
+text files, these authors will be in a pickle in 2020.
 

### 2011-03-21 workflow.Rnw
COntinued editing

@@
 
-\title{A Few Easy Pieces: \\ How to enhance your relationship with your future self.}
+\title{How to improve your relationship with your future self.}
 
@@
 
-An idea is born in a coffee shop, a seminar, a quiet walk. It is a
-gray winter in 2011, but the idea dispells February's doldrums and
-energizes the student.% Now he knows why he went to graduate school.
+An idea is born in a coffee shop, a seminar, a quiet walk. On this
+gray day in 2011, the idea dispells February's doldrums. The student
+rushes home, mind racing, the cold ignored.
+% Now he knows why he went to graduate school.
 
-This idea inspires a seminar paper in the spring. And a conference
+This idea inspires a seminar paper in the spring. A conference
 paper arises from the seminar paper in collaboration with another
@@
 research. In 2020 a first year graduate student in a coffee shop has
-an idea that challenges the analysis in the now famous paper. What
-would happen if they had controlled for X? Or run one more MCMC chain?
+an idea that challenges the results in the now famous paper. What
+would happen if the authors had controlled for X? Or run one more MCMC chain?
 Or chosen a different likelihood function?  Will the United Nations
@@
 The first author convenes a three way video conference with the other
-collaborators during his homeward commute while his flying car is in
-auto-drive mode (video chatting during manual driving having been
-outlawed in his state in 2017). The group must go back to the
-analyses. Which ones?  The ones from 2011? Or 2012? Or 2018? Where are
-the files? The next day, one member of the group who has kept some
-hard-drives around out of nostalgia (the same guy who still owns
-cassette tapes) finds some of the files. Now the checks and
-re-analyses should be easy. Right? The student, now professor, is the
-same person and should remember the reason for those bits of code (or
-at least should remember which series of mouse clicks were used to
-produce the numbers for that crucial table --- and those mouse clicks
-should not be time consuming to redo exactly as they were done in
-2011, or was it 2015?). Right? And, of course, the way Microsoft
-Word/Stata/SPSS/R/LISREL stores files and the way that current machines in 2020
-read and write them is the same --- since Windows and Mac OS
-X have always existed and will always continue to exist more or less
-as they currently exist? Right? And the group knows exactly which bit of code
-produced which table and which figure, right? And they wrote their
-code following Nagler's Maxims \citep{nagler1995coding} and King's
-Replication Standard \citep{king1995replication} right?
+collaborators during his homeward commute after putting his flying car
+in auto-drive mode.\footnote{One assumes that video chatting during
+  manual driving of flying cars will have been outlawed in his state
+  in 2017.} The group must go back to the analyses. Which ones?  The
+ones from 2011?  Or 2012? Or 2018? Where are the files? The next day,
+one member of the group who has kept some hard-drives around out of
+nostalgia finds some of the files.\footnote{This is the same guy who
+  still owns cassette tapes and compact discs.} Now the checks and
+re-analyses should be easy. Right? The student, now professor, should
+remember the reason for those bits of code (or at least should
+remember which series of mouse clicks were used to produce the numbers
+for that crucial table --- and that mousing should not be time
+consuming to redo exactly as they were done in 2011, or was it
+2015?). Right? And, of course, the way Microsoft
+Word/Stata/SPSS/R/LISREL stores files and the way that current
+machines in 2020 read and write them is the same --- since Windows and
+Mac OS X have always existed and will always continue to exist more or
+less as they currently exist? Right? And the group knows exactly which
+bit of code produced which table and which figure, right? And they
+wrote their code following Nagler's Maxims \citep{nagler1995coding}
+and King's Replication Standard \citep{king1995replication} right?
 
-Right? Wrong? If the collaborators find themselves answering ``wrong'' to many of those questions
-
-If the original analyses were not done via typing in commands in plain
-text files, these authors will be in a pickle in 2020.
-
-Just in case one has some doubt about answering these questions, this
-piece aims to amplify some of what we already ought to know from
+Right? Wrong? If the collaborators find themselves answering ``wrong''
+to many of the questions posed here then reproducing the analyses will
+take a lot of time if such reproduction is possible at all. This essay
+provides some suggestions for practices which will make such
+reproduction occur much more easily and quickly in the event that
+famous papers require special scrutiny. Specifically, this piece aims
+to amplify some of what we already ought to know from
 \citet{nagler1995coding} and \citet{king1995replication}, and to
 update some of those ideas given current practices, platforms, and
-propensities
+propensities.
 
@@
 %%% Putting it up top, so we can use it in the comments section.
-
 <<read.data,echo=FALSE,results=hide>>=
@@
 dem.nations.df<-read.csv(url("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv",encoding="ISO_8859-1"),as.is=TRUE)
+row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful
+good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
+@
 
-row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful
+\section{Data analysis is computer programming.}
 
-good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
-@.
-\section{Type, don't click.}
-
-All tables and graphs should be generated from code, not clicking
-(let alone copying and pasting). If I wanted to re-create the figure
-you created but with red lines, including a new variable or specification, I
-should be able to do so with just a few edits to the code.
+All results (numbers, comparisons, tables, figures) should arise
+generated from code, not series of mouse clicks or copying and
+pasting. If I wanted to re-create the figure you created but including
+a new variable or specification, I should be able to do so with just a
+few edits to the code rather than knowledge of how you used your
+pointing device in your graphical user interface.
 
 Using R, for example, I might specify that the file \Verb+fig1.pdf+
-  was produced by the command, where the \Verb+please-plot+ function was
-loaded elsewhere.
+was produced by the following commands in a file called \Verb+fig1.R+.
 
-\begin{Verbatim}[fontsize=\small]
+<<eval=FALSE>>=
+thedata<-read.csv("Data/thedata.csv")
 pdf('fig1.pdf')
-please-plot(outcome by explanatory)
+please-plot(outcome by explanatory using thedata)
 dev.off()
-\end{Verbatim}
+@
 
@@
 
-\section{Enjoy literate data analysis.}
+\section{Data analysis is commnication.}
 
@@
 
-\section{Enabling revision to prevent clobbering and record history.}
+\section{Version control prevents clobbering and record history.}
 
@@
 
-\section{Take advantage of file names to convey information to your future self.}
+\section{File names send messages to your future self.}
 
@@
 \begin{quote}
-         [I]f the empirical basis for an article or book cannot
-         be reproduced, of what use to the discipline are its conclusions?
-         What purpose does an article like this serve?
+  [I]f the empirical basis for an article or book cannot be
+  reproduced, of what use to the discipline are its
+  conclusions?  What purpose does an article like this serve?
   \cite[445]{king1995replication}
@@
   Ask your advisers how long it took them from idea to dissertation to
-                                                        book (articles).}
-         The time-frames over which collaboration are required --- whether
-         among a group of people working together or within a single scholar's
-productive life or probably both --- are much longer than any given
-version of any given software will easily exist. \href{http://en.wikipedia.org/wiki/Plain_text#Usage}{Plain text} is the
-exception. Thus, even as we extol version control systems, one must
-have a way to ensure future access to them in a form that will still
-be around when sentient cockroaches finally join political science
-departments (by then dominated by cetaceans after humans are mostly
-uploads).\footnote{The arrival of the six-legged social scientists
-  revives Emacs and finally makes Ctrl-a Ctrl-x
+  book (articles).}  The time-frames over which collaboration are
+required --- whether among a group of people working together or
+within a single scholar's productive life or probably both --- are
+much longer than any given version of any given software will easily
+exist. \href{http://en.wikipedia.org/wiki/Plain_text#Usage}{Plain
+  text} is the exception. Thus, even as we extol version control
+systems, one must have a way to ensure future access to them in a form
+that will still be around when sentient cockroaches finally join
+political science departments (by then dominated by cetaceans after
+humans are mostly uploads).\footnote{The arrival of the six-legged
+  social scientists revives Emacs and finally makes Ctrl-a Ctrl-x
   Esc-x Ctrl-c a \href{http://kieran.healy.usesthis.com/}{reasonable

### 2011-03-21 workflow.Rnw
Continued editing

@@
 
-All results (numbers, comparisons, tables, figures) should arise
-generated from code, not series of mouse clicks or copying and
-pasting. If I wanted to re-create the figure you created but including
-a new variable or specification, I should be able to do so with just a
-few edits to the code rather than knowledge of how you used your
-pointing device in your graphical user interface.
+All results (numbers, comparisons, tables, figures) should arise from
+code, not series of mouse clicks or copying and pasting. If I wanted
+to re-create the figure you created but including a new variable or
+specification, I should be able to do so with just a few edits to the
+code rather than knowledge of how you used your pointing device in
+your graphical user interface.
 
 Using R, for example, I might specify that the file \Verb+fig1.pdf+
-was produced by the following commands in a file called \Verb+fig1.R+.
+was produced by the following commands in a file called
+\Verb+fig1.R+.\footnote{The command \Verb+please-plot+ and some other
+  R functions used in this essay come from the
+  \Verb+MayIPleaseDoStatistics+ package which emphasizes politeness in
+  data analysis. The short hand commands like \Verb+please-plot+ can
+  be blocked and \Verb+may-I-please-have-a-plot+ can be required using
+  \Verb+options(politeness=99)+ }
 
-<<eval=FALSE>>=
+\begin{Verbatim}[fontsize=\small]
 thedata<-read.csv("Data/thedata.csv")
 pdf('fig1.pdf')
-please-plot(outcome by explanatory using thedata)
+please-plot(outcome by explanatory using thedata. red lines please.)
 dev.off()
-@
+\end{Verbatim}
 
 Now, in the future if I wonder how ``that plot on page 10'' was
-created, if I suitably commented my manuscript or inserted text into
-my MANIFEST.txt file I will know (1) ``that plot'' is from a file
-called ``fig1.pdf'' and (2) fig1.pdf was created in
-\Verb+figs1and2.R+. In the future, if I wanted to quickly change
-fig1.pdf, I could edit the commands quickly and easily to do
-so. Another good practice is to name the files descriptively, so that
-we don't in fact have ``fig1.pdf'' but
-``political-participation-by-education-in-the-south.pdf''. Now, if Figure is
-moved to an appendix, the file name is still meaningful.
+created: (1) ``that plot'' is from a file called ``fig1.pdf'' and
+fig1.pdf was created in \Verb+fig1.R+. In a future where R still
+exists, if I wanted to change the figure, I could edit the commands
+quickly and easily to do so (since R commands are written in plain
+text, and plain text is a format that will be around for a long time
+to come).
 
-% An even nicer way to ensure that figures and tables are reproducible
-% or easily modifiable by your older and wiser but more harried self is
-% to use literate programming. For example, to produce the
-% Figure~\ref{fig:giniprot}, I insert the following chunk of code into
-% the LaTeX document, and then I asked R to process the document, and
-% then produce a pdf file in which the code has disappeared to be
-% replaced with an actual figure.\footnote{One can insert the same code
-%   chunk in an OpenOffice document and use the odfWeave package to
-%   produce OpenOffice documents in which figures and tables replace
-%   code. One may also use the LyX interface to LaTeX to write LaTeX
-%   documents within a WYSIWYG environment.}
-
-<<fig1code,echo=TRUE,fig=FALSE>>=
-##  Make a scatterplot of Protest by Inequality
-with(good.df,plot(Gini2004/100,protac2000,
-                  xlab='Gini Coefficient 2004 (UNDP)',
-                  ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
-                  cex=.8))
-
-##  Label a few interesting points
-with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
-     text(Gini2004/100,protac2000,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
-@
-
-\begin{figure}[H]
-  \begin{center}
-<<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>=
-par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
-<<fig1code>>
-@
-    \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}\label{fig:giniprot}
-  \end{center}
-\end{figure}
-
-The principle here is to know where tables and figures came from ---
-to know what commands created which figures and tables.
+\paragraph{Principle} Know the provenance of your results so that your
+future self or current collaborators can quickly and easily reproduce
+your work (and make changes).% --- to know what
+% commands created which figures and tables.
 
 \section{No data analyst is an island.}
-% MMF: My original title before I saw you had you own comment section
-% \section{Comment on your own work. }
+Data analysis involves a long series of decisions. Each decision
+requires justification.  The body of the paper or book or report will
+justify the most important decisions. Some decisions will be too small
+and technical for inclusion in the paper itself. These need to be
+documented in the code itself \citep{nagler1995coding}.
 
-Comments, or text inside of a file that is not executed, are a message to your
-future self and other consumers of your work. 
-Read \citet{nagler1995coding}. Comment your code. Luckily, if you are
-using a literate programming practice (Sweave, odfWeave, using R2HTML,
+Comments --- unexecuted text inside of a script --- are a message to
+your future self and other consumers of your work.  Read
+. Comment your code. Luckily, if you are using
+a literate programming practice (Sweave, odfWeave, using R2HTML,
 etc..), you can write paragraphs to surround your code as well a
@@
 \end{verbatim}
-
-A given analysis of a given collection is a long series of
-decisions. You need to be able to justify them. Often you justify them
-in the body of the paper you are writing. Some decisions will be too small and
-technical for inclusion in the paper itself. These need to be
-documented in the code itself.
 
@@
 despaired. What is the meaning of \Verb+please-fit(var1 with var2)+ if
-I do not know where var1 var2 come from.\footnote{The command
-  \Verb+please-fit+ and some other R functions used in this essay come
-  from the \Verb+MayIPleaseDoStatistics+ package which emphasizes
-  politeness in data analysis. The commands \Verb+please-fit+ and
-  \Verb+please-plot+ can be blocked and \Verb+would-you-please-fit+ or
-  \Verb+may-I-please-have-a-plot+ can be required using
-  \Verb+options(politeness=99)+ }
+I do not know where var1 var2 come from.
 

### 2011-03-22 workflow.Rnw
More writing

@@
   \Verb+MayIPleaseDoStatistics+ package which emphasizes politeness in
-  data analysis. The short hand commands like \Verb+please-plot+ can
-  be blocked and \Verb+may-I-please-have-a-plot+ can be required using
+  data analysis. Functions like \Verb+please-plot+ can
+  be blocked and more polite versions such as \Verb+may-I-please-have-a-plot+ can be required using
   \Verb+options(politeness=99)+ }
@@
 \begin{Verbatim}[fontsize=\small]
-thedata<-read.csv("Data/thedata.csv")
-pdf('fig1.pdf')
+thedata<-read.csv("Data/thedata-15-03-2011.csv") ## Read the data
+pdf('fig1.pdf') ## begin writing to the pdf file
 please-plot(outcome by explanatory using thedata. red lines please.)
-dev.off()
+dev.off() ## stop writing to the pdf file
 \end{Verbatim}
@@
 
-\section{No data analyst is an island.}
+\section{No data analyst is an island for long.}
 Data analysis involves a long series of decisions. Each decision
-requires justification.  The body of the paper or book or report will
+requires justification.  The text itself will
 justify the most important decisions. Some decisions will be too small
@@
 Comments --- unexecuted text inside of a script --- are a message to
-your future self and other consumers of your work.  Read
-. Comment your code. Luckily, if you are using
-a literate programming practice (Sweave, odfWeave, using R2HTML,
-etc..), you can write paragraphs to surround your code as well a
-technical comments in the code itself. If you are not using a strictly
-literate programming practice then use whatever commenting protocol
-exists in your analysis language. Here are some examples:
-\begin{verbatim}
-# This is a comment in R.
-# With a a second line.
-
-* This is a comment in a Stata .do file
-* You can also embed comments at the ends of lines using /* ... */
-
-/* SAS uses a similar convention to Stata
-   and allows multi-line comments like this.
-*/
-
-% Finally, the percent symbol is a comment in TeX.
-\end{verbatim}
-
-Using R, for example, the comment is marked using the
-pound sign. Notice that I've left in a regression not run for the
-paper directly, but run as a part of the due diligence process of
-writing the paper.
+your future self and other consumers of your work. Luckily, if you are
+using a literate programming practice (ex. Sweave or odfWeave), you
+can write paragraphs to surround your code as well as technical
+comments in the code itself. If you are not using a strictly literate
+programming practice then use whatever commenting protocol exists in
+your analysis language. R, for example, considers text marked with
+``\#'' as a comment. Here for example I comment about a linear model
+run for a paper and also include comments about another model run for
+due diligence.
 
 <<reg1,echo=TRUE,results=verb>>=
-
 ##  Repeat the regression run for Table~\ref{tab:protest}
 ##  Do political rights predict protest reporting?
- lm1<-lm(protac2000 ~ I(Gini2004/100) + meanpr, data=good.df)
+lm1 <- lm(protac2000 ~ I(Gini2004/100) + meanpr, data=good.df)
 ##  Q: Was it worthwhile controlling for political rights?
- lm2<-lm(protac2000~I(Gini2004/100),data=good.df)
+lm2 <- lm(protac2000~I(Gini2004/100),data=good.df)
 coef(lm2)[2]-coef(lm1)[2]
@@
 
-If you need to track down something more than a small detail, but yet
-the sleuthing is really a footnote to the article, then you can create
-another file to explore and report on those questions for the research
-group. The writing in such a memo can be more informal, but the norms
-of reproducibility ought to be the same as for the main paper. After
-all, you will be releasing both documents (probably) publicly upon
-publication of the article if not before.
+Why leave messages for yourself (or others)? These messages help you
+retrace and justify the steps of your data-analytic journey in case
+you ever need to revist them as the work moves from seminar paper to
+conference paper to poster back to paper to dissertation and
+onwards. Notice one another benefit of coding for an audience: we
+learn by teaching. By assuming that others will look at your code, you
+will be more likely to write clearer code, or perhaps even to learn
+more about what you are doing as you do it.
 
-Notice one another benefit of coding for an audience: we learn by
-teaching. By assuming that others will look at your code, you will be
-more likely to write clearer code, or perhaps even to learn more about
-what you are doing as you do it.
+Sometimes simple comments in the main command file are not enough:
+imagine if the authors had arrived at the bootstrap only after running
+simulations to assess the coverage of that procedure compared to a
+profiled likelihood based approach. Rather than include time consuming
+simulations in the main body of their Sweave document, they might have
+written some memos that discuss the tradeoffs and implementation
+decisions involved in such a comparison. In the main text they would
+have just read in the results of the simulation study with an
+appropriate comment:
 
-Comment liberally. Comments are discarded when R runs analysis or LaTeX turns
-your poignant writings into a PDF, so only those who dig into your work will
-see them. They are footnotes on the process of producing your research.
+<<bsreg1,echo=TRUE,results=verb,eval=FALSE>>=
+## We assessed the bootstrap against direct inspection of the likelihood profile
+## in the file Memos/shouldwebootstrap.Rnw. Here we read in the results of the bootstrap procedure:
+load("Memos/bsresults.rda") ## using a compressed binary format here for speed
+@
+
+Comment liberally. Comments are discarded when R runs analysis or
+LaTeX turns text into a pdf, so only those who dig into your work will
+see them. They are footnotes on the process of producing your
+research.
+
+% If you need to track down something more than a small detail, but yet
+% the sleuthing is really a footnote to the article, then you can create
+% another file to explore and report on those questions for the research
+% group. The writing in such a memo can be more informal, but the norms
+% of reproducibility ought to be the same as for the main paper. After
+% all, you will be releasing both documents (probably) publicly upon
+% publication of the article if not before.
 

### 2011-03-23 workflow.Rnw
Version that just needs one more edit.

@@
 
-\section{Data analysis is computer programming.}
+\section{Data analysis is computer programming.}\label{sec:data-analys-comp}
 
@@
 Now, in the future if I wonder how ``that plot on page 10'' was
-created: (1) ``that plot'' is from a file called ``fig1.pdf'' and
-fig1.pdf was created in \Verb+fig1.R+. In a future where R still
+created: (1) ``that plot'' is from a file called \Verb+fig1.pdf+ and
+\Verb+fig1.pdf+ was created in \Verb+fig1.R+. In a future where R still
 exists, if I wanted to change the figure, I could edit the commands
@@
 
+\subsection{Comment your code.}
 Comments --- unexecuted text inside of a script --- are a message to
@@
 
-Why leave messages for yourself (or others)? These messages help you
+Messages left for your future self (or near-future others) help
 retrace and justify the steps of your data-analytic journey in case
@@
 
+Comment liberally. Comments are discarded when R runs analysis or
+LaTeX turns text into a pdf, so only those who dig into your work will
+see them.
+
+\subsection{Programming can be literate.}
+
 Sometimes simple comments in the main command file are not enough:
@@
 profiled likelihood based approach. Rather than include time consuming
-simulations in the main body of their Sweave document, they might have
+simulations in the body of their main Sweave document, they might have
 written some memos that discuss the tradeoffs and implementation
@@
 <<bsreg1,echo=TRUE,results=verb,eval=FALSE>>=
-## We assessed the bootstrap against direct inspection of the likelihood profile
-## in the file Memos/shouldwebootstrap.Rnw. Here we read in the results of the bootstrap procedure:
+## We assessed the coverage of the bootstrap intervals against 
+## direct inspection of the likelihood profile in the file
+## Memos/shouldwebootstrap.Rnw which produced bsresults.rda
+
+##Here we read in the results of the bootstrap procedure:
 load("Memos/bsresults.rda") ## using a compressed binary format here for speed
@@
 
-Comment liberally. Comments are discarded when R runs analysis or
-LaTeX turns text into a pdf, so only those who dig into your work will
-see them. They are footnotes on the process of producing your
-research.
+Now that memo itself (or perhaps the main document) probably involves
+sentences and paragraphs explaining an analysis. Literate programming
+or literate data analysis happens when the code of the analysis is
+mixed side by side with the paragraphs explaining, interpreting,
+foreshadowing, discussing the analysis.\footnote{This document is
+  written in such a way using \LaTeX and R code mixed together in what
+  is known as an Sweave file \citep{Leis:2005}. You can download and
+  view and edit this document as a plain text file from
+  \url{https://github.com/jwbowers/workflow}.}
 
-% If you need to track down something more than a small detail, but yet
-% the sleuthing is really a footnote to the article, then you can create
-% another file to explore and report on those questions for the research
-% group. The writing in such a memo can be more informal, but the norms
-% of reproducibility ought to be the same as for the main paper. After
-% all, you will be releasing both documents (probably) publicly upon
-% publication of the article if not before.
+If comments are guideposts, literate programs are full fledged
+travelogues. Why might one care to use a literate programming
+practice?
 
-\section{Data analysis is commnication.}
+Say you run some command, discover something new (or confirm something
+old) and then make a claim backed up by table or a figure. You produce
+a nice little report on your work for use in discussions of the
+working group. The report itself is a pdf file or some other format
+which emphasizes reading.\footnote{like html, postscript, dvi or even
+  the wordprocessor formats like rtf or doc or odt, or the web formats
+  like html.}  Eventually we want to use pieces of that report
+(tables, graphs, paragraphs) in a publishable paper. It would be a
+waste of time if you had to re-create those analyses by pointing,
+clicking, copying, or pasting. It would also be a waste of time if you
+couldn't remember why you decided to use a quadratic instead of a
+cubic term and had to re-create that argument. So, we might use
+literate programming because we don't want to waste time.
 
-The source code of any pdf document exchanged by the group must be
-available and executable. Say you run some command and then make a
-table or a graph. Then you produce a nice little report on your work
-for use among the working group. Eventually we want to use pieces of
-that report (tables, graphs, paragraphs) in a publishable paper. Thus,
-that report must not have been created using copying and pasting. It
-must be created using LaTeX (or HTML or OpenOffice or something else
-that the working group standardizes on). And, in fact, it must be
-created using the reproducible markup system like Sweave (if you are
-using LaTeX and R as your working group tools)
-\citep{Leis:2005}. Alternatively, you can put comments into your
-report detailing where the figures and tables came from (i.e. which
-command file produced them). Or include details about which code
-produced which figures in your MANIFEST.txt.
+For example, in \S~\ref{sec:data-analys-comp}, I suggested that we
+know where ``that plot on page 10'' comes from by making sure we had a
+``fig1.pdf'' file produced from a clearly commented plain text file
+called something like ``fig1.R.'' If one were using Sweave, one could
+even more easily have written the following to produce the desired
+figure.
 
-In order to know that ``that plot on page 10'' is from a file called
-``fig1.pdf'' you need to document it somehow. I can imagine a system
-with very disciplined use of a word processor (like Open Office or
-Word) --- perhaps using special auxiliary ``listoffigures.txt'' files
-or something. In my experience, it is much easier just to use LaTeX
-(which can be used in a very user friendly way via LyX).
+\begin{Verbatim}[fontsize=\footnotesize]
+\begin{figure}[H]
+\begin{center}
+ <<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>=
+ par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
+ ##  Make a scatterplot of Protest by Inequality
+ with(good.df,plot(Gini2004/100,protac2000,
+                   xlab='Gini Coefficient 2004 (UNDP)',
+                   ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
+                   cex=.8))
 
-The Sweave system is even nicer because you can have a LaTeX file with
-R code right inside of it! I showed Sweave with plotting in
-Figure~\ref{fig:giniprot}. Those who view the code for this essay will
-see how Table~\ref{tab:protest} was also generated directly from a
-regression object.
+ ##  Label a few interesting points
+ with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
+      text(Gini2004/100,protac2000,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
+ @
+\caption{Protest activity by income inequality \citep[from][]{norris2009data}.}\label{fig:giniprot}
+\end{center}
+\end{figure}
+\end{Verbatim}
+
+<<fig1code,echo=FALSE,fig=FALSE>>=
+##  Make a scatterplot of Protest by Inequality
+with(good.df,plot(Gini2004/100,protac2000,
+                  xlab='Gini Coefficient 2004 (UNDP)',
+                  ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
+                  cex=.8))
+
+##  Label a few interesting points
+with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
+     text(Gini2004/100,protac2000,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
+@
+
+\begin{figure}[H]
+  \begin{center}
+<<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>=
+par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
+<<fig1code>>
+@
+    \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}\label{fig:giniprot}
+  \end{center}
+\end{figure}
+
+By using \Verb+\label{fig:giniprot}+, I do not need to keep track of
+the figure number. Nor do I need a separate \Verb+fig1.R+ file or
+\Verb+fig1.pdf+ file.  Tables and other numerical results are also
+possible to generate within ``code chunks'' in the source code of a
+scholarly paper. Those who view the code for this essay will see how
+Table~\ref{tab:protest} was also generated directly from a regression
+object.\footnote{See [cite to Beck TPM] for the inspiration for this
+  particular presentation of a linear model.}
 
@@
 
-Right now it is hard to beat Sweave as a system for enhancing the
-collaboration of research groups doing quantitative work in political
+If your workflow does not involve \LaTeX and R, you can still
+implement some of the principles here (and if you use OpenOffice or
+LyX and R you can use \Verb+odfWeave+ do directly do literate data
+analysis in a WYSIWYG environment).\footnote{A quick Google search of
+  ``Sweave for Stata'' turned up lots of resources for literate
+  programming with Stata.}
+
+Literate data analysis is not the same as Sweave, even if Sweave is a
+nice implementation.\footnote{See
+  \url{http://cran.r-project.org/web/views/ReproducibleResearch.html}
+  for many of the different approaches to literature programming for
+  R.} For example, the person most often credited with
+these ideas used plain \TeX as the ``document formatting'' language
+and Pascal as the programming language [cite]. See [X] for an example
+of using the Emacs org-mode as literate programming. And, one could
+imagine a Word or HTML file in which code chunks are inserted by
+commented out and run one at a time (or perhaps extracted by a clever
+script from the XML source file of a modern Microsoft Word Document)
+--- or an HTML file in which hyperlinks are inserted to plain text
+files containing the relevant code chunks.
+
+\paragraph{Principle} Keep in mind the distinction between the
+``source code'' of a document (i.e. what computation was required to
+produce it) and the visible, type-set page image. Page images are
+great for reading, but not great for reproducing or collaborating. The
+source code of any document exchanged by the group must be available
+and executable.
+
+In order to know that ``that plot on page 10'' is from a file called
+``fig1.pdf'' you need to document it somehow. I can imagine a system
+with very disciplined use of a word processor perhaps using special
+auxiliary ``listoffigures.txt'' or ``MANIFEST.txt'' or ``README'' or
+``Makefile'' files or something. Right now it is hard to beat Sweave
+or some plain text combination of \LaTeX (or some other markup based
+document formatting system with support for citations, etc... ) with
+some plain text based statistical programming environment (like R or
+Stata or Matlab, etc.) as a system for enhancing the collaboration of
+research groups doing quantitative work in political
 science. Something other than than Sweave will make our lives easier
-in the future. Then we'll change.
+in the future. Then we'll change.\footnote{For example, X uses the
+  org-mode markup system in this issue.}
+
+\subsection{File names send messages to your future self.}
+
+Name your files with evocative and descriptive names. Your
+collaborators are less likely to call you at midnight asking for help
+if your files are named ``inequality-and-protest-analyses.R'' than if
+your files are called ``temp9'' or
+``supercalifragilisticexpialidocious.'' (Note the use of the extension
+.R to tell us that the file contains R commands. Use extensions like
+this as a standard practice to help you and your computer get along.)
+
+\paragraph{Principle} Comments and other documentation enable you to
+learn through teaching, code efficiently, and avoid duplicating your
+effort. Literate programming also enables you to avoid the kinds of
+errors which occur when you re-type numbers in to tables or
+graphic-producing systems or even into a document itself. Finally, and
+most importantly, a document that can be ``run'' to reproduce all of
+the results is a document that can more effectively spur discussion
+and learning and cumulation of research as people no longer need spend
+weeks attempting to reproduce research on which they desire to build.
+
+\url{http://en.wikipedia.org/wiki/Literate_programming}
 
@@
 \Verb+read.csv('http://www.mywebsite.org/Data/thedata.csv')+ or some
-equivalent line in them, and `thedata.csv' should be stored in some
-place easy to find (like in the same directory as the file or perhaps
-                    in \Verb+'Data/thedata.rda'+).
-It never hurts to drop in a comment about where the data is, if in doubt.
+equivalent line in them, and \Verb+thedata.csv+ should be stored in
+some place easy to find (like in the same directory as the file or
+perhaps in \Verb+'Data/thedata.rda'+). Of course, it never hurts to
+drop in a comment pointing to the data file.
 
-Where should one store data files? A very obvious solution is
-always to make sure that the data file used by a command file is in the
-same directory as the command file. More elegant solutions (pointed to
-                                                            below) require all co-authors to have the same directory structure so
-that \Verb+load('Data/thedata.rda')+ means the same thing on all computers
-used to work on said project.
+Where should one store data files? A obvious solution is always to
+make sure that the data file used by a command file is in the same
+directory as the command file. More elegant solutions require all
+co-authors to have the same directory structure so that
+\Verb+load('Data/thedata.rda')+ means the same thing on all computers
+used to work on the project.
 
-The principle of modularity \cite{nagler1995coding} suggests that you
-separate data cleaning, processing, recoding, and merging from
-analysis in different files. So, perhaps your analysis oriented files
-will \Verb+load('cleandata.rda')+ or something but with a comment in the code
-telling the future you (among others) that cleandata.rda was created
-from create-cleandata.R which in turn begins with
-\Verb+read.csv('dirtydata.csv')+ which, from a comment in the file, was
-downloaded from
-\Verb+http://www.greatfreedata.gov/dirtydata.csv+ on April 1,
-2011. Such a data processing file will typically end with something
-like \Verb+save('cleandata.rda')+ so that we are doubly certain about
-the provenance of the data.
+The principle of modularity suggests that you separate data cleaning,
+processing, recoding, and merging from analysis in different files
+\citep{nagler1995coding}. So, perhaps your analysis oriented files
+will \Verb+load('cleandata.rda')+ and a comment in the code will alert
+the future you (among others) that \Verb+cleandata.rda+ was created
+from \Verb+create-cleandata.R+ which in turn begins with
+\Verb+read.csv(\url('http://www.greatfreedata.gov/dirtydata.csv'))+. Such
+a data processing file will typically end with something like
+\Verb+save('cleandata.rda')+ so that we are doubly certain about the
+provenance of the data.
 
-Now, if in 5 years we wonder where `cleandata.rda' came from, we might
-\Verb+ grep cleandata *.R+ to search for occurrences of this file on
-our Unix/OS X system. However, if such searching among files is a
-burden, an even nicer solution is to maintain a file for each project
-called ``MANIFEST.txt'' or ``INDEX.txt'' or ``README.txt'' which lists
-the data and command files with brief descriptions of their functions
-and relations.
+Now, if in the future we wonder where \Verb+cleandata.rda+ came from, we
+might search for occurrences of 'cleandata' in the files our
+system. However, if such searching among files is a burden, an even
+nicer solution is to maintain a file for each project called
+``MANIFEST.txt'' or ``INDEX.txt'' or ``README.txt'' which lists the
+data and command files with brief descriptions of their functions and
+relations.
 
-The principle regarding data is to know where the data came from and
-what operations were performed on which set of data. I have seen
-command files which do not begin with a call to load data, and I have
-despaired. What is the meaning of \Verb+please-fit(var1 with var2)+ if
-I do not know where var1 var2 come from.
+\paragraph{Principle} We should know data where the data came from and
+what operations were performed on which set of data.
 
-\section{Version control prevents clobbering and record history.}
+Let me note that this principle was required when all our data
+analyses occurred in batch mode on VAX (and later Unix) machines. The
+fact that I need to articulate this principle at all arises because of
+the rise of interactive data analysis and graphical user interfaces:
+it is all too easy to use the mouse to load a data file into memory
+and then to write a script to analyze this file without ever noting
+the actual name or location of the data file.
+
+\section{Version control prevents clobbering and reconciles history.}
 
@@
   issue for more discussion of what version control is.} Many people
-are familiar with the ``Track Changes'' feature in modern WYSIWYG word
+are familiar with the ``track changes'' feature in modern WYSIWYG word
 processors or the fact that Dropbox allows one to recover previous
@@
 collaborating with yourself or others, it is useful to see what has
-changed, to feel free to experiment and then to dump the experiment in
-favor of previous work, to have multiple ``releases'' of the same
-document (one to MPSA, one to APSR, one to your parents) without
-requiring that spawn many copies of the same document and risk
-confusion and clobbering. Clobbering is what happens when your future
-self or your other collaborators saves an old version of a file over a
-new version --- erasing good work by accident.
+changed, to feel free to experiment and then to dump parts of the
+experiment in favor of previous work while merging the successful
+parts of the experiment into the main body of the paper, and to have
+multiple ``releases'' of the same document (one to MPSA, one to APSR,
+one to your parents) without spawning many possibly conflicting copies
+of the same document, risking confusion and clobbering. Clobbering is
+what happens when your future self or your current collaborator saves
+an old version of a file over a new version --- erasing good work by
+accident.
 
@@
 getting lost when you both try to save the same file on top of each
-other. If you find that you need to work on the same files at
-the same time, then you should work on establishing your own shared
-version control system. Free options include launchpad, github,
-sourceforge for open source projects (i.e. papers you are writing
-which you are happy to share with others). Each of those services
-include paid versions too. One may also use Dropbox as a kind of
-server for version control: checking out files from the Dropbox
-directory into a local working directory. (Notice that this is
-different from directly working on files within your Dropbox managed
-directories.)
+other. If you find that you need to work on the same files at the same
+time, then you should work on establishing your own shared version
+control system. Free options include launchpad, github, sourceforge
+for open source projects (i.e. papers you are writing which you are
+happy to share with others). Each of those services include paid
+versions too. One may also use Dropbox as a kind of server for version
+control: for example, one may copy files from the Dropbox directory
+into a local working directory so as to avoid clobbering and then
+working on merging changes by hand before copying over existing
+files. (Notice that this is different from directly working on files
+within your Dropbox managed directories.)
 
 We use subversion with our own research group, and I use it for all of
-my own projects. Subversion and bazaar and git are all great. They
-mainly differ in the extent to which you need to run a
-server. Subversion requires a server and we are lucky that the
-National Center for Supercomputing Applications provides such a server
-for us. However, the price of such hosting may not be that much using
-one of the many webhosting services out there. [See Y in this issue
-for much more about version control and collaboration.]
+my own projects (except this one, for which I am experimenting with
+git). Subversion and bazaar and git are all great. They mainly differ
+in the extent to which you need to run a server. Subversion requires a
+server and we are lucky that the NCSA provides such a server for
+us. However, the price of such hosting may not be that much using one
+of the many webhosting services out there or perhaps may be available
+for free at your university.\footnote{For example, if you already pay
+  to have a website, you may already have the right to run a
+  subversion server there.}
 
-Fancy version control systems are not required, however. I suspect
-that Google Docs allows a kind of version tracking and collaboration
-as well.  An excellent, simple, and robust version control system is
-to merely rename your files with the date and time of saving them:
-thedoc.tex becomes thedoc25-12-2011-23:50.tex.  Be sure to include
-year in the file names --- remember, the life of an idea is measured
-in years. Also, if you use this method, spend a little extra time to
+Fancy version control systems are not required, however, to get many
+of the benefits of formal version control. I suspect that Google Docs
+allows a kind of version tracking and collaboration as well.  An
+excellent, simple, and robust version control system is to merely
+rename your files with the date and time of saving them: thedoc.tex
+becomes thedoc25-12-2011-23:50.tex.  Be sure to include year in the
+file names --- remember, the life of an idea is measured in years. If
+you are wise enough to have saved your documents as plain text then
+you can easily compare documents using the many utilities available
+for comparing text files. Adobe Acrobat allows one to compare
+differences in pdf files. OpenOffice supports a ``Compare Documents''
+option.
+
+Also, if you use this method, spend a little extra time to
 ensure that you do not clobber files when you make typos in the
@@
 reconciling changes made by different collaborators by hand that
-modern version control systems (particularly the distributed kind
-these days) take care of quickly and easily.
-
-If you are wise enough to have saved your documents as plain text
-(Such as LaTeX source (with or without R/Stata code chunks)
- \footnote{A quick Google search of ``Sweave for Stata'' turned up lots
-           of resources for literate programming with Stata.} then you can
- easily compare documents using the many utilities available for
- comparing text files [FileMerge on OS X is pretty, but ediff and diff
-                       for Unix are very useful, I am sure that Windows has many other
-                       options]. Adobe Acrobat allows one to compare differences in pdf
- files. OpenOffice supports a ``Compare Documents'' option.
+modern version control systems take care of quickly and easily.
 
@@
 thedocAPSR2015.tex --- for the version eventually sent to the APSR six
- years after you presented it at APSA. The systems I mentioned above
- all allow this kind of thing and are much more elegant and capable,
- but you can do it by hand too as long as you don't mind taking up a
-lot of disk space and having many many ``thedoc...''  files around.
+years after you presented it at APSA. The formal version control
+systems I mentioned above (and which are described in more depth in X
+and Y) all allow this kind of thing and are much more elegant and
+capable, but you can do it by hand too as long as you don't mind
+taking up a lot of disk space and having  many ``thedoc...''
+files around.
+
+\paragraph{Principle} Writing is rewriting. Thus, all writing involves
+versions. When we collaborate with ourselves and others we want to
+avoid clobbering and we want to enable graceful reconciliation of
+rewriting. One can do these things with formal systems of software
+(like subversion, git, etc...) or with formal systems of file naming,
+file comparing, and communication or, even better, with both.
+
+% This is so true that the
+% source of the quote is unclear.\footnote{Could it be Paul Abbott?
+%   \url{http://www.bbc.co.uk/writersroom/insight/paul_abbott.shtml}}
+% Here is a quote with a known source: ``Writing and rewriting are a constant search for what it is one is
+% saying'' (John Updike but ??Writing With Style: Conversations on the Art of Writing, 1975 
+% by John Trimble ??)
 
@@
 
- Imagine that you reported bootstrap confidence intervals in your
- famous article. Now the statisticians at the UN are worrying about
- your bootstrap confidence interval. You'd like to evaluate your
- bootstrap procedure. Although nice code exists for bootstrapping
- linear models, no nice code exists to bootstrap the bootstrap. Of
- course, the code required is not complex, but since you are writing
- your own code you worry about getting it right. Now, 9 years after
- the idea, you've had lots of time to appreciate problems arising from
- bugs and errors in data analysis and code.
+Now, back to that confidence interval in the famous article of
+2018. The statisticians at the UN worry about your use of the
+bootstrap. The authors would like to evaluate their bootstrap
+procedure --- and they did so originally in their memo comparing it to
+a profile likelihood based approach. Although nice code exists for
+bootstrapping linear models, no nice code exists to bootstrap the
+bootstrap. Of course, the code required is not complex, but since they
+are are writing custom code they worry about getting it right. Now, 9
+years after the idea, they've had lots of time to appreciate problems
+arising from bugs and errors in data analysis and code.
 
- Now, if you had a moment to think in between teaching that new class,
+Now, if they had a moment to think in between teaching that new class,
 reading books for the awards committee, reading application files for
 the admissions committee, staying home with a sick child, and
- undertaking the odd bit of your own current research, you might say
- to yourself, ``Before I write new code, I should write a test of the
+undertaking the odd bit of your own current research, they might say to
+themselves, ``Before I write new code, I should write a test of the
 code. I should write a little bit of code that let's me know that my
@@
 collaborating with her future self and/or a small group of
- collaborators.
-
-Here is an example of this idea in a very simple form.
-I want to write a function to multiply a number by 2.
-If my function works, when I give it the number 4, I should see it return the number 8 and when I give it -4, I should get -8.
+collaborators here is an example of this idea in a very simple form. I
+want to write a function to multiply a number by 2. If my function
+works, when I give it the number 4, I should see it return the number
+8 and when I give it -4, I should get -8.
 
 <<unit.test>>=
-
 ##  The test function:
@@
   ##  This function tests times.2.fn
-    if(times.2.fn(thenumber=4) == 8 &
-       times.2.fn(thenumber=-4) == -8){
+  if (times.2.fn(thenumber=4) == 8 &
+     times.2.fn(thenumber=-4) == -8) { 
     print("It works!")
-    }else{print("It does not work!")}
+  } else { print("It does not work!")
+         }
 }
@@
 
-\section{File names send messages to your future self.}
-
-Name your files with evocative and descriptive names. Your
-collaborators are less likely to call you at midnight asking for help
-if your files are named ``inequality-and-protest-analyses.R'' than if your files are
-called ``temp9'' or ``supercalifragilisticexpialidocious.'' (Note the
-                                                            use of the extension .R to tell us that the file contains R
-                                                            commands. Use extensions like this as a standard practice to help you
-                                                            and your computer get along.)
+\paragraph{Principle} You cannot forsee all of the ways that your code
+could be used, but you can at least make sure it does what it is
+supposed to do in your particular case.
 
@@
 simply included a Sweave document and data files into a compressed
-archive \cite{bowers2005dataverse}; (2) for another more computing
+archive \citep{bowers2005dataverse}; (2) for another more computing
 intensive paper we assembled a set of files which enabled reproduction
@@
 \cite{bowers2011dataverse}. The benefit of this last system is that
-one is not required to have access to unix: the compendium is
-downloadable from within R using \Verb+install.packages()+ and it
-viewable using the \Verb+vignette()+ function. The idea that one ought
-to be able to install and run and use an academic paper just as one
-installs and uses statistical software packages is very attractive and
-I anticipate that it will become ever easier to turn papers into R
-packages as creative and energetic folks turn their attention to the
-question of reproducible research.
+one is not required to have access to a command line for \Verb+make+:
+the compendium is downloadable from within R using
+\Verb+install.packages()+ and it viewable using the \Verb+vignette()+
+function. The idea that one ought to be able to install and run and
+use an academic paper just as one installs and uses statistical
+software packages is very attractive and I anticipate that it will
+become ever easier to turn papers into R packages as creative and
+energetic folks turn their attention to the question of reproducible
+research.
 
@@
   Ask your advisers how long it took them from idea to dissertation to
-  book (articles).}  The time-frames over which collaboration are
+  publication.}  The time-frames over which collaboration are
 required --- whether among a group of people working together or
@@
 of opinion; and, (2) other people should be able to reproduce the findings. That is,
-what the work represents is not a matter of opinion but is a shared
+the work represents a shared
 experience --- and an experience shared without respect to the
-identities of others (although requiring some shared technical
+identities of others (although requiring some common technical
 training and research resources).
@@
 notes, images, ideas, etc \ldots The point is that we are making some
-claims about patterns in some collection of stuff. Now, it
-might be easy to convince others that ``this collection of stuff
-is different from this collection of stuff'' if those people were
-looking over our shoulders the whole time that we made decisions about
-collecting the stuff and broke it up into understandable
-parts and reorganized and summarized it. Unfortunately, we can't assume that people are willing to
-         shadow a researcher throughout her career. Rather, we do our work
-         alone or in small groups and want to convince other distant and future people
-         about our analyses.
+claims about patterns in some collection of stuff. Now, it might be
+easy to convince others that ``this collection of stuff is different
+from this collection of stuff'' if those people were looking over our
+shoulders the whole time that we made decisions about collecting the
+stuff and broke it up into understandable parts and reorganized and
+summarized it. Unfortunately, we can't assume that people are willing
+to shadow a researcher throughout her career. Rather, we do our work
+alone or in small groups and want to convince other distant and future
+people about our analyses.
 
@@
 in this issue allows your computerized analyses of your collections
-stuff of to be credible.  Finally, if the UN quibbles with your
+of stuff to be credible.  Finally, if the UN quibbles with your
 analyses, your future self can shoot the archive required to reproduce

### 2011-03-24 workflow.Rnw
More work

@@
 
-\title{How to improve your relationship with your future self.}
+\title{Six steps to a better relationship with your future self.}
 
 \author{Jake Bowers\thanks{I owe many thanks to Mark Fredrickson,
-    Brian Gaines, Mark Fredrickson, Kieran Healy, Kevin Quinn and Cara
+    Brian Gaines, Kieran Healy, Kevin Quinn and Cara
     Wong for direct help on this document and to Mika LaVaque-Manty
@@
 \usepackage[noae,nogin]{Sweave} %Declaring this explicitly so that I can modify the Sweave environment here rather in Sweave.sty which will be c
-\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=.5em,fontsize=\small}
-\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=.5em,fontsize=\small}
-\DefineVerbatimEnvironment{Scode}{Verbatim}  {xleftmargin=.5em,fontsize=\small}
+\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=.5em,fontsize=\footnotesize}
+\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=.5em,fontsize=\footnotesize}
+\DefineVerbatimEnvironment{Scode}{Verbatim}  {xleftmargin=.5em,fontsize=\footnotesize}
 \fvset{listparameters={\setlength{\topsep}{0pt}}}
@@
 \SweaveOpts{keep.source=TRUE}  %Show comments
-\fvset{fontsize=\small} %slightly smaller verbatim font
+\fvset{fontsize=\footnotesize} %slightly smaller verbatim font
 
@@
 
-This idea inspires a seminar paper in the spring. A conference
-paper arises from the seminar paper in collaboration with another
-student in 2012. A dissertation chapter descends from the conference
-paper in 2013. Other dissertation chapters take up 2014. A submission
-to a journal with the original co-author and a new collaborator
-happens in 2015. Revision and resubmission wait until 2017 while
-harried editors, reviewers and authors strive to balance teaching,
-service, and life. By now, the three lucky collaborators work as
-professors in three different universities. In 2018 a child is born
-and a paper is published. The United Nations takes an interest in the
-paper in 2019 and hosts a conference to discuss implications of the
-research. In 2020 a first year graduate student in a coffee shop has
-an idea that challenges the results in the now famous paper. What
-would happen if the authors had controlled for X? Or run one more MCMC chain?
-Or chosen a different likelihood function?  Will the United Nations
-(now eager to act based the paper) make a wrong move?
+This idea inspires a seminar paper in the spring. A conference paper
+arises from the seminar paper in collaboration with another student in
+2012. A dissertation chapter descends from the conference paper in
+2013. Other dissertation chapters take up 2014. A submission to a
+journal with the original co-author and a new collaborator happens in
+2015. Revision and resubmission wait until 2017 while harried editors,
+reviewers and authors strive to balance research, teaching, service,
+and life. By now, the three lucky collaborators work as professors in
+three different universities. In 2018 a child is born and a paper is
+published. The United Nations takes an interest in the paper in 2019
+and hosts a conference to discuss implications of the research. In
+2020 a first year graduate student in a coffee shop has an idea that
+challenges the results in the now famous paper. What would happen if
+the authors had controlled for X? Or included information now
+available but missing in 2012?  Or chosen a different likelihood
+function?  Will the United Nations (now eager to act based the paper)
+make a wrong move?
 
@@
   manual driving of flying cars will have been outlawed in his state
-  in 2017.} The group must go back to the analyses. Which ones?  The
+  by 2020.} The group must go back to the analyses. Which ones?  The
 ones from 2011?  Or 2012? Or 2018? Where are the files? The next day,
@@
 nostalgia finds some of the files.\footnote{This is the same guy who
-  still owns cassette tapes and compact discs.} Now the checks and
+  still owns cassette tapes and compact discs.} Now
 re-analyses should be easy. Right? The student, now professor, should
@@
 for that crucial table --- and that mousing should not be time
-consuming to redo exactly as they were done in 2011, or was it
+consuming to redo exactly as it was done in 2011 \ldots or was it
 2015?). Right? And, of course, the way Microsoft
-Word/Stata/SPSS/R/LISREL stores files and the way that current
+Word/Stata/SPSS/R/LISREL understands files and the way that 
 machines in 2020 read and write them is the same --- since Windows and
@@
 Right? Wrong? If the collaborators find themselves answering ``wrong''
-to many of the questions posed here then reproducing the analyses will
-take a lot of time if such reproduction is possible at all. This essay
-provides some suggestions for practices which will make such
-reproduction occur much more easily and quickly in the event that
-famous papers require special scrutiny. Specifically, this piece aims
-to amplify some of what we already ought to know from
-\citet{nagler1995coding} and \citet{king1995replication}, and to
-update some of those ideas given current practices, platforms, and
-propensities.
+to the questions posed here then reproducing, updating, or changing
+the analyses will take a lot of time. If reproduction is hard to do,
+then the reputations of the scholars will suffer and, more
+importantly, the UN will be unable to implement a master plan for
+world peace. This essay provides some suggestions for practices which
+will make such reproduction occur much more easily and quickly in the
+event that famous papers require special scrutiny. Specifically, this
+piece aims to amplify some of what we already ought to know from
+\citet{nagler1995coding} and \citet{king1995replication}, and to add
+to some of those ideas given current practices, platforms, and
+possibilities.
 
@@
 
-\begin{Verbatim}[fontsize=\small]
+\begin{Verbatim}[fontsize=\footnotesize]
 thedata<-read.csv("Data/thedata-15-03-2011.csv") ## Read the data
@@
 please-plot(outcome by explanatory using thedata. red lines please.)
+please-add-a-line(using model1)
+## Note to self: the following is no different from model1
+## model2<-please-fit(outcome by explanatory+explanatory^2 using thedata
+## summary(abs(fitted(model1)-fitted(model2)))
 dev.off() ## stop writing to the pdf file
@@
 Now, in the future if I wonder how ``that plot on page 10'' was
-created: (1) ``that plot'' is from a file called \Verb+fig1.pdf+ and
-\Verb+fig1.pdf+ was created in \Verb+fig1.R+. In a future where R still
-exists, if I wanted to change the figure, I could edit the commands
-quickly and easily to do so (since R commands are written in plain
-text, and plain text is a format that will be around for a long time
-to come).
+created, I will know: (1) ``that plot'' is from a file called
+\Verb+fig1.pdf+ and (2) \Verb+fig1.pdf+ was created in
+\Verb+fig1.R+. In a future where R still exists, if I wanted to change
+the figure, I could edit the commands quickly and easily to do so
+(since R commands are written in
+\href{http://en.wikipedia.org/wiki/Plain_text}{plain text}, and plain
+text is a format that will be around as long as computer programmers
+write computer programs). In a future where R does not exist, I will
+at least be able to read the plain text R commands and use them to
+write code in my new favorite statistical computing
+language.\footnote{A most, since R is open source, I will be able to
+  download R, download an old fashioned open-source operating systems
+  (like Ubuntu 10), and run the old-fashioned statistical computing
+  environment in the old-fashioned operating system in a virtual
+  machine on my new-fashioned actual machine.}
 
-\paragraph{Principle} Know the provenance of your results so that your
+\paragraph{Step 1} Know the provenance of your results so that your
 future self or current collaborators can quickly and easily reproduce
@@
 Data analysis involves a long series of decisions. Each decision
-requires justification.  The text itself will
-justify the most important decisions. Some decisions will be too small
-and technical for inclusion in the paper itself. These need to be
-documented in the code itself \citep{nagler1995coding}.
+requires justification.  Some decisions will be too small and
+technical for inclusion in the paper itself. These need to be
+documented in the code itself \citep{nagler1995coding}. Paragraphs and
+citations in the article itself will justify the most important
+decisions.
 
-\subsection{Comment your code.}
+\subsection{Code to communicate: Comment your code.}
 Comments --- unexecuted text inside of a script --- are a message to
-your future self and other consumers of your work. Luckily, if you are
-using a literate programming practice (ex. Sweave or odfWeave), you
-can write paragraphs to surround your code as well as technical
-comments in the code itself. If you are not using a strictly literate
-programming practice then use whatever commenting protocol exists in
-your analysis language. R, for example, considers text marked with
-``\#'' as a comment. Here for example I comment about a linear model
-run for a paper and also include comments about another model run for
-due diligence.
+collaborators (including your future self) and other consumers of your
+work. In the above code, I used comments to explain the lines to
+readers unfamiliar with R and to remember that I had tried a different
+specification but decided not to use it because adding the squared
+term did not really change the predictions.\footnote{R considers text
+  marked with ``\#'' as a comment.}
 
-<<reg1,echo=TRUE,results=verb>>=
-##  Repeat the regression run for Table~\ref{tab:protest}
-##  Do political rights predict protest reporting?
-lm1 <- lm(protac2000 ~ I(Gini2004/100) + meanpr, data=good.df)
-##  Q: Was it worthwhile controlling for political rights?
-lm2 <- lm(protac2000~I(Gini2004/100),data=good.df)
-coef(lm2)[2]-coef(lm1)[2]
-##  A: Yes (around 3/4 an act of protest difference)
-@
+% Luckily, if you are
+% using a literate programming practice (ex. Sweave or odfWeave), you
+% can write paragraphs to surround your code as well as technical
+% comments in the code itself. If you are not using a strictly literate
+% programming practice then use whatever commenting protocol exists in
+% your analysis language. R, for example, considers text marked with
+% ``\#'' as a comment. Here for example I comment about a linear model
+% run for a paper and also include comments about another model run for
+% due diligence.
+
+% <<reg1,echo=TRUE,results=verb>>=
+% ##  Repeat the regression run for Table~\ref{tab:protest}
+% ##  Do political rights predict protest reporting?
+% lm1 <- lm(protac2000 ~ I(Gini2004/100) + meanpr, data=good.df)
+% ##  Q: Was it worthwhile controlling for political rights?
+% lm2 <- lm(protac2000~I(Gini2004/100),data=good.df)
+% coef(lm2)[2]-coef(lm1)[2]
+% ##  A: Yes (around 3/4 an act of protest difference)
+% @
 
 Messages left for your future self (or near-future others) help
-retrace and justify the steps of your data-analytic journey in case
-you ever need to revist them as the work moves from seminar paper to
-conference paper to poster back to paper to dissertation and
-onwards. Notice one another benefit of coding for an audience: we
-learn by teaching. By assuming that others will look at your code, you
-will be more likely to write clearer code, or perhaps even to learn
-more about what you are doing as you do it.
+retrace and justify your decisions. If you ever need to revist
+them as the work moves from seminar paper to conference paper to
+poster back to paper to dissertation and onwards. Notice one another
+benefit of coding for an audience: we learn by teaching. By assuming
+that others will look at your code, you will be more likely to write
+clearer code, or perhaps even to learn more about what you are doing
+as you do it.
 

### 2011-03-25 workflow.Rnw
More work

@@
 \usepackage[colorlinks=TRUE]{hyperref}
+\usepackage{microtype}
 
@@
 \begin{document}
+\frenchspacing
+\VerbatimFootnotes
 \maketitle
@@
           scipen=8,
-          width=100,
+          width=50,
           error=function(){options(prompt="> ",continue="+ ");NULL}
@@
 
-Right? Wrong? If the collaborators find themselves answering ``wrong''
-to the questions posed here then reproducing, updating, or changing
-the analyses will take a lot of time. If reproduction is hard to do,
-then the reputations of the scholars will suffer and, more
-importantly, the UN will be unable to implement a master plan for
-world peace. This essay provides some suggestions for practices which
-will make such reproduction occur much more easily and quickly in the
-event that famous papers require special scrutiny. Specifically, this
-piece aims to amplify some of what we already ought to know from
-\citet{nagler1995coding} and \citet{king1995replication}, and to add
-to some of those ideas given current practices, platforms, and
-possibilities.
+Right? Wrong? If the collaborators find themselves say ``Wrong'' in
+answer to the questions posed here then reproducing, updating, or
+changing the original analyses will take a lot of time. If
+reproduction is hard to do, then the reputations of the scholars will
+suffer and, more importantly, world peace will have been delayed. This
+essay provides some suggestions for practices which will make such
+reproduction occur much more easily and quickly in the event that
+famous papers require special scrutiny. Specifically, this piece aims
+to amplify some of what we already ought to know
+\citep{king1995replication,nagler1995coding}, and to add to some of
+those ideas given current practices, platforms, and possibilities.
 
@@
 good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
+good.df$gini04<-good.df$Gini2004/100
+good.df$protac00<-good.df$protac2000
 @
@@
 
-Using R, for example, I might specify that the file \Verb+fig1.pdf+
+Using R \citep{R:2011}, for example, I might specify that the file \Verb+fig1.pdf+
 was produced by the following commands in a file called
@@
 \begin{Verbatim}[fontsize=\footnotesize]
-thedata<-read.csv("Data/thedata-15-03-2011.csv") ## Read the data
+thedata <- read.csv("Data/thedata-15-03-2011.csv") ## Read the data
 pdf('fig1.pdf') ## begin writing to the pdf file
@@
 please-add-a-line(using model1)
-## Note to self: the following is no different from model1
+## Note to self: a quadratic term does not add to the substance
 ## model2<-please-fit(outcome by explanatory+explanatory^2 using thedata
@@
 \Verb+fig1.pdf+ and (2) \Verb+fig1.pdf+ was created in
-\Verb+fig1.R+. In a future where R still exists, if I wanted to change
-the figure, I could edit the commands quickly and easily to do so
-(since R commands are written in
+\Verb+fig1.R+. In a future where R still exists, changing the figure
+will require quick edits of commands already written.  In a future
+where R does not exist, I will at least be able to read the plain text
+R commands and use them to write code in my new favorite statistical
+computing language: R scripts are written in
 \href{http://en.wikipedia.org/wiki/Plain_text}{plain text}, and plain
 text is a format that will be around as long as computer programmers
-write computer programs). In a future where R does not exist, I will
-at least be able to read the plain text R commands and use them to
-write code in my new favorite statistical computing
-language.\footnote{A most, since R is open source, I will be able to
-  download R, download an old fashioned open-source operating systems
-  (like Ubuntu 10), and run the old-fashioned statistical computing
-  environment in the old-fashioned operating system in a virtual
-  machine on my new-fashioned actual machine.}
+write computer programs. \footnote{What is more, since R is open
+  source, I will be able to download an old version of R, download an
+  old fashioned open-source operating system (like Ubuntu 10), and
+  run the old-fashioned statistical computing environment in the
+  old-fashioned operating system in a virtual machine on my
+  new-fashioned actual machine.}
+
+Moreover, realize that file names send messages to your future
+self. Name your files with evocative and descriptive names. Your
+collaborators are less likely to call you at midnight asking for help
+if your files are named \Verb+inequality-and-protest-figures.R+ than
+if your files are called \Verb+temp9+ or
+\Verb+supercalifragilisticexpialidocious+. The use of the extension
+\Verb+.R+ tells us and the operating system that the file contains R
+commands. This part of the filename enables us to quickly search our
+antique hard drives for files containing R scripts.
 
@@
 future self or current collaborators can quickly and easily reproduce
-your work (and make changes).% --- to know what
+and thus change your work.% --- to know what
 % commands created which figures and tables.
@@
 requires justification.  Some decisions will be too small and
-technical for inclusion in the paper itself. These need to be
-documented in the code itself \citep{nagler1995coding}. Paragraphs and
-citations in the article itself will justify the most important
-decisions.
+technical for inclusion in the published article itself. These need to
+be documented in the code itself \citep{nagler1995coding}. Paragraphs
+and citations in the publication will justify the most important
+decisions. So, one must code to communicate with yourself and
+others. There are two main ways to avoid forgetting the reasons you
+did something with data: comment your code and tightly link your code
+with your writing.\footnote{One can also try the R command
+  \Verb+put-it-in-my-mind(reason,importance='high')+ to firmly place a
+  reason for a decision into the mind of the analyst. I myself have
+  not had much luck with this function.}
 
@@
 collaborators (including your future self) and other consumers of your
-work. In the above code, I used comments to explain the lines to
+work. In the above code chunk, I used comments to explain the lines to
 readers unfamiliar with R and to remember that I had tried a different
 specification but decided not to use it because adding the squared
-term did not really change the predictions.\footnote{R considers text
-  marked with ``\#'' as a comment.}
+term did not really change the substantive story arising from the
+model.\footnote{R considers text marked with \Verb+#+ as a comment.}
+Messages left for your future self (or near-future others) help
+retrace and justify your decisions as the work moves from seminar
+paper to conference paper to poster back to paper to dissertation and
+onwards.
 
@@
 % ##  Do political rights predict protest reporting?
-% lm1 <- lm(protac2000 ~ I(Gini2004/100) + meanpr, data=good.df)
+% lm1 <- lm(protac00 ~ gini04 + meanpr, data=good.df)
 % ##  Q: Was it worthwhile controlling for political rights?
-% lm2 <- lm(protac2000~I(Gini2004/100),data=good.df)
+% lm2 <- lm(protac00~gini04,data=good.df)
 % coef(lm2)[2]-coef(lm1)[2]
@@
 
-Messages left for your future self (or near-future others) help
-retrace and justify your decisions. If you ever need to revist
-them as the work moves from seminar paper to conference paper to
-poster back to paper to dissertation and onwards. Notice one another
-benefit of coding for an audience: we learn by teaching. By assuming
-that others will look at your code, you will be more likely to write
-clearer code, or perhaps even to learn more about what you are doing
-as you do it.
+Notice one another benefit of coding for an audience: we learn by
+teaching. By assuming that others will look at your code, you will be
+more likely to write clearer code, or perhaps even to learn more about
+what you are doing as you do it.
 
 Comment liberally. Comments are discarded when R runs analysis or
-LaTeX turns text into a pdf, so only those who dig into your work will
-see them.
+\LaTeX~turns plain text into page images, so only those who dig into
+your work will see them.
 
-\subsection{Programming can be literate.}
+\subsection{Code to communicate: Literate programming.}
 
-Sometimes simple comments in the main command file are not enough:
-imagine if the authors had arrived at the bootstrap only after running
-simulations to assess the coverage of that procedure compared to a
-profiled likelihood based approach. Rather than include time consuming
-simulations in the body of their main Sweave document, they might have
-written some memos that discuss the tradeoffs and implementation
-decisions involved in such a comparison. In the main text they would
-have just read in the results of the simulation study with an
-appropriate comment:
+Imagine you discover something new (or confirm something old). You
+produce a nice little report on your work for use in discussions of
+your working group or as a memo for a web or reviewer appendix. The
+report itself is a pdf file or some other format which emphasizes
+reading. Eventually pieces of that report (tables, graphs, paragraphs)
+ought to contribute to the publishable paper. Re-create those analyses
+by pointing, clicking, copying, or pasting would invite typing error
+and waste time. Re-creating your arguments justifying why you dropped
+China but not India from your analyses or why you used a quadratic
+term but not a cubic term would also waste time. More importantly, we
+and others want to know why we did what we did. Such explanations may
+not be very clear if we have some pages of printed code in one hand
+and a manuscript in the other.
 
-<<bsreg1,echo=TRUE,results=verb,eval=FALSE>>=
-## We assessed the coverage of the bootstrap intervals against 
-## direct inspection of the likelihood profile in the file
-## Memos/shouldwebootstrap.Rnw which produced bsresults.rda
+How might one avoid these problems?
+\href{http://en.wikipedia.org/wiki/Literate_programming}{Literate
+  programming} is the practice of weaving code into a document ---
+paragraphs, equations, and diagrams can explain the code, and the code
+can numbers, figures, and tables (and diagrams and even equations and
+paragraphs). Literate programming is not merely fancy commenting but
+is about enabling the practice of programming itself to enable easy
+reproduction and communication.
 
-##Here we read in the results of the bootstrap procedure:
-load("Memos/bsresults.rda") ## using a compressed binary format here for speed
-@
-
-Now that memo itself (or perhaps the main document) probably involves
-sentences and paragraphs explaining an analysis. Literate programming
-or literate data analysis happens when the code of the analysis is
-mixed side by side with the paragraphs explaining, interpreting,
-foreshadowing, discussing the analysis.\footnote{This document is
-  written in such a way using \LaTeX and R code mixed together in what
-  is known as an Sweave file \citep{Leis:2005}. You can download and
-  view and edit this document as a plain text file from
-  \url{https://github.com/jwbowers/workflow}.}
-
-If comments are guideposts, literate programs are full fledged
-travelogues. Why might one care to use a literate programming
-practice?
-
-Say you run some command, discover something new (or confirm something
-old) and then make a claim backed up by table or a figure. You produce
-a nice little report on your work for use in discussions of the
-working group. The report itself is a pdf file or some other format
-which emphasizes reading.\footnote{like html, postscript, dvi or even
-  the wordprocessor formats like rtf or doc or odt, or the web formats
-  like html.}  Eventually we want to use pieces of that report
-(tables, graphs, paragraphs) in a publishable paper. It would be a
-waste of time if you had to re-create those analyses by pointing,
-clicking, copying, or pasting. It would also be a waste of time if you
-couldn't remember why you decided to use a quadratic instead of a
-cubic term and had to re-create that argument. So, we might use
-literate programming because we don't want to waste time.
-
-For example, in \S~\ref{sec:data-analys-comp}, I suggested that we
-know where ``that plot on page 10'' comes from by making sure we had a
-``fig1.pdf'' file produced from a clearly commented plain text file
-called something like ``fig1.R.'' If one were using Sweave, one could
-even more easily have written the following to produce the desired
-figure.
+For example, in \S\ref{sec:data-analys-comp}, I suggested that we know
+where ``that plot on page 10'' comes from by making sure we had a
+\Verb+fig1.pdf+ file produced from a clearly commented plain text file
+called something like \Verb+fig1.R+.  An even easier solution would be to
+directly include a chunk of code to produce the figure inside of the
+paper itself. This paper, for example, was written in plain text using
+\LaTeX~ markup with R code chunks to make things like
+Figure~\ref{fig:giniprot}: this combination of \LaTeX and R is called
+Sweave \citep{Leis:2005}.\footnote{Support for Sweave is included with
+  R.}
 
 \begin{Verbatim}[fontsize=\footnotesize]
+This paper, for example, was written in plain text using \LaTeX
+markup with R code chunks to make things like
+Figure~\ref{fig:giniprot}: this combination of \LaTeX and R is
+called Sweave \citep{Leis:2005}.\footnote{Support for Sweave is
+  included with R.}
 \begin{figure}[H]
-\begin{center}
- <<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>=
- par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
+\centering
+ <<fig1plot,echo=FALSE,fig=TRUE>>=
  ##  Make a scatterplot of Protest by Inequality
- with(good.df,plot(Gini2004/100,protac2000,
+ with(good.df,plot(gini04,protac00,xlab='Gini Coefficient 2004 (UNDP)',
+                   ylab='Mean Protest Activities\n(World Values Survey 1980-2000)'))
+ ##  Label a few interesting points
+ with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
+      text(gini04,protac00,labels=Nation))
+ @
+\caption{Protest activity by income inequality \citep[from][]{norris2009data}.}
+\label{fig:giniprot}
+\end{figure}
+\end{Verbatim}
+
+<<fig1code,echo=FALSE,fig=FALSE>>=
+##  Make a scatterplot of Protest by Inequality
+with(good.df,plot(gini04,protac00,
                   xlab='Gini Coefficient 2004 (UNDP)',
@@
 with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
-      text(Gini2004/100,protac2000,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
- @
-\caption{Protest activity by income inequality \citep[from][]{norris2009data}.}\label{fig:giniprot}
-\end{center}
-\end{figure}
-\end{Verbatim}
-
-<<fig1code,echo=FALSE,fig=FALSE>>=
-##  Make a scatterplot of Protest by Inequality
-with(good.df,plot(Gini2004/100,protac2000,
-                  xlab='Gini Coefficient 2004 (UNDP)',
-                  ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
-                  cex=.8))
-
-##  Label a few interesting points
-with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
-     text(Gini2004/100,protac2000,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
+     text(gini04,protac00,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
 @
@@
   \begin{center}
-<<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>=
+<<fig1plot,echo=FALSE,fig=TRUE,width=4,height=4>>=
 par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
@@
 @
-    \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}\label{fig:giniprot}
+    \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}
+    \label{fig:giniprot}
   \end{center}
@@
 \Verb+fig1.pdf+ file.  Tables and other numerical results are also
-possible to generate within ``code chunks'' in the source code of a
+possible to generate within in the source code of a
 scholarly paper. Those who view the code for this essay will see how
 Table~\ref{tab:protest} was also generated directly from a regression
-object.\footnote{See [cite to Beck TPM] for the inspiration for this
+object.\footnote{\citet{beck2010reg} inspired this
   particular presentation of a linear model.}
@@
 <<maketable, echo=FALSE, results=hide>>=
-  lm1<-lm(protac2000~I(Gini2004/100)+meanpr,data=good.df)  #  Run the regression
+  lm1<-lm(protac00~gini04+meanpr,data=good.df)  #  Run the regression
 
@@
 library(xtable)
-thextab<-xtable(themat,align=c("l",rep(".",4)),
+thextab<-xtable(themat,align=c("l",rep(".",4)),digits=1,
                 caption="People living in countries with unequal
@@
             paste("& \\multicolumn{4}{c}{n: ",n,",",
-                  "   resid.sd:",signif(est.sigma),
-                  ", R$^2$:",r2,"} \\\\ \\bottomrule",sep="")
+                  "   resid.sd: ",signif(est.sigma),
+                  ", R$^2$: ",r2,"} \\\\ \\bottomrule",sep="")
             )
@@
 
-If your workflow does not involve \LaTeX and R, you can still
-implement some of the principles here (and if you use OpenOffice or
-LyX and R you can use \Verb+odfWeave+ do directly do literate data
-analysis in a WYSIWYG environment).\footnote{A quick Google search of
-  ``Sweave for Stata'' turned up lots of resources for literate
-  programming with Stata.}
+Literate data analysis is not the same as Sweave, even if Sweave is a
+nice implementation.\footnote{The R project has a task view devoted to
+  \href{http://cran.r-project.org/web/views/ReproducibleResearch.html}{reproducible
+    research} listing many of the different approaches to literate
+  programming for R.} If your workflow does not involve \LaTeX~ and R,
+you can still implement some of the principles
+here. \href{http://www.lyx.org/}{LyX} offers a WYSIWYG environment for
+\LaTeX~ which supports Sweave. And the \Verb+odfWeave+ package in R
+allows the use of OpenOffice documents in exactly the same
+way.\footnote{A quick Google search of ``Sweave for Stata'' turned up
+  lots of resources for literate programming with Stata.} Something
+other than than Sweave will make our lives easier in the future. Then
+we'll change.
 
-Literate data analysis is not the same as Sweave, even if Sweave is a
-nice implementation.\footnote{See
-  \url{http://cran.r-project.org/web/views/ReproducibleResearch.html}
-  for many of the different approaches to literature programming for
-  R.} For example, the person most often credited with
-these ideas used plain \TeX as the ``document formatting'' language
-and Pascal as the programming language [cite]. See [X] for an example
-of using the Emacs org-mode as literate programming. And, one could
-imagine a Word or HTML file in which code chunks are inserted by
-commented out and run one at a time (or perhaps extracted by a clever
-script from the XML source file of a modern Microsoft Word Document)
---- or an HTML file in which hyperlinks are inserted to plain text
-files containing the relevant code chunks.
+\paragraph{Step \thesection}
 
-\paragraph{Principle} Keep in mind the distinction between the
+\begin{quote}
+  Let us change our traditional attitude to the construction of
+  programs: Instead of imagining that our main task is to instruct a
+  computer what to do, let us concentrate rather on explaining to
+  human beings what we want a computer to do.
+  \citep[][p. 97]{knuth1984literate}
+\end{quote}
+
+If we focus on explaining our data analysis to human beings, we will
+do a better job with the data analysis itself: we will learn as we
+focus on teaching, and we will avoid errors and wasting time as we
+ensure that others (including our future selves) can retrace our
+steps.
+
+Keep in mind the distinction between the
 ``source code'' of a document (i.e. what computation was required to
@@
 
-\subsection{File names send messages to your future self.}
-
-Name your files with evocative and descriptive names. Your
-collaborators are less likely to call you at midnight asking for help
-if your files are named ``inequality-and-protest-analyses.R'' than if
-your files are called ``temp9'' or
-``supercalifragilisticexpialidocious.'' (Note the use of the extension
-.R to tell us that the file contains R commands. Use extensions like
-this as a standard practice to help you and your computer get along.)
-
-\paragraph{Principle} Comments and other documentation enable you to
+\paragraph{Step \thesection} Comments and other documentation enable you to
 learn through teaching, code efficiently, and avoid duplicating your
@@
 
-\paragraph{Principle} We should know data where the data came from and
+\paragraph{Step \thesection} We should know data where the data came from and
 what operations were performed on which set of data.
@@
 
-\paragraph{Principle} Writing is rewriting. Thus, all writing involves
+\paragraph{Step \thesection} Writing is rewriting. Thus, all writing involves
 versions. When we collaborate with ourselves and others we want to
@@
 
-\paragraph{Principle} You cannot forsee all of the ways that your code
+\paragraph{Step \thesection} You cannot forsee all of the ways that your code
 could be used, but you can at least make sure it does what it is
@@
   Ask your advisers how long it took them from idea to dissertation to
-  publication.}  The time-frames over which collaboration are
-required --- whether among a group of people working together or
-within a single scholar's productive life or probably both --- are
-much longer than any given version of any given software will easily
-exist. \href{http://en.wikipedia.org/wiki/Plain_text#Usage}{Plain
-  text} is the exception. Thus, even as we extol version control
-systems, one must have a way to ensure future access to them in a form
-that will still be around when sentient cockroaches finally join
-political science departments (by then dominated by cetaceans after
-humans are mostly uploads).\footnote{The arrival of the six-legged
-  social scientists revives Emacs and finally makes Ctrl-a Ctrl-x
-  Esc-x Ctrl-c a \href{http://kieran.healy.usesthis.com/}{reasonable
-    key combination}.}
+  publication.}  The time-frames over which collaboration are required
+--- whether among a group of people working together or within a
+single scholar's productive life or probably both --- are much longer
+than any given version of any given software will easily exist. Plain
+text is the exception. Thus, even as we extol version control systems,
+one must have a way to ensure future access to them in a form that
+will still be around when sentient cockroaches finally join political
+science departments (by then dominated by cetaceans after humans are
+mostly uploads).\footnote{The arrival of the six-legged social
+  scientists revives Emacs and finally makes Ctrl-a Ctrl-x Esc-x
+  Ctrl-c a \href{http://kieran.healy.usesthis.com/}{reasonable key
+    combination}.}
 

### 2011-03-25 workflow.Rnw
Almost done

@@
 2020 a first year graduate student in a coffee shop has an idea that
-challenges the results in the now famous paper. What would happen if
-the authors had controlled for X? Or included information now
-available but missing in 2012?  Or chosen a different likelihood
-function?  Will the United Nations (now eager to act based the paper)
-make a wrong move?
+challenges the results in the now famous paper. She presents her paper
+at a conference in 2021. What would happen if the authors had
+controlled for X? Or included information now available but missing in
+2012?  Or chosen a different likelihood function?  Will the United
+Nations (now eager to act based the paper) make a wrong move?
 
@@
 Word/Stata/SPSS/R/LISREL understands files and the way that 
-machines in 2020 read and write them is the same --- since Windows and
+machines in 2021 read and write them is the same --- since Windows and
 Mac OS X have always existed and will always continue to exist more or
@@
 if your files are called \Verb+temp9+ or
-\Verb+supercalifragilisticexpialidocious+. The use of the extension
+\Verb+supercalifragilisticexpialidocious+. The extension
 \Verb+.R+ tells us and the operating system that the file contains R
@@
 
-Notice one another benefit of coding for an audience: we learn by
+Notice one other benefit of coding for an audience: we learn by
 teaching. By assuming that others will look at your code, you will be
-more likely to write clearer code, or perhaps even to learn more about
-what you are doing as you do it.
+more likely to write clearer code, or perhaps even to think more
+deeply about what you are doing as you do it.
 
@@
 your working group or as a memo for a web or reviewer appendix. The
-report itself is a pdf file or some other format which emphasizes
-reading. Eventually pieces of that report (tables, graphs, paragraphs)
-ought to contribute to the publishable paper. Re-create those analyses
-by pointing, clicking, copying, or pasting would invite typing error
-and waste time. Re-creating your arguments justifying why you dropped
-China but not India from your analyses or why you used a quadratic
-term but not a cubic term would also waste time. More importantly, we
-and others want to know why we did what we did. Such explanations may
-not be very clear if we have some pages of printed code in one hand
-and a manuscript in the other.
+report itself is a pdf file or some other format which displays page
+images to ease reading rather than to encourage reanalysis and
+rewriting. Eventually pieces of that report (tables, graphs,
+paragraphs) ought to show up in, or at least inform, the publishable
+paper. Re-creating those analyses by pointing, clicking, copying, or
+pasting would invite typing error and waste time. Re-creating your
+arguments justifying your analysis decisions would also waste
+time. More importantly, we and others want to know why we did what we
+did. Such explanations may not be very clear if we have some pages of
+printed code in one hand and a manuscript in the other.
 
@@
   included with R.}
-\begin{figure}[H]
+\begin{figure}[h]
 \centering
- <<fig1plot,echo=FALSE,fig=TRUE>>=
+ <<fig1plot,fig=TRUE>>=
  ##  Make a scatterplot of Protest by Inequality
@@
 By using \Verb+\label{fig:giniprot}+, I do not need to keep track of
-the figure number. Nor do I need a separate \Verb+fig1.R+ file or
-\Verb+fig1.pdf+ file.  Tables and other numerical results are also
-possible to generate within in the source code of a
-scholarly paper. Those who view the code for this essay will see how
-Table~\ref{tab:protest} was also generated directly from a regression
-object.\footnote{\citet{beck2010reg} inspired this
+the figure number, nor do extra work when I reorganize the document in
+response to reviewer suggestions. Nor do I need a separate
+\Verb+fig1.R+ file or \Verb+fig1.pdf+ file.  Tables and other
+numerical results are also possible to generate within the source
+code of a scholarly paper. Those who view the code for this essay will
+see how Table~\ref{tab:protest} was also generated directly from a
+regression object.\footnote{\citet{beck2010reg} inspired this
   particular presentation of a linear model.}
@@
     research} listing many of the different approaches to literate
-  programming for R.} If your workflow does not involve \LaTeX~ and R,
-you can still implement some of the principles
-here. \href{http://www.lyx.org/}{LyX} offers a WYSIWYG environment for
-\LaTeX~ which supports Sweave. And the \Verb+odfWeave+ package in R
-allows the use of OpenOffice documents in exactly the same
-way.\footnote{A quick Google search of ``Sweave for Stata'' turned up
-  lots of resources for literate programming with Stata.} Something
-other than than Sweave will make our lives easier in the future. Then
-we'll change.
+  programming for R.}  \href{http://www.lyx.org/}{LyX} offers a
+WYSIWYG environment for \LaTeX~ which supports Sweave. And the
+\Verb+odfWeave+ package in R allows the use of OpenOffice documents in
+exactly the same way.\footnote{A quick Google search of ``Sweave for
+  Stata'' turned up lots of resources for literate programming with
+  Stata.} If your workflow does not involve \LaTeX~ and R, you can
+still implement some of the principles here. Imagine creating a style
+in Microsoft Word called ``code'' which hides your code when you print
+your document, but which allows you do at least run each code chunk
+piece by piece [or perhaps there are ways to extract all text of style
+``code'' from a Microsoft Word document]. Or imagine just using some
+other kind of indication linking paragraphs to specific places in code
+files. There are many ways that creative people can program in a
+literate way.
+
+Literate programming need not go against the principle of modular data
+analysis \citep{nagler1995coding}: in my own work I routinely have
+several different Sweave files that fulfill different functions, some
+of them create \LaTeX code that I can \Verb+\input+ into my
+\Verb+main.tex+ file, others setup the data, run simulations, or allow
+me to record my journeys down blind alleys. Of course, when we have
+flying cars running on autopilot, perhaps something other than than
+Sweave will make our lives even easier. Then we'll change.
 
@@
 
-If we focus on explaining our data analysis to human beings, we will
-do a better job with the data analysis itself: we will learn as we
-focus on teaching, and we will avoid errors and wasting time as we
-ensure that others (including our future selves) can retrace our
-steps.
+We analyze data in order to explain something about world to other
+scholars and policy makers. If we focus on explaining how we got our
+computers to do data analysis to human beings, we will do a better job
+with the data analysis itself: we will learn as we focus on teaching,
+and we will avoid errors and save time as we ensure that others
+(including our future selves) can retrace our steps. A document than
+can be ``run'' to reproduce all of the analyses also instills
+confidence in readers and can more effectively spur discussion and
+learning and cumulation of research as people no longer need spend
+weeks attempting to reproduce research which inspired them and on
+which they desire to build.
 
-Keep in mind the distinction between the
-``source code'' of a document (i.e. what computation was required to
-produce it) and the visible, type-set page image. Page images are
-great for reading, but not great for reproducing or collaborating. The
-source code of any document exchanged by the group must be available
-and executable.
-
-In order to know that ``that plot on page 10'' is from a file called
-``fig1.pdf'' you need to document it somehow. I can imagine a system
-with very disciplined use of a word processor perhaps using special
-auxiliary ``listoffigures.txt'' or ``MANIFEST.txt'' or ``README'' or
-``Makefile'' files or something. Right now it is hard to beat Sweave
-or some plain text combination of \LaTeX (or some other markup based
-document formatting system with support for citations, etc... ) with
-some plain text based statistical programming environment (like R or
-Stata or Matlab, etc.) as a system for enhancing the collaboration of
-research groups doing quantitative work in political
-science. Something other than than Sweave will make our lives easier
-in the future. Then we'll change.\footnote{For example, X uses the
-  org-mode markup system in this issue.}
-
-\paragraph{Step \thesection} Comments and other documentation enable you to
-learn through teaching, code efficiently, and avoid duplicating your
-effort. Literate programming also enables you to avoid the kinds of
-errors which occur when you re-type numbers in to tables or
-graphic-producing systems or even into a document itself. Finally, and
-most importantly, a document that can be ``run'' to reproduce all of
-the results is a document that can more effectively spur discussion
-and learning and cumulation of research as people no longer need spend
-weeks attempting to reproduce research on which they desire to build.
-
-\url{http://en.wikipedia.org/wiki/Literate_programming}
+Keep in mind the distinction between the ``source code'' of a document
+(i.e. what computation was required to produce it) and the visible,
+type-set page image. Page images are great for reading, but not great
+for reproducing or collaborating. The source code of any document
+exchanged by the group must be available and executable.
 
@@
 Using R, for example, all analysis files should have
-\Verb+load('thedata.rda')+ or
-\Verb+read.csv('http://www.mywebsite.org/Data/thedata.csv')+ or some
+\Verb+load('thedata.rda')+ or \Verb+read.csv('thedata.csv')+ or some
 equivalent line in them, and \Verb+thedata.csv+ should be stored in
@@
 \Verb+load('Data/thedata.rda')+ means the same thing on all computers
-used to work on the project.
+used to work on the project. This kind of solution is one of the
+things that formal version control systems do well (as discussed
+a bit more in \S\ref{sec:vers-contr-prev}).
 
@@
 \Verb+save('cleandata.rda')+ so that we are doubly certain about the
-provenance of the data.
+provenance of the data.\footnote{Of course, if you need math or
+  paragraphs to explain what is happening in these files, you might
+  prefer to make them into Sweave files, for which the conventional
+  extension is .Rnw. So you'd have \Verb+create-cleandata.Rnw+ which
+  might explain and explore the different coding decisions you made,
+  perhaps involving a factor analysis and diagnostic plots.}
 
 Now, if in the future we wonder where \Verb+cleandata.rda+ came from, we
-might search for occurrences of 'cleandata' in the files our
+might search for occurrences of `cleandata' in the files our
 system. However, if such searching among files is a burden, an even
@@
 
-Let me note that this principle was required when all our data
-analyses occurred in batch mode on VAX (and later Unix) machines. The
-fact that I need to articulate this principle at all arises because of
-the rise of interactive data analysis and graphical user interfaces:
-it is all too easy to use the mouse to load a data file into memory
-and then to write a script to analyze this file without ever noting
-the actual name or location of the data file.
+In the good old days, when we executed our LISREL code in batch mode,
+we had no choice but to tell the machine clearly, in a few easy to
+understand and informative lines, what files (with filenames no longer
+than 8 characters) to use. For example,
 
-\section{Version control prevents clobbering and reconciles history.}
+\begin{Verbatim}[fontsize=\footnotesize]
+DA NI=19 NO=199 MA=CM
+LA=basic.lab
+CM FI=basic.cov
+\end{Verbatim}
 
-Group work requires version control.\footnote{See X and Y in this
-  issue for more discussion of what version control is.} Many people
-are familiar with the ``track changes'' feature in modern WYSIWYG word
-processors or the fact that Dropbox allows one to recover previous
-versions of files. These are both kinds of version control. When
-collaborating with yourself or others, it is useful to see what has
-changed, to feel free to experiment and then to dump parts of the
-experiment in favor of previous work while merging the successful
-parts of the experiment into the main body of the paper, and to have
+The fact that I need to articulate this idea at all arises because of
+graphical user interfaces: it is all too easy to use the mouse to load
+a data file into memory and then to write a script to analyze this
+file without ever noting the actual name or location of the data file.
+
+\section{Version control prevents clobbering and reconciles history.}\label{sec:vers-contr-prev}
+
+Group work requires version control.\footnote{
+  \citet{fredrickson2011tpm} and \citet{healy2011tpm} in this issue
+  also explain what version control is and why we might want to use
+  it.} Many people are familiar with the ``track changes'' feature in
+modern WYSIWYG word processors or the fact that Dropbox allows one to
+recover previous versions of files. These are both kinds of version
+control. More generall, when we collaborate, we'd like to do a variety
+of actions with our shared files. Collaboration on data analytic
+projects is more productive and easier when (1) it is easy to see what
+has changed between versions of files; (2) members of the team feel
+free to experiment and then to dump parts of the experiment in favor
+of previous work while merging the successful parts of the experiment
+into the main body of the paper; (3) the team can produce have
 multiple ``releases'' of the same document (one to MPSA, one to APSR,
 one to your parents) without spawning many possibly conflicting copies
-of the same document, risking confusion and clobbering. Clobbering is
-what happens when your future self or your current collaborator saves
-an old version of a file over a new version --- erasing good work by
-accident.
+of the same document; (4) when people can work on the same files at
+the same time without conflicting with one another (and can reconcile
+their changes without too much confusion and clobbering).  Clobbering
+is what happens when your future self or your current collaborator
+saves an old version of a file over a new version --- erasing good
+work by accident.
 
@@
 for open source projects (i.e. papers you are writing which you are
-happy to share with others). Each of those services include paid
-versions too. One may also use Dropbox as a kind of server for version
-control: for example, one may copy files from the Dropbox directory
-into a local working directory so as to avoid clobbering and then
-working on merging changes by hand before copying over existing
-files. (Notice that this is different from directly working on files
-within your Dropbox managed directories.)
+happy to share with others as you write). Each of those services
+include paid versions too. One may also use Dropbox as a kind of
+server for version control: for example, one may copy files from the
+Dropbox directory into a local working directory so as to avoid
+clobbering and then working on merging changes by hand before copying
+back to the Dropbox directory and replacing existing files.
 
@@
 in the extent to which you need to run a server. Subversion requires a
-server and we are lucky that the NCSA provides such a server for
-us. However, the price of such hosting may not be that much using one
-of the many webhosting services out there or perhaps may be available
-for free at your university.\footnote{For example, if you already pay
-  to have a website, you may already have the right to run a
-  subversion server there.}
+server. \footnote{If you already pay to a hosting provider for
+  website, you may already have the right to run a subversion or git
+  server there. Your university or institute may have a version
+  control system running somewhere on campus. And google will direct
+  you to many helpful people who have installed such servers on their
+  own diverse desktop machines. Github requires that you pay to host
+  private repositories.}
 
-Fancy version control systems are not required, however, to get many
-of the benefits of formal version control. I suspect that Google Docs
-allows a kind of version tracking and collaboration as well.  An
-excellent, simple, and robust version control system is to merely
-rename your files with the date and time of saving them: thedoc.tex
-becomes thedoc25-12-2011-23:50.tex.  Be sure to include year in the
-file names --- remember, the life of an idea is measured in years. If
-you are wise enough to have saved your documents as plain text then
-you can easily compare documents using the many utilities available
-for comparing text files. Adobe Acrobat allows one to compare
-differences in pdf files. OpenOffice supports a ``Compare Documents''
-option.
+Of course, one may take advantage of many of the benefits of formal
+version control systems with some disciplined systems for file and
+directory organization.  An excellent, simple, and robust version
+control system is to merely rename your files with the date and time
+of saving them: thedoc.tex becomes thedoc25-12-2011-23:50.tex.  Be
+sure to include year in the file names --- remember, the life of an
+idea is measured in years. If you are wise enough to have saved your
+documents as plain text then you can easily compare documents using
+the many utilities available for comparing text files.\footnote{Adobe
+  Acrobat allows one to compare differences in pdf files. OpenOffice
+  supports a ``Compare Documents'' option. And Google Docs will report
+  on the version history of a document.} When you reach certain
+milestones you can rename the file accordingly: thedocAPSA2009.tex ---
+for the one sent to discussants at APSA --- or thedocAPSR2015.tex ---
+for the version eventually sent to the APSR six years after you
+presented it at APSA. The formal version control systems I mentioned
+above all allow this kind of thing and are much more elegant and
+capable, but you can do it by hand too as long as you don't mind
+taking up a lot of disk space and having many ``thedoc...''  files
+around. If you do version control by hand, spend a little extra time
+to ensure that you do not clobber files when you make mistakes typing
+in the file-names. And, you will find yourself spending extra time
+reconciling changes made by different collaborators by hand; a task
+that modern version control systems take care of quickly and easily.
 
-Also, if you use this method, spend a little extra time to
-ensure that you do not clobber files when you make typos in the
-file-names. And, you will find yourself spending extra time
-reconciling changes made by different collaborators by hand that
-modern version control systems take care of quickly and easily.
-
-When you reach certain milestones you can rename the file accordingly:
-thedocAPSA2009.tex --- for the one sent to discussants at APSA --- or
-thedocAPSR2015.tex --- for the version eventually sent to the APSR six
-years after you presented it at APSA. The formal version control
-systems I mentioned above (and which are described in more depth in X
-and Y) all allow this kind of thing and are much more elegant and
-capable, but you can do it by hand too as long as you don't mind
-taking up a lot of disk space and having  many ``thedoc...''
-files around.
-
-\paragraph{Step \thesection} Writing is rewriting. Thus, all writing involves
-versions. When we collaborate with ourselves and others we want to
-avoid clobbering and we want to enable graceful reconciliation of
-rewriting. One can do these things with formal systems of software
-(like subversion, git, etc...) or with formal systems of file naming,
-file comparing, and communication or, even better, with both.
+\paragraph{Step \thesection} Writing is rewriting. Thus, all writing
+involves versions. When we collaborate with ourselves and others we
+want to avoid clobbering and we want to enable graceful reconciliation
+of rewriting. One can do these things with formal systems of software
+(like subversion or git or bazaar) or with formal systems of file
+naming, file comparing, and communication or, even better, with
+both. In either case, plain text files will make such tasks easier,
+will take up less disk space, and be easier to read for the future
+you.
 
@@
 
-Now, back to that confidence interval in the famous article of
-2018. The statisticians at the UN worry about your use of the
-bootstrap. The authors would like to evaluate their bootstrap
-procedure --- and they did so originally in their memo comparing it to
-a profile likelihood based approach. Although nice code exists for
-bootstrapping linear models, no nice code exists to bootstrap the
-bootstrap. Of course, the code required is not complex, but since they
-are are writing custom code they worry about getting it right. Now, 9
-years after the idea, they've had lots of time to appreciate problems
-arising from bugs and errors in data analysis and code.
+Now, back to that famous article of 2018. After reading the conference
+paper critique of 2021 (that came from the seminar paper of 2020), the
+statisticians at the UN worry about the bootstrap confidence intervals
+presented in the original paper.\footnote{Perhaps they should be
+  worried about the deeper substantive critiques offered by the
+  student, but they are statisticians and so focus on the stats. The
+  policy makers of 2021 were cowed by the methodological virtuosity of
+  the 2018 article, and so, even though they had the same substantive
+  concerns as the student, they kept their mouths shut at the
+  mini-conference to avoid looking dumb in front of their bosses.} So,
+now authors would like to evaluate their bootstrap procedure. Although
+nice code exists for bootstrapping linear models, no nice code exists
+to bootstrap the bootstrap. Of course, the code required is not
+complex, but since they are writing custom code they worry about
+getting it right. As they've struggled to respond to the critiques of
+their paper, they've had lots of time to appreciate problems arising
+from bugs, errors, and typos in data analysis and code.
 
 Now, if they had a moment to think in between teaching that new class,
-reading books for the awards committee, reading application files for
-the admissions committee, staying home with a sick child, and
-undertaking the odd bit of your own current research, they might say to
-themselves, ``Before I write new code, I should write a test of the
-code. I should write a little bit of code that let's me know that my
-double-bootstrap procedure actually does what it is supposed to do.''
+reading books for the awards committee, evaluating application files
+for the admissions committee, feeding popsicles to a sick child, and
+undertaking the odd bit of research, they might say to themselves,
+``Before I write new code, I should write a test of the code. I should
+write a little bit of code that let's me know that my double-bootstrap
+procedure actually does what it is supposed to do.''
 
@@
 collaborating with her future self and/or a small group of
-collaborators here is an example of this idea in a very simple form. I
-want to write a function to multiply a number by 2. If my function
-works, when I give it the number 4, I should see it return the number
-8 and when I give it -4, I should get -8.
+collaborators here is an example of this idea in a very simple
+form: Say, I want to write a function to multiply a number by 2. If my
+function works, when I give it the number 4, I should see it return
+the number 8 and when I give it -4, I should get -8.
 
@@
 ##  The test function:
-test.times.2.fn<-function(){
+test.times.2.fn <- function(){
   ##  This function tests times.2.fn
@@
 ##  The function:
-times.2.fn<-function(thenumber){
+times.2.fn <- function(thenumber){
   ##  This function multiplies a scalar number by 2
@@
 
-##Use the test function
+##  Use the test function
 test.times.2.fn()
@@
 
-Ack! I mistyped ``+'' for ``*''. Good thing I wrote the test!
+Ack! I mistyped ``+'' for ``*''. Good thing I wrote the
+test!\footnote{A more common example of this kind of testing occur
+  everyday when we recode variables into new forms but look at a
+  crosstab of the old vs. new variable before proceeding.}
 
-\paragraph{Step \thesection} You cannot forsee all of the ways that your code
-could be used, but you can at least make sure it does what it is
-supposed to do in your particular case.
+\paragraph{Step \thesection} No one can forsee all of the ways that a
+computer program can faile. One can, however, at least make sure that
+it succeeds in doing the task motivating the writing of the code in
+the first place.
 
@@
 intensive paper we assembled a set of files which enabled reproduction
-of our results using the ``make'' system \citep{bowers2008dataverse};
-and (3) recently I have tried the ``compendium'' approach
+of our results using the \texttt{make} system
+\citep{bowers2008dataverse}; and (3) recently I have tried the
+``compendium'' approach
 \citep{gentleman2005reproducible,gentleman2007statistical} which
-embeds an academic paper with the R package system
-\cite{bowers2011dataverse}. The benefit of this last system is that
+embeds an academic paper within the R package system
+\citep{bowers2011dataverse}. The benefit of this last approach is that
 one is not required to have access to a command line for \Verb+make+:
@@
 \Verb+install.packages()+ and it viewable using the \Verb+vignette()+
-function. The idea that one ought to be able to install and run and
-use an academic paper just as one installs and uses statistical
-software packages is very attractive and I anticipate that it will
-become ever easier to turn papers into R packages as creative and
-energetic folks turn their attention to the question of reproducible
-research.
+function in any operating system than runs R.\footnote{Notice that my
+  reproduction archives and/or instructions for using them are hosted
+  on the \href{http://thedata.org/book/learn-about-project}{Dataverse}
+  which is another system to enhance academic collaboration across
+  time and space.}  The idea that one ought to be able to install and
+run and use an academic paper just as one installs and uses
+statistical software packages is very attractive and I anticipate that
+it will become ever easier to turn papers into R packages as creative
+and energetic folks turn their attention to the question of
+reproducible research.
+
+\paragraph{Step \thesection} We all learn by doing. When we share
+reproduction materials we improve both cumulation of knowledge as
+articulated in \citep{freese2007replication,king1995replication} and
+our methods for doing social science.  As we copy and improve upon
+each others modes of doing work we enhance our collective ability to
+believe each other for future scholars to believe us, too.
 
@@
   time-span for which to plan for self-collaboration on a single idea?
-  Ask your advisers how long it took them from idea to dissertation to
+  Ask your advisers how long it took them to move from idea to dissertation to
   publication.}  The time-frames over which collaboration are required
@@
 essay) make collaboration within research groups more effective.
-Believeability comes in part from reproducibility and research groups
+Believeability comes in part from reproducibility and researchers
 often need to be able to reproduce in part or in whole what different
-people in the group have done.
+people in the group have done or what they, themselves, did in the past.
 
-In the end, following these practices and those recommended by X and Y
-in this issue allows your computerized analyses of your collections
-of stuff to be credible.  Finally, if the UN quibbles with your
-analyses, your future self can shoot the archive required to reproduce
-your work (in still intelligible plain text, analyzed using commented
-code so that folks can translate to whatever system succeeds R, or
-since you used R, you can include a copy of R and all of the R
-packages you used in your final analyses in 2018 in the archive
-itself.) You can say, ``Here is everything you need to reproduce my
-work." To be extra helpful you can add ``Read the README file for
-futher instructions." And then you can get on with your life: maybe
-the next great idea will occur when your 4-year-old asks a wacky
+In the end, following these practices and those recommended by
+\citet{fredrickson2011tpm} and \citet{healy2011tpm} in this issue
+allows your computerized analyses of your collections of stuff to be
+credible.  Finally, if the UN quibbles with your analyses, your future
+self can shoot them the archive required to reproduce your
+work.\footnote{Since you used plain text, the files will still be
+  intelligible, analyzed using commented code so that folks can
+  translate to whatever system succeeds R, or since you used R, you
+  can include a copy of R and all of the R packages you used in your
+  final analyses in 2018 in the archive itself. You can even throw in
+  a copy of Ubuntu 10 and an open source virtual machine running the
+  whole environment.} You can say, ``Here is everything you need to
+reproduce my work." To be extra helpful you can add ``Read the README
+file for futher instructions." And then you can get on with your life:
+maybe the next great idea will occur when your 4-year-old asks a wacky
 question after stripping and painting her overly cooperative

### 2011-03-26 workflow.Rnw
Almost done

@@
 An idea is born in a coffee shop, a seminar, a quiet walk. On this
-gray day in 2011, the idea dispells February's doldrums. The student
+gray day in 2011, the idea dispels February's doldrums. The student
 rushes home, mind racing, the cold ignored.
@@
 remember which series of mouse clicks were used to produce the numbers
-for that crucial table --- and that mousing should not be time
-consuming to redo exactly as it was done in 2011 \ldots or was it
+for that crucial table as it was done in 2011 \ldots or was it
 2015?). Right? And, of course, the way Microsoft
@@
 Mac OS X have always existed and will always continue to exist more or
-less as they currently exist? Right? And the group knows exactly which
+less as they currently exist. Right? And the group knows exactly which
 bit of code produced which table and which figure, right? And they
 wrote their code following Nagler's Maxims \citep{nagler1995coding}
-and King's Replication Standard \citep{king1995replication} right?
+and King's Replication Standard \citep{king1995replication}, right?
 
-Right? Wrong? If the collaborators find themselves say ``Wrong'' in
+If the collaborators find themselves say ``Wrong'' in
 answer to the questions posed here then reproducing, updating, or
@@
 suffer and, more importantly, world peace will have been delayed. This
-essay provides some suggestions for practices which will make such
+essay provides some suggestions for practices that will make such
 reproduction occur much more easily and quickly in the event that
@@
 All results (numbers, comparisons, tables, figures) should arise from
-code, not series of mouse clicks or copying and pasting. If I wanted
-to re-create the figure you created but including a new variable or
+code, not from a series of mouse clicks or copying and pasting. If I wanted
+to re-create the figure from 2011 but include a new variable or
 specification, I should be able to do so with just a few edits to the
-code rather than knowledge of how you used your pointing device in
-your graphical user interface.
+code rather than knowledge of how I (or you) used a pointing device in
+your graphical user interface some years ago.
 
-Using R \citep{R:2011}, for example, I might specify that the file \Verb+fig1.pdf+
-was produced by the following commands in a file called
-\Verb+fig1.R+.\footnote{The command \Verb+please-plot+ and some other
-  R functions used in this essay come from the
+Using R \citep{R:2011}, for example, I might specify that the file
+\Verb+fig1.pdf+ was produced by the following commands in a file
+called \Verb+fig1.R+.\footnote{The command \Verb+please-plot+ and some
+  other R functions used in this essay come from the
   \Verb+MayIPleaseDoStatistics+ package which emphasizes politeness in
-  data analysis. Functions like \Verb+please-plot+ can
-  be blocked and more polite versions such as \Verb+may-I-please-have-a-plot+ can be required using
-  \Verb+options(politeness=99)+ }
+  data analysis. Functions like \Verb+please-plot+ can be blocked and
+  more polite versions such as \Verb+may-I-please-have-a-plot+ can be
+  required using \Verb+options(politeness=99)+ }
 
@@
 text is a format that will be around as long as computer programmers
-write computer programs. \footnote{What is more, since R is open
-  source, I will be able to download an old version of R, download an
-  old fashioned open-source operating system (like Ubuntu 10), and
+write computer programs.\footnote{Since R is open
+  source, I will also be able to download an old version of R, download an
+  old-fashioned open-source operating system (like Ubuntu 10), and
   run the old-fashioned statistical computing environment in the
@@
 
-\paragraph{Step 1} Know the provenance of your results so that your
-future self or current collaborators can quickly and easily reproduce
-and thus change your work.% --- to know what
+\paragraph{Step 1} Know the provenance of results so that 
+future  or current collaborators can quickly and easily reproduce
+and thus change and improve upon the work.% --- to know what
 % commands created which figures and tables.
@@
 
+\begin{quote}
+  Let us change our traditional attitude to the construction of
+  programs: Instead of imagining that our main task is to instruct a
+  computer what to do, let us concentrate rather on explaining to
+  human beings what we want a computer to do.
+  \citep[][p. 97]{knuth1984literate}
+\end{quote}
+
 Imagine you discover something new (or confirm something old). You
@@
 did. Such explanations may not be very clear if we have some pages of
-printed code in one hand and a manuscript in the other.
+printed code in one hand and a manuscript in the other. Keep in mind
+the distinction between the ``source code'' of a document (i.e. what
+computation was required to produce it) and the visible, type-set page
+image. Page images are great for reading, but not great for
+reproducing or collaborating. The source code of any document
+exchanged by the group must be available and executable.
 
@@
 paragraphs, equations, and diagrams can explain the code, and the code
-can numbers, figures, and tables (and diagrams and even equations and
+can produce numbers, figures, and tables (and diagrams and even equations and
 paragraphs). Literate programming is not merely fancy commenting but
-is about enabling the practice of programming itself to enable easy
+is about enabling the practice of programming itself to facilitate easy
 reproduction and communication.
@@
 \LaTeX~ markup with R code chunks to make things like
-Figure~\ref{fig:giniprot}: this combination of \LaTeX and R is called
+Figure~\ref{fig:giniprot}. This combination of \LaTeX and R is called
 Sweave \citep{Leis:2005}.\footnote{Support for Sweave is included with
@@
 markup with R code chunks to make things like
-Figure~\ref{fig:giniprot}: this combination of \LaTeX and R is
+Figure~\ref{fig:giniprot}. This combination of \LaTeX and R is
 called Sweave \citep{Leis:2005}.\footnote{Support for Sweave is
@@
   programming for R.}  \href{http://www.lyx.org/}{LyX} offers a
-WYSIWYG environment for \LaTeX~ which supports Sweave. And the
+WYSIWYG environment for \LaTeX~ that supports Sweave. And the
 \Verb+odfWeave+ package in R allows the use of OpenOffice documents in
@@
 Literate programming need not go against the principle of modular data
-analysis \citep{nagler1995coding}: in my own work I routinely have
+analysis \citep{nagler1995coding}. In my own work I routinely have
 several different Sweave files that fulfill different functions, some
@@
 me to record my journeys down blind alleys. Of course, when we have
-flying cars running on autopilot, perhaps something other than than
+flying cars running on autopilot, perhaps something other than
 Sweave will make our lives even easier. Then we'll change.
@@
 
-\begin{quote}
-  Let us change our traditional attitude to the construction of
-  programs: Instead of imagining that our main task is to instruct a
-  computer what to do, let us concentrate rather on explaining to
-  human beings what we want a computer to do.
-  \citep[][p. 97]{knuth1984literate}
-\end{quote}
-
-We analyze data in order to explain something about world to other
+We analyze data in order to explain something about the world to other
 scholars and policy makers. If we focus on explaining how we got our
@@
 confidence in readers and can more effectively spur discussion and
-learning and cumulation of research as people no longer need spend
-weeks attempting to reproduce research which inspired them and on
-which they desire to build.
-
-Keep in mind the distinction between the ``source code'' of a document
-(i.e. what computation was required to produce it) and the visible,
-type-set page image. Page images are great for reading, but not great
-for reproducing or collaborating. The source code of any document
-exchanged by the group must be available and executable.
+learning and cumulation of research.
 
@@
 
-Where should one store data files? A obvious solution is always to
+Where should one store data files? Anobvious solution is always to
 make sure that the data file used by a command file is in the same
@@
 Now, if in the future we wonder where \Verb+cleandata.rda+ came from, we
-might search for occurrences of `cleandata' in the files our
+might search for occurrences of `cleandata' in the files on our
 system. However, if such searching among files is a burden, an even
@@
 understand and informative lines, what files (with filenames no longer
-than 8 characters) to use. For example,
+than 8 characters) to use:
 
@@
 recover previous versions of files. These are both kinds of version
-control. More generall, when we collaborate, we'd like to do a variety
+control. More generally, when we collaborate, we'd like to do a variety
 of actions with our shared files. Collaboration on data analytic
-projects is more productive and easier when (1) it is easy to see what
+projects is more productive and better when (1) it is easy to see what
 has changed between versions of files; (2) members of the team feel
@@
 into the main body of the paper; (3) the team can produce have
-multiple ``releases'' of the same document (one to MPSA, one to APSR,
+``releases'' of the same document (one to MPSA, one to APSR,
 one to your parents) without spawning many possibly conflicting copies
-of the same document; (4) when people can work on the same files at
+of the same document; (4) people can work on the same files at
 the same time without conflicting with one another (and can reconcile
@@
 is what happens when your future self or your current collaborator
-saves an old version of a file over a new version --- erasing good
+saves an old version of a file over a new version, erasing good
 work by accident.
@@
 Dropbox directory into a local working directory so as to avoid
-clobbering and then working on merging changes by hand before copying
+clobbering and then work on merging changes by hand before copying
 back to the Dropbox directory and replacing existing files.
@@
 in the extent to which you need to run a server. Subversion requires a
-server. \footnote{If you already pay to a hosting provider for
-  website, you may already have the right to run a subversion or git
-  server there. Your university or institute may have a version
-  control system running somewhere on campus. And google will direct
-  you to many helpful people who have installed such servers on their
-  own diverse desktop machines. Github requires that you pay to host
-  private repositories.}
+server.\footnote{If you already pay to host a website, you may already
+  have the right to run a subversion or git server there. Your
+  university or institute may have a version control system running
+  somewhere on campus. And Google will direct you to many helpful
+  people who have installed such servers on their own diverse desktop
+  machines. Github requires that you pay to host private
+  repositories.}
 
@@
 directory organization.  An excellent, simple, and robust version
-control system is to merely rename your files with the date and time
-of saving them: thedoc.tex becomes thedoc25-12-2011-23:50.tex.  Be
-sure to include year in the file names --- remember, the life of an
-idea is measured in years. If you are wise enough to have saved your
-documents as plain text then you can easily compare documents using
-the many utilities available for comparing text files.\footnote{Adobe
-  Acrobat allows one to compare differences in pdf files. OpenOffice
-  supports a ``Compare Documents'' option. And Google Docs will report
-  on the version history of a document.} When you reach certain
-milestones you can rename the file accordingly: thedocAPSA2009.tex ---
-for the one sent to discussants at APSA --- or thedocAPSR2015.tex ---
-for the version eventually sent to the APSR six years after you
-presented it at APSA. The formal version control systems I mentioned
-above all allow this kind of thing and are much more elegant and
-capable, but you can do it by hand too as long as you don't mind
-taking up a lot of disk space and having many ``thedoc...''  files
-around. If you do version control by hand, spend a little extra time
-to ensure that you do not clobber files when you make mistakes typing
-in the file-names. And, you will find yourself spending extra time
-reconciling changes made by different collaborators by hand; a task
-that modern version control systems take care of quickly and easily.
+control system is to rename your files with the date and time of
+saving them: thedoc.tex becomes thedoc25-12-2011-23:50.tex.  Be sure
+to include year in the file names --- remember, the life of an idea is
+measured in years. If you are wise enough to have saved your documents
+as plain text then you can easily compare documents using the many
+utilities available for comparing text files.\footnote{Adobe Acrobat
+  allows one to compare differences in pdf files. OpenOffice supports
+  a ``Compare Documents'' option. And Google Docs will report on the
+  version history of a document.} When you reach certain milestones
+you can rename the file accordingly: thedocAPSA2009.tex --- for the
+one sent to discussants at APSA --- or thedocAPSR2015.tex --- for the
+version eventually sent to the APSR six years after you presented it
+at APSA. The formal version control systems I mentioned above all
+allow this kind of thing and are much more elegant and capable, but
+you can do it by hand too as long as you don't mind taking up a lot of
+disk space and having many ``thedoc...''  files around. If you do
+version control by hand, spend a little extra time to ensure that you
+do not clobber files when you make mistakes typing in the
+file-names. And, if you find yourself spending extra time reconciling
+changes made by different collaborators by hand, remember this is a
+task that modern version control systems take care of quickly and
+easily.
 
@@
   mini-conference to avoid looking dumb in front of their bosses.} So,
-now authors would like to evaluate their bootstrap procedure. Although
+now the authors would like to evaluate their bootstrap procedure. Although
 nice code exists for bootstrapping linear models, no nice code exists
@@
 Now, if they had a moment to think in between teaching that new class,
-reading books for the awards committee, evaluating application files
+reading books for an awards committee, evaluating application files
 for the admissions committee, feeding popsicles to a sick child, and
@@
 ``Before I write new code, I should write a test of the code. I should
-write a little bit of code that let's me know that my double-bootstrap
+write a little bit of code that lets me know that my double-bootstrap
 procedure actually does what it is supposed to do.''
 
-Of course, this idea, like most others, is not new. When large groups
-of programmers write code for multi-million dollar programs the
-question about avoiding error looms large. The idea of
+Of course, this idea, like most others, is not new. The desire to
+avoid error looms large when large groups of programmers write code
+for multi-million dollar programs. The idea of
 \href{http://en.wikipedia.org/wiki/Test-driven_development}{test
@@
 collaborating with her future self and/or a small group of
-collaborators here is an example of this idea in a very simple
-form: Say, I want to write a function to multiply a number by 2. If my
+collaborators, here is an example of this idea in a very simple form:
+Say I want to write a function to multiply a number by 2. If my
 function works, when I give it the number 4, I should see it return
@@
 \paragraph{Step \thesection} No one can forsee all of the ways that a
-computer program can faile. One can, however, at least make sure that
+computer program can fail. One can, however, at least make sure that
 it succeeds in doing the task motivating the writing of the code in
@@
 Lots of people are thinking about ``reproducible research'' and
-``literate programming'' these days. Google those terms. Of course the
-devil is the details: Here I list a few of my own attempts at enabling
+``literate programming'' these days. Google those terms. Of course, the
+devil is in the details: Here I list a few of my own attempts at enabling
 reproducible research. You'll find many other inspiring examples on
@@
 archive \citep{bowers2005dataverse}; (2) for another more computing
-intensive paper we assembled a set of files which enabled reproduction
+intensive paper we assembled a set of files that enabled reproduction
 of our results using the \texttt{make} system
@@
 the compendium is downloadable from within R using
-\Verb+install.packages()+ and it viewable using the \Verb+vignette()+
+\Verb+install.packages()+ and is viewable using the \Verb+vignette()+
 function in any operating system than runs R.\footnote{Notice that my
   reproduction archives and/or instructions for using them are hosted
-  on the \href{http://thedata.org/book/learn-about-project}{Dataverse}
+  on the \href{http://thedata.org/book/learn-about-project}{Dataverse},
   which is another system to enhance academic collaboration across
@@
 run and use an academic paper just as one installs and uses
-statistical software packages is very attractive and I anticipate that
+statistical software packages is very attractive, and I anticipate that
 it will become ever easier to turn papers into R packages as creative
@@
 
-\section{Remember that research ought to be credibile communication.}
+\section{Remember that research ought to be credible communication.}
 \begin{quote}
@@
 characteristics: (1) the findings of the work should not be a matter
-of opinion; and, (2) other people should be able to reproduce the findings. That is,
+of opinion; and (2) other people should be able to reproduce the findings. That is,
 the work represents a shared
@@
 easy to convince others that ``this collection of stuff is different
-from this collection of stuff'' if those people were looking over our
+from that collection of stuff'' if those people were looking over our
 shoulders the whole time that we made decisions about collecting the
@@
 chosen tools of analyses are computer programs. How can we convince
-people that what we did with some data with some program is credible:
+people that what we did with some data with some program is credible,
 not a matter of whim or opinion, and reproducible by others who didn't
 shadow us as we wrote our papers? This essay has suggested a few
-concrete ways to enhance the believeability of such scholarly work. In
+concrete ways to enhance the believability of such scholarly work. In
 addition, these actions (as summarized in the section headings of this
 essay) make collaboration within research groups more effective.
-Believeability comes in part from reproducibility and researchers
+Believability comes in part from reproducibility and researchers
 often need to be able to reproduce in part or in whole what different
@@
 reproduce my work." To be extra helpful you can add ``Read the README
-file for futher instructions." And then you can get on with your life:
+file for further instructions." And then you can get on with your life:
 maybe the next great idea will occur when your 4-year-old asks a wacky

### 2011-03-26 workflow.tex
Version given to Ashly. Perhaps the final version.

--- 2011-03-26 workflow.Rnw
@@
 \renewenvironment{Schunk}{\vspace{\topsep}}{\vspace{\topsep}}
-\SweaveOpts{keep.source=TRUE}  %Show comments
+  %Show comments
 \fvset{fontsize=\footnotesize} %slightly smaller verbatim font
@@
 \maketitle
-
-<<initialize,echo=FALSE,results=hide>>=
-  ##First, just setup the \R environment:
-  options(SweaveHooks=list(
-            fig=function(){
-              par(mar=c(3.5, 3, 1.1, 0),
-                  pty="s",
-                  mgp=c(1.5,0.5,0),
-                  oma=c(0,0,0,0))},#,
-            echo=function(){options(continue=" ") ##Don't show "+" prompts,
-                            options(prompt=" ")}
-            ),
-          digits=3,
-          scipen=8,
-          width=50,
-          error=function(){options(prompt="> ",continue="+ ");NULL}
-          )
-
-@
 
@@
 %%% Putting it up top, so we can use it in the comments section.
-<<read.data,echo=FALSE,results=hide>>=
-##  This read.csv line is slightly more complex than needed because of
-##  the particular encoding used for characters in the source file.
-dem.nations.df<-read.csv(url("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv",encoding="ISO_8859-1"),as.is=TRUE)
-row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful
-good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
-good.df$gini04<-good.df$Gini2004/100
-good.df$protac00<-good.df$protac2000
-@
 
@@
 
-\paragraph{Step 1} Know the provenance of results so that 
+\paragraph{Step \thesection} If we know the provenance of results, 
 future  or current collaborators can quickly and easily reproduce
-and thus change and improve upon the work.% --- to know what
-% commands created which figures and tables.
+and thus change and improve upon the work.
 
@@
 
-<<fig1code,echo=FALSE,fig=FALSE>>=
-##  Make a scatterplot of Protest by Inequality
-with(good.df,plot(gini04,protac00,
-                  xlab='Gini Coefficient 2004 (UNDP)',
-                  ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
-                  cex=.8))
-
-##  Label a few interesting points
-with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
-     text(gini04,protac00,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
-@
-
 \begin{figure}[H]
   \begin{center}
-<<fig1plot,echo=FALSE,fig=TRUE,width=4,height=4>>=
-par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
-<<fig1code>>
-@
+\includegraphics{workflow-fig1plot}
     \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}
@@
 
-<<maketable, echo=FALSE, results=hide>>=
-  lm1<-lm(protac00~gini04+meanpr,data=good.df)  #  Run the regression
-
-##  The next lines extract relevant and useful output from the
-##  regression object, following Neal Beck's TPM article 2010.
-themat<-cbind(summary(lm1)$coef[,c("Estimate","Std. Error")],confint(lm1))
-n<-length(lm1$fitted.values)
-r2<-signif(summary(lm1)$r.squared,2)
-est.sigma<-signif(summary(lm1)$sigma,2)
-colnames(themat)[3:4]<-c("~{}","~{}")
-row.names(themat)<-c("Intercept","Income Inequality (lower=more equal)","Mean Political Rights (lower=more rights)")
-
-library(xtable)
-thextab<-xtable(themat,align=c("l",rep(".",4)),digits=1,
-                caption="People living in countries with unequal
+% latex table generated in R 2.12.2 by xtable 1.5-6 package
+% Sat Mar 26 09:57:24 2011
+\begin{table}[H]
+\begin{center}
+\begin{tabular}{l....}
+   \toprule 
+ &
+             \multicolumn{1}{c}{Coef} &
+             \multicolumn{1}{c}{Std. Err.} &
+             \multicolumn{2}{c}{95\% CI} \\ \cmidrule(r){2-5} Intercept & 1.5 & 0.2 & 1.2 & 1.8 \\ 
+  Income Inequality (lower=more equal) & -1.0 & 0.4 & -1.9 & -0.2 \\ 
+  Mean Political Rights (lower=more rights) & -0.2 & 0.0 & -0.2 & -0.1 \\ 
+   & \multicolumn{4}{c}{n: 53,   resid.sd: 0.28, R$^2$: 0.57} \\ \bottomrule\end{tabular}
+\caption{People living in countries with unequal
   income distributions report less protest activity to World Values
@@
   more equal income distributions, adjusting for average political
-  rights as measured by Freedom House 1980--2000. Data from \\citep{norris2009data}.",label="tab:protest")
-@
-
-<<printtable,echo=FALSE,results=tex>>=
-  print(thextab,include.colnames=FALSE,
-        hline.after=NULL,
-        table.placement="H",
-        add.to.row=list(
-          pos=list(0,3),
-          command=c(
-            "\\toprule \n &
-             \\multicolumn{1}{c}{Coef} &
-             \\multicolumn{1}{c}{Std. Err.} &
-             \\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ",
-            paste("& \\multicolumn{4}{c}{n: ",n,",",
-                  "   resid.sd: ",signif(est.sigma),
-                  ", R$^2$: ",r2,"} \\\\ \\bottomrule",sep="")
-            )
-          ))
-@
+  rights as measured by Freedom House 1980--2000. Data from \citep{norris2009data}.}
+\label{tab:protest}
+\end{center}
+\end{table}
 
@@
 
-\paragraph{Step \thesection} We should know data where the data came from and
-what operations were performed on which set of data.
+\paragraph{Step \thesection} We should know data where the data came
+from and what operations were performed on which set of data.
 
@@
 
-<<unit.test>>=
+\begin{Schunk}
+\begin{Sinput}
  ##  The test function:
@@
  }
-
  ##  The function:
@@
  }
-
  ##  Use the test function
  test.times.2.fn()
-@
+\end{Sinput}
+\begin{Soutput}
+[1] "It does not work!"
+\end{Soutput}
+\end{Schunk}
 
@@
   reproduction archives and/or instructions for using them are hosted
-  on the \href{http://thedata.org/book/learn-about-project}{Dataverse},
-  which is another system to enhance academic collaboration across
+  on the
+  \href{http://thedata.org/book/learn-about-project}{Dataverse}, which
+  is another system designed to enhance academic collaboration across
   time and space.}  The idea that one ought to be able to install and
 run and use an academic paper just as one installs and uses
-statistical software packages is very attractive, and I anticipate that
-it will become ever easier to turn papers into R packages as creative
-and energetic folks turn their attention to the question of
+statistical software packages is very attractive, and I anticipate
+that it will become ever easier to turn papers into R packages as
+creative and energetic folks turn their attention to the question of
 reproducible research.
@@
 \paragraph{Step \thesection} We all learn by doing. When we share
-reproduction materials we improve both cumulation of knowledge as
-articulated in \citep{freese2007replication,king1995replication} and
-our methods for doing social science.  As we copy and improve upon
-each others modes of doing work we enhance our collective ability to
-believe each other for future scholars to believe us, too.
+reproduction materials we improve both cumulation of knowledge and our
+methods for doing social science
+\citep{freese2007replication,king1995replication}.  As we copy and
+improve upon each other's modes of doing work we enhance our collective
+ability to believe each other and for future scholars to believe us, too.
 
@@
 \end{document}
-<<>>=
+\begin{Schunk}
+\begin{Sinput}
  options(prompt="> ",continue="+ ")
-@
+\end{Sinput}
+\end{Schunk}

### 2011-03-26 workflow.Rnw
Fixed typos.

@@
   still owns cassette tapes and compact discs.} Now
-re-analyses should be easy. Right? The student, now professor, should
+re-analysis should be easy. Right? The student, now professor, should
 remember the reason for those bits of code (or at least should
@@
 
-If the collaborators find themselves say ``Wrong'' in
+If the collaborators find themselves saying ``Wrong'' in
 answer to the questions posed here then reproducing, updating, or
@@
 term did not really change the substantive story arising from the
-model.\footnote{R considers text marked with \Verb+#+ as a comment.}
+model.\footnote{R considers text marked with \texttt{\#} as a comment.}
 Messages left for your future self (or near-future others) help
@@
 in Microsoft Word called ``code'' which hides your code when you print
-your document, but which allows you do at least run each code chunk
+your document, but which allows you to at least run each code chunk
 piece by piece [or perhaps there are ways to extract all text of style
@@
 
-Where should one store data files? Anobvious solution is always to
+Where should one store data files? An obvious solution is always to
 make sure that the data file used by a command file is in the same
@@
 
-\paragraph{Step \thesection} We should know data where the data came from and
-what operations were performed on which set of data.
+\paragraph{Step 3} We should know where the data came
+from and what operations were performed on which set of data.
 
@@
 
-But what if the UN never hears of your work, or, by some cruel fate, that your
+But what if the UN never hears of your work, or, by some cruel fate, your
 article does not spawn debate? Why then would you spend time to

### 2016-05-20 workflow.Rmd
Starting to transition to Rmd

--- 2011-03-26 workflow.Rnw
@@
-\documentclass[12pt]{article}
-\usepackage{fancyvrb,natbib,url,array,dcolumn,float,booktabs,listings}
-\usepackage[colorlinks=TRUE]{hyperref}
-\usepackage{microtype}
-
-\title{Six steps to a better relationship with your future self.}
-
-\author{Jake Bowers\thanks{I owe many thanks to Mark Fredrickson,
+---
+title: How to improve your relationship with your future self.
+author: 
+- name: Jake Bowers\thanks{I owe many thanks to Mark Fredrickson,
     Brian Gaines, Kieran Healy, Kevin Quinn and Cara
@@
     source code for this document may be freely downloaded and
-    modified from \url{https://github.com/jwbowers/workflow}.}}
+    modified from \url{https://github.com/jwbowers/workflow}.}
+  affiliation: <jwbowers@illinois.edu>, University of Illinois at Urbana-Champaign and White House Social and Behavioral Sciences Team
+- name: Maarten Voors
+  affiliation: Wageningen University
+date: 20 May 2016
+...
 
-\date{\today}
-
-\newcolumntype{.}{D{.}{.}{1.2}}
-
-\usepackage[noae,nogin]{Sweave} %Declaring this explicitly so that I can modify the Sweave environment here rather in Sweave.sty which will be c
-\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=.5em,fontsize=\footnotesize}
-\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=.5em,fontsize=\footnotesize}
-\DefineVerbatimEnvironment{Scode}{Verbatim}  {xleftmargin=.5em,fontsize=\footnotesize}
-\fvset{listparameters={\setlength{\topsep}{0pt}}}
-\renewenvironment{Schunk}{\vspace{\topsep}}{\vspace{\topsep}}
-\SweaveOpts{keep.source=TRUE}  %Show comments
-\fvset{fontsize=\footnotesize} %slightly smaller verbatim font
-
-%% These next lines tell latex that it is ok to have a single graphic
-%% taking up most of a page, and they also decrease the space arou
-%% figures and tables.
-\renewcommand\floatpagefraction{.9}
-\renewcommand\topfraction{.9}
-\renewcommand\bottomfraction{.9}
-\renewcommand\textfraction{.1}
-\setcounter{totalnumber}{50}
-\setcounter{topnumber}{50}
-\setcounter{bottomnumber}{50}
-\setlength{\intextsep}{2ex}
-\setlength{\floatsep}{2ex}
-\setlength{\textfloatsep}{2ex}
-
-\begin{document}
-\frenchspacing
-\VerbatimFootnotes
-\maketitle
-
-<<initialize,echo=FALSE,results=hide>>=
+```{r initialize,echo=FALSE,results=hide}
   ##First, just setup the \R environment:
@@
 
-@
+```
 
@@
 %%% Putting it up top, so we can use it in the comments section.
-<<read.data,echo=FALSE,results=hide>>=
+```{r read.data,echo=FALSE,results=hide}
 ##  This read.csv line is slightly more complex than needed because of
@@
 good.df$protac00<-good.df$protac2000
-@
+```
 
@@
 
-% <<reg1,echo=TRUE,results=verb>>=
+% ```{r reg1,echo=TRUE,results=verb}
 % ##  Repeat the regression run for Table~\ref{tab:protest}
@@
 % ##  A: Yes (around 3/4 an act of protest difference)
-% @
+% ```
 
@@
 \centering
- <<fig1plot,fig=TRUE>>=
+ ```{r fig1plot,fig=TRUE}
  ##  Make a scatterplot of Protest by Inequality
@@
       text(gini04,protac00,labels=Nation))
- @
+ ```
 \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}
@@
 
-<<fig1code,echo=FALSE,fig=FALSE>>=
+```{r fig1code,echo=FALSE,fig=FALSE}
 ##  Make a scatterplot of Protest by Inequality
@@
      text(gini04,protac00,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
-@
+```
 
@@
   \begin{center}
-<<fig1plot,echo=FALSE,fig=TRUE,width=4,height=4>>=
+```{r fig1plot,echo=FALSE,fig=TRUE,width=4,height=4}
 par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
-<<fig1code>>
-@
+```{r fig1code>>
+```
     \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}
@@
 
-<<maketable, echo=FALSE, results=hide>>=
+```{r maketable, echo=FALSE, results=hide}
   lm1<-lm(protac00~gini04+meanpr,data=good.df)  #  Run the regression
@@
   rights as measured by Freedom House 1980--2000. Data from \\citep{norris2009data}.",label="tab:protest")
-@
+```
 
-<<printtable,echo=FALSE,results=tex>>=
+```{r printtable,echo=FALSE,results=tex}
   print(thextab,include.colnames=FALSE,
@@
           ))
-@
+```
 
@@
 
-<<unit.test>>=
+```{r unit.test}
 ##  The test function:
@@
 test.times.2.fn()
-@
+```
 
@@
 \end{document}
-<<>>=
+```{r }
 options(prompt="> ",continue="+ ")
-@
+```

### 2016-05-31 workflow.Rmd
Some progress: new intro

@@
   affiliation: Wageningen University
-date: 20 May 2016
+bibliography: ../BIB/references.bib
+csl: american-political-science-review.csl
+date: 19 May 2016
 ...
 
-```{r initialize,echo=FALSE,results=hide}
-  ##First, just setup the \R environment:
-  options(SweaveHooks=list(
-            fig=function(){
+```{r include=FALSE, cache=FALSE}
+# Some customization.  You can alter or delete as desired (if you know what you are doing).
+# knitr settings to control how R chunks work.
+
+## To make the html file do
+## render("exploration4.Rmd",output_format=html_document(fig_retina=FALSE))
+## To make the pdf file do
+## render("exploration4.Rmd",output_format=pdf_document())
+
+require(knitr)
+opts_chunk$set(
+  tidy=FALSE,     # display code as typed
+  size="small",    # slightly smaller font for code
+  echo=TRUE,
+  results='markup',
+  strip.white=TRUE,
+  fig.path='figs/fig',
+  cache=FALSE,
+  highlight=TRUE,
+  width.cutoff=132,
+  size='footnotesize',
+  out.width='.9\\textwidth',
+  message=FALSE,
+  comment=NA)
+```
+
+```{r initialize,echo=FALSE}
+##First, just setup the R environment for today:
+if(!file.exists('figs')) dir.create('figs')
+
+options(SweaveHooks=list(fig=function(){
 			   par(mar=c(3.5, 3, 1.1, 0),
@@
 			       mgp=c(1.5,0.5,0),
-                  oma=c(0,0,0,0))},#,
+			       oma=c(0,0,0,0))},
 			 echo=function(){options(continue=" ") ##Don't show "+" prompts,
-                            options(prompt=" ")}
-            ),
-          digits=3,
+			 options(prompt=" ")
+			 }),
+	digits=4,
 	scipen=8,
-          width=50,
-          error=function(){options(prompt="> ",continue="+ ");NULL}
+	width=132
 	)
-
+options(error=function(){options(prompt="> ",continue="+ ");NULL})
 ```
 
-\begin{verse}
-Do I contradict myself? \\
-Very well then I contradict myself, \\
-(I am large, I contain multitudes.) \\
-\citep{whitman1855song}
-\end{verse}
+| Do I contradict myself?
+| Very well then I contradict myself,
+| (I am large, I contain multitudes.)
+| [@whitman1855song]
 
-An idea is born in a coffee shop, a seminar, a quiet walk. On this
-gray day in 2011, the idea dispels February's doldrums. The student
-rushes home, mind racing, the cold ignored.
-% Now he knows why he went to graduate school.
+> If you tell the truth, you don't have to remember anything. [@twain1976book]^[<https://www.marktwainhouse.org/man/famous_twain_quotes.php>]
 
-This idea inspires a seminar paper in the spring. A conference paper
-arises from the seminar paper in collaboration with another student in
-2012. A dissertation chapter descends from the conference paper in
-2013. Other dissertation chapters take up 2014. A submission to a
-journal with the original co-author and a new collaborator happens in
-2015. Revision and resubmission wait until 2017 while harried editors,
-reviewers and authors strive to balance research, teaching, service,
-and life. By now, the three lucky collaborators work as professors in
-three different universities. In 2018 a child is born and a paper is
-published. The United Nations takes an interest in the paper in 2019
-and hosts a conference to discuss implications of the research. In
-2020 a first year graduate student in a coffee shop has an idea that
-challenges the results in the now famous paper. She presents her paper
-at a conference in 2021. What would happen if the authors had
-controlled for X? Or included information now available but missing in
-2012?  Or chosen a different likelihood function?  Will the United
-Nations (now eager to act based the paper) make a wrong move?
+> Never memorize what you can look up in books -Albert Einstein^[<https://en.wikiquote.org/wiki/Albert_Einstein>]
 
-The first author convenes a three way video conference with the other
-collaborators during his homeward commute after putting his flying car
-in auto-drive mode.\footnote{One assumes that video chatting during
-  manual driving of flying cars will have been outlawed in his state
-  by 2020.} The group must go back to the analyses. Which ones?  The
-ones from 2011?  Or 2012? Or 2018? Where are the files? The next day,
-one member of the group who has kept some hard-drives around out of
-nostalgia finds some of the files.\footnote{This is the same guy who
-  still owns cassette tapes and compact discs.} Now
-re-analysis should be easy. Right? The student, now professor, should
-remember the reason for those bits of code (or at least should
-remember which series of mouse clicks were used to produce the numbers
-for that crucial table as it was done in 2011 \ldots or was it
-2015?). Right? And, of course, the way Microsoft
-Word/Stata/SPSS/R/LISREL understands files and the way that 
-machines in 2021 read and write them is the same --- since Windows and
-Mac OS X have always existed and will always continue to exist more or
-less as they currently exist. Right? And the group knows exactly which
-bit of code produced which table and which figure, right? And they
-wrote their code following Nagler's Maxims \citep{nagler1995coding}
-and King's Replication Standard \citep{king1995replication}, right?
+Memory is tricky.
+When we really want to remember something we should practice it until it becomes internalized.
+In fact, learning requires effort.
+When we do not practice, we tend to be overconfident in our recall abilities [@koriat2005illusions] and false memories can be implanted.^[See the following site for a nice overview of what we know about memory including the fact that learning requires practice: <http://www.spring.org.uk/2012/10/how-memory-works-10-things-most-people-get-wrong.php>.
+On false memory see Wikipedia and linked studies <https://en.wikipedia.org/wiki/False_memory>.]
 
-If the collaborators find themselves saying ``Wrong'' in
-answer to the questions posed here then reproducing, updating, or
-changing the original analyses will take a lot of time. If
-reproduction is hard to do, then the reputations of the scholars will
-suffer and, more importantly, world peace will have been delayed. This
-essay provides some suggestions for practices that will make such
-reproduction occur much more easily and quickly in the event that
-famous papers require special scrutiny. Specifically, this piece aims
-to amplify some of what we already ought to know
-\citep{king1995replication,nagler1995coding}, and to add to some of
-those ideas given current practices, platforms, and possibilities.
+How long does it take from planning a study to publication and to the first reproduction of the study after publication?
+Is three years too short?
+Is ten years too long?
+We suspect that few of our colleagues in the social and behavioral sciences conceive of, field, write, and publish a study in faster than about three years.
+We also suspect that, if some other scholar decides to re-examine the analyses of a published study, that it will occur after publication, and that this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, years past the initial publication of the article.^[We note that sometimes this reproduction occurs within the context of classes and learning about statistics and data analysis.
+Othertimes, this process is organized more broadly (cite IIIE program, cite Open Science Foundation on psychology).]
 
-%%% This data frame will be used throughout the analyses.
-%%% Putting it up top, so we can use it in the comments section.
-```{r read.data,echo=FALSE,results=hide}
+If we cannot count on our memories about why we made such and such a design or analysis decision, then what should we do?
+How can we minimize regret with our past decisions?
+How can we improve our relationship with our future self?
+This essay is a heavily revised version of @bowers2011tpm and provides some suggestions for practices that will make such reproduction occur much more easily and quickly in the event that famous papers require special scrutiny.
+Specifically, this piece aims to amplify some of what we already ought to know,  and to add to some of those ideas
+given current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists.
+Now, the efforts to encourage transparency and ease of learning from the data and analyses of others has become institutionalized with teh DA-RT inititiave <http://www.dartstatement.org/> @lupia2014openness.
+These ideas are spreading beyond political science as well @freese1007replication and @asendorpf2013recommendations.] This essay is an attempt at providing practical advice about how to *do work* such that one complies with such recommendations as a matter of course and, in so doing, can focus personal regret on other bad past decisions but not on decisions about data analysis and the production of scholarly papers.
+
+%```{r read.data,echo=FALSE,results=hide}
+## This data frame will be used throughout the analyses.
+## Putting it up top, so we can use it in the comments section.
 ##  This read.csv line is slightly more complex than needed because of

### 2016-06-05 workflow.Rmd
Workflow piece almost ready for Maarten to edit

@@
 author:
-- name: Jake Bowers\thanks{I owe many thanks to Mark Fredrickson,
-    Brian Gaines, Kieran Healy, Kevin Quinn and Cara
-    Wong for direct help on this document and to Mika LaVaque-Manty
-    and Ben Hansen for many useful discussions on this topic. The
-    source code for this document may be freely downloaded and
-    modified from \url{https://github.com/jwbowers/workflow}.}
-  affiliation: <jwbowers@illinois.edu>, University of Illinois at Urbana-Champaign and White House Social and Behavioral Sciences Team
+- name: Jake Bowers
+  affiliation: <jwbowers@illinois.edu>, University of Illinois at Urbana-Champaign, Departments of Political Science and Statistics; Fellow, White House Social and Behavioral Sciences Team.
+Many thanks to Mark Fredrickson, Brian Gaines, Kieran Healy, Kevin Quinn and Cara Wong for direct help on this document and to Mika LaVaque-Manty and Ben Hansen for many useful discussions on this topic.
+The source code for this document may be freely downloaded and modified from <https://github.com/jwbowers/workflow>.
 - name: Maarten Voors
-  affiliation: Wageningen University
-bibliography: ../BIB/references.bib
+  affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics
+bibliography: references.bib
 csl: american-political-science-review.csl
-date: 19 May 2016
+date: '`r format(Sys.Date(), "%B %d, %Y")`'
+output:
+    pdf_document:
+      number_sections: true
+      fig_width: 7
+      fig_height: 7
+      fig_caption: true
+      template: bowersarticletemplate.latex
+      pandoc_args: [
+      "--filter", "/usr/local/bin/pandoc-crossref"
+    ]
 ...
@@
 ## render("exploration4.Rmd",output_format=pdf_document())
+## to debug, try render("thefile.Rmd", clean=FALSE) so that you can see the intermediate files
 
@@
 	)
-options(error=function(){options(prompt="> ",continue="+ ");NULL})
+
+#options(error=function(){options(prompt="> ",continue="+ ");NULL})
 ```
 
-| Do I contradict myself?
+> | Do I contradict myself?
 | Very well then I contradict myself,
@@
 
-> If you tell the truth, you don't have to remember anything. [@twain1976book]^[<https://www.marktwainhouse.org/man/famous_twain_quotes.php>]
-
-> Never memorize what you can look up in books -Albert Einstein^[<https://en.wikiquote.org/wiki/Albert_Einstein>]
+> If you tell the truth, you don't have to remember anything. [@twain1975mark]
 
@@
 When we do not practice, we tend to be overconfident in our recall abilities [@koriat2005illusions] and false memories can be implanted.^[See the following site for a nice overview of what we know about memory including the fact that learning requires practice: <http://www.spring.org.uk/2012/10/how-memory-works-10-things-most-people-get-wrong.php>.
-On false memory see Wikipedia and linked studies <https://en.wikipedia.org/wiki/False_memory>.]
+On false memory see Wikipedia and linked studies <https://en.wikipedia.org/wiki/False_memory>.] If we cannot count on memory, then how can we do science?
 
@@
 given current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists.
-Now, the efforts to encourage transparency and ease of learning from the data and analyses of others has become institutionalized with teh DA-RT inititiave <http://www.dartstatement.org/> @lupia2014openness.
-These ideas are spreading beyond political science as well @freese1007replication and @asendorpf2013recommendations.] This essay is an attempt at providing practical advice about how to *do work* such that one complies with such recommendations as a matter of course and, in so doing, can focus personal regret on other bad past decisions but not on decisions about data analysis and the production of scholarly papers.
+Now, the efforts to encourage transparency and ease of learning from the data and analyses of others has become institutionalized with the DA-RT inititiave <http://www.dartstatement.org/> @lupia2014openness.
+These ideas are spreading beyond political science as well @freese2007replication and @asendorpf2013recommendations.] This essay is an attempt at providing practical advice about how to *do work* such that one complies with such recommendations as a matter of course and, in so doing, can focus personal regret on other bad past decisions but not on decisions about data analysis and the production of scholarly papers.
 
-%```{r read.data,echo=FALSE,results=hide}
+```{r read.data,echo=FALSE,results='hide', cache=TRUE}
 ## This data frame will be used throughout the analyses.
 ## Putting it up top, so we can use it in the comments section.
-##  This read.csv line is slightly more complex than needed because of
-##  the particular encoding used for characters in the source file.
-dem.nations.df<-read.csv(url("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv",encoding="ISO_8859-1"),as.is=TRUE)
+library(readstata13)
+dem.nations.df<-read.dta13("https://www.dropbox.com/s/huhqag5rudwtbno/Democracy%20Cross-National%20Data%20V4.1%2009092015.dta?dl=1&pv=1&raw=1",convert.factors=FALSE)
+
 row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful
-good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
+good.df<-na.omit(dem.nations.df[,c("Gini2004","WVS5_protest","Nation","meanpr")])
 good.df$gini04<-good.df$Gini2004/100
-good.df$protac00<-good.df$protac2000
+good.df$protest05<-good.df$WVS5_protest
 ```
 
-\section{Data analysis is computer programming.}\label{sec:data-analys-comp}
+We organize the paper around a series of homilies that lead to certain concrete actions.
+
+# Data analysis is computer programming. {#sec:data-analys-comp}
 
@@
 
-Using R \citep{R:2011}, for example, I might specify that the file
-\Verb+fig1.pdf+ was produced by the following commands in a file
-called \Verb+fig1.R+.\footnote{The command \Verb+please-plot+ and some
-  other R functions used in this essay come from the
-  \Verb+MayIPleaseDoStatistics+ package which emphasizes politeness in
-  data analysis. Functions like \Verb+please-plot+ can be blocked and
-  more polite versions such as \Verb+may-I-please-have-a-plot+ can be
-  required using \Verb+options(politeness=99)+ }
+Using R [@R:2011], for example, I might specify that the file `fig1.pdf` was
+produced by the following commands in a file called `fig1.R`.^[The command
+`please-plot`and some other R functions used in this essay come from the
+`MayIPleaseDoStatistics` package which emphasizes politeness in data analysis.
+Functions like `please-plot`can be blocked and more polite versions such as
+`may-I-please-have-a-plot`can be required using `options(politeness=99)`. We do
+not recommend setting negative values for the politeness option should you use this package.]
 
-\begin{Verbatim}[fontsize=\footnotesize]
+```{r eval=FALSE}
+## This file produces a plot relating the explanatory variable to the outcome.
 thedata <- read.csv("Data/thedata-15-03-2011.csv") ## Read the data
-pdf('fig1.pdf') ## begin writing to the pdf file
+pdf("fig1.pdf") ## begin writing to the pdf file
 please-plot(outcome by explanatory using thedata. red lines please.)
@@
 ## Note to self: a quadratic term does not add to the substance
-## model2<-please-fit(outcome by explanatory+explanatory^2 using thedata
+## model2 <- please-fit(outcome by explanatory+explanatory^2 using thedata
 ## summary(abs(fitted(model1)-fitted(model2)))
 dev.off() ## stop writing to the pdf file
-\end{Verbatim}
+```
 
-Now, in the future if I wonder how ``that plot on page 10'' was
-created, I will know: (1) ``that plot'' is from a file called
-\Verb+fig1.pdf+ and (2) \Verb+fig1.pdf+ was created in
-\Verb+fig1.R+. In a future where R still exists, changing the figure
+Now, in the future if I wonder how 'that plot on page 10' was
+created, I will know: (1) 'that plot' is from a file called
+`fig1.pdf`and (2) `fig1.pdf`was created in
+`fig1.R`. In a future where R still exists, changing the figure
 will require quick edits of commands already written.  In a future
@@
 computing language: R scripts are written in
-\href{http://en.wikipedia.org/wiki/Plain_text}{plain text}, and plain
+[plain text](http://en.wikipedia.org/wiki/Plain_text), and plain
 text is a format that will be around as long as computer programmers
-write computer programs.\footnote{Since R is open
+write computer programs.^[Since R is open
   source, I will also be able to download an old version of R, download an
@@
   old-fashioned operating system in a virtual machine on my
-  new-fashioned actual machine.}
+  new-fashioned actual machine.]
 
-Moreover, realize that file names send messages to your future
-self. Name your files with evocative and descriptive names. Your
-collaborators are less likely to call you at midnight asking for help
-if your files are named \Verb+inequality-and-protest-figures.R+ than
-if your files are called \Verb+temp9+ or
-\Verb+supercalifragilisticexpialidocious+. The extension
-\Verb+.R+ tells us and the operating system that the file contains R
+Moreover, realize that *file names send messages* to your future
+self. This means that if you name your files with evocative and descriptive names, your
+collaborators are less likely to call you at midnight asking for help and you will remove some regret from your future self.
+For example, if you are studying inequalit and protest, you might try naming a file something like  `inequality-and-protest-figures.R` instead of  `temp9`or
+`supercalifragilisticexpialidocious`.  By the way, the extension
+`.R` tells us and the operating system that the file contains R
 commands. This part of the filename enables us to quickly search our
@@
 
-\paragraph{Step \thesection} If we know the provenance of results, 
+**Step 1** If we know the provenance of results,
 future  or current collaborators can quickly and easily reproduce
@@
 
-\section{No data analyst is an island for long.}
+# No data analyst is an island for long.
+
 Data analysis involves a long series of decisions. Each decision
@@
 technical for inclusion in the published article itself. These need to
-be documented in the code itself \citep{nagler1995coding}. Paragraphs
+be documented in the code itself [@nagler1995coding]. Paragraphs
 and citations in the publication will justify the most important
@@
 did something with data: comment your code and tightly link your code
-with your writing.\footnote{One can also try the R command
-  \Verb+put-it-in-my-mind(reason,importance='high')+ to firmly place a
+with your writing.^[One can also try the R command
+  `put-it-in-my-mind(reason,importance="high")` to firmly place a
   reason for a decision into the mind of the analyst. I myself have
-  not had much luck with this function.}
+  not had much luck with this function.]
 
-\subsection{Code to communicate: Comment your code.}
+## Code is communication: Comment code
+
 Comments --- unexecuted text inside of a script --- are a message to
@@
 term did not really change the substantive story arising from the
-model.\footnote{R considers text marked with \texttt{\#} as a comment.}
+model.^[R considers text marked with `#` as a comment.]
 Messages left for your future self (or near-future others) help
@@
 onwards.
-
-% Luckily, if you are
-% using a literate programming practice (ex. Sweave or odfWeave), you
-% can write paragraphs to surround your code as well as technical
-% comments in the code itself. If you are not using a strictly literate
-% programming practice then use whatever commenting protocol exists in
-% your analysis language. R, for example, considers text marked with
-% ``\#'' as a comment. Here for example I comment about a linear model
-% run for a paper and also include comments about another model run for
-% due diligence.
-
-% ```{r reg1,echo=TRUE,results=verb}
-% ##  Repeat the regression run for Table~\ref{tab:protest}
-% ##  Do political rights predict protest reporting?
-% lm1 <- lm(protac00 ~ gini04 + meanpr, data=good.df)
-% ##  Q: Was it worthwhile controlling for political rights?
-% lm2 <- lm(protac00~gini04,data=good.df)
-% coef(lm2)[2]-coef(lm1)[2]
-% ##  A: Yes (around 3/4 an act of protest difference)
-% ```
 
@@
 
-Comment liberally. Comments are discarded when R runs analysis or
-\LaTeX~turns plain text into page images, so only those who dig into
+Comment liberally. Comments are discarded when R runs analysis or LaTeX or pandoc
+turns plain text into pdf or html files so only those who dig into
 your work will see them.
 
-\subsection{Code to communicate: Literate programming.}
+## Code to communicate: Literate programming.
 
-\begin{quote}
-  Let us change our traditional attitude to the construction of
+> Let us change our traditional attitude to the construction of
   programs: Instead of imagining that our main task is to instruct a
@@
   human beings what we want a computer to do.
-  \citep[][p. 97]{knuth1984literate}
-\end{quote}
+  [@knuth1984literate, p.97]
 
@@
 arguments justifying your analysis decisions would also waste
-time. More importantly, we and others want to know why we did what we
+time.
+
+More importantly, we and others want to know why we did what we
 did. Such explanations may not be very clear if we have some pages of
 printed code in one hand and a manuscript in the other. Keep in mind
-the distinction between the ``source code'' of a document (i.e. what
+the distinction between the 'source code' of a document (i.e. what
 computation was required to produce it) and the visible, type-set page
@@
 reproducing or collaborating. The source code of any document
-exchanged by the group must be available and executable.
+exchanged by the group must be available and executable.^[As of the 2016 version of this paper, this idea is now widespread and made much easier than before via online services for code sharing and collaboration such as [GitHub](http://github.com) and [BitBucket](http://bitbucket.org).]
 
 How might one avoid these problems?
-\href{http://en.wikipedia.org/wiki/Literate_programming}{Literate
-  programming} is the practice of weaving code into a document ---
+[Literate programming](http://en.wikipedia.org/wiki/Literate_programming) is the practice of weaving code into a document ---
 paragraphs, equations, and diagrams can explain the code, and the code
@@
 
-For example, in \S\ref{sec:data-analys-comp}, I suggested that we know
-where ``that plot on page 10'' comes from by making sure we had a
-\Verb+fig1.pdf+ file produced from a clearly commented plain text file
-called something like \Verb+fig1.R+.  An even easier solution would be to
+For example, in [@sec:data-analys-comp], I suggested that we know
+where 'that plot on page 10 'comes from by making sure we had a
+`fig1.pdf` file produced from a clearly commented plain text file
+called something like `fig1.R`.  An even easier solution would be to
 directly include a chunk of code to produce the figure inside of the
@@
 \LaTeX~ markup with R code chunks to make things like
-Figure~\ref{fig:giniprot}. This combination of \LaTeX and R is called
-Sweave \citep{Leis:2005}.\footnote{Support for Sweave is included with
-  R.}
+Figure~\ref{fig:giniprot}. This combination of LaTeX and R is called
+Sweave [@leis:2005].^[Support for Sweave is included with R.]
 
-\begin{Verbatim}[fontsize=\footnotesize]
-This paper, for example, was written in plain text using \LaTeX
+```
+This paper, for example, was written in plain text using Markdown
 markup with R code chunks to make things like
-Figure~\ref{fig:giniprot}. This combination of \LaTeX and R is
-called Sweave \citep{Leis:2005}.\footnote{Support for Sweave is
-  included with R.}
+Figure~\ref{fig:giniprot}. This combination of Markdown and R is
+called [R markdown](http://rmarkdown.rstudio.com/).\footnote{Support for R markdown is
+  included with R. For nicer documents, use the `rmarkdown` R package or click `Knit to PDF` or `Knit to HTML` using the free and open-source [RStudio IDE](http://rstudio.com) for R.}
 \begin{figure}[h]
 \centering
- ```{r fig1plot,fig=TRUE}
+\`\`\`{r fig1plota,fig=TRUE}
 ##  Make a scatterplot of Protest by Inequality
- with(good.df,plot(gini04,protac00,xlab='Gini Coefficient 2004 (UNDP)',
-                   ylab='Mean Protest Activities\n(World Values Survey 1980-2000)'))
+with(good.df,plot(gini04,protest05,xlab='Gini Coefficient 2004 (UNDP)',
+                  ylab='Mean Protest Activities\n(World Values Survey 2005--2010)'))
 ##  Label a few interesting points
 with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
-      text(gini04,protac00,labels=Nation))
- ```
-\caption{Protest activity by income inequality \citep[from][]{norris2009data}.}
+     text(gini04,protest05,labels=Nation))
+\`\`\`
+\caption{Protest activity by income inequality [@norris2009data].}
 \label{fig:giniprot}
 \end{figure}
-\end{Verbatim}
+```
 
@@
 ##  Make a scatterplot of Protest by Inequality
-with(good.df,plot(gini04,protac00,
+with(good.df,plot(gini04,protest05,
                   xlab='Gini Coefficient 2004 (UNDP)',
@@
 with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
-     text(gini04,protac00,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
+     text(gini04,protest05,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
 ```
@@
   \begin{center}
-```{r fig1plot,echo=FALSE,fig=TRUE,width=4,height=4}
+```{r fig1plot, ref.label="fig1code",echo=FALSE,fig=TRUE,width=4,height=4}
 par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
-```{r fig1code>>
 ```
-    \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}
+    \caption{Protest activity by income inequality @norris2009data.}
     \label{fig:giniprot}
@@
 
-By using \Verb+\label{fig:giniprot}+, I do not need to keep track of
+By using `\label{fig:giniprot}`, I do not need to keep track of
 the figure number, nor do extra work when I reorganize the document in
 response to reviewer suggestions. Nor do I need a separate
-\Verb+fig1.R+ file or \Verb+fig1.pdf+ file.  Tables and other
+`fig1.R`file or `fig1.pdf` file.  Tables and other
 numerical results are also possible to generate within the source
@@
 see how Table~\ref{tab:protest} was also generated directly from a
-regression object.\footnote{\citet{beck2010reg} inspired this
+regression object.\footnote{@beck2010reg inspired this
   particular presentation of a linear model.}
 
-```{r maketable, echo=FALSE, results=hide}
-  lm1<-lm(protac00~gini04+meanpr,data=good.df)  #  Run the regression
+```{r maketable, echo=FALSE, results='hide'}
+  lm1<-lm(protest05~gini04+meanpr,data=good.df)  #  Run the regression
 
@@
   more equal income distributions, adjusting for average political
-  rights as measured by Freedom House 1980--2000. Data from \\citep{norris2009data}.",label="tab:protest")
+  rights as measured by Freedom House 1980--2000. Data from [@norris2009data].",label="tab:protest")
 ```
 
-```{r printtable,echo=FALSE,results=tex}
+```{r printtable,echo=FALSE,results='asis'}
   print(thextab,include.colnames=FALSE,
@@
 Literate data analysis is not the same as Sweave, even if Sweave is a
-nice implementation.\footnote{The R project has a task view devoted to
-  \href{http://cran.r-project.org/web/views/ReproducibleResearch.html}{reproducible
-    research} listing many of the different approaches to literate
-  programming for R.}  \href{http://www.lyx.org/}{LyX} offers a
-WYSIWYG environment for \LaTeX~ that supports Sweave. And the
-\Verb+odfWeave+ package in R allows the use of OpenOffice documents in
-exactly the same way.\footnote{A quick Google search of ``Sweave for
-  Stata'' turned up lots of resources for literate programming with
-  Stata.} If your workflow does not involve \LaTeX~ and R, you can
+nice implementation.^[The R project has a task view devoted to
+  [reproducible research](http://cran.r-project.org/web/views/ReproducibleResearch.html) listing many of the different approaches to literate
+  programming for R.}  If your workflow does not involve R, you can
 still implement some of the principles here. Imagine creating a style
-in Microsoft Word called ``code'' which hides your code when you print
+in Microsoft Word called 'code 'which hides your code when you print
 your document, but which allows you to at least run each code chunk
-piece by piece [or perhaps there are ways to extract all text of style
-``code'' from a Microsoft Word document]. Or imagine just using some
+piece by piece (or perhaps there are ways to extract all text of style
+code from a Microsoft Word document). Or imagine just using some
 other kind of indication linking paragraphs to specific places in code
@@
 Literate programming need not go against the principle of modular data
-analysis \citep{nagler1995coding}. In my own work I routinely have
+analysis [@nagler1995coding]. In my own work I routinely have
 several different Sweave files that fulfill different functions, some
-of them create \LaTeX code that I can \Verb+\input+ into my
-\Verb+main.tex+ file, others setup the data, run simulations, or allow
+of them create \LaTeX code that I can `\input` into my
+`main.tex` file, others setup the data, run simulations, or allow
 me to record my journeys down blind alleys. Of course, when we have
@@
 
-\paragraph{Step \thesection}
+** Step 2 **
 
@@
 (including our future selves) can retrace our steps. A document than
-can be ``run'' to reproduce all of the analyses also instills
+can be 'run 'to reproduce all of the analyses also instills
 confidence in readers and can more effectively spur discussion and
@@
 
-\section{Meaningful code requires data.}
+# Meaningful code requires data.
 
@@
 file. A reference to a data file is a line of code the analysis
-program will use to operate on (``load''/ ``open'' / ``get'' /
-``use'') the data file. One should not have to edit this line on
+program will use to operate on ('load'/ 'open '/ 'get '/
+'use') the data file. One should not have to edit this line on
 different computers or platforms in order to execute this command.
 Using R, for example, all analysis files should have
-\Verb+load('thedata.rda')+ or \Verb+read.csv('thedata.csv')+ or some
-equivalent line in them, and \Verb+thedata.csv+ should be stored in
+`load("thedata.rda")`or `read.csv("thedata.csv")`or some
+equivalent line in them, and `thedata.csv`should be stored in
 some place easy to find (like in the same directory as the file or
-perhaps in \Verb+'Data/thedata.rda'+). Of course, it never hurts to
+perhaps in `"Data/thedata.rda"`). Of course, it never hurts to
 drop in a comment pointing to the data file.
@@
 co-authors to have the same directory structure so that
-\Verb+load('Data/thedata.rda')+ means the same thing on all computers
+`load("Data/thedata.rda")` means the same thing on all computers
 used to work on the project. This kind of solution is one of the
 things that formal version control systems do well (as discussed
-a bit more in \S\ref{sec:vers-contr-prev}).
+a bit more in [@sec:vers-contr-prev]).
 
@@
 processing, recoding, and merging from analysis in different files
-\citep{nagler1995coding}. So, perhaps your analysis oriented files
-will \Verb+load('cleandata.rda')+ and a comment in the code will alert
-the future you (among others) that \Verb+cleandata.rda+ was created
-from \Verb+create-cleandata.R+ which in turn begins with
-\Verb+read.csv(\url('http://www.greatfreedata.gov/dirtydata.csv'))+. Such
+[@nagler1995coding]. So, perhaps your analysis oriented files
+will `load("cleandata.rda")`and a comment in the code will alert
+the future you (among others) that `cleandata.rda`was created
+from `create-cleandata.R` which in turn begins with
+`read.csv(\url("http://www.greatfreedata.gov/dirtydata.csv"))`. Such
 a data processing file will typically end with something like
-\Verb+save('cleandata.rda')+ so that we are doubly certain about the
+`save("cleandata.rda")`so that we are doubly certain about the
 provenance of the data.\footnote{Of course, if you need math or
@@
   prefer to make them into Sweave files, for which the conventional
-  extension is .Rnw. So you'd have \Verb+create-cleandata.Rnw+ which
+  extension is .Rnw. So you'd have `create-cleandata.Rnw` which
   might explain and explore the different coding decisions you made,
@@
 
-Now, if in the future we wonder where \Verb+cleandata.rda+ came from, we
-might search for occurrences of `cleandata' in the files on our
+Now, if in the future we wonder where `cleandata.rda`came from, we
+might search for occurrences of `cleandata` in the files on our
 system. However, if such searching among files is a burden, an even
 nicer solution is to maintain a file for each project called
-``MANIFEST.txt'' or ``INDEX.txt'' or ``README.txt'' which lists the
+`MANIFEST.txt` or INDEX.txt or README.txt which lists the
 data and command files with brief descriptions of their functions and
@@
 
-\paragraph{Step 3} We should know where the data came
+** Step 3** We should know where the data came
 from and what operations were performed on which set of data.
@@
 
-\begin{Verbatim}[fontsize=\footnotesize]
+```
 DA NI=19 NO=199 MA=CM
@@
 CM FI=basic.cov
-\end{Verbatim}
+```
 
@@
 
-\section{Version control prevents clobbering and reconciles history.}\label{sec:vers-contr-prev}
+# Version control prevents clobbering and reconciles history. {#sec:vers-contr-prev}
 
-Group work requires version control.\footnote{
-  \citet{fredrickson2011tpm} and \citet{healy2011tpm} in this issue
-  also explain what version control is and why we might want to use
-  it.} Many people are familiar with the ``track changes'' feature in
-modern WYSIWYG word processors or the fact that Dropbox allows one to
-recover previous versions of files. These are both kinds of version
-control. More generally, when we collaborate, we'd like to do a variety
-of actions with our shared files. Collaboration on data analytic
-projects is more productive and better when (1) it is easy to see what
-has changed between versions of files; (2) members of the team feel
-free to experiment and then to dump parts of the experiment in favor
-of previous work while merging the successful parts of the experiment
-into the main body of the paper; (3) the team can produce have
-``releases'' of the same document (one to MPSA, one to APSR,
-one to your parents) without spawning many possibly conflicting copies
-of the same document; (4) people can work on the same files at
-the same time without conflicting with one another (and can reconcile
-their changes without too much confusion and clobbering).  Clobbering
-is what happens when your future self or your current collaborator
-saves an old version of a file over a new version, erasing good
-work by accident.
+Group work requires version control.^[@fredrickson2011tpm and
+@healy2011tpm in this issue also explain what version control is and why we
+might want to use it.] Many people are familiar with the track changes
+feature in modern WYSIWYG word processors or the fact that Dropbox allows one
+to recover previous versions of files. These are both kinds of version control.
+More generally, when we collaborate, we'd like to do a variety of actions with
+our shared files. Collaboration on data analytic projects is more productive
+and better when (1) it is easy to see what has changed between versions of
+files; (2) members of the team feel free to experiment and then to dump parts
+of the experiment in favor of previous work while merging the successful parts
+of the experiment into the main body of the paper; (3) the team can produce
+have 'releases 'of the same document (one to MPSA, one to APSR, one to your
+parents) without spawning many possibly conflicting copies of the same
+document; (4) people can work on the same files at the same time without
+conflicting with one another (and can reconcile their changes without too much
+confusion and clobbering).  Clobbering is what happens when your future self or
+your current collaborator saves an old version of a file over a new version,
+erasing good work by accident.
 
-Of course if you rely on Dropbox or ``track changes'' for version
+Of course if you rely on Dropbox or 'track changes 'for version
 control, you must communicate with other folks in your group before
@@
 
-We use subversion with our own research group, and I use it for all of
-my own projects (except this one, for which I am experimenting with
-git). Subversion and bazaar and git are all great. They mainly differ
-in the extent to which you need to run a server. Subversion requires a
-server.\footnote{If you already pay to host a website, you may already
-  have the right to run a subversion or git server there. Your
-  university or institute may have a version control system running
-  somewhere on campus. And Google will direct you to many helpful
-  people who have installed such servers on their own diverse desktop
-  machines. Github requires that you pay to host private
-  repositories.}
+We use git with our own research group, and I use it for all of
+my own projects.MORE ON GITHUB
 
@@
 as plain text then you can easily compare documents using the many
-utilities available for comparing text files.\footnote{Adobe Acrobat
+utilities available for comparing text files.^[Adobe Acrobat
   allows one to compare differences in pdf files. OpenOffice supports
-  a ``Compare Documents'' option. And Google Docs will report on the
-  version history of a document.} When you reach certain milestones
+  a 'Compare Documents 'option. And Google Docs will report on the
+  version history of a document.] When you reach certain milestones
 you can rename the file accordingly: thedocAPSA2009.tex --- for the
@@
 you can do it by hand too as long as you don't mind taking up a lot of
-disk space and having many ``thedoc...''  files around. If you do
+disk space and having many 'thedoc... ' files around. If you do
 version control by hand, spend a little extra time to ensure that you
@@
 
-\paragraph{Step \thesection} Writing is rewriting. Thus, all writing
+** Step 5** Writing is rewriting. Thus, all writing
 involves versions. When we collaborate with ourselves and others we
@@
 
-% This is so true that the
-% source of the quote is unclear.\footnote{Could it be Paul Abbott?
-%   \url{http://www.bbc.co.uk/writersroom/insight/paul_abbott.shtml}}
-% Here is a quote with a known source: ``Writing and rewriting are a constant search for what it is one is
-% saying'' (John Updike but ??Writing With Style: Conversations on the Art of Writing, 1975 
-% by John Trimble ??)
-
-\section{Minimize error by testing.}
+# Minimize error by testing.
 
@@
 statisticians at the UN worry about the bootstrap confidence intervals
-presented in the original paper.\footnote{Perhaps they should be
+presented in the original paper.^[Perhaps they should be
   worried about the deeper substantive critiques offered by the
@@
   concerns as the student, they kept their mouths shut at the
-  mini-conference to avoid looking dumb in front of their bosses.} So,
+  mini-conference to avoid looking dumb in front of their bosses.] So,
 now the authors would like to evaluate their bootstrap procedure. Although
@@
 undertaking the odd bit of research, they might say to themselves,
-``Before I write new code, I should write a test of the code. I should
+"Before I write new code, I should write a test of the code. I should
 write a little bit of code that lets me know that my double-bootstrap
-procedure actually does what it is supposed to do.''
+procedure actually does what it is supposed to do."
 
@@
 for multi-million dollar programs. The idea of
-\href{http://en.wikipedia.org/wiki/Test-driven_development}{test
-  driven development} and the idea that one ought to create tests of
-\href{http://en.wikipedia.org/wiki/Unit_testing}{small parts of one's
-  code} arose to address such concerns. For the social scientist
+[test driven development](http://en.wikipedia.org/wiki/Test-driven_development) and the idea that one ought to create tests of
+[small parts of one's code](http://en.wikipedia.org/wiki/Unit_testing) arose to address such concerns. For the social scientist
 collaborating with her future self and/or a small group of
@@
 
-Ack! I mistyped ``+'' for ``*''. Good thing I wrote the
-test!\footnote{A more common example of this kind of testing occur
+Ack! I mistyped '+' for '*'. Good thing I wrote the
+test!^[A more common example of this kind of testing occur
   everyday when we recode variables into new forms but look at a
-  crosstab of the old vs. new variable before proceeding.}
+  crosstab of the old vs. new variable before proceeding.]
 
-\paragraph{Step \thesection} No one can forsee all of the ways that a
+**Step 6** No one can forsee all of the ways that a
 computer program can fail. One can, however, at least make sure that
@@
 
-\section{Copy and improve on others' examples.}
+# Copy and improve on others' examples.
 
-Lots of people are thinking about ``reproducible research'' and
-``literate programming'' these days. Google those terms. Of course, the
+Lots of people are thinking about 'reproducible research'and
+'literate programming' these days. Google those terms. Of course, the
 devil is in the details: Here I list a few of my own attempts at enabling
@@
 
-I have experimented with three systems so far: (1) for one paper we
-simply included a Sweave document and data files into a compressed
-archive \citep{bowers2005dataverse}; (2) for another more computing
-intensive paper we assembled a set of files that enabled reproduction
-of our results using the \texttt{make} system
-\citep{bowers2008dataverse}; and (3) recently I have tried the
-``compendium'' approach
-\citep{gentleman2005reproducible,gentleman2007statistical} which
-embeds an academic paper within the R package system
-\citep{bowers2011dataverse}. The benefit of this last approach is that
-one is not required to have access to a command line for \Verb+make+:
-the compendium is downloadable from within R using
-\Verb+install.packages()+ and is viewable using the \Verb+vignette()+
-function in any operating system than runs R.\footnote{Notice that my
-  reproduction archives and/or instructions for using them are hosted
-  on the
-  \href{http://thedata.org/book/learn-about-project}{Dataverse}, which
-  is another system designed to enhance academic collaboration across
-  time and space.}  The idea that one ought to be able to install and
-run and use an academic paper just as one installs and uses
-statistical software packages is very attractive, and I anticipate
-that it will become ever easier to turn papers into R packages as
-creative and energetic folks turn their attention to the question of
+I have experimented with three systems so far: (1) for one paper we simply
+included a Sweave document and data files into a compressed archive
+[@bowers2005dataverse]; (2) for another more computing intensive paper we
+assembled a set of files that enabled reproduction of our results using the
+`make` system [@bowers2008dataverse]; and (3) recently I have tried the
+"compendium" approach [@gentleman2005reproducible,@gentleman2007statistical]
+which embeds an academic paper within the R package system
+[@bowers2011dataverse]. The benefit of this last approach is that one is not
+required to have access to a command line for `make`: the compendium is
+downloadable from within R using `install.packages()` and is viewable using the
+`vignette()` function in any operating system than runs R.^[Notice that my
+reproduction archives and/or instructions for using them are hosted on the
+[Dataverse](http://thedata.org/book/learn-about-project), which is another
+system designed to enhance academic collaboration across time and space.]  The
+idea that one ought to be able to install and run and use an academic paper
+just as one installs and uses statistical software packages is very attractive,
+and I anticipate that it will become ever easier to turn papers into R packages
+as creative and energetic folks turn their attention to the question of
 reproducible research.
 
-\paragraph{Step \thesection} We all learn by doing. When we share
+**Step 6** We all learn by doing. When we share
 reproduction materials we improve both cumulation of knowledge and our
 methods for doing social science
-\citep{freese2007replication,king1995replication}.  As we copy and
+[@freese2007replication,king1995replication}.  As we copy and
 improve upon each other's modes of doing work we enhance our collective
@@
 
-\section{Remember that research ought to be credible communication.}
-\begin{quote}
-  [I]f the empirical basis for an article or book cannot be
+# Remember that research ought to be credible communication.
+
+> [I]f the empirical basis for an article or book cannot be
   reproduced, of what use to the discipline are its
   conclusions?  What purpose does an article like this serve?
-  \cite[445]{king1995replication}
-\end{quote}
+  [@king1995replication, p. 445]
 
@@
 science departments (by then dominated by cetaceans after humans are
-mostly uploads).\footnote{The arrival of the six-legged social
+mostly uploads).^[The arrival of the six-legged social
   scientists revives Emacs and finally makes Ctrl-a Ctrl-x Esc-x
-  Ctrl-c a \href{http://kieran.healy.usesthis.com/}{reasonable key
-    combination}.}
+  Ctrl-c a [reasonable key combination](http://kieran.healy.usesthis.com/).]
 
@@
 training and research resources).
+
 % Such work should make us change how
@@
 narrowly, assume we want other people to believe us when we say
-something about data: ``data'' here can be words, numbers, musical
+something about data: 'data 'here can be words, numbers, musical
 notes, images, ideas, etc \ldots The point is that we are making some
 claims about patterns in some collection of stuff. Now, it might be
-easy to convince others that ``this collection of stuff is different
-from that collection of stuff'' if those people were looking over our
+easy to convince others that 'this collection of stuff is different
+from that collection of stuff 'if those people were looking over our
 shoulders the whole time that we made decisions about collecting the
@@
 In the end, following these practices and those recommended by
-\citet{fredrickson2011tpm} and \citet{healy2011tpm} in this issue
+@fredrickson2011tpm and @healy2011tpm in this issue
 allows your computerized analyses of your collections of stuff to be
@@
 self can shoot them the archive required to reproduce your
-work.\footnote{Since you used plain text, the files will still be
+work.^[Since you used plain text, the files will still be
   intelligible, analyzed using commented code so that folks can
@@
   a copy of Ubuntu 10 and an open source virtual machine running the
-  whole environment.} You can say, ``Here is everything you need to
-reproduce my work." To be extra helpful you can add ``Read the README
-file for further instructions." And then you can get on with your life:
+  whole environment.] You can say, 'Here is everything you need to
+reproduce my work.' To be extra helpful you can add 'Read the README
+file for further instructions.' And then you can get on with your life:
 maybe the next great idea will occur when your 4-year-old asks a wacky
@@
 
-\bibliographystyle{apsr}
-\bibliography{/Users/jwbowers/Documents/BIB/trunk/big}
-
-\end{document}
-```{r }
-options(prompt="> ",continue="+ ")
-```
+#References

### 2016-06-06 workflow.Rmd
Version ready for Maarten

@@
     pdf_document:
+      keep_tex: true
       number_sections: true
@@
       pandoc_args: [
+      "-M", "secPrefix=\\S",
       "--filter", "/usr/local/bin/pandoc-crossref"
       ]
+      md_extensions: +autolink_bare_uris+ascii_identifiers+tex_math_single_backslash
 ...
@@
 
-row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful
+row.names(dem.nations.df)<-dem.nations.df$Nation ##make the row names more helpful
 good.df<-na.omit(dem.nations.df[,c("Gini2004","WVS5_protest","Nation","meanpr")])
@@
 
-Comments --- unexecuted text inside of a script --- are a message to
+Comments, unexecuted text inside of a script, are a message to
 collaborators (including your future self) and other consumers of your
@@
 paper itself. This paper, for example, was written in plain text using
-\LaTeX~ markup with R code chunks to make things like
-Figure~\ref{fig:giniprot}. This combination of LaTeX and R is called
-Sweave [@leis:2005].^[Support for Sweave is included with R.]
+markdown markup with R code chunks to make things like
+Figure \ref{fig:giniprot}. This combination of markdown and R is called
+[R Markdown](http://rmarkdown.rstudio.org).
 
-```
-This paper, for example, was written in plain text using Markdown
-markup with R code chunks to make things like
-Figure~\ref{fig:giniprot}. This combination of Markdown and R is
-called [R markdown](http://rmarkdown.rstudio.com/).\footnote{Support for R markdown is
-  included with R. For nicer documents, use the `rmarkdown` R package or click `Knit to PDF` or `Knit to HTML` using the free and open-source [RStudio IDE](http://rstudio.com) for R.}
-\begin{figure}[h]
-\centering
-\`\`\`{r fig1plota,fig=TRUE}
+```{r fig1plotcode,eval=FALSE,echo=TRUE}
 ##  Make a scatterplot of Protest by Inequality
-with(good.df,plot(gini04,protest05,xlab='Gini Coefficient 2004 (UNDP)',
-                  ylab='Mean Protest Activities\n(World Values Survey 2005--2010)'))
-##  Label a few interesting points
-with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
-     text(gini04,protest05,labels=Nation))
-\`\`\`
-\caption{Protest activity by income inequality [@norris2009data].}
-\label{fig:giniprot}
-\end{figure}
-```
-
-```{r fig1code,echo=FALSE,fig=FALSE}
-##  Make a scatterplot of Protest by Inequality
+par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
 with(good.df,plot(gini04,protest05,
                   xlab='Gini Coefficient 2004 (UNDP)',
-                  ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
+                  ylab='Proportion Reporting Protest Activities\n(World Values Survey 2005)',
                   cex=.8))
-
 ##  Label a few interesting points
-with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
+with(good.df[c("Brazil","United Kingdom","United States","Sweden","Chile"),],
      text(gini04,protest05,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
@@
 
-\begin{figure}[H]
-  \begin{center}
-```{r fig1plot, ref.label="fig1code",echo=FALSE,fig=TRUE,width=4,height=4}
+```{r giniprot,message=FALSE,echo=FALSE,fig=TRUE,include=TRUE,width=4,height=4,fig.cap="Average number of Protest activity by income inequality.",out.width=".5\\textwidth",fig.env="figure",fig.lp="fig:",fig.pos="H",fig.align="center"}
+##  Make a scatterplot of Protest by Inequality
 par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
+with(good.df,plot(gini04,protest05,
+                  xlab='Gini Coefficient 2004 (UNDP)',
+                  ylab='Proportion Reporting Protest Activities\n(World Values Survey 2005)',
+                  cex=.8))
+##  Label a few interesting points
+with(good.df[c("Brazil","United Kingdom","United States","Sweden","Chile"),],
+     text(gini04,protest05,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
 ```
-    \caption{Protest activity by income inequality @norris2009data.}
-    \label{fig:giniprot}
-  \end{center}
-\end{figure}
 
-By using `\label{fig:giniprot}`, I do not need to keep track of
-the figure number, nor do extra work when I reorganize the document in
-response to reviewer suggestions. Nor do I need a separate
-`fig1.R`file or `fig1.pdf` file.  Tables and other
-numerical results are also possible to generate within the source
-code of a scholarly paper. Those who view the code for this essay will
-see how Table~\ref{tab:protest} was also generated directly from a
-regression object.\footnote{@beck2010reg inspired this
-  particular presentation of a linear model.}
+Markdown (and LaTeX and HTML) all have ways to cross-reference within a
+document. For example, by using `\label{fig:giniprot}` in LaTeX or
+`{#fig:giniprot}` in the pandoc version of markdown, I do not need to keep
+track of the figure number, nor do extra work when I reorganize the document in
+response to reviewer suggestions. Nor do I need a separate `fig1.R`file or
+`fig1.pdf` file.  Tables and other numerical results are also possible to
+generate within the source code of a scholarly paper. Those who view the code
+for this essay will see how Table \ref{tab:protest} was also generated directly
+from a regression object.\footnote{@beck2010reg inspired this particular
+presentation of a linear model.}
 
@@
   more equal income distributions, adjusting for average political
-  rights as measured by Freedom House 1980--2000. Data from [@norris2009data].",label="tab:protest")
+  rights as measured by Freedom House 1980--2000. Data from [@norris2015data].",label="tab:protest")
 ```
 
-```{r printtable,echo=FALSE,results='asis'}
+```{r printtable,echo=FALSE,results='asis',message=FALSE}
   print(thextab,include.colnames=FALSE,
@@
         table.placement="H",
+	comment=FALSE,
         add.to.row=list(
@@
 
-** Step 2 **
-
+**Step 2**
 We analyze data in order to explain something about the world to other
@@
 (including our future selves) can retrace our steps. A document than
-can be 'run 'to reproduce all of the analyses also instills
+can be 'run' to reproduce all of the analyses also instills
 confidence in readers and can more effectively spur discussion and
@@
 
+[MAARTEN IDEAS HERE? PLUS ANOTHER SECTION ABOUT ORGANIZATION?]
+
 Where should one store data files? An obvious solution is always to
@@
 processing, recoding, and merging from analysis in different files
-[@nagler1995coding]. So, perhaps your analysis oriented files
-will `load("cleandata.rda")`and a comment in the code will alert
-the future you (among others) that `cleandata.rda`was created
-from `create-cleandata.R` which in turn begins with
-`read.csv(\url("http://www.greatfreedata.gov/dirtydata.csv"))`. Such
-a data processing file will typically end with something like
-`save("cleandata.rda")`so that we are doubly certain about the
-provenance of the data.\footnote{Of course, if you need math or
-  paragraphs to explain what is happening in these files, you might
-  prefer to make them into Sweave files, for which the conventional
-  extension is .Rnw. So you'd have `create-cleandata.Rnw` which
-  might explain and explore the different coding decisions you made,
-  perhaps involving a factor analysis and diagnostic plots.}
+[@nagler1995coding]. So, perhaps your analysis oriented files will
+`load("cleandata.rda")`and a comment in the code will alert the future you
+(among others) that `cleandata.rda`was created from `create-cleandata.R` which
+in turn begins with
+`read.csv(\url("http://www.greatfreedata.gov/dirtydata.csv"))`. Such a data
+processing file will typically end with something like
+`save("cleandata.rda")`so that we are doubly certain about the provenance of
+the data.\footnote{Of course, if you need math or paragraphs to explain what is
+happening in these files, you might prefer to make them into R markdown or
+Sweave files, for which the conventional extension is .Rmd or .Rnw
+respectively. So you'd have `create-cleandata.Rnw` written as a mixture of
+laTeX and R which might explain and explore the different coding decisions you
+made, perhaps involving a factor analysis and diagnostic plots.}
 
@@
 
-** Step 3** We should know where the data came
+**Step 3** We should know where the data came
 from and what operations were performed on which set of data.
@@
 understand and informative lines, what files (with filenames no longer
-than 8 characters) to use:
+than 8 characters) to use. See for example, this extremely clear code snippet:
 
@@
 
-# Version control prevents clobbering and reconciles history. {#sec:vers-contr-prev}
+# Version control prevents clobbering, reconciles history, and helps organize work. {#sec:vers-contr-prev}
 
@@
 of the experiment into the main body of the paper; (3) the team can produce
-have 'releases 'of the same document (one to MPSA, one to APSR, one to your
+have 'releases' of the same document (one to MPSA, one to APSR, one to your
 parents) without spawning many possibly conflicting copies of the same
@@
 
-Of course if you rely on Dropbox or 'track changes 'for version
-control, you must communicate with other folks in your group before
-you edit existing files. Only one of you can edit and save a given
-file at a time. This prevents your work (or your colleagues work) from
-getting lost when you both try to save the same file on top of each
-other. If you find that you need to work on the same files at the same
-time, then you should work on establishing your own shared version
-control system. Free options include launchpad, github, sourceforge
-for open source projects (i.e. papers you are writing which you are
-happy to share with others as you write). Each of those services
-include paid versions too. One may also use Dropbox as a kind of
-server for version control: for example, one may copy files from the
-Dropbox directory into a local working directory so as to avoid
-clobbering and then work on merging changes by hand before copying
-back to the Dropbox directory and replacing existing files.
+Of course if you rely on Dropbox or 'track changes' in Word for version
+control, you must communicate with other folks in your group before you edit
+existing files. Only one of you can edit and save a given file at a time. This
+prevents your work (or your colleagues work) from getting lost when you both
+try to save the same file on top of each other. If you find that you need to
+work on the same files at the same time, then you should work on establishing
+your own shared version control system. Free options include launchpad, github,
+sourceforge for open source projects (i.e. papers you are writing which you are
+happy to share with others as you write). As of this writing, Google Docs does
+an excellent job with shared editing of files as long as your are online. Each
+of those services include paid versions too. One may also use Dropbox as a kind
+of server for version control: for example, one may copy files from the Dropbox
+directory into a local working directory so as to avoid clobbering and then
+work on merging changes by hand before copying back to the Dropbox directory
+and replacing existing files. MORE PERHAPS ON GOOGLE? LESS ON DROPBOX?
 
 We use git with our own research group, and I use it for all of
-my own projects.MORE ON GITHUB
+my own projects. MORE ON GITHUB
 
@@
 
-** Step 5** Writing is rewriting. Thus, all writing
+**Step 5** Writing is rewriting. Thus, all writing
 involves versions. When we collaborate with ourselves and others we
@@
 
-Now, back to that famous article of 2018. After reading the conference
-paper critique of 2021 (that came from the seminar paper of 2020), the
-statisticians at the UN worry about the bootstrap confidence intervals
-presented in the original paper.^[Perhaps they should be
-  worried about the deeper substantive critiques offered by the
-  student, but they are statisticians and so focus on the stats. The
-  policy makers of 2021 were cowed by the methodological virtuosity of
-  the 2018 article, and so, even though they had the same substantive
-  concerns as the student, they kept their mouths shut at the
-  mini-conference to avoid looking dumb in front of their bosses.] So,
-now the authors would like to evaluate their bootstrap procedure. Although
-nice code exists for bootstrapping linear models, no nice code exists
-to bootstrap the bootstrap. Of course, the code required is not
-complex, but since they are writing custom code they worry about
-getting it right. As they've struggled to respond to the critiques of
-their paper, they've had lots of time to appreciate problems arising
-from bugs, errors, and typos in data analysis and code.
+Anyone writing custom code should worry about getting it right. The more code one writes, the more time one has to appreciate problems arising from bugs, errors, and typos in data analysis
+and code.
 
-Now, if they had a moment to think in between teaching that new class,
+Now, if one has a moment to think in between teaching that new class,
 reading books for an awards committee, evaluating application files
 for the admissions committee, feeding popsicles to a sick child, and
-undertaking the odd bit of research, they might say to themselves,
+undertaking the odd bit of research, one might say to oneself,
 "Before I write new code, I should write a test of the code. I should
@@
 [test driven development](http://en.wikipedia.org/wiki/Test-driven_development) and the idea that one ought to create tests of
-[small parts of one's code](http://en.wikipedia.org/wiki/Unit_testing) arose to address such concerns. For the social scientist
+[small parts of one's code](http://en.wikipedia.org/wiki/Unit_testing) arose to address such concerns.^[There are now R packages to help R package writers do this, see for example <https://github.com/hadley/testthat> and the article on it <https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf>.] For the social scientist
 collaborating with her future self and/or a small group of
@@
 
+MAYBE SOMETHING HERE ON `stopifnot` and recoding data.
+
 **Step 6** No one can forsee all of the ways that a
@@
 assembled a set of files that enabled reproduction of our results using the
-`make` system [@bowers2008dataverse]; and (3) recently I have tried the
-"compendium" approach [@gentleman2005reproducible,@gentleman2007statistical]
-which embeds an academic paper within the R package system
-[@bowers2011dataverse]. The benefit of this last approach is that one is not
-required to have access to a command line for `make`: the compendium is
-downloadable from within R using `install.packages()` and is viewable using the
-`vignette()` function in any operating system than runs R.^[Notice that my
-reproduction archives and/or instructions for using them are hosted on the
+`make` system [@bowers2008dataverse]; (3) I have tried the "compendium"
+approach [@gentleman2005reproducible,@gentleman2007statistical] which embeds an
+academic paper within the R package system [@bowers2011dataverse]; and (4)
+recently I have started organizing my work on github which enables me to make
+"releases" of the project which should enable smooth reproduction of all of the
+products of the research (simulations, data analyses, tables, figures, etc..).
+The benefit of the compendium approach is that one is not required to have
+access to a command line for `make`: the compendium is downloadable from within
+R using `install.packages()` and is viewable using the `vignette()` function in
+any operating system than runs R.^[Notice that my reproduction archives and/or
+instructions for using them are hosted on the
 [Dataverse](http://thedata.org/book/learn-about-project), which is another
@@
 as creative and energetic folks turn their attention to the question of
-reproducible research.
+reproducible research.^[See for example <http://r-pkgs.had.co.nz/>]
 
@@
 
-But what if the UN never hears of your work, or, by some cruel fate, your
+But what if no one ever hears of your work, or, by some cruel fate, your
 article does not spawn debate? Why then would you spend time to
@@
 
-% Such work should make us change how
-% we act --- or at least, it ought to stand on stronger epistemological
-% ground than other claims about experiences which ought to be shared or
-% shareable.
-
-Assume we want others to believe us when we say something. More
-narrowly, assume we want other people to believe us when we say
-something about data: 'data 'here can be words, numbers, musical
-notes, images, ideas, etc \ldots The point is that we are making some
-claims about patterns in some collection of stuff. Now, it might be
-easy to convince others that 'this collection of stuff is different
-from that collection of stuff 'if those people were looking over our
-shoulders the whole time that we made decisions about collecting the
-stuff and broke it up into understandable parts and reorganized and
-summarized it. Unfortunately, we can't assume that people are willing
-to shadow a researcher throughout her career. Rather, we do our work
-alone or in small groups and want to convince other distant and future
-people about our analyses.
+Assume we want others to believe us when we say something. More narrowly,
+assume we want other people to believe us when we say something about data:
+'data' here can be words, numbers, musical notes, images, ideas, etc \ldots The
+point is that we are making some claims about patterns in some collection of
+stuff. For example, when I was invited into the homes and offices of ordinary
+people in Chile in 1991, the "stuff" was recordings of conversations that we
+had about life during the first year of democracy and I was comparing the
+responses of people who had experienced the Pinochet dictatorship differently.
+Now, it might be easy to convince others that 'this collection of stuff is
+different from that collection of stuff' if those people were looking over our
+shoulders the whole time that we made decisions about collecting the stuff and
+broke it up into understandable parts and reorganized and summarized it.
+Unfortunately, we can't assume that people are willing to shadow a researcher
+throughout her career. Rather, we do our work alone or in small groups and want
+to convince other distant and future people about our analyses.
 
@@
 In the end, following these practices and those recommended by
-@fredrickson2011tpm and @healy2011tpm in this issue
+@fredrickson2011tpm and @healy2011tpm among others working on these topics
 allows your computerized analyses of your collections of stuff to be
-credible.  Finally, if the UN quibbles with your analyses, your future
+credible.  Finally, if the someone quibbles with your analyses, your future
 self can shoot them the archive required to reproduce your
@@
   can include a copy of R and all of the R packages you used in your
-  final analyses in 2018 in the archive itself. You can even throw in
-  a copy of Ubuntu 10 and an open source virtual machine running the
+  final analyses in the archive itself. You can even throw in
+  a copy of whatever version of linux you used and an open source virtual machine running the
   whole environment.] You can say, 'Here is everything you need to

### 2016-06-09 workflow.Rmd
Update workflow.Rmd

@@
 - name: Maarten Voors
-  affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics
+  affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics Group
 bibliography: references.bib
@@
 collaborators are less likely to call you at midnight asking for help and you will remove some regret from your future self.
-For example, if you are studying inequalit and protest, you might try naming a file something like  `inequality-and-protest-figures.R` instead of  `temp9`or
+For example, if you are studying inequality and protest, you might try naming a file something like  `inequality-and-protest-figures.R` instead of  `temp9`or
 `supercalifragilisticexpialidocious`.  By the way, the extension

### 2016-06-24 workflow.Rmd
Update workflow.Rmd
fixed an error, will continue soon!
made some changes, not done yet, but loving it!

@@
 
-How long does it take from planning a study to publication and to the first reproduction of the study after publication?
+How long does it take from planning a study to publication?
+And subsequently, how much time passes to the first reproduction of a study after publication?
 Is three years too short?
@@
 We suspect that few of our colleagues in the social and behavioral sciences conceive of, field, write, and publish a study in faster than about three years.
-We also suspect that, if some other scholar decides to re-examine the analyses of a published study, that it will occur after publication, and that this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, years past the initial publication of the article.^[We note that sometimes this reproduction occurs within the context of classes and learning about statistics and data analysis.
-Othertimes, this process is organized more broadly (cite IIIE program, cite Open Science Foundation on psychology).]
+We also suspect that, if some other scholar decides to re-examine the analyses of a published study, it will occur after publication, and that this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, years past the initial publication of the article.^[We note that sometimes this reproduction occurs within the context of classes and learning about statistics and data analysis.
+Other times, this process is organized more broadly (cite IIIE program, cite Open Science Foundation on psychology, cite the Camerer 2016 Science paper?, cite the cool "Crowdsourcing Analysis" project by Silberzahn and Uhlmann Nature 2015?).]
 
@@
 How can we improve our relationship with our future self?
-This essay is a heavily revised version of @bowers2011tpm and provides some suggestions for practices that will make such reproduction occur much more easily and quickly in the event that famous papers require special scrutiny.
-Specifically, this piece aims to amplify some of what we already ought to know,  and to add to some of those ideas
-given current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists.
+This essay is a heavily revised and updated version of @bowers2011tpm and provides some suggestions for practices that will make such reproduction occur much more easily and quickly in the event that famous papers require special scrutiny.
+Specifically, this piece aims to amplify some of what we already ought to know, and to add to some of those ideas given current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists.
 Now, the efforts to encourage transparency and ease of learning from the data and analyses of others has become institutionalized with the DA-RT inititiave <http://www.dartstatement.org/> @lupia2014openness.
@@
 All results (numbers, comparisons, tables, figures) should arise from
-code, not from a series of mouse clicks or copying and pasting. If I wanted
-to re-create the figure from 2011 but include a new variable or
-specification, I should be able to do so with just a few edits to the
-code rather than knowledge of how I (or you) used a pointing device in
-your graphical user interface some years ago.
+code, not from a series of mouse clicks or copying and pasting. If you wanted
+to re-create the figure from say 2011 but include a new variable or
+specification, you should be able to do so with just a few edits to the
+code rather than knowledge of how you used a pointing device in
+your graphical user interface all those years ago.
 
-Using R [@R:2011], for example, I might specify that the file `fig1.pdf` was
-produced by the following commands in a file called `fig1.R`.^[The command
+For example, using R [@R:2011], a open-source statistical software program, I might specify that a file, called `fig1.pdf` was
+produced by the following set of commands in a file called `makefig1.R`.^[The command
 `please-plot`and some other R functions used in this essay come from the
@@
 not recommend setting negative values for the politeness option should you use this package.]
+Lets look at some annotated R code you could produce:
 
@@
 
-Now, in the future if I wonder how 'that plot on page 10' was
-created, I will know: (1) 'that plot' is from a file called
+Now, in the future if you wonder how 'that plot on page 10' was
+created, you will know: (1) 'that plot' is from a file called
 `fig1.pdf`and (2) `fig1.pdf`was created in
-`fig1.R`. In a future where R still exists, changing the figure
-will require quick edits of commands already written.  In a future
-where R does not exist, I will at least be able to read the plain text
-R commands and use them to write code in my new favorite statistical
+`makefig1.R`. Any time in a future (provided R still exists), changing the figure
+will require some quick edits of commands already written.  In a future
+where R does not exist, you (or someone else) will at least be able to read the plain text
+R commands and use them to write code in a new favorite statistical
 computing language: R scripts are written in
@@
 write computer programs.^[Since R is open
-  source, I will also be able to download an old version of R, download an
+  source, you will also be able to download an old version of R, download an
   old-fashioned open-source operating system (like Ubuntu 10), and
@@
 
-Moreover, realize that *file names send messages* to your future
-self. This means that if you name your files with evocative and descriptive names, your
+Moreover, coding saves time. Often, a lot of time, especially if projects increase (in the number of files, partners, analysis). Manually importing
+a data file into R, may be effortless. Importing 100 files is another issue and the costs of each manual action add up quickly.
+
+Also, realize that *file names send messages* to your future self. This means that if you name your files with evocative and descriptive names, your
 collaborators are less likely to call you at midnight asking for help and you will remove some regret from your future self.
 For example, if you are studying inequality and protest, you might try naming a file something like  `inequality-and-protest-figures.R` instead of  `temp9`or
-`supercalifragilisticexpialidocious`.  By the way, the extension
-`.R` tells us and the operating system that the file contains R
-commands. This part of the filename enables us to quickly search our
+`supercalifragilisticexpialidocious`.  More on this below. By the way, the extension
+`.R` tells us and the operating system that the file contains R commands. This part of the filename enables us to quickly search our
 antique hard drives for files containing R scripts.
 
-**Step 1** If we know the provenance of results,
-future  or current collaborators can quickly and easily reproduce
-and thus change and improve upon the work.
+Finally, coding helps us to avoid making mistakes.
+For example, say in our example above we are interested in how many people protest.
+We may use a data file containing all protests for several years.
+We can then create a table with the number of protesters by copying the results manually to a document you use to write your paper.
+In copying we can make mistakes.
+So a better approach is to ask R to automatically create the table (in the format you like, with horizontal lines, 3 decimals, a title, etc) and output this in whatever file type you need (.pdf, .rtf, etc).
+Now when we obtain new data (for additional years, or updated figures for existing data) a new table can be created quickly, so we make less mistakes and save time!
+
+NOTE we could include some code here to provide an example of automation of results?
+
+**Step 1** Code everything that can be coded.
+If we know the provenance of results, future or current collaborators make less mistakes and can quickly and easily reproduce (and thus change and improve) upon the work.
 
@@
 Data analysis involves a long series of decisions. Each decision
-requires justification.  Some decisions will be too small and
+requires a justification.  Some decisions will be too small and
 technical for inclusion in the published article itself. These need to

### 2016-06-24 workflow.Rmd
more changes, time for dinner!

@@
 Moreover, coding saves time. Often, a lot of time, especially if projects increase (in the number of files, partners, analysis). Manually importing
-a data file into R, may be effortless. Importing 100 files is another issue and the costs of each manual action add up quickly.
+a data file into R, may be effortless. Importing 100 files is another issue, and the costs of each manual action add up quickly.
 
@@
 
-NOTE we could include some code here to provide an example of automation of results?
+[NOTE we could include some code here to provide an example of automation of results?]
 
@@
   reason for a decision into the mind of the analyst. I myself have
-  not had much luck with this function.]
+  not had much luck with this function.] [I FEEL THIS COMMENT IS FUNNY BUT CONFUSING FOR OUR LATIN AMERICAN READERS PROBABLY]
 
@@
 collaborators (including your future self) and other consumers of your
-work. In the above code chunk, I used comments to explain the lines to
-readers unfamiliar with R and to remember that I had tried a different
+work. In the above code chunk, we used comments to explain the lines to
+readers unfamiliar with R and to remember that we had tried a different
 specification but decided not to use it because adding the squared
@@
 exchanged by the group must be available and executable.^[As of the 2016 version of this paper, this idea is now widespread and made much easier than before via online services for code sharing and collaboration such as [GitHub](http://github.com) and [BitBucket](http://bitbucket.org).]
+[I FEEL WE CAN MOVE THIS TO THE MAIN TEXT AND (BRIEFLY) INTRODUCE GITHUB]
 
@@
 `fig1.pdf` file produced from a clearly commented plain text file
-called something like `fig1.R`.  An even easier solution would be to
+called something like `makefig1.R`.  An even easier solution would be to
 directly include a chunk of code to produce the figure inside of the
@@
 Literate programming need not go against the principle of modular data
-analysis [@nagler1995coding]. In my own work I routinely have
+analysis [@nagler1995coding]. Jake routinely uses
 several different Sweave files that fulfill different functions, some
-of them create \LaTeX code that I can `\input` into my
+of them create \LaTeX code that he can `\input` into his
 `main.tex` file, others setup the data, run simulations, or allow
-me to record my journeys down blind alleys. Of course, when we have
+him to record his journeys down blind alleys. Of course, when we have
 flying cars running on autopilot, perhaps something other than
@@
 scholars and policy makers. If we focus on explaining how we got our
-computers to do data analysis to human beings, we will do a better job
+computers to do data analysis, we will do a better job
 with the data analysis itself: we will learn as we focus on teaching,
@@
 some place easy to find (like in the same directory as the file or
-perhaps in `"Data/thedata.rda"`). Of course, it never hurts to
-drop in a comment pointing to the data file.
+perhaps in `"Data/thedata.rda"`). Of course, its even better to
+include a comment pointing to the data file.
 
@@
 
-The fact that I need to articulate this idea at all arises because of
+The fact that we need to articulate this idea at all arises because of
 graphical user interfaces: it is all too easy to use the mouse to load
@@
 
-Anyone writing custom code should worry about getting it right. The more code one writes, the more time one has to appreciate problems arising from bugs, errors, and typos in data analysis
-and code.
+Anyone writing custom code should worry about getting it right.
+The more code one writes, the more time one has to appreciate problems arising from bugs, errors, and typos in data analysis and code.
+Mistakes will always be made we should just recognize we need more standardization of practice to minimise the mistakes and help others find them.
+One thing we can do is to include testing in our coding.
 
-Now, if one has a moment to think in between teaching that new class,
+If one has a moment to think in between teaching that new class,
 reading books for an awards committee, evaluating application files
@@
 collaborators, here is an example of this idea in a very simple form:
-Say I want to write a function to multiply a number by 2. If my
-function works, when I give it the number 4, I should see it return
-the number 8 and when I give it -4, I should get -8.
+Say you want to write a function to multiply a number by 2. If the
+function works, when you give it the number 4, you should see it return
+the number 8 and when you give it -4, you should get -8.
 
@@
 
-Ack! I mistyped '+' for '*'. Good thing I wrote the
+Ack! we mistyped '+' for '*'. Good thing I wrote the
 test!^[A more common example of this kind of testing occur
@@
 
-Lots of people are thinking about 'reproducible research'and
+[IS THIS SECTION MORE MORE ON CREATING AN INTEGRATED REPRODUCABLE WORKFLOW RATHER THAN COPYING AND IMPROVING?]
+Lots of people are thinking about 'reproducible research', creating a 'reproducable workflow' and
 'literate programming' these days. Google those terms. Of course, the
@@
 
-**Step 6** We all learn by doing. When we share
+[ADD OPEN SCIENCE FRAMEWORK, COLLABORATING ON DROPBOX AND NEW INITIATIVES LIKE Dynamic Documents (R Markdown/KnitR in R Studio,
+Markdoc in Stata), Jupyter Notebook, Docker for Cross-Platform Reproducibility)]
+
+**Step 6** We all learn by doing. When we create a reproducible workflow and share
 reproduction materials we improve both cumulation of knowledge and our
@@
 ability to believe each other and for future scholars to believe us, too.
+[I FEEL LIKE THE COPY COMMENT IS PERHAPS ANOTHER POINT TO MAKE (OR LEAVE)]
 
@@
 article does not spawn debate? Why then would you spend time to
-communicate with your future self and others? My own answer to this
-question is that I want my work to be credible and useful to myself
-and other scholars even if each article does not immediately change
-the world.  What I report in my data analyses should have two main
+communicate with your future self and others? Our own answer to this
+question is that we want our work to be credible and useful to ourself
+and other scholars regardless.  What we report in our data analyses should have two main
 characteristics: (1) the findings of the work should not be a matter
 of opinion; and (2) other people should be able to reproduce the findings. That is,
-the work represents a shared
-experience --- and an experience shared without respect to the
+the work represents a shared experience --- and an experience shared without respect to the
 identities of others (although requiring some common technical
@@
 point is that we are making some claims about patterns in some collection of
-stuff. For example, when I was invited into the homes and offices of ordinary
+stuff. For example, when Jake was invited into the homes and offices of ordinary
 people in Chile in 1991, the "stuff" was recordings of conversations that we
-had about life during the first year of democracy and I was comparing the
+had about life during the first year of democracy and Jake was comparing the
 responses of people who had experienced the Pinochet dictatorship differently.
-Now, it might be easy to convince others that 'this collection of stuff is
-different from that collection of stuff' if those people were looking over our
+Now, it might be easy to convince others that 'this collection of stuff' is
+different from 'that collection of stuff' if those people were looking over our
 shoulders the whole time that we made decisions about collecting the stuff and
@@
 throughout her career. Rather, we do our work alone or in small groups and want
-to convince other distant and future people about our analyses.
+to convince other distant and future people about our analyses and findings.
 

### 2016-06-24 workflow.Rmd
some more edits
ok ready for a discussion with you

@@
   affiliation: <jwbowers@illinois.edu>,  University of Illinois at Urbana-Champaign, Departments of Political Science and Statistics; Fellow, White House Social and Behavioral Sciences Team.
+Nothing in this paper reflect the official position of the US Government.
 Many thanks to Mark Fredrickson, Brian Gaines, Kieran Healy, Kevin Quinn and Cara Wong for direct help on this document and to Mika LaVaque-Manty and Ben Hansen for many useful discussions on this topic.
@@
 - name: Maarten Voors
-  affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics Group
+  affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics Group.
+Special thanks to the BITSS team and the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during 2016.
 bibliography: references.bib
@@
 In fact, learning requires effort.
-When we do not practice, we tend to be overconfident in our recall abilities [@koriat2005illusions] and false memories can be implanted.^[See the following site for a nice overview of what we know about memory including the fact that learning requires practice: <http://www.spring.org.uk/2012/10/how-memory-works-10-things-most-people-get-wrong.php>.
+And when we do not practice, we tend to be overconfident in our recall abilities [@koriat2005illusions] and false memories can be implanted.^[See the following site for a nice overview of what we know about memory including the fact that learning requires practice: <http://www.spring.org.uk/2012/10/how-memory-works-10-things-most-people-get-wrong.php>.
 On false memory see Wikipedia and linked studies <https://en.wikipedia.org/wiki/False_memory>.] If we cannot count on memory, then how can we do science?
@@
 How long does it take from planning a study to publication?
-And subsequently, how much time passes to the first reproduction of a study after publication?
+And subsequently, how much time passes to the first reproduction of a study?
 Is three years too short?
 Is ten years too long?
-We suspect that few of our colleagues in the social and behavioral sciences conceive of, field, write, and publish a study in faster than about three years.
-We also suspect that, if some other scholar decides to re-examine the analyses of a published study, it will occur after publication, and that this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, years past the initial publication of the article.^[We note that sometimes this reproduction occurs within the context of classes and learning about statistics and data analysis.
+We suspect that few of our colleagues in the social and behavioral sciences conceive of, field, write, and publish a data driven study faster than about three years.
+We also suspect that, if some other scholar decides to re-examine the analyses of a published study, it will occur after publication.
+Moreover, this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, years past the initial publication of the article.^[We note that sometimes this reproduction occurs within the context of classes and learning about statistics and data analysis.
 Other times, this process is organized more broadly (cite IIIE program, cite Open Science Foundation on psychology, cite the Camerer 2016 Science paper?, cite the cool "Crowdsourcing Analysis" project by Silberzahn and Uhlmann Nature 2015?).]
@@
 How can we improve our relationship with our future self?
-This essay is a heavily revised and updated version of @bowers2011tpm and provides some suggestions for practices that will make such reproduction occur much more easily and quickly in the event that famous papers require special scrutiny.
+This essay is a heavily revised and updated version of @bowers2011tpm and provides some suggestions for practices that will make such reproduction occur much more easily and quickly.
 Specifically, this piece aims to amplify some of what we already ought to know, and to add to some of those ideas given current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists.
@@
 reproducing or collaborating. The source code of any document
-exchanged by the group must be available and executable.^[As of the 2016 version of this paper, this idea is now widespread and made much easier than before via online services for code sharing and collaboration such as [GitHub](http://github.com) and [BitBucket](http://bitbucket.org).]
-[I FEEL WE CAN MOVE THIS TO THE MAIN TEXT AND (BRIEFLY) INTRODUCE GITHUB]
+exchanged by the group must be available and executable.^[As of the 2016 version of this paper, this idea is now widespread and made much easier than before via online services for code sharing and collaboration such as [GitHub](http://github.com), [Open Science Framework](https://osf.io/) and [BitBucket](http://bitbucket.org), see below.]
 
@@
 
-For example, in [@sec:data-analys-comp], I suggested that we know
-where 'that plot on page 10 'comes from by making sure we had a
+For example, in [@sec:data-analys-comp], we suggested that we know
+where 'that plot on page 10' comes from by making sure we had a
 `fig1.pdf` file produced from a clearly commented plain text file
@@
 track of the figure number, nor do extra work when I reorganize the document in
-response to reviewer suggestions. Nor do I need a separate `fig1.R`file or
+response to reviewer suggestions. Nor do I need a separate `makefig1.R`file or
 `fig1.pdf` file.  Tables and other numerical results are also possible to
@@
 
-Group work requires version control.^[@fredrickson2011tpm and
-@healy2011tpm in this issue also explain what version control is and why we
-might want to use it.] Many people are familiar with the track changes
-feature in modern WYSIWYG word processors or the fact that Dropbox allows one
+Group work requires version control. Many people are familiar with the track changes
+feature in modern WYSIWYG (what you see is what you get) word processors or the fact that Dropbox allows one
 to recover previous versions of files. These are both kinds of version control.
@@
 files; (2) members of the team feel free to experiment and then to dump parts
-of the experiment in favor of previous work while merging the successful parts
-of the experiment into the main body of the paper; (3) the team can produce
+of the experimentation in favor of previous work while merging the successful parts
+ into the main body of the paper; (3) the team can produce
 have 'releases' of the same document (one to MPSA, one to APSR, one to your
@@
 document; (4) people can work on the same files at the same time without
-conflicting with one another (and can reconcile their changes without too much
-confusion and clobbering).  Clobbering is what happens when your future self or
+conflicting with one another, and can reconcile their changes without too much
+confusion and clobbering. Clobbering is what happens when your future self or
 your current collaborator saves an old version of a file over a new version,
@@
 work on the same files at the same time, then you should work on establishing
-your own shared version control system. Free options include launchpad, github,
+your own shared version control system. Free options include launchpad, github, Open Science Framework,
 sourceforge for open source projects (i.e. papers you are writing which you are
-happy to share with others as you write). As of this writing, Google Docs does
-an excellent job with shared editing of files as long as your are online. Each
-of those services include paid versions too. One may also use Dropbox as a kind
-of server for version control: for example, one may copy files from the Dropbox
+happy to share with others as you write, see more below).
+Each has a slightly difference language (some say folders other repositories, some say download othere clone, etc) and it takes a bit of time adjusting.
+But as we'll argue: it is well worth the time investment as it will save (lots!) of time later on.
+Currently Google Docs does
+an excellent job with shared editing of files. The downside is: you have to be online. For people with irregular or slow internet connections, this may not be an option. MORE ON GOOGLE?
+One may also use Dropbox as a kind of server for version control (though the downside is if your collaborations or project grows large (in terms of MBs) you need to switch to a paid service).
+Maarten has been using this for most of his collaborations.
+You can copy files from the Dropbox
 directory into a local working directory so as to avoid clobbering and then
 work on merging changes by hand before copying back to the Dropbox directory
-and replacing existing files. MORE PERHAPS ON GOOGLE? LESS ON DROPBOX?
+and replacing existing files. You will however need to agree on an iron file and directory naming rule with your collaborators.
 
-We use git with our own research group, and I use it for all of
-my own projects. MORE ON GITHUB
+Jake uses github within his research group, and used it for all of
+his own projects. MORE ON GITHUB
+You may even obtain a two year unlimited repositories with an academic affiliation.
 
-Of course, one may take advantage of many of the benefits of formal
-version control systems with some disciplined systems for file and
-directory organization.  An excellent, simple, and robust version
+ An excellent, simple, and robust version
 control system is to rename your files with the date and time of
-saving them: thedoc.tex becomes thedoc25-12-2011-23:50.tex.  Be sure
-to include year in the file names --- remember, the life of an idea is
+saving them: yyyymmdd_project.docx (for changes within the same day, just before a deadline for example, you may even add a time, so it becomes 20160623_4PM_inequalitypaper.docx).
+Be sure to include year in the file names --- remember, the life of an idea is
 measured in years. If you are wise enough to have saved your documents
@@
   allows one to compare differences in pdf files. OpenOffice supports
-  a 'Compare Documents 'option. And Google Docs will report on the
+  a 'Compare Documents 'option. Word now does the same. And Google Docs will report on the
   version history of a document.] When you reach certain milestones
@@
 version eventually sent to the APSR six years after you presented it
-at APSA. The formal version control systems I mentioned above all
+at APSA. The formal version control systems mentioned above all
 allow this kind of thing and are much more elegant and capable, but
 you can do it by hand too as long as you don't mind taking up a lot of
-disk space and having many 'thedoc... ' files around. If you do
+disk space and having many 'thedoc...
+' files around.
+Indeed, the number of files can add up quickly (especially around submission time!) and you may want to create an "0_Archive" folder to store older versions.
+If you do
 version control by hand, spend a little extra time to ensure that you
@@
 
-[IS THIS SECTION MORE MORE ON CREATING AN INTEGRATED REPRODUCABLE WORKFLOW RATHER THAN COPYING AND IMPROVING?]
-Lots of people are thinking about 'reproducible research', creating a 'reproducable workflow' and
+[IS THIS SECTION MORE MORE ON CREATING AN INTEGRATED REPRODUCIBLE WORKFLOW RATHER THAN COPYING AND IMPROVING?]
+Lots of people are thinking about 'reproducible research', creating a 'reproducible workflow' and
 'literate programming' these days. Google those terms. Of course, the
@@
 
-I have experimented with three systems so far: (1) for one paper we simply
-included a Sweave document and data files into a compressed archive
-[@bowers2005dataverse]; (2) for another more computing intensive paper we
+We have experimented with several systems so far: (1) we wrote this paper in R markdown (Jake) and Atom (Maarten) and organized our files and collaboration on Github.
+The source code can be accessed, downloaded and modified from <https://github.com/jwbowers/workflow>.
+The program also allowed us to send notes to each other.
+The nice thing about github is that is enables you to make "releases" of a project which enable you to smooth the reproduction of all of the
+products of your research (simulations, data analyses, tables, figures, etc..).
+Similar features are offered by the Open Science Framework (https://osf.io/), Maarten recently participated in a workshop on research transparency organized by the Berkeley Institute for Transparency in the Social Sciences (www.bitss.org), all course material was accessible through OSF (https://osf.io/qh2nr/).
+Both platforms, OSF and Github nicely integrate and a key benefit is that they have been created to remain for a long time (unlike perhaps your university drive); (2) for one paper Jake simply included a Sweave document and data files into a compressed archive [@bowers2005dataverse] [SAY ABIT MORE ABOUT WHAT SWEAVE IS IN PRINCIPLE]; (3) for another more computing intensive paper Jake
 assembled a set of files that enabled reproduction of our results using the
-`make` system [@bowers2008dataverse]; (3) I have tried the "compendium"
+`make` system [@bowers2008dataverse] [EXPLAIN A BIT MORE OR INCLUDE REF TO WEBSITE?]; (4) Jake also tried the "compendium"
 approach [@gentleman2005reproducible,@gentleman2007statistical] which embeds an
-academic paper within the R package system [@bowers2011dataverse]; and (4)
-recently I have started organizing my work on github which enables me to make
-"releases" of the project which should enable smooth reproduction of all of the
-products of the research (simulations, data analyses, tables, figures, etc..).
-The benefit of the compendium approach is that one is not required to have
+academic paper within the R package system [@bowers2011dataverse]; The benefit of the compendium approach is that one is not required to have
 access to a command line for `make`: the compendium is downloadable from within
 R using `install.packages()` and is viewable using the `vignette()` function in
-any operating system than runs R.^[Notice that my reproduction archives and/or
+any operating system than runs R.^[Notice that Jake's reproduction archives and/or
 instructions for using them are hosted on the
@@
 just as one installs and uses statistical software packages is very attractive,
-and I anticipate that it will become ever easier to turn papers into R packages
+and we anticipate that it will become ever easier to turn papers into R packages
 as creative and energetic folks turn their attention to the question of
 reproducible research.^[See for example <http://r-pkgs.had.co.nz/>]
+Finally, (5) Maarten has been using Dropbox for most of his collaborative projects (https://www.dropbox.com/). ALREADY ABOVE?
 
-[ADD OPEN SCIENCE FRAMEWORK, COLLABORATING ON DROPBOX AND NEW INITIATIVES LIKE Dynamic Documents (R Markdown/KnitR in R Studio,
-Markdoc in Stata), Jupyter Notebook, Docker for Cross-Platform Reproducibility)]
+There are some new noteworthy developments also.
+For example, and similar to R Markdown/KnitR in R Studio, there is Markdoc for STATA users (https://github.com/haghish/MarkDoc).
+People at UC Berkeley have been developing a is a web app, Jupyter Notebook (http://jupyter.org/), that enables you to make and share documents that include text, code, equations and visualizations in one.
+For researchers working across difference platforms, there is Docker (https://www.docker.com/).
 
@@
 methods for doing social science
-[@freese2007replication,king1995replication}.  As we copy and
+[@freese2007replication,king1995replication}.
+
+ As we copy and
 improve upon each other's modes of doing work we enhance our collective

### 2016-06-24 workflow.Rmd
Added a comment
small change

@@
 ...
+
+<!-- An example comment -->
 
@@
 Similar features are offered by the Open Science Framework (https://osf.io/), Maarten recently participated in a workshop on research transparency organized by the Berkeley Institute for Transparency in the Social Sciences (www.bitss.org), all course material was accessible through OSF (https://osf.io/qh2nr/).
-Both platforms, OSF and Github nicely integrate and a key benefit is that they have been created to remain for a long time (unlike perhaps your university drive); (2) for one paper Jake simply included a Sweave document and data files into a compressed archive [@bowers2005dataverse] [SAY ABIT MORE ABOUT WHAT SWEAVE IS IN PRINCIPLE]; (3) for another more computing intensive paper Jake
+Both platforms, OSF and Github nicely integrate and a key benefit is that they have been created to remain for a long time (unlike perhaps your university drive) JAKES PROJECT; (2) for one paper Jake simply included a Sweave document and data files into a compressed archive [@bowers2005dataverse] [SAY ABIT MORE ABOUT WHAT SWEAVE IS IN PRINCIPLE]; (3) for another more computing intensive paper Jake
 assembled a set of files that enabled reproduction of our results using the
@@
 reproducible research.^[See for example <http://r-pkgs.had.co.nz/>]
-Finally, (5) Maarten has been using Dropbox for most of his collaborative projects (https://www.dropbox.com/). ALREADY ABOVE?
+Finally, (5) Maarten has been using Dropbox for most of his collaborative projects (https://www.dropbox.com/). ALREADY ABOVE? NOTE BACK
 
@@
 For example, and similar to R Markdown/KnitR in R Studio, there is Markdoc for STATA users (https://github.com/haghish/MarkDoc).
-People at UC Berkeley have been developing a is a web app, Jupyter Notebook (http://jupyter.org/), that enables you to make and share documents that include text, code, equations and visualizations in one.
+People at UC Berkeley have been developing a is a web app, Jupyter Notebook (http://jupyter.org/), that enables you to make and share documents that include text, code, equations and visualizations in one INCLUDE REF TO GITHUB - USING PYTHON.
 For researchers working across difference platforms, there is Docker (https://www.docker.com/).

### 2016-06-24 workflow.Rmd
some more

@@
   affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics Group.
-Special thanks to the BITSS team and the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during 2016.
+Special thanks, the EGAP Learning Days 2016 participants in Santiago Chile,  to the BITSS team and the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during spring 2016.
+  <!-- Separate the thanks sections from affiliation and integrate? Wasnt sure how to do this here-->
 bibliography: references.bib
@@
 Moreover, this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, years past the initial publication of the article.^[We note that sometimes this reproduction occurs within the context of classes and learning about statistics and data analysis.
-Other times, this process is organized more broadly (cite IIIE program, cite Open Science Foundation on psychology, cite the Camerer 2016 Science paper?, cite the cool "Crowdsourcing Analysis" project by Silberzahn and Uhlmann Nature 2015?).]
+Other times, this process is organized more broadly (cite IIIE program, researchers have started to implement psychological experiments using the same protocols across many locations, see Open Science Foundation 2014, 2015, within experimental economics a team did the same, see Camerer et al 2016 Science paper.
+Another example is to share analysis across many researchers.
+in a their 'crowdsourcing analysis" Silberzahn and Uhlmann (Nature 2015) asked 29 research teams to run the analysis for a project that looked at the impact of skin color on the probability of getting a foul card in a football match.<!-- Can you do the correct citing here -->]
 
@@
 These ideas are spreading beyond political science as well @freese2007replication and @asendorpf2013recommendations.] This essay is an attempt at providing practical advice about how to *do work* such that one complies with such recommendations as a matter of course and, in so doing, can focus personal regret on other bad past decisions but not on decisions about data analysis and the production of scholarly papers.
+
+<!-- JAKE: we dont say anything about pre registration, should we? -->
 
@@
 Finally, coding helps us to avoid making mistakes.
-For example, say in our example above we are interested in how many people protest.
+For example, in our example above we may be interested in how many people protest.
 We may use a data file containing all protests for several years.
@@
 
-[NOTE we could include some code here to provide an example of automation of results?]
+<!-- Can you include some code here to provide an example of automation of results? -->
 
@@
 
- An excellent, simple, and robust version
-control system is to rename your files with the date and time of
+An excellent, simple, and robust version control system is to rename your files with the date and time of
 saving them: yyyymmdd_project.docx (for changes within the same day, just before a deadline for example, you may even add a time, so it becomes 20160623_4PM_inequalitypaper.docx).
+This communicates very clear what version you are working.
+Remember the days when you received a file called inequalityMV4_APSRsubmit_reallyfinal2.tex?
+That should quickly become something of the past!
+We find ourselves preaching the gospel to our collaborators frequently and will keep renaming files until a collaborator has been successfully pacified.
 Be sure to include year in the file names --- remember, the life of an idea is
@@
 The more code one writes, the more time one has to appreciate problems arising from bugs, errors, and typos in data analysis and code.
-Mistakes will always be made we should just recognize we need more standardization of practice to minimise the mistakes and help others find them.
+Mistakes will always be made we should just recognize we need more standardization of practice to minimize the mistakes and help others find them.
 One thing we can do is to include testing in our coding.
@@
 [test driven development](http://en.wikipedia.org/wiki/Test-driven_development) and the idea that one ought to create tests of
-[small parts of one's code](http://en.wikipedia.org/wiki/Unit_testing) arose to address such concerns.^[There are now R packages to help R package writers do this, see for example <https://github.com/hadley/testthat> and the article on it <https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf>.] For the social scientist
-collaborating with her future self and/or a small group of
+[small parts of one's code](http://en.wikipedia.org/wiki/Unit_testing) arose to address such concerns.^[There are now R packages to help R package writers do this, see for example <https://github.com/hadley/testthat> and the article on it <https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf>.] For the social scientist collaborating with her future self and/or a small group of
 collaborators, here is an example of this idea in a very simple form:
@@
 
-Ack! we mistyped '+' for '*'. Good thing I wrote the
+Ack! we mistyped '+' for '*'. Good thing we wrote the
 test!^[A more common example of this kind of testing occur
@@
 
-# Copy and improve on others' examples.
+# Creating a reproducible workflow
 
-[IS THIS SECTION MORE MORE ON CREATING AN INTEGRATED REPRODUCIBLE WORKFLOW RATHER THAN COPYING AND IMPROVING?]
 Lots of people are thinking about 'reproducible research', creating a 'reproducible workflow' and
@@
 Similar features are offered by the Open Science Framework (https://osf.io/), Maarten recently participated in a workshop on research transparency organized by the Berkeley Institute for Transparency in the Social Sciences (www.bitss.org), all course material was accessible through OSF (https://osf.io/qh2nr/).
-Both platforms, OSF and Github nicely integrate and a key benefit is that they have been created to remain for a long time (unlike perhaps your university drive) JAKES PROJECT; (2) for one paper Jake simply included a Sweave document and data files into a compressed archive [@bowers2005dataverse] [SAY ABIT MORE ABOUT WHAT SWEAVE IS IN PRINCIPLE]; (3) for another more computing intensive paper Jake
-assembled a set of files that enabled reproduction of our results using the
-`make` system [@bowers2008dataverse] [EXPLAIN A BIT MORE OR INCLUDE REF TO WEBSITE?]; (4) Jake also tried the "compendium"
+Both platforms, OSF and Github nicely integrate and a key benefit is that they have been created to remain for a long time (unlike perhaps your university drive) <!-- ADD REFERENCE TO YOUR OSF PROJECT -->; (2) for one paper Jake simply included a Sweave document and data files into a compressed archive [@bowers2005dataverse] <!-- PERHAPS SAY ABIT MORE ABOUT WHAT SWEAVE IS IN PRINCIPLE-->; (3) for another more computing intensive paper Jake assembled a set of files that enabled reproduction of results using the
+`make` system [@bowers2008dataverse] <!-- AGAIN A FEW WORDS ON WHAT PERHAPS LINK TO A WEBSITE? -->; (4) Jake also tried the "compendium"
 approach [@gentleman2005reproducible,@gentleman2007statistical] which embeds an
@@
 reproducible research.^[See for example <http://r-pkgs.had.co.nz/>]
-Finally, (5) Maarten has been using Dropbox for most of his collaborative projects (https://www.dropbox.com/). ALREADY ABOVE? NOTE BACK
+Finally, (5) Maarten has been using Dropbox for most of his collaborative projects (https://www.dropbox.com/).
+See comments above.
+As a tool for making files publicly available Dropbox is less useful.
+You can of course make replication files available through creating an public folder, but you will still need to refer to this folder on a particular project website (an example is Maarten's project on conflict and football https://www.dropbox.com/sh/wqosq99019mf04d/AAASTgJLkn6lmZJI_1veUK7ca?dl=0)
 
-There are some new noteworthy developments also.
+There are some noteworthy new developments also.
 For example, and similar to R Markdown/KnitR in R Studio, there is Markdoc for STATA users (https://github.com/haghish/MarkDoc).
 People at UC Berkeley have been developing a is a web app, Jupyter Notebook (http://jupyter.org/), that enables you to make and share documents that include text, code, equations and visualizations in one INCLUDE REF TO GITHUB - USING PYTHON.
-For researchers working across difference platforms, there is Docker (https://www.docker.com/).
+For researchers working across different platforms, there is Docker (https://www.docker.com/).
+There are also are great notes programs that help you communicate with collaborators.
+For example, Simplenote is very basic but allows you to share and change notes.
 
@@
 [@freese2007replication,king1995replication}.
-
- As we copy and
-improve upon each other's modes of doing work we enhance our collective
-ability to believe each other and for future scholars to believe us, too.
-[I FEEL LIKE THE COPY COMMENT IS PERHAPS ANOTHER POINT TO MAKE (OR LEAVE)]
 

### 2016-06-24 workflow.Rmd
Removing comment from top section becasue of problems making the pdf

@@
 Special thanks, the EGAP Learning Days 2016 participants in Santiago Chile,  to the BITSS team and the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during spring 2016.
-  <!-- Separate the thanks sections from affiliation and integrate? Wasnt sure how to do this here-->
 bibliography: references.bib
@@
 
+  <!-- Separate the thanks sections from affiliation and integrate? Wasnt sure how to do this here-->
 ```{r include=FALSE, cache=FALSE}

### 2016-06-24 workflow.Rmd
some more again, time for lunch!

@@
 
-Finally, coding helps us to avoid making mistakes.
+Coding helps us to avoid making mistakes.
 For example, in our example above we may be interested in how many people protest.
@@
 did something with data: comment your code and tightly link your code
-with your writing.^[One can also try the R command
-  `put-it-in-my-mind(reason,importance="high")` to firmly place a
-  reason for a decision into the mind of the analyst. I myself have
-  not had much luck with this function.] [I FEEL THIS COMMENT IS FUNNY BUT CONFUSING FOR OUR LATIN AMERICAN READERS PROBABLY]
+with your writing.
 
@@
 reproducing or collaborating. The source code of any document
-exchanged by the group must be available and executable.^[As of the 2016 version of this paper, this idea is now widespread and made much easier than before via online services for code sharing and collaboration such as [GitHub](http://github.com), [Open Science Framework](https://osf.io/) and [BitBucket](http://bitbucket.org), see below.]
+exchanged by the group must be available and executable.^[As of the 2016 version of this paper, this idea is now widespread and made much easier than before via online services for code sharing and collaboration such as [GitHub](http://github.com), [Open Science Framework](https://osf.io/) and [BitBucket](http://bitbucket.org), see more below.]
 
@@
 
-[MAARTEN IDEAS HERE? PLUS ANOTHER SECTION ABOUT ORGANIZATION?]
-
 Where should one store data files? An obvious solution is always to
@@
 things that formal version control systems do well (as discussed
-a bit more in [@sec:vers-contr-prev]).
+a bit more below in [@sec:vers-contr-prev]).
 
@@
 
+Best practice is to create a file system where you separate folders by function and separate input from output files.
+For example, below we show a folder structure for a paper where we look at inequality.
+The paper is written in .tex, data analysis in STATA (that Maarten prefers, so we see .do and .dta file extensions) in R (the program of choice for Jake, see the .R extensions):
+
+<!-- JAKE please check if I made no mistake, also there may be a prettier way to show a folder structure (a screen shot?),feel free to change-->
+00_archive
+01_paper
+  20160622_inequality.tex
+02_data
+  00_archive
+  01_rawdata
+  02_cleandata
+  20160622_analysis.dta
+  20160622_analysis.rda
+03_analysis
+  00_archive
+  01_temp
+  02_output
+    01_tables
+    02_figures
+  20160622_analysis.do
+  20160622_analysis.R
+  20160622_prep_data.do
+  20160622_prep_data.R
+
+There are a couple of things to note.
+See that files start with a number to give them an ordering.
+Note also we create an 00_archive folder, here you can store older files without having to delete them permanently and keeping the number of files in your main folders to a minimum.
+Also there is a clear separation between raw data (data that came straight from the field, downloaded data, someone else's replication files, etc) and clean data (data ready to use).
+Note that in the analysis folder we created two files, one for cleaning and merging data called 20160622_prep_data and another for running the analysis called 20160622_analysis.
+See also the output folders, with subfolders for graphs and tables.
+These are place holder for R or STATA to output the results from the analysis and keeps things organized.
+We discuss integrated programs (that combine code and text below).
+
 **Step 3** We should know where the data came
 from and what operations were performed on which set of data.
-
-In the good old days, when we executed our LISREL code in batch mode,
-we had no choice but to tell the machine clearly, in a few easy to
-understand and informative lines, what files (with filenames no longer
-than 8 characters) to use. See for example, this extremely clear code snippet:
-
-```
-DA NI=19 NO=199 MA=CM
-LA=basic.lab
-CM FI=basic.cov
-```
-
-The fact that we need to articulate this idea at all arises because of
-graphical user interfaces: it is all too easy to use the mouse to load
-a data file into memory and then to write a script to analyze this
-file without ever noting the actual name or location of the data file.
 
@@
 Jake uses github within his research group, and used it for all of
-his own projects. MORE ON GITHUB
+his own projects. <!-- JAKE do you want to add something else here? -->
 You may even obtain a two year unlimited repositories with an academic affiliation.
@@
 For researchers working across different platforms, there is Docker (https://www.docker.com/).
+
 There are also are great notes programs that help you communicate with collaborators.
-For example, Simplenote is very basic but allows you to share and change notes.
+Remember, while email is fantastic for fast communication, it is not a tool designed for project management.
+It may work for small projects with few co-authors but when the number of collaborators increase it becomes very difficult to keep track.
+There are some great online systems for task management, most of which have mobile apps also, examples are Flow (www.getflow.com), Asana (www.asana.com), Wrike (www.wrike.com) <!-- JAKE do you know others?
+--> .
+There also are great programs for notes.
+For example, Simplenote is very basic but allows you to share and edit notes.
+Others are Evernote (www.evernote.com) and for Windows users there is OneNote.
 

### 2016-06-24 workflow.Rmd
ok thats me for now, over to you!
Feel free to change back, undelete, modify, etc at will!

@@
 Special thanks, the EGAP Learning Days 2016 participants in Santiago Chile,  to the BITSS team and the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during spring 2016.
+Maarten gratefully acknowledges financial support from N.W.O.
+grant 451-14-001.
 bibliography: references.bib
@@
 And when we do not practice, we tend to be overconfident in our recall abilities [@koriat2005illusions] and false memories can be implanted.^[See the following site for a nice overview of what we know about memory including the fact that learning requires practice: <http://www.spring.org.uk/2012/10/how-memory-works-10-things-most-people-get-wrong.php>.
-On false memory see Wikipedia and linked studies <https://en.wikipedia.org/wiki/False_memory>.] If we cannot count on memory, then how can we do science?
+On false memory see Wikipedia and linked studies <https://en.wikipedia.org/wiki/False_memory>.] Moreover, most of us live busy lives.
+We write text and code behind our new standing desk at home while the kids run around, cry about an elder sister taking away a favorite new yellow school bus toy, drinks spill on the floor, the laundry needs going, the kettle boils...
+Our minds are full and we need to efficiently move from task to task.
+If we cannot count on memory, then how can we do science?
 
 How long does it take from planning a study to publication?
-And subsequently, how much time passes to the first reproduction of a study?
+How much time passes to the first reproduction of a study?
 Is three years too short?
@@
 This essay is a heavily revised and updated version of @bowers2011tpm and provides some suggestions for practices that will make such reproduction occur much more easily and quickly.
-Specifically, this piece aims to amplify some of what we already ought to know, and to add to some of those ideas given current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists.
+Specifically, this piece aims to amplify some of what we already ought to know and do, and highlight some current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists.
 Now, the efforts to encourage transparency and ease of learning from the data and analyses of others has become institutionalized with the DA-RT inititiave <http://www.dartstatement.org/> @lupia2014openness.
-These ideas are spreading beyond political science as well @freese2007replication and @asendorpf2013recommendations.] This essay is an attempt at providing practical advice about how to *do work* such that one complies with such recommendations as a matter of course and, in so doing, can focus personal regret on other bad past decisions but not on decisions about data analysis and the production of scholarly papers.
+These ideas are spreading beyond political science as well @freese2007replication and @asendorpf2013recommendations.] We aim to provide practical advice about how to *do work* such that one complies with such recommendations as a matter of course and, in so doing, can focus personal regret on other bad past decisions but not on decisions about data analysis and the production of scholarly papers.
 
@@
 All results (numbers, comparisons, tables, figures) should arise from
-code, not from a series of mouse clicks or copying and pasting. If you wanted
+code. Not from a series of mouse clicks or copying and pasting. If you wanted
 to re-create the figure from say 2011 but include a new variable or
@@
 
-For example, using R [@R:2011], a open-source statistical software program, I might specify that a file, called `fig1.pdf` was
-produced by the following set of commands in a file called `makefig1.R`.^[The command
-`please-plot`and some other R functions used in this essay come from the
-`MayIPleaseDoStatistics` package which emphasizes politeness in data analysis.
-Functions like `please-plot`can be blocked and more polite versions such as
-`may-I-please-have-a-plot`can be required using `options(politeness=99)`. We do
-not recommend setting negative values for the politeness option should you use this package.]
-Lets look at some annotated R code you could produce:
+For example, using R [@R:2011], a open-source statistical software program, you might specify that a file, called `fig1.pdf` was
+produced by the following set of commands in a file called `makefig1.R`. Lets look at some annotated R code you could produce:
 
@@
 `fig1.pdf`and (2) `fig1.pdf`was created in
-`makefig1.R`. Any time in a future (provided R still exists), changing the figure
+`makefig1.R`. Any time in a future, provided R still exists, changing the figure
 will require some quick edits of commands already written.  In a future
@@
 
-Moreover, coding saves time. Often, a lot of time, especially if projects increase (in the number of files, partners, analysis). Manually importing
-a data file into R, may be effortless. Importing 100 files is another issue, and the costs of each manual action add up quickly.
+Moreover, coding saves time. Often, a lot of time, especially if projects grow (in the number of files, partners, analysis). Manually importing
+a data file into R, may be effortless. Importing 100 files is another issue, and the time costs of each manual action add up quickly (and the probability for error increases).
 
@@
 collaborators are less likely to call you at midnight asking for help and you will remove some regret from your future self.
-For example, if you are studying inequality and protest, you might try naming a file something like  `inequality-and-protest-figures.R` instead of  `temp9`or
-`supercalifragilisticexpialidocious`.  More on this below. By the way, the extension
+For example, if you are studying inequality and protest, you might try naming a file something like  `inequality-and-protest-figures.R` instead of `temp9.R` or
+`supercalifragilisticexpialidocious.R`.  More on this below. By the way, the extension
 `.R` tells us and the operating system that the file contains R commands. This part of the filename enables us to quickly search our
@@
 In copying we can make mistakes.
-So a better approach is to ask R to automatically create the table (in the format you like, with horizontal lines, 3 decimals, a title, etc) and output this in whatever file type you need (.pdf, .rtf, etc).
+So a better approach is to ask R (or STATA, or another program of choice) to automatically create the table (in the format you like, with horizontal lines, 3 decimals, a title, etc) and output this in whatever file type you need (.pdf, .rtf, etc).
 Now when we obtain new data (for additional years, or updated figures for existing data) a new table can be created quickly, so we make less mistakes and save time!
@@
 requires a justification.  Some decisions will be too small and
-technical for inclusion in the published article itself. These need to
+technical for inclusion in the published article itself. Still, these need to
 be documented in the code itself [@nagler1995coding]. Paragraphs
@@
 term did not really change the substantive story arising from the
-model.^[R considers text marked with `#` as a comment.]
+model.^[R considers text marked with `#` as a comment. For STATA simply add a `*` before the text]
 Messages left for your future self (or near-future others) help
@@
 
-Comment liberally. Comments are discarded when R runs analysis or LaTeX or pandoc
+Comment liberally. Comments are discarded when R or STATA runs analysis, or LaTeX or pandoc
 turns plain text into pdf or html files so only those who dig into
@@
 
+<!-- JAKE in the example below can you also add a line of text to people see the combination of text and table? -->
+
 ```{r fig1plotcode,eval=FALSE,echo=TRUE}
@@
 presentation of a linear model.}
+
+For STATA users, there now is Markdoc (https://github.com/haghish/MarkDoc) which can be used similarly.
 
@@
 ```
+
+<!-- I how many of our readers will know what Sweave is.. drop or change? -->
 
@@
 used to work on the project. This kind of solution is one of the
-things that formal version control systems do well (as discussed
+things that Dropbox and more formal version control systems do well (as discussed
 a bit more below in [@sec:vers-contr-prev]).
@@
 There are some noteworthy new developments also.
-For example, and similar to R Markdown/KnitR in R Studio, there is Markdoc for STATA users (https://github.com/haghish/MarkDoc).
 People at UC Berkeley have been developing a is a web app, Jupyter Notebook (http://jupyter.org/), that enables you to make and share documents that include text, code, equations and visualizations in one INCLUDE REF TO GITHUB - USING PYTHON.
@@
 allows your computerized analyses of your collections of stuff to be
-credible.  Finally, if the someone quibbles with your analyses, your future
+credible.  If then someone quibbles with your analyses, your future
 self can shoot them the archive required to reproduce your
@@
 #References
+
+<!-- dont forget Camerer 2016 Science and  Silberzahn and Uhlmann (Nature 2015)

### 2016-06-27 workflow.Rmd
Jake working on Maartens edits

@@
 ---
-title: How to improve your relationship with your future self.
+title: How to improve your relationship with your future self, V 2.0
 author:
@@
   affiliation: <jwbowers@illinois.edu>,  University of Illinois at Urbana-Champaign, Departments of Political Science and Statistics; Fellow, White House Social and Behavioral Sciences Team.
-Nothing in this paper reflect the official position of the US Government.
-Many thanks to Mark Fredrickson, Brian Gaines, Kieran Healy, Kevin Quinn and Cara Wong for direct help on this document and to Mika LaVaque-Manty and Ben Hansen for many useful discussions on this topic.
-The source code for this document may be freely downloaded and modified from <https://github.com/jwbowers/workflow>.
 - name: Maarten Voors
   affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics Group.
-Special thanks, the EGAP Learning Days 2016 participants in Santiago Chile,  to the BITSS team and the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during spring 2016.
+\emph{Acknowledgements:} Many thanks to the EGAP Learning Days 2016 participants in Santiago Chile, to the BITSS team and the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during spring 2016.
 Maarten gratefully acknowledges financial support from N.W.O.
 grant 451-14-001.
+The previous version of this paper benefited from comments and discussions with Mark Fredrickson, Brian Gaines, Kieran Healy, Kevin Quinn, Cara Wong, Mika LaVaque-Manty and Ben Hansen.
+The source code for this document may be freely downloaded and modified from <https://github.com/jwbowers/workflow>.
 bibliography: references.bib
@@
 <!-- An example comment -->
+<!-- Separate the thanks sections from affiliation and integrate? Wasnt sure how to do this here-->
 
-  <!-- Separate the thanks sections from affiliation and integrate? Wasnt sure how to do this here-->
 ```{r include=FALSE, cache=FALSE}
@@
 opts_chunk$set(
-  tidy=FALSE,     # display code as typed
-  size="small",    # slightly smaller font for code
+  tidy=TRUE,     # display code as typed
+  size="footnotesize",    # slightly smaller font for code
   echo=TRUE,
@@
   highlight=TRUE,
-  width.cutoff=132,
-  size='footnotesize',
+  width.cutoff=100,
   out.width='.9\\textwidth',
@@
 	scipen=8,
-	width=132
+	width=100
 	)
@@
 Memory is tricky.
-When we really want to remember something we should practice it until it becomes internalized.
-In fact, learning requires effort.
-And when we do not practice, we tend to be overconfident in our recall abilities [@koriat2005illusions] and false memories can be implanted.^[See the following site for a nice overview of what we know about memory including the fact that learning requires practice: <http://www.spring.org.uk/2012/10/how-memory-works-10-things-most-people-get-wrong.php>.
+Learning requires effort.
+When we do not practice and repeat something that we want to remember, most people forget quickly, are overconfident in their abilities to recall future information [@koriat2005illusions], and may even recall events that never happened.^[See the following site for a nice overview of what we know about memory including the fact that learning requires practice: <http://www.spring.org.uk/2012/10/how-memory-works-10-things-most-people-get-wrong.php>.
 On false memory see Wikipedia and linked studies <https://en.wikipedia.org/wiki/False_memory>.] Moreover, most of us live busy lives.
-We write text and code behind our new standing desk at home while the kids run around, cry about an elder sister taking away a favorite new yellow school bus toy, drinks spill on the floor, the laundry needs going, the kettle boils...
-Our minds are full and we need to efficiently move from task to task.
+We write text and code behind our new standing desk at home while the kids run around, cry about an elder sister taking away a favorite new yellow school bus toy, drinks spill on the floor, the laundry needs doing, the kettle boils \ldots Our lives and minds are full and we need to efficiently move from task to task.
 If we cannot count on memory, then how can we do science?
@@
 We also suspect that, if some other scholar decides to re-examine the analyses of a published study, it will occur after publication.
-Moreover, this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, years past the initial publication of the article.^[We note that sometimes this reproduction occurs within the context of classes and learning about statistics and data analysis.
-Other times, this process is organized more broadly (cite IIIE program, researchers have started to implement psychological experiments using the same protocols across many locations, see Open Science Foundation 2014, 2015, within experimental economics a team did the same, see Camerer et al 2016 Science paper.
-Another example is to share analysis across many researchers.
-in a their 'crowdsourcing analysis" Silberzahn and Uhlmann (Nature 2015) asked 29 research teams to run the analysis for a project that looked at the impact of skin color on the probability of getting a foul card in a football match.<!-- Can you do the correct citing here -->]
+Moreover, this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, many years past the initial publication of the article.^[The process of reproducing past findings can occur when one researcher wants to build on the work of another.
+It can also occur within the context of classes --- some professors assign reproduction tasks to students to aid learning about data analysis and statistics.
+In addition to those models, recently reproduction of research has been organized to enhance the quality of public policy in the field of economic development by the 3ie Replication Program (@brown2014quality) and to assess the quality of scientific research at the level of subdisciplines in social psychology (@open2015estimating) and within experimental economics (@camerer2016evaluating).
+Another of 29 research teams recently collaborated on a project focusing on applied statistics to see if the same answers would emerge from re-analyses of the same data set (@silberzahn2015crowdsourced).]
 
@@
 How can we improve our relationship with our future self?
-This essay is a heavily revised and updated version of @bowers2011tpm and provides some suggestions for practices that will make such reproduction occur much more easily and quickly.
+This essay is a heavily revised and updated version of @bowers2011tpm and provides some suggestions for practices that will make reproducible data analysis easy and quick.
 Specifically, this piece aims to amplify some of what we already ought to know and do, and highlight some current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists.
 Now, the efforts to encourage transparency and ease of learning from the data and analyses of others has become institutionalized with the DA-RT inititiave <http://www.dartstatement.org/> @lupia2014openness.
-These ideas are spreading beyond political science as well @freese2007replication and @asendorpf2013recommendations.] We aim to provide practical advice about how to *do work* such that one complies with such recommendations as a matter of course and, in so doing, can focus personal regret on other bad past decisions but not on decisions about data analysis and the production of scholarly papers.
-
-<!-- JAKE: we dont say anything about pre registration, should we? -->
+These ideas are spreading beyond political science as well @freese2007replication and @asendorpf2013recommendations, <http://osf.io> and <http://www.bitss.org/>.] We aim to provide practical advice about how to *do work* such that one complies with such recommendations as a matter of course and, in so doing, can focus personal regret on bad past decisions that do not have to do with data analysis and the production of scholarly papers.
 
@@
 
-All results (numbers, comparisons, tables, figures) should arise from
-code. Not from a series of mouse clicks or copying and pasting. If you wanted
-to re-create the figure from say 2011 but include a new variable or
-specification, you should be able to do so with just a few edits to the
-code rather than knowledge of how you used a pointing device in
-your graphical user interface all those years ago.
+All results (numbers, comparisons, tables, plots, figures) should arise from
+code not from a series of mouse clicks or copying and pasting.^[To our future
+selves: both of us are using a computer control device called a track-pad, and
+haven't used an older device called a
+[mouse](https://en.wikipedia.org/wiki/Computer_mouse) in some years. We no
+longer hear a clicking sound when we operate the control device either. We are
+not sure why we talked about mouse clicks just now.] If you wanted to re-create
+the figure from say 2011 but include a new variable or specification, you
+should be able to do so with just a few edits to the code rather than knowledge
+of how you used a pointing device in your graphical user interface all those
+years ago.
 
-For example, using R [@R:2011], a open-source statistical software program, you might specify that a file, called `fig1.pdf` was
-produced by the following set of commands in a file called `makefig1.R`. Lets look at some annotated R code you could produce:
+For example, using R [@R:2011], a open-source statistical software program, you
+might specify that a file, called `fig1.pdf` was produced by the following set
+of commands in a file called `makefig1.R`. Lets look at some annotated R code
+you could produce:
 
-```{r eval=FALSE}
+```{r eval=FALSE,tidy=FALSE}
 ## This file produces a plot relating the explanatory variable to the outcome.
@@
 
-Now, in the future if you wonder how 'that plot on page 10' was
-created, you will know: (1) 'that plot' is from a file called
-`fig1.pdf`and (2) `fig1.pdf`was created in
-`makefig1.R`. Any time in a future, provided R still exists, changing the figure
-will require some quick edits of commands already written.  In a future
-where R does not exist, you (or someone else) will at least be able to read the plain text
-R commands and use them to write code in a new favorite statistical
-computing language: R scripts are written in
-[plain text](http://en.wikipedia.org/wiki/Plain_text), and plain
-text is a format that will be around as long as computer programmers
-write computer programs.^[Since R is open
-  source, you will also be able to download an old version of R, download an
-  old-fashioned open-source operating system (like Ubuntu 10), and
-  run the old-fashioned statistical computing environment in the
-  old-fashioned operating system in a virtual machine on my
-  new-fashioned actual machine.]
+Now, in the future if you wonder how 'that plot on page 10' was created, you
+will know: (1) 'that plot' is from a file called `fig1.pdf`and (2)
+`fig1.pdf`was created in `makefig1.R`. Any time in a future, provided R still
+exists, changing the figure will require some quick edits of commands already
+written.^[We use data from @norris2015data throughout this paper.]  In a future where R does not exist, you (or someone else) will at
+least be able to read the plain text R commands and use them to write code in a
+new favorite statistical computing language: R scripts are written in [plain
+text](http://en.wikipedia.org/wiki/Plain_text), and plain text is a format that
+will be around as long as computer programmers write computer programs.^[Since
+R is open source, you will also be able to download an old version of R,
+download an old-fashioned open-source operating system (like Ubuntu 22), and
+run the old-fashioned statistical computing environment in the old-fashioned
+operating system in a virtual machine on your new-fashioned actual machine.]
 
-Moreover, coding saves time. Often, a lot of time, especially if projects grow (in the number of files, partners, analysis). Manually importing
-a data file into R, may be effortless. Importing 100 files is another issue, and the time costs of each manual action add up quickly (and the probability for error increases).
+Moreover, coding saves time. Often, a lot of time, especially if projects grow
+in the number of files, partners, or complexity of analysis. Manually importing
+one data file into R, may be effortless. Importing 100 files is another issue,
+and the time costs of each manual action add up quickly (and the probability
+for error increases).
 
-Also, realize that *file names send messages* to your future self. This means that if you name your files with evocative and descriptive names, your
-collaborators are less likely to call you at midnight asking for help and you will remove some regret from your future self.
-For example, if you are studying inequality and protest, you might try naming a file something like  `inequality-and-protest-figures.R` instead of `temp9.R` or
-`supercalifragilisticexpialidocious.R`.  More on this below. By the way, the extension
-`.R` tells us and the operating system that the file contains R commands. This part of the filename enables us to quickly search our
-antique hard drives for files containing R scripts.
+Also, realize that *file names send messages* to your future self. This means
+that if you name your files with evocative and descriptive names, your
+collaborators are less likely to call you at midnight asking for help and you
+will remove some regret from your future self and protect your friendships and
+working relationships. For example, if you are studying inequality and protest,
+you might try naming a file something like `inequality-and-protest-figures.R`
+instead of `temp9.R` or `supercalifragilisticexpialidocious.R`.  By the way,
+the extension `.R` tells us and the operating system that the file contains R
+commands. This part of the filename enables us to quickly search our antique
+hard drives for files containing R scripts.^[We think that some method of
+tagging files by the purpose of the file will continue help analysts find and organize their files for long after the idea of a computer mouse ceases to make sense.]
 
-Coding helps us to avoid making mistakes.
-For example, in our example above we may be interested in how many people protest.
-We may use a data file containing all protests for several years.
-We can then create a table with the number of protesters by copying the results manually to a document you use to write your paper.
-In copying we can make mistakes.
-So a better approach is to ask R (or STATA, or another program of choice) to automatically create the table (in the format you like, with horizontal lines, 3 decimals, a title, etc) and output this in whatever file type you need (.pdf, .rtf, etc).
-Now when we obtain new data (for additional years, or updated figures for existing data) a new table can be created quickly, so we make less mistakes and save time!
+Coding helps us avoid making mistakes. For example, in our example above we may
+be interested in how many people protest. We may use a data file containing all
+protests for several years. People often create a tabular display of this data
+by copying the results manually to the working paper document. In copying we
+can make mistakes. So a better approach is to ask your favorite data analysis
+program to automatically create the table (in the format you like, with
+horizontal lines, 3 decimals, a title, etc) and output this in whatever file
+type you need (.pdf, .rtf, etc). Now when we obtain new data (for additional
+years, or updated figures for existing data) a new table can be created
+quickly, so we make less mistakes and save time!
 
-<!-- Can you include some code here to provide an example of automation of results? -->
+For example, the following table was created entirely with code using R and the `xtable` package @dahl2016aa.\footnote{@beck2010reg inspired this particular
+presentation of a linear model.}
+
+```{r loadlibraries, echo=FALSE}
+makebecktable<-function(lmobj,vars,thecaption,thelabel,filename){
+	require(xtable)
+	##  The next lines extract relevant and useful output from the
+	##  regression object, following Neal Beck's TPM article 2010.
+	themat<-cbind(summary(lmobj)$coef[,c("Estimate","Std. Error")],confint(lmobj))
+	n<-length(lmobj$fitted.values)
+	r2<-signif(summary(lmobj)$r.squared,2)
+	est.sigma<-signif(summary(lmobj)$sigma,2)
+	colnames(themat)[3:4]<-c("~{}","~{}")
+	row.names(themat)<-vars
+	thextab<-xtable(themat,align=c("l",rep(".",4)),digits=1,
+			caption=thecaption,
+			label=thelabel)
+	print(thextab,
+	      file=filename,
+	      include.colnames=FALSE,
+	      hline.after=NULL,
+	      table.placement="H",
+	      comment=FALSE,
+	      sanitize.text.function=function(x){x},
+	      add.to.row=list(
+			      pos=list(0,3),
+			      command=c(
+					"\\toprule \n &
+					\\multicolumn{1}{c}{Coef} &
+					\\multicolumn{1}{c}{Std. Err.} &
+					\\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ",
+				paste("& \\multicolumn{4}{c}{n: ",n,",",
+				      "   resid.sd: ",signif(est.sigma),
+				      ", R$^2$: ",r2,"} \\\\ \\bottomrule",sep="")
+				)
+			      ))
+
+}
+
+```
+
+```{r maketable1, echo=TRUE, tidy=FALSE}
+lm1<-lm(protest05~gini04+meanpr,data=good.df)  #  Run the regression
+makebecktable(lmobj=lm1, # make the table file
+	      vars=c("Intercept","Income Inequality (lower=more equal)",
+		     "Mean Political Rights (lower=more rights)"),
+	      thecaption="People living in countries with unequal
+	      income distributions report more protest activity to World Values
+	      Survey interviewers than people living in countries with relatively
+	      more equal income distributions, adjusting for average political
+	      rights as measured by Freedom House 1980--2010.",
+	      thelabel="tab:protest",
+	      filename="protesttable.tex")
+```
+
+```
+\input{protesttable.tex} ## read in the table file
+```
+
+\input{protesttable.tex}
 
@@
 
-Data analysis involves a long series of decisions. Each decision
-requires a justification.  Some decisions will be too small and
-technical for inclusion in the published article itself. Still, these need to
-be documented in the code itself [@nagler1995coding]. Paragraphs
-and citations in the publication will justify the most important
-decisions. So, one must code to communicate with yourself and
-others. There are two main ways to avoid forgetting the reasons you
-did something with data: comment your code and tightly link your code
-with your writing.
+Data analysis involves a long series of decisions. Each decision requires a
+justification.  Some decisions will be too small and technical for inclusion in
+the published article itself. Still, these need to be documented in the code
+itself [@nagler1995coding]. Paragraphs and citations in the publication will
+justify the most important decisions but the code itself documents the smaller
+but still consequential decisions. So, one must code to communicate with
+yourself and others. There are two main ways to avoid forgetting the reasons
+you did something with data: comment your code and tightly link your code with
+your writing --- making your code literate.
 
@@
 
-Comments, unexecuted text inside of a script, are a message to
-collaborators (including your future self) and other consumers of your
-work. In the above code chunk, we used comments to explain the lines to
-readers unfamiliar with R and to remember that we had tried a different
-specification but decided not to use it because adding the squared
-term did not really change the substantive story arising from the
-model.^[R considers text marked with `#` as a comment. For STATA simply add a `*` before the text]
-Messages left for your future self (or near-future others) help
-retrace and justify your decisions as the work moves from seminar
-paper to conference paper to poster back to paper to dissertation and
-onwards.
+Comments, unexecuted text inside of a script, are a message to collaborators
+(including your future self) and other consumers of your work. In the above
+code chunk, we used comments to explain the lines to readers unfamiliar with R
+and to remember that we had tried a different specification but decided not to
+use it because adding the squared term did not really change the substantive
+story arising from the model.^[R considers text marked with `#` as a comment.
+For Stata simply add a `*` before the text.] Messages left for your future self
+(or near-future others) help retrace and justify your decisions as the work
+moves from seminar paper to conference paper to poster back to paper to
+dissertation and onwards maybe even to publication.
 
-Notice one other benefit of coding for an audience: we learn by
-teaching. By assuming that others will look at your code, you will be
-more likely to write clearer code, or perhaps even to think more
-deeply about what you are doing as you do it.
+Notice one other benefit of coding for an audience: we learn by teaching. By
+assuming that others will look at your code, you will be more likely to write
+clearer code, or perhaps even to think more deeply about what you are doing as
+you do it since you are explaining even as you write.
 
-Comment liberally. Comments are discarded when R or STATA runs analysis, or LaTeX or pandoc
-turns plain text into pdf or html files so only those who dig into
-your work will see them.
+Comment liberally. Comments are discarded when R or STATA runs analysis, so
+only those who dig into the source code of your work will see them.
 
@@
 your working group or as a memo for a web or reviewer appendix. The
-report itself is a pdf file or some other format which displays page
+report itself is a pdf or html file or some other format which displays page
 images to ease reading rather than to encourage reanalysis and
@@
 
-More importantly, we and others want to know why we did what we
-did. Such explanations may not be very clear if we have some pages of
-printed code in one hand and a manuscript in the other. Keep in mind
-the distinction between the 'source code' of a document (i.e. what
-computation was required to produce it) and the visible, type-set page
-image. Page images are great for reading, but not great for
-reproducing or collaborating. The source code of any document
-exchanged by the group must be available and executable.^[As of the 2016 version of this paper, this idea is now widespread and made much easier than before via online services for code sharing and collaboration such as [GitHub](http://github.com), [Open Science Framework](https://osf.io/) and [BitBucket](http://bitbucket.org), see more below.]
+More importantly, we and others want to know why we did what we did. Such
+explanations may not be very clear if we have some pages of printed code in one
+hand and a manuscript in the other. Keep in mind the distinction between the
+'source code' of a document (i.e. what computation was required to produce it)
+and the visible, type-set page image. Page images are great for reading, but
+not great for reproducing or collaborating. The source code of any document
+exchanged by the group must be available and executable.^[As of the 2016
+version of this paper, this idea is now widespread and made much easier than
+before via online services for code sharing and collaboration such as
+[GitHub](http://github.com), [Open Science Framework](https://osf.io/) and
+[BitBucket](http://bitbucket.org). As more data analysis moves online, it will
+become easier for cross-platform and geographically distant collaboration to
+occur, see for only one set of examples, [Docker](http://docker.com) or other
+services that make what is currently called 'cloud-computing' easier and more
+accessible.]
 
-How might one avoid these problems?
-[Literate programming](http://en.wikipedia.org/wiki/Literate_programming) is the practice of weaving code into a document ---
-paragraphs, equations, and diagrams can explain the code, and the code
-can produce numbers, figures, and tables (and diagrams and even equations and
-paragraphs). Literate programming is not merely fancy commenting but
-is about enabling the practice of programming itself to facilitate easy
-reproduction and communication.
+How might one avoid these problems?  [Literate
+programming](http://en.wikipedia.org/wiki/Literate_programming) is the practice
+of weaving code into a document --- paragraphs, equations, and diagrams can
+explain the code, and the code can produce numbers, figures, and tables (and
+diagrams and even equations and paragraphs). Literate programming is not merely
+fancy commenting but is about enabling the practice of programming itself to
+facilitate easy reproduction and communication.
 
-For example, in [@sec:data-analys-comp], we suggested that we know
-where 'that plot on page 10' comes from by making sure we had a
-`fig1.pdf` file produced from a clearly commented plain text file
-called something like `makefig1.R`.  An even easier solution would be to
-directly include a chunk of code to produce the figure inside of the
-paper itself. This paper, for example, was written in plain text using
-markdown markup with R code chunks to make things like
-Figure \ref{fig:giniprot}. This combination of markdown and R is called
-[R Markdown](http://rmarkdown.rstudio.org).
-
-<!-- JAKE in the example below can you also add a line of text to people see the combination of text and table? -->
+For example, in [@sec:data-analys-comp], we suggested that we knew where 'that
+plot on page 10' comes from by making sure we had a `fig1.pdf` file produced
+from a clearly commented plain text file called something like `makefig1.R`.
+An even easier solution would be to directly include a chunk of code to produce
+the figure inside of the paper itself. This paper, for example, was written in
+plain text using markdown markup with R code chunks to make things like Figure
+\ref{fig:giniprot}. This combination of markdown and R is called [R
+Markdown](http://rmarkdown.rstudio.org). Here, is an example of embedding a
+plot inside of a document. In a document sent to a journal one would tend to hide the code chunks and thus one would only see the plot.
 
@@
 
-```{r giniprot,message=FALSE,echo=FALSE,fig=TRUE,include=TRUE,width=4,height=4,fig.cap="Average number of Protest activity by income inequality.",out.width=".5\\textwidth",fig.env="figure",fig.lp="fig:",fig.pos="H",fig.align="center"}
+```{r giniprot,message=FALSE,echo=FALSE,fig=TRUE,include=TRUE,width=4,height=4,fig.cap="Average number of protest activities by income inequality across countries in 2004--2005.",out.width=".5\\textwidth",fig.env="figure",fig.lp="fig:",fig.pos="H",fig.align="center"}
 ##  Make a scatterplot of Protest by Inequality
@@
 Markdown (and LaTeX and HTML) all have ways to cross-reference within a
-document. For example, by using `\label{fig:giniprot}` in LaTeX or
-`{#fig:giniprot}` in the pandoc version of markdown, I do not need to keep
-track of the figure number, nor do extra work when I reorganize the document in
-response to reviewer suggestions. Nor do I need a separate `makefig1.R`file or
-`fig1.pdf` file.  Tables and other numerical results are also possible to
-generate within the source code of a scholarly paper. Those who view the code
-for this essay will see how Table \ref{tab:protest} was also generated directly
-from a regression object.\footnote{@beck2010reg inspired this particular
-presentation of a linear model.}
+document. For example, by using `\label{fig:giniprot}` in \LaTeX or
+`{#fig:giniprot}` in the pandoc version of markdown, we do not need to keep
+track of the figure number, nor do extra work when we reorganize the document
+in response to reviewer suggestions. Nor do we need a separate `makefig1.R`file
+or `fig1.pdf` file.  Tables and other numerical results are also possible to
+generate within the source code of a scholarly paper. For example, if we had
+ommitted the filename in `makebecktable` above, we would have seen a LaTeX
+table be generated within this document itself. For Stata users, there now is
+[Markdoc](https://github.com/haghish/MarkDoc) which is very similar to R
+markdown.
 
-For STATA users, there now is Markdoc (https://github.com/haghish/MarkDoc) which can be used similarly.
+The R project has a task view devoted to [reproducible
+research](http://cran.r-project.org/web/views/ReproducibleResearch.html)
+listing many of the different approaches to literate programming for R.  If
+your workflow does not involve R, you can still implement some of the
+principles here. Imagine creating a style in Microsoft Word called 'code 'which
+hides your code when you print your document, but which allows you to at least
+run each code chunk piece by piece (or perhaps there are ways to extract all
+text of style code from a Microsoft Word document). Or imagine just using some
+other kind of indication linking paragraphs to specific places in code files.
+There are many ways that creative people can program in a literate way.
 
-```{r maketable, echo=FALSE, results='hide'}
-  lm1<-lm(protest05~gini04+meanpr,data=good.df)  #  Run the regression
+Literate programming need not go against the principle of modular data analysis
+[@nagler1995coding]. Jake routinely uses several different files that fulfill
+different functions, some of them create \LaTeX code that he can `\input` into
+his `main.tex` file, others setup the data, run simulations, or allow him to
+record his journeys down blind alleys. Of course, when we have flying cars
+running on autopilot, perhaps something other than combining R and Markdown or
+LaTeX will make our lives even easier. Then we'll change.
 
-##  The next lines extract relevant and useful output from the
-##  regression object, following Neal Beck's TPM article 2010.
-themat<-cbind(summary(lm1)$coef[,c("Estimate","Std. Error")],confint(lm1))
-n<-length(lm1$fitted.values)
-r2<-signif(summary(lm1)$r.squared,2)
-est.sigma<-signif(summary(lm1)$sigma,2)
-colnames(themat)[3:4]<-c("~{}","~{}")
-row.names(themat)<-c("Intercept","Income Inequality (lower=more equal)","Mean Political Rights (lower=more rights)")
-
-library(xtable)
-thextab<-xtable(themat,align=c("l",rep(".",4)),digits=1,
-                caption="People living in countries with unequal
-  income distributions report less protest activity to World Values
-  Survey interviewers than people living in countries with relatively
-  more equal income distributions, adjusting for average political
-  rights as measured by Freedom House 1980--2000. Data from [@norris2015data].",label="tab:protest")
-```
-
-```{r printtable,echo=FALSE,results='asis',message=FALSE}
-  print(thextab,include.colnames=FALSE,
-        hline.after=NULL,
-        table.placement="H",
-	comment=FALSE,
-        add.to.row=list(
-          pos=list(0,3),
-          command=c(
-            "\\toprule \n &
-             \\multicolumn{1}{c}{Coef} &
-             \\multicolumn{1}{c}{Std. Err.} &
-             \\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ",
-            paste("& \\multicolumn{4}{c}{n: ",n,",",
-                  "   resid.sd: ",signif(est.sigma),
-                  ", R$^2$: ",r2,"} \\\\ \\bottomrule",sep="")
-            )
-          ))
-```
-
-<!-- I how many of our readers will know what Sweave is.. drop or change? -->
-
-Literate data analysis is not the same as Sweave, even if Sweave is a
-nice implementation.^[The R project has a task view devoted to
-  [reproducible research](http://cran.r-project.org/web/views/ReproducibleResearch.html) listing many of the different approaches to literate
-  programming for R.}  If your workflow does not involve R, you can
-still implement some of the principles here. Imagine creating a style
-in Microsoft Word called 'code 'which hides your code when you print
-your document, but which allows you to at least run each code chunk
-piece by piece (or perhaps there are ways to extract all text of style
-code from a Microsoft Word document). Or imagine just using some
-other kind of indication linking paragraphs to specific places in code
-files. There are many ways that creative people can program in a
-literate way.
-
-Literate programming need not go against the principle of modular data
-analysis [@nagler1995coding]. Jake routinely uses
-several different Sweave files that fulfill different functions, some
-of them create \LaTeX code that he can `\input` into his
-`main.tex` file, others setup the data, run simulations, or allow
-him to record his journeys down blind alleys. Of course, when we have
-flying cars running on autopilot, perhaps something other than
-Sweave will make our lives even easier. Then we'll change.
-
-**Step 2**
-We analyze data in order to explain something about the world to other
-scholars and policy makers. If we focus on explaining how we got our
-computers to do data analysis, we will do a better job
-with the data analysis itself: we will learn as we focus on teaching,
-and we will avoid errors and save time as we ensure that others
-(including our future selves) can retrace our steps. A document than
-can be 'run' to reproduce all of the analyses also instills
-confidence in readers and can more effectively spur discussion and
-learning and cumulation of research.
+**Step 2** We analyze data in order to explain something about the world to
+other scholars and policy makers. If we focus on explaining how we got our
+computers to do data analysis, we will do a better job with the data analysis
+itself: we will learn as we focus on teaching others about why we did what we
+did, and we will avoid errors and save time as we ensure that others (including
+our future selves) can retrace our steps. A document than can be 'run' to
+reproduce all of the analyses also instills confidence in readers and can more
+effectively spur discussion and learning and cumulation of research.
 
@@
 
-All files containing commands operating on data must refer to a data
-file. A reference to a data file is a line of code the analysis
-program will use to operate on ('load'/ 'open '/ 'get '/
-'use') the data file. One should not have to edit this line on
-different computers or platforms in order to execute this command.
-Using R, for example, all analysis files should have
-`load("thedata.rda")`or `read.csv("thedata.csv")`or some
-equivalent line in them, and `thedata.csv`should be stored in
-some place easy to find (like in the same directory as the file or
-perhaps in `"Data/thedata.rda"`). Of course, its even better to
-include a comment pointing to the data file.
+All files containing commands operating on data must refer to a data file. A
+reference to a data file is a line of code the analysis program will use to
+operate on ('load'/ 'open '/ 'get '/ 'use') the data file. One should not have
+to edit this line on different computers or platforms in order to execute this
+command.  Using R, for example, all analysis files should have
+`load("thedata.rda")`or `read.csv("thedata.csv")`or some equivalent line in
+them, and `thedata.csv`should be stored in some easy to find place (like in the
+same directory as the file or perhaps in `"Data/thedata.rda"`). Of course, its
+even better to include a comment pointing to the data file in addition to the
+line loading the file itself.
 
-Where should one store data files? An obvious solution is always to
+Where should one store data files? An obvious solution is to
 make sure that the data file used by a command file is in the same
@@
 `save("cleandata.rda")`so that we are doubly certain about the provenance of
-the data.\footnote{Of course, if you need math or paragraphs to explain what is
-happening in these files, you might prefer to make them into R markdown or
-Sweave files, for which the conventional extension is .Rmd or .Rnw
+the data.^[Of course, if you need math or paragraphs to explain what is
+happening in these files, you might prefer to make them into R+Markdown or
+R+LaTeX files, for which the conventional extension is .Rmd or .Rnw
 respectively. So you'd have `create-cleandata.Rnw` written as a mixture of
-laTeX and R which might explain and explore the different coding decisions you
-made, perhaps involving a factor analysis and diagnostic plots.}
+LaTeX and R which might explain and explore the different coding decisions you
+made, perhaps involving some diagnostic plots.]
 
@@
 nicer solution is to maintain a file for each project called
-`MANIFEST.txt` or INDEX.txt or README.txt which lists the
+MANIFEST.txt or INDEX.txt or README.txt which lists the
 data and command files with brief descriptions of their functions and
@@
 For example, below we show a folder structure for a paper where we look at inequality.
-The paper is written in .tex, data analysis in STATA (that Maarten prefers, so we see .do and .dta file extensions) in R (the program of choice for Jake, see the .R extensions):
+The paper is written in .tex, data analysis in Stata (that Maarten prefers, so we see .do and .dta file extensions) in R (the program of choice for Jake, see the .R extensions).
+Maarten likes to make sure that the directory structure in his projects is the same across projects (so, for example, the numbers on the directory names allow for easy default sorting).
+The file naming with full dates also allows for easy sorting:
 
-<!-- JAKE please check if I made no mistake, also there may be a prettier way to show a folder structure (a screen shot?),feel free to change-->
-00_archive
-01_paper
+```{bash ,eval=FALSE, tidy=FALSE}
+00_archive/
+01_paper/
   20160622_inequality.tex
-02_data
-  00_archive
-  01_rawdata
-  02_cleandata
+02_data/
+  00_archive/
+  01_rawdata/
+  02_cleandata/
   20160622_analysis.dta
   20160622_analysis.rda
-03_analysis
-  00_archive
-  01_temp
-  02_output
-    01_tables
-    02_figures
+03_analysis/
+  00_archive/
+  01_temp/
+  02_output/
+    01_tables/
+    02_figures/
   20160622_analysis.do
@@
   20160622_prep_data.R
+ ```
 
-There are a couple of things to note.
-See that files start with a number to give them an ordering.
-Note also we create an 00_archive folder, here you can store older files without having to delete them permanently and keeping the number of files in your main folders to a minimum.
+There are a couple of things to note: See that files start with a number to give them an ordering.
+Note also we create an 00_archive folder, here you can store older files without having to delete them permanently and can reduce the number of files in your main folders.
 Also there is a clear separation between raw data (data that came straight from the field, downloaded data, someone else's replication files, etc) and clean data (data ready to use).
-Note that in the analysis folder we created two files, one for cleaning and merging data called 20160622_prep_data and another for running the analysis called 20160622_analysis.
+Note that in the analysis folder we created two files, one for cleaning and merging data called `20160622_prep_data` and another for running the analysis called `20160622_analysis`.
 See also the output folders, with subfolders for graphs and tables.
 These are place holder for R or STATA to output the results from the analysis and keeps things organized.
-We discuss integrated programs (that combine code and text below).
 
-**Step 3** We should know where the data came
-from and what operations were performed on which set of data.
+Jake is less organized about directory structure and tends to have directories with structures like this:
+
+```{bash, tidy=FALSE, eval=FALSE}
+
+manuscript\
+
+```
+
+**Step 3** We should know where the data came from and what operations were performed on which set of data.
 

### 2016-06-28 workflow.Rmd
Almost done

@@
 
-Jake is less organized about directory structure and tends to have directories with structures like this:
+Jake tends to be less organized than Maarten and relies on README.md files to tell his future self what is going on in a given directory, a [version control system](http://github.com) to keep track of versions of files, and a [Makefile](http://kbroman.org/minimal_make/) to keep track of relationships between files.
 
 ```{bash, tidy=FALSE, eval=FALSE}
-
 manuscript\
-
+figures\
+analysis\
+  README.md
+data\
+build\
+libraries\
+README.md
+Makefile
 ```
@@
 
-Group work requires version control. Many people are familiar with the track changes
-feature in modern WYSIWYG (what you see is what you get) word processors or the fact that Dropbox allows one
-to recover previous versions of files. These are both kinds of version control.
-More generally, when we collaborate, we'd like to do a variety of actions with
-our shared files. Collaboration on data analytic projects is more productive
-and better when (1) it is easy to see what has changed between versions of
-files; (2) members of the team feel free to experiment and then to dump parts
-of the experimentation in favor of previous work while merging the successful parts
- into the main body of the paper; (3) the team can produce
-have 'releases' of the same document (one to MPSA, one to APSR, one to your
-parents) without spawning many possibly conflicting copies of the same
-document; (4) people can work on the same files at the same time without
-conflicting with one another, and can reconcile their changes without too much
-confusion and clobbering. Clobbering is what happens when your future self or
-your current collaborator saves an old version of a file over a new version,
-erasing good work by accident.
+Group work requires version control. Many people are familiar with the 'track
+changes' feature in modern WYSIWYG (what you see is what you get) word
+processors or the fact that Dropbox allows one to recover previous versions of
+files. These are both kinds of version control.  More generally, when we
+collaborate, we'd like to do a variety of actions with our shared files.
+Collaboration on data analytic projects is more productive and better when (1)
+it is easy to see what has changed between versions of files; (2) members of
+the team feel free to experiment and then to dump parts of the experimentation
+in favor of previous work while merging the successful parts into the main body
+of the paper; (3) the team can produce 'releases' of the same document
+(one to MPSA, one to APSR, one to their parents) without spawning many possibly
+conflicting copies of the same document; (4) people can work on the same files
+at the same time without conflicting with one another, and can reconcile their
+changes without too much confusion and clobbering. Clobbering is what happens
+when your future self or your current collaborator saves an old version of a
+file over a new version, erasing good work by accident.
 
-Of course if you rely on Dropbox or 'track changes' in Word for version
-control, you must communicate with other folks in your group before you edit
-existing files. Only one of you can edit and save a given file at a time. This
-prevents your work (or your colleagues work) from getting lost when you both
-try to save the same file on top of each other. If you find that you need to
-work on the same files at the same time, then you should work on establishing
-your own shared version control system. Free options include launchpad, github, Open Science Framework,
-sourceforge for open source projects (i.e. papers you are writing which you are
-happy to share with others as you write, see more below).
-Each has a slightly difference language (some say folders other repositories, some say download othere clone, etc) and it takes a bit of time adjusting.
-But as we'll argue: it is well worth the time investment as it will save (lots!) of time later on.
-Currently Google Docs does
-an excellent job with shared editing of files. The downside is: you have to be online. For people with irregular or slow internet connections, this may not be an option. MORE ON GOOGLE?
-One may also use Dropbox as a kind of server for version control (though the downside is if your collaborations or project grows large (in terms of MBs) you need to switch to a paid service).
-Maarten has been using this for most of his collaborations.
-You can copy files from the Dropbox
+Many rely on Dropbox, 'track changes' in Word, or Google Docs for
+collaboration. Those tools tend to require you to communicate with other folks
+in your group before you edit existing files. On Dropbox, only one of you can
+edit and save a given file at a time. Tracking changes with Word tends to
+function best if two people are not editing the same document (merging the
+changes between two documents is not simple or builtin). Google Docs allows for
+two people to edit the same document at the same time: but you must be online
+with a fairly fast internet connection.^[Note to future selves: change this
+section every five years or so. Stop saying 'internet' in 2026.] Communication
+and trading off editing a file prevents your work (or your colleagues' work)
+from getting lost when you both try to save the same file on top of each other. One may also use Dropbox as a kind of server for version
+control (though the downside is if your collaborations or project grows large
+(in terms of MBs) you need to switch to Dropbox's paid service). Maarten has been using
+this for most of his collaborations. You can copy files from the Dropbox
 directory into a local working directory so as to avoid clobbering and then
 work on merging changes by hand before copying back to the Dropbox directory
-and replacing existing files. You will however need to agree on an iron file and directory naming rule with your collaborators.
+and replacing existing files. You will however need to agree on an iron file
+and directory naming rule with your collaborators to ensure that the files and directories are always named the same across machines and versions.
 
-Jake uses github within his research group, and used it for all of
-his own projects. <!-- JAKE do you want to add something else here? -->
-You may even obtain a two year unlimited repositories with an academic affiliation.
+Another approach to organizing work on many files, including files for code and
+for writing, is to use a formal [version
+control](https://en.wikipedia.org/wiki/Version_control) strategy. As of 2016,
+the standard for managing large collaborations with many files is the
+[Git](https://en.wikipedia.org/wiki/Git_(software)) system. User friendly free
+front-ends for Git currently make the process of learning and using Git very
+convenient and integrated into the kind of workflow that your future self will
+be proud of: [Open Science Framework](http://osf.io)
+[bitbucket](http://bitbucket.com), [github](http://github.com). Jake uses
+[github](http://github.com) for all of his collaborations with others as well
+as with his future self: his collaborators appreciate some of the nice extra
+features of github, such as the ability to keep a [shared task
+list](https://guides.github.com/features/issues/). Learning about version
+control systems takes a bit of time. We suggest, however, it is well worth the
+time investment as it will save (lots!) of time later on.
 
-An excellent, simple, and robust version control system is to rename your files with the date and time of
-saving them: yyyymmdd_project.docx (for changes within the same day, just before a deadline for example, you may even add a time, so it becomes 20160623_4PM_inequalitypaper.docx).
-This communicates very clear what version you are working.
-Remember the days when you received a file called inequalityMV4_APSRsubmit_reallyfinal2.tex?
-That should quickly become something of the past!
-We find ourselves preaching the gospel to our collaborators frequently and will keep renaming files until a collaborator has been successfully pacified.
-Be sure to include year in the file names --- remember, the life of an idea is
-measured in years. If you are wise enough to have saved your documents
-as plain text then you can easily compare documents using the many
-utilities available for comparing text files.^[Adobe Acrobat
-  allows one to compare differences in pdf files. OpenOffice supports
-  a 'Compare Documents 'option. Word now does the same. And Google Docs will report on the
-  version history of a document.] When you reach certain milestones
-you can rename the file accordingly: thedocAPSA2009.tex --- for the
-one sent to discussants at APSA --- or thedocAPSR2015.tex --- for the
-version eventually sent to the APSR six years after you presented it
-at APSA. The formal version control systems mentioned above all
-allow this kind of thing and are much more elegant and capable, but
-you can do it by hand too as long as you don't mind taking up a lot of
-disk space and having many 'thedoc...
-' files around.
-Indeed, the number of files can add up quickly (especially around submission time!) and you may want to create an "0_Archive" folder to store older versions.
-If you do
-version control by hand, spend a little extra time to ensure that you
-do not clobber files when you make mistakes typing in the
-file-names. And, if you find yourself spending extra time reconciling
-changes made by different collaborators by hand, remember this is a
-task that modern version control systems take care of quickly and
-easily.
+An excellent, simple, and robust version control system is to rename your files
+with the date and time of saving them: yyyymmdd_project.docx (for changes
+within the same day, just before a deadline for example, you may even add a
+time, so it becomes 20160623_4PM_inequalitypaper.docx). This communicates very
+clear what version you are working. Remember the days when you received a file
+called inequalityMV4_APSRsubmit_reallyfinal2.tex? That should quickly become
+something of the past! We find ourselves preaching the gospel to our
+collaborators frequently and will keep renaming files until a collaborator has
+been successfully pacified. Be sure to include year in the file names ---
+remember, the life of an idea is measured in years. If you are wise enough to
+have saved your documents as plain text then you can easily compare documents
+using the many utilities available for comparing text files.^[Adobe Acrobat
+allows one to compare differences in pdf files. OpenOffice supports a 'Compare
+Documents' option. Word now does the same. And Google Docs will report on the
+version history of a document.] When you reach certain milestones you can
+rename the file accordingly: thedocAPSA2009.tex --- for the one sent to
+discussants at APSA --- or thedocAPSR2015.tex --- for the version eventually
+sent to the APSR six years after you presented it at APSA. The formal version
+control systems mentioned above all allow this kind of thing and are much more
+elegant and capable, but you can do it by hand too as long as you don't mind
+taking up a lot of disk space and having many 'thedoc... ' files around.
+Indeed, the number of files can add up quickly (especially around submission
+time!) and you may want to create an "0_Archive" folder to store older
+versions. If you do version control by hand, spend a little extra time to
+ensure that you do not clobber files when you make mistakes typing in the
+file-names. And, if you find yourself spending extra time reconciling changes
+made by different collaborators by hand, remember this is a task that modern
+version control systems take care of quickly and easily.
 
-**Step 5** Writing is rewriting. Thus, all writing
-involves versions. When we collaborate with ourselves and others we
-want to avoid clobbering and we want to enable graceful reconciliation
-of rewriting. One can do these things with formal systems of software
-(like subversion or git or bazaar) or with formal systems of file
-naming, file comparing, and communication or, even better, with
-both. In either case, plain text files will make such tasks easier,
-will take up less disk space, and be easier to read for the future
-you.
+**Step 5** Writing is rewriting. Thus, all writing involves versions. When we
+collaborate with ourselves and others we want to avoid clobbering and we want
+to enable graceful reconciliation of rewriting. One can do these things with
+formal systems of software (like git) or with formal systems of file naming,
+file comparing, and communication or, even better, with both. In either case,
+plain text files will make such tasks easier, will take up less disk space, and
+be easier to read for the future you.
 
@@
 
-Anyone writing custom code should worry about getting it right.
-The more code one writes, the more time one has to appreciate problems arising from bugs, errors, and typos in data analysis and code.
-Mistakes will always be made we should just recognize we need more standardization of practice to minimize the mistakes and help others find them.
-One thing we can do is to include testing in our coding.
+Anyone writing custom code should worry about getting it right. The more code
+one writes, the more time one has to appreciate problems arising from bugs,
+errors, and typos in data analysis and code. Since mistakes will always be
+made, we should just recognize we need more standardization of practice to
+minimize the mistakes and help others find them. One thing we can do to
+minimize and catch the inevitable mistakes is to include testing in our coding.
 
-If one has a moment to think in between teaching that new class,
-reading books for an awards committee, evaluating application files
-for the admissions committee, feeding popsicles to a sick child, and
-undertaking the odd bit of research, one might say to oneself,
-"Before I write new code, I should write a test of the code. I should
-write a little bit of code that lets me know that my double-bootstrap
-procedure actually does what it is supposed to do."
+If one has a moment to think in between teaching that new class, reading books
+for an awards committee, evaluating application files for the admissions
+committee, feeding popsicles to a sick child, and undertaking the odd bit of
+research, one might say to oneself, "Before I write new code, I should write a
+test of the code. I should write a little bit of code that lets me know that my
+double-bootstrap procedure actually does what it is supposed to do."
 
-Of course, this idea, like most others, is not new. The desire to
-avoid error looms large when large groups of programmers write code
-for multi-million dollar programs. The idea of
-[test driven development](http://en.wikipedia.org/wiki/Test-driven_development) and the idea that one ought to create tests of
-[small parts of one's code](http://en.wikipedia.org/wiki/Unit_testing) arose to address such concerns.^[There are now R packages to help R package writers do this, see for example <https://github.com/hadley/testthat> and the article on it <https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf>.] For the social scientist collaborating with her future self and/or a small group of
-collaborators, here is an example of this idea in a very simple form:
-Say you want to write a function to multiply a number by 2. If the
-function works, when you give it the number 4, you should see it return
-the number 8 and when you give it -4, you should get -8.
+Of course, this idea, is not new. The desire to avoid error
+looms large when large groups of programmers write code for multi-million
+dollar programs. The idea of [test driven
+development](http://en.wikipedia.org/wiki/Test-driven_development) and the idea
+that one ought to create tests of [small parts of one's
+code](http://en.wikipedia.org/wiki/Unit_testing) arose to address such
+concerns.^[There are now R packages to help R package writers do this, see for
+example <https://github.com/hadley/testthat> and the article on it
+<https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf>.]
 
-```{r unit.test}
+For the social scientist collaborating with her future self and/or a small
+group of collaborators, here is an example of this idea in a very simple form:
+Say you want to write a function to multiply a number by 2. If the function
+works, when you give it the number 4, you should see it return the number 8 and
+when you give it -4, you should get -8.
+
+```{r unit.test, tidy=FALSE}
 ##  The test function:
@@
 
-##  The function:
+##  The actual function is:
 times.2.fn <- function(thenumber){
@@
 }
+```
+Here we use the test function to make sure it works.
 
-##  Use the test function
+```{r results='asis'}
 test.times.2.fn()
@@
 Ack! we mistyped '+' for '*'. Good thing we wrote the
-test!^[A more common example of this kind of testing occur
+test!^[A more common example of this kind of testing occurs
 everyday when we recode variables into new forms but look at a
@@
 
-MAYBE SOMETHING HERE ON `stopifnot` and recoding data.
+That approach works well when one is paying close attention to messages within
+the code. Another approach that we like better is to make the code stop with an
+error whenever it doesn't pass a test. In R we use the `stopifnot` function for
+this. For example, say we wanted to re-run our analyses excluding the countries
+with extreme values of protest. The following code makes a new dataset
+excluding places where more than 90% of the respondents to the World Values
+survey reported some protest activity. If all of the values of `protest05` are
+not less than 90, the code will stop with an error.
+
+```{r}
+smalldat <- subset(good.df,subset=protest05<90)
+stopifnot(all(smalldat$protest05<90))
+```
 
@@
 
-Lots of people are thinking about 'reproducible research', creating a 'reproducible workflow' and
-'literate programming' these days. Google those terms. Of course, the
-devil is in the details: Here I list a few of my own attempts at enabling
-reproducible research. You'll find many other inspiring examples on
-the web. Luckily, the open source ethos aligns nicely with academic
-incentives, so we are beginning to find more and more people offering
-their files online for copying and improvement. By the way, if you do
-copy and improve, it is polite to alert the person from whom you made
-the copy about your work.
+Lots of people are thinking about 'reproducible research', creating a
+'reproducible workflow' and 'literate programming' these days. Google those
+terms. Of course, the devil is in the details: Here we list a few of our own
+attempts at enabling reproducible research. You'll find many other inspiring
+examples on the web. Luckily, the open source ethos aligns nicely with academic
+incentives, so we are beginning to find more and more people offering their
+files online for copying and improvement. By the way, if you do copy and
+improve, it is polite to alert the person from whom you made the copy about
+your work and to credit them in some way.
 
-We have experimented with several systems so far: (1) we wrote this paper in R markdown (Jake) and Atom (Maarten) and organized our files and collaboration on Github.
-The source code can be accessed, downloaded and modified from <https://github.com/jwbowers/workflow>.
-The program also allowed us to send notes to each other.
-The nice thing about github is that is enables you to make "releases" of a project which enable you to smooth the reproduction of all of the
-products of your research (simulations, data analyses, tables, figures, etc..).
-Similar features are offered by the Open Science Framework (https://osf.io/), Maarten recently participated in a workshop on research transparency organized by the Berkeley Institute for Transparency in the Social Sciences (www.bitss.org), all course material was accessible through OSF (https://osf.io/qh2nr/).
-Both platforms, OSF and Github nicely integrate and a key benefit is that they have been created to remain for a long time (unlike perhaps your university drive) <!-- ADD REFERENCE TO YOUR OSF PROJECT -->; (2) for one paper Jake simply included a Sweave document and data files into a compressed archive [@bowers2005dataverse] <!-- PERHAPS SAY ABIT MORE ABOUT WHAT SWEAVE IS IN PRINCIPLE-->; (3) for another more computing intensive paper Jake assembled a set of files that enabled reproduction of results using the
-`make` system [@bowers2008dataverse] <!-- AGAIN A FEW WORDS ON WHAT PERHAPS LINK TO A WEBSITE? -->; (4) Jake also tried the "compendium"
+We have experimented with several systems so far:
+
+(1) We wrote this paper in the [R markdown](http://rmarkdown.rstudio.com/)
+literate programming format using two different text editors (Jake used vim)
+and (Maarten used [atom](https://atom.io/)). We organized our files and
+collaboration on Github. The source code for this document can be accessed,
+downloaded and modified from <https://github.com/jwbowers/workflow>. We also
+used Github Issues to send notes to each other and maintain a task list. The
+nice thing about github is that is enables you to make "releases" of a project
+which enable you to smooth the reproduction of all of the products of your
+research (simulations, data analyses, tables, figures, etc..).  Similar
+features are offered by the [Open Science Framework (OSF)](https://osf.io/).
+Maarten recently participated in a workshop on research transparency organized
+by the [Berkeley Institute for Transparency in the Social
+Sciences](www.bitss.org) in which all [course material](https://osf.io/qh2nr/)
+was accessible through OSF.  OSF integrates with Github nicely and a key
+benefit is that both systems have been created to remain for a long time
+(unlike perhaps your university drive).  (2) For one paper Jake simply mixed R
+and \LaTeX code in one document (called the "Sweave" format of literate
+programming) and then added that document and data into into a compressed
+archive [@bowers2005dataverse]; (3) For another more computing intensive paper
+Jake assembled a set of files that enabled reproduction of results using the
+`make` system [@bowers2008dataverse]; (4) Jake also tried the "compendium"
 approach [@gentleman2005reproducible,@gentleman2007statistical] which embeds an
-academic paper within the R package system [@bowers2011dataverse]; The benefit of the compendium approach is that one is not required to have
-access to a command line for `make`: the compendium is downloadable from within
-R using `install.packages()` and is viewable using the `vignette()` function in
-any operating system than runs R.^[Notice that Jake's reproduction archives and/or
+academic paper within the R package system [@bowers2011dataverse]; The benefit
+of the compendium approach is that one is not required to have access to a
+command line for `make`: the compendium is downloadable from within R using
+`install.packages()` and is viewable using the `vignette()` function in any
+operating system than runs R.^[Notice that Jake's reproduction archives and/or
 instructions for using them are hosted on the
@@
 just as one installs and uses statistical software packages is very attractive,
-and we anticipate that it will become ever easier to turn papers into R packages
-as creative and energetic folks turn their attention to the question of
-reproducible research.^[See for example <http://r-pkgs.had.co.nz/>]
-Finally, (5) Maarten has been using Dropbox for most of his collaborative projects (https://www.dropbox.com/).
-See comments above.
-As a tool for making files publicly available Dropbox is less useful.
-You can of course make replication files available through creating an public folder, but you will still need to refer to this folder on a particular project website (an example is Maarten's project on conflict and football https://www.dropbox.com/sh/wqosq99019mf04d/AAASTgJLkn6lmZJI_1veUK7ca?dl=0)
+and we anticipate that it will become ever easier to turn papers into R
+packages as creative and energetic folks turn their attention to the question
+of reproducible research.^[See for example <http://r-pkgs.had.co.nz/>] Finally,
+(5) Maarten has been using Dropbox for most of his collaborative projects. See
+comments above. As a tool for making files publicly available Dropbox is less
+useful. You can of course make replication files available through creating an
+public folder, but you will still need to refer to this folder on a particular
+project website (an example is [Maarten's project on conflict and football](
+https://www.dropbox.com/sh/wqosq99019mf04d/AAASTgJLkn6lmZJI_1veUK7ca?dl=0)).
 
-There are some noteworthy new developments also.
-People at UC Berkeley have been developing a is a web app, Jupyter Notebook (http://jupyter.org/), that enables you to make and share documents that include text, code, equations and visualizations in one INCLUDE REF TO GITHUB - USING PYTHON.
-For researchers working across different platforms, there is Docker (https://www.docker.com/).
+There are some noteworthy new developments also. People at UC Berkeley have
+been developing a web app, Jupyter Notebook (http://jupyter.org/), that enables
+you to make and share documents that include text, code, equations and
+visualizations in one [literate programming
+environment](http://blog.jupyter.org/2015/05/07/rendering-notebooks-on-github/).
+For researchers working across different platforms, there is
+[Docker](https://www.docker.com/) which promises to enable all of the
+collaborators on one project to use the same computing environment even if they
+are using different laptops running different operating systems.
 
@@
 It may work for small projects with few co-authors but when the number of collaborators increase it becomes very difficult to keep track.
-There are some great online systems for task management, most of which have mobile apps also, examples are Flow (www.getflow.com), Asana (www.asana.com), Wrike (www.wrike.com) <!-- JAKE do you know others?
---> .
+There are some great online systems for task management, most of which have mobile apps also, examples include [Flow](http://www.getflow.com), [Asana](http://www.asana.com), [Wrike](http://www.wrike.com), [Basecamp](http://basecamp.com) and many others.
 There also are great programs for notes.
 For example, Simplenote is very basic but allows you to share and edit notes.
-Others are Evernote (www.evernote.com) and for Windows users there is OneNote.
+Others include [Evernote](www.evernote.com) and for Windows users there is OneNote.
 
@@
 methods for doing social science
-[@freese2007replication,king1995replication}.
+[@freese2007replication,king1995replication].
 

### 2016-06-28 workflow.Rmd
Version sent to Fernando and Thad

@@
 
-# Creating a reproducible workflow
+# Create a reproducible workflow
 
@@
 
-We all always collaborate. Many of us collaborate with groups of
-people at one moment in time as we race against a deadline. All of us
-collaborate with ourselves over time.\footnote{What is a reasonable
-  time-span for which to plan for self-collaboration on a single idea?
-  Ask your advisers how long it took them to move from idea to dissertation to
-  publication.}  The time-frames over which collaboration are required
---- whether among a group of people working together or within a
-single scholar's productive life or probably both --- are much longer
-than any given version of any given software will easily exist. Plain
-text is the exception. Thus, even as we extol version control systems,
-one must have a way to ensure future access to them in a form that
-will still be around when sentient cockroaches finally join political
-science departments (by then dominated by cetaceans after humans are
-mostly uploads).^[The arrival of the six-legged social
-  scientists revives Emacs and finally makes Ctrl-a Ctrl-x Esc-x
-  Ctrl-c a [reasonable key combination](http://kieran.healy.usesthis.com/).]
+We all always collaborate. Many of us collaborate with groups of people at one
+moment in time as we race against a deadline. All of us collaborate with
+ourselves over time.^[What is a reasonable time-span for which to plan
+for self-collaboration on a single idea?  Ask your advisers how long it took
+them to move from idea to dissertation to publication.]  The time-frames over
+which collaboration are required --- whether among a group of people working
+together or within a single scholar's productive life or probably both --- are
+much longer than any given version of any given software will easily exist.
+Plain text is the exception. Thus, even as we extol version control systems,
+one must have a way to ensure future access to them in a form that will still
+be around when sentient cockroaches finally join political science departments
+(by then dominated by cetaceans after humans are mostly uploads).^[The arrival
+of the six-legged social scientists revives Emacs and finally makes Ctrl-a
+Ctrl-x Esc-x Ctrl-c a [reasonable key
+combination](http://kieran.healy.usesthis.com/). So far, in terms of version
+control systems, Jake has used RCS, CVS, bazaar, subversion, and now mostly
+uses git.]
 
 But what if no one ever hears of your work, or, by some cruel fate, your
-article does not spawn debate? Why then would you spend time to
-communicate with your future self and others? Our own answer to this
-question is that we want our work to be credible and useful to ourself
-and other scholars regardless.  What we report in our data analyses should have two main
-characteristics: (1) the findings of the work should not be a matter
-of opinion; and (2) other people should be able to reproduce the findings. That is,
-the work represents a shared experience --- and an experience shared without respect to the
-identities of others (although requiring some common technical
-training and research resources).
+article does not spawn debate? Why then would you spend time to communicate
+with your future self and others? Our own answer to this question is that we
+want our work to be credible and useful to ourselves and other scholars
+regardless.  What we report in our data analyses should have two main
+characteristics: (1) the findings of the work should not be a matter of
+opinion; and (2) other people should be able to reproduce the findings. That
+is, the work represents a shared experience --- and an experience shared
+without respect to the identities of others (although requiring some common
+technical training and research resources).
 
@@
 assume we want other people to believe us when we say something about data:
-'data' here can be words, numbers, musical notes, images, ideas, etc \ldots The
-point is that we are making some claims about patterns in some collection of
-stuff. For example, when Jake was invited into the homes and offices of ordinary
-people in Chile in 1991, the "stuff" was recordings of conversations that we
-had about life during the first year of democracy and Jake was comparing the
-responses of people who had experienced the Pinochet dictatorship differently.
-Now, it might be easy to convince others that 'this collection of stuff' is
-different from 'that collection of stuff' if those people were looking over our
-shoulders the whole time that we made decisions about collecting the stuff and
-broke it up into understandable parts and reorganized and summarized it.
-Unfortunately, we can't assume that people are willing to shadow a researcher
-throughout her career. Rather, we do our work alone or in small groups and want
-to convince other distant and future people about our analyses and findings.
+'data' here can be words, numbers, musical notes, images, ideas, etc. The point
+is that we are making some claims about patterns in some collection of stuff.
+For example, when Jake was invited into the homes and offices of ordinary
+people in Chile in 1991, the "stuff" was recordings of conversations that he
+had about life during the first year of democracy. Now, it might be easy to
+convince others that 'this collection of stuff' is different from 'that
+collection of stuff' if those people were looking over our shoulders the whole
+time that we made decisions about collecting the stuff and broke it up into
+understandable parts and reorganized and summarized it.  Unfortunately, we
+can't assume that people are willing to shadow a researcher throughout her
+career. Rather, we do our work alone or in small groups and want to convince
+other distant and future people to take our analyses and findings seriously.
 
@@
 chosen tools of analyses are computer programs. How can we convince
-people that what we did with some data with some program is credible,
+people that what we did with some data with some code is credible,
 not a matter of whim or opinion, and reproducible by others who didn't
@@
 @fredrickson2011tpm and @healy2011tpm among others working on these topics
-allows your computerized analyses of your collections of stuff to be
-credible.  If then someone quibbles with your analyses, your future
-self can shoot them the archive required to reproduce your
-work.^[Since you used plain text, the files will still be
-  intelligible, analyzed using commented code so that folks can
-  translate to whatever system succeeds R, or since you used R, you
-  can include a copy of R and all of the R packages you used in your
-  final analyses in the archive itself. You can even throw in
-  a copy of whatever version of linux you used and an open source virtual machine running the
-  whole environment.] You can say, 'Here is everything you need to
-reproduce my work.' To be extra helpful you can add 'Read the README
-file for further instructions.' And then you can get on with your life:
-maybe the next great idea will occur when your 4-year-old asks a wacky
-question after stripping and painting her overly cooperative
-1-year-old brother purple, or teaching a class, or in a coffee shop,
-or on a quiet walk.
+allows your computerized analyses of your collections of stuff to be credible.
+If then someone quibbles with your analyses, your future self can shoot them
+the archive required to reproduce your work.^[Since you used plain text, the
+files will still be intelligible, analyzed using commented code so that folks
+can translate to whatever system succeeds R, or since you used R, you can
+include a copy of R and all of the R packages you used in your final analyses
+in the archive itself. You can even throw in a copy of whatever version of
+linux you used and an open source virtual machine running the whole environment
+using say, docker.] You can say, 'Here is everything you need to reproduce my
+work.' To be extra helpful you can add 'Read the README file for further
+instructions.' And then you can get on with your life: maybe the next great
+idea will occur when your 4-year-old asks a wacky question after stripping and
+painting her overly cooperative 1-year-old brother purple, or teaching a class,
+or in a coffee shop, or on a quiet walk.
 
 #References
-
-<!-- dont forget Camerer 2016 Science and  Silberzahn and Uhlmann (Nature 2015)

### 2016-10-11 workflow.Rmd
Anonymous version of the MS. We should be able to ignore this branch except in case reviewers have line/page specific edits.

@@
 author:
-- name: Jake Bowers
-  affiliation: <jwbowers@illinois.edu>,  University of Illinois at Urbana-Champaign, Departments of Political Science and Statistics; Fellow, White House Social and Behavioral Sciences Team.
-- name: Maarten Voors
-  affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics Group.
-\emph{Acknowledgements:} Many thanks to the EGAP Learning Days 2016 participants in Santiago Chile, to the BITSS team and the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during spring 2016.
-Maarten gratefully acknowledges financial support from N.W.O.
-grant 451-14-001.
-The previous version of this paper benefited from comments and discussions with Mark Fredrickson, Brian Gaines, Kieran Healy, Kevin Quinn, Cara Wong, Mika LaVaque-Manty and Ben Hansen.
-The source code for this document may be freely downloaded and modified from <https://github.com/jwbowers/workflow>.
+- name: Anon 1
+  affiliation: Anon 1 Affiliation
+- name: Anon 2
+  affiliation: Anon 2 Affiliation and Acknowledgements
 bibliography: references.bib
@@
 How can we improve our relationship with our future self?
-This essay is a heavily revised and updated version of @bowers2011tpm and provides some suggestions for practices that will make reproducible data analysis easy and quick.
+This essay is a heavily revised and updated version of anonwhitepaper and provides some suggestions for practices that will make reproducible data analysis easy and quick.
 Specifically, this piece aims to amplify some of what we already ought to know and do, and highlight some current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists.
@@
 Literate programming need not go against the principle of modular data analysis
-[@nagler1995coding]. Jake routinely uses several different files that fulfill
+[@nagler1995coding]. Anon1 routinely uses several different files that fulfill
 different functions, some of them create \LaTeX code that he can `\input` into
@@
 For example, below we show a folder structure for a paper where we look at inequality.
-The paper is written in .tex, data analysis in Stata (that Maarten prefers, so we see .do and .dta file extensions) in R (the program of choice for Jake, see the .R extensions).
-Maarten likes to make sure that the directory structure in his projects is the same across projects (so, for example, the numbers on the directory names allow for easy default sorting).
+The paper is written in .tex, data analysis in Stata (that Anon2 prefers, so we see .do and .dta file extensions) in R (the program of choice for Anon1, see the .R extensions).
+Anon2 likes to make sure that the directory structure in his projects is the same across projects (so, for example, the numbers on the directory names allow for easy default sorting).
 The file naming with full dates also allows for easy sorting:
@@
 
-Jake tends to be less organized than Maarten and relies on README.md files to tell his future self what is going on in a given directory, a [version control system](http://github.com) to keep track of versions of files, and a [Makefile](http://kbroman.org/minimal_make/) to keep track of relationships between files.
+Anon1 tends to be less organized than Anon2 and relies on README.md files to tell his future self what is going on in a given directory, a [version control system](http://github.com) to keep track of versions of files, and a [Makefile](http://kbroman.org/minimal_make/) to keep track of relationships between files.
 
@@
 control (though the downside is if your collaborations or project grows large
-(in terms of MBs) you need to switch to Dropbox's paid service). Maarten has been using
+(in terms of MBs) you need to switch to Dropbox's paid service). Anon2 has been using
 this for most of his collaborations. You can copy files from the Dropbox
@@
 be proud of: [Open Science Framework](http://osf.io)
-[bitbucket](http://bitbucket.com), [github](http://github.com). Jake uses
+[bitbucket](http://bitbucket.com), [github](http://github.com). Anon1 uses
 [github](http://github.com) for all of his collaborations with others as well
@@
 (1) We wrote this paper in the [R markdown](http://rmarkdown.rstudio.com/)
-literate programming format using two different text editors (Jake used vim)
-and (Maarten used [atom](https://atom.io/)). We organized our files and
+literate programming format using two different text editors (Anon1 used vim)
+and (Anon2 used [atom](https://atom.io/)). We organized our files and
 collaboration on Github. The source code for this document can be accessed,
-downloaded and modified from <https://github.com/jwbowers/workflow>. We also
+downloaded and modified from <https://github.com/anon1/workflow>. We also
 used Github Issues to send notes to each other and maintain a task list. The
@@
 features are offered by the [Open Science Framework (OSF)](https://osf.io/).
-Maarten recently participated in a workshop on research transparency organized
+Anon2 recently participated in a workshop on research transparency organized
 by the [Berkeley Institute for Transparency in the Social
@@
 benefit is that both systems have been created to remain for a long time
-(unlike perhaps your university drive).  (2) For one paper Jake simply mixed R
+(unlike perhaps your university drive).  (2) For one paper Anon1 simply mixed R
 and \LaTeX code in one document (called the "Sweave" format of literate
 programming) and then added that document and data into into a compressed
-archive [@bowers2005dataverse]; (3) For another more computing intensive paper
-Jake assembled a set of files that enabled reproduction of results using the
-`make` system [@bowers2008dataverse]; (4) Jake also tried the "compendium"
+archive [anon1refA]; (3) For another more computing intensive paper
+Anon1 assembled a set of files that enabled reproduction of results using the
+`make` system [anon1refB]; (4) Anon1 also tried the "compendium"
 approach [@gentleman2005reproducible,@gentleman2007statistical] which embeds an
-academic paper within the R package system [@bowers2011dataverse]; The benefit
+academic paper within the R package system [anon1refC]; The benefit
 of the compendium approach is that one is not required to have access to a
@@
 `install.packages()` and is viewable using the `vignette()` function in any
-operating system than runs R.^[Notice that Jake's reproduction archives and/or
+operating system than runs R.^[Notice that Anon1's reproduction archives and/or
 instructions for using them are hosted on the
@@
 of reproducible research.^[See for example <http://r-pkgs.had.co.nz/>] Finally,
-(5) Maarten has been using Dropbox for most of his collaborative projects. See
+(5) Anon2 has been using Dropbox for most of his collaborative projects. See
 comments above. As a tool for making files publicly available Dropbox is less
@@
 public folder, but you will still need to refer to this folder on a particular
-project website (an example is [Maarten's project on conflict and football](
+project website (an example is [Anon2's project on conflict and football](
 https://www.dropbox.com/sh/wqosq99019mf04d/AAASTgJLkn6lmZJI_1veUK7ca?dl=0)).
@@
 combination](http://kieran.healy.usesthis.com/). So far, in terms of version
-control systems, Jake has used RCS, CVS, bazaar, subversion, and now mostly
+control systems, Anon1 has used RCS, CVS, bazaar, subversion, and now mostly
 uses git.]
@@
 is that we are making some claims about patterns in some collection of stuff.
-For example, when Jake was invited into the homes and offices of ordinary
+For example, when Anon1 was invited into the homes and offices of ordinary
 people in Chile in 1991, the "stuff" was recordings of conversations that he

### 2016-11-29 responsetoreviewers.md
Adding draft response to reviewers

Dear Cristian and Diego,

Thanks for giving us the chance to revise the paper on workflow.

In response to your comments and those of the reviewer, we have added XX
mentions about experiments, although we realize that the issue is really about
transparency and not experiments in particular.

About the comment that much of the material appeared in an non-peer reviewed
journal previously: When I was approached to participate in this journal, wI
volunteered to send this previously non-peer reviewed article as a submission
to be peer-reviewed. That idea was accepted. As I worked to update the
previously published version, I invited Maarten Voors to work on the paper
with me. So the version that we sent differs a bit more from the 2011 version
than I had anticipated. I don't currently see us writing a whole new paper on
this topic. We really appreciate that you allowed our paper to go through the
peer-review process and very much hope that the version that we attach here is
satisfactory. It is still more of a version 2.0 than a completely new set of
thoughts on the topic. We are happy to specifically note the differences from
the first piece and have done so here.

We have attached the file here in pdf and LaTeX format. Don't hesitate to let
us know if you would like other revisions.

Best

Jake and Maarten

Regarding the comments of Reviewer 1:

We very much appreciate the thoughtful and supportive and helpful review. We
respond to the points made here.

 - We have now pointed out explicitly where we have updated and changed the
   paper to reflect our own changes thinking and the changes in the technology
landscape.

 - We have kept the primary lessons the same: we submitted this manuscript to
   RCP in order to gain more impact from the non-peer reviewed version, and
this idea was approved by the folks working on the project at that time. We
also feel like the primary lessons stand the test of time even as the specific
methods for achieving them might change.

 - The editors highlighted to us that the issue is not primarily about
   experiments but rather about transparency, so, although we have mentioned
experiments more in this version, we have not entirely changed the focus on
the piece.

 - We hope that we have clarified the role of Dropbox, and about makefiles and
   the make system.

 - We also have edited the paper to try to emphasize that there may be many
   best practices as long as they all are created to keep in mind the central
lessons of the paper. (For example, we believe that one can do transparent
scientific workflow using many different programming languages, computing
systems, and processes as long as the central lessons are followed.)

 - We have tried to quash the typos and formatting problems.

 - We have downplayed the parenting references.

Reviewer 1

This is a helpful paper that guides readers through a set of lessons to improve the way they organize their work.  If this piece appears in RCP, the authors should do more to highlight the new contributions over the very similar piece in The Political Methodologist. The primary lessons and many of the examples are identical or very similar. The paper should also be more closely tied to experimental work, rather than data analysis in general.  Are any of these problems specific to experiments?  Are there other issues that come up in experiments that are unique to that research environment?

The authors’ perspective on the role of Dropbox is unclear.  It seems that there are a lot of potential problems here.  In this way, the paper wavers from being about best practices to being about a diversity of approaches.  The authors may want to take a harder line, at the expense of trying to be comprehensive.  Also, if it mentions them, the paper should be more clear about the content and role of the “makefile” (p. 7) and “make” (p. 10).

The paper has many typos and mechanical problems like missing spaces, inconsistent formatting of “GitHub” and “LaTeX”, a missing figure number (“??”), inconsistent spacing around operators in code (see the tests on p.9 vs. the modeling on pp. 2 and 3), etc.  There are also some odd phrasings and interjections that I recommend removing (footnote 12, “Google those terms”, “People at UC Berkeley”, etc.  I appreciate and enjoyed the paper’s  casual tone, but it goes too far in my opinion. The parenting references, for example, are frequent and detailed enough to distract the reader.

### 2016-12-07 workflow.Rmd
made some edits

--- 2016-06-28 workflow.Rmd
@@
 ---
-title: How to improve your relationship with your future self, V 2.0
+title: How to improve your relationship with your future self
 author:
@@
   affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics Group.
-\emph{Acknowledgements:} Many thanks to the EGAP Learning Days 2016 participants in Santiago Chile, to the BITSS team and the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during spring 2016.
+\emph{Acknowledgements:} Many thanks to the EGAP Learning Days 2016 participants in Santiago Chile, to the BITSS team, the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during spring 2016.
 Maarten gratefully acknowledges financial support from N.W.O.
 grant 451-14-001.
-The previous version of this paper benefited from comments and discussions with Mark Fredrickson, Brian Gaines, Kieran Healy, Kevin Quinn, Cara Wong, Mika LaVaque-Manty and Ben Hansen.
+A previous version of this paper benefited from comments and discussions with Mark Fredrickson, Brian Gaines, Kieran Healy, Kevin Quinn, Cara Wong, Mika LaVaque-Manty and Ben Hansen.
 The source code for this document may be freely downloaded and modified from <https://github.com/jwbowers/workflow>.
+This paper extends a previous version by Jake @bowers2011tpm.
+While some text remains the same we have expanded and updated the essay to reflect current developments.
 bibliography: references.bib
@@
 
-> | Do I contradict myself?
-| Very well then I contradict myself,
-| (I am large, I contain multitudes.)
-| [@whitman1855song]
-
 > If you tell the truth, you don't have to remember anything. [@twain1975mark]
@@
 It can also occur within the context of classes --- some professors assign reproduction tasks to students to aid learning about data analysis and statistics.
-In addition to those models, recently reproduction of research has been organized to enhance the quality of public policy in the field of economic development by the 3ie Replication Program (@brown2014quality) and to assess the quality of scientific research at the level of subdisciplines in social psychology (@open2015estimating) and within experimental economics (@camerer2016evaluating).
-Another of 29 research teams recently collaborated on a project focusing on applied statistics to see if the same answers would emerge from re-analyses of the same data set (@silberzahn2015crowdsourced).]
+In addition to those models, recently reproduction of research has been organized to enhance the quality of public policy in the field of economic development by the 3IE Replication Program (@brown2014quality) and to assess the quality of scientific research at the level of sub-disciplines in social psychology (@open2015estimating) and within experimental economics (@camerer2016evaluating).
+In another study 29 research teams recently collaborated on a project focusing on applied statistics to see if the same answers would emerge from re-analyses of the same data set (@silberzahn2015crowdsourced).]
 
@@
 Specifically, this piece aims to amplify some of what we already ought to know and do, and highlight some current practices, platforms, and possibilities.^[@king1995replication and @nagler1995coding were two of the first pieces introducing these kinds of ideas to political scientists.
-Now, the efforts to encourage transparency and ease of learning from the data and analyses of others has become institutionalized with the DA-RT inititiave <http://www.dartstatement.org/> @lupia2014openness.
+Now, the efforts to encourage transparency and ease of learning from the data and analyses of others has become institutionalized with the DA-RT initiative <http://www.dartstatement.org/> @lupia2014openness.
 These ideas are spreading beyond political science as well @freese2007replication and @asendorpf2013recommendations, <http://osf.io> and <http://www.bitss.org/>.] We aim to provide practical advice about how to *do work* such that one complies with such recommendations as a matter of course and, in so doing, can focus personal regret on bad past decisions that do not have to do with data analysis and the production of scholarly papers.
@@
 
-All results (numbers, comparisons, tables, plots, figures) should arise from
-code not from a series of mouse clicks or copying and pasting.^[To our future
+All your results (numbers, comparisons, tables, plots, figures) should be produced from
+code, not from a series of mouse clicks or copying and pasting.^[To our future
 selves: both of us are using a computer control device called a track-pad, and
@@
 longer hear a clicking sound when we operate the control device either. We are
-not sure why we talked about mouse clicks just now.] If you wanted to re-create
-the figure from say 2011 but include a new variable or specification, you
-should be able to do so with just a few edits to the code rather than knowledge
+not sure why we talked about mouse clicks just now.] Imagine you wanted to re-create
+a figure and include a new variable, you should be able to do so with just a few edits to the code rather than knowledge
 of how you used a pointing device in your graphical user interface all those
@@
 
-For example, using R [@R:2011], a open-source statistical software program, you
-might specify that a file, called `fig1.pdf` was produced by the following set
-of commands in a file called `makefig1.R`. Lets look at some annotated R code
-you could produce:
+Lets look at an example. Using R [@R:2011], a open-source statistical software program, you
+might specify that a file, called `fig1.pdf` is produced by the following set
+of commands in a file called `makefig1.R`. Lets look at some annotated R code:
 
@@
 will know: (1) 'that plot' is from a file called `fig1.pdf`and (2)
-`fig1.pdf`was created in `makefig1.R`. Any time in a future, provided R still
-exists, changing the figure will require some quick edits of commands already
-written.^[We use data from @norris2015data throughout this paper.]  In a future where R does not exist, you (or someone else) will at
-least be able to read the plain text R commands and use them to write code in a
+`fig1.pdf`was created in `makefig1.R`. Any changes to the file in the future (provided R still
+exists), will just require some quick edits of commands already
+written.^[We use data from @norris2015data throughout this paper.]  Even if in the future R ceases to exist, you (or someone else) will at
+least be able to read the plain text commands and use them to write code in a
 new favorite statistical computing language: R scripts are written in [plain
 text](http://en.wikipedia.org/wiki/Plain_text), and plain text is a format that
-will be around as long as computer programmers write computer programs.^[Since
-R is open source, you will also be able to download an old version of R,
-download an old-fashioned open-source operating system (like Ubuntu 22), and
-run the old-fashioned statistical computing environment in the old-fashioned
-operating system in a virtual machine on your new-fashioned actual machine.]
+will be around as long as computer programmers write computer programs.
 
@@
 and the time costs of each manual action add up quickly (and the probability
-for error increases).
+for error increases too).
 
-Also, realize that *file names send messages* to your future self. This means
+Also, realize that *file names send messages* to coauthors and to your future self. This means
 that if you name your files with evocative and descriptive names, your
@@
 
-Coding helps us avoid making mistakes. For example, in our example above we may
+Coding helps us to avoid making mistakes. For example, in our example above we may
 be interested in how many people protest. We may use a data file containing all
@@
 by copying the results manually to the working paper document. In copying we
-can make mistakes. So a better approach is to ask your favorite data analysis
-program to automatically create the table (in the format you like, with
+can make mistakes. So a better approach is to automatically create the table (in the format you like, with
 horizontal lines, 3 decimals, a title, etc) and output this in whatever file
-type you need (.pdf, .rtf, etc). Now when we obtain new data (for additional
-years, or updated figures for existing data) a new table can be created
+type you need (.tex. .pdf, .rtf, etc). Now when we obtain new data a new table can be created
 quickly, so we make less mistakes and save time!
@@
 you did something with data: comment your code and tightly link your code with
-your writing --- making your code literate.
+your writing -- making your code literate.
 
@@
 when your future self or your current collaborator saves an old version of a
-file over a new version, erasing good work by accident.
+file over a new version, erasing good work by accident (or creating a conflicted copy resulting in an arduous back and forth between versions).
 
-Many rely on Dropbox, 'track changes' in Word, or Google Docs for
-collaboration. Those tools tend to require you to communicate with other folks
-in your group before you edit existing files. On Dropbox, only one of you can
-edit and save a given file at a time. Tracking changes with Word tends to
-function best if two people are not editing the same document (merging the
+Using Dropbox, 'track changes' in Word, or Google Docs is one way to manage your collaborative efforts. They are popular tools, but
+have some downsides as they require you to communicate with other folks
+in your group before you can edit existing files. On Dropbox, you cannot
+edit and save a given file at the same time. Tracking changes with Word has similar limitations (merging the
 changes between two documents is not simple or builtin). Google Docs allows for
@@
 with a fairly fast internet connection.^[Note to future selves: change this
-section every five years or so. Stop saying 'internet' in 2026.] Communication
+section every five years or so. Probably stop saying 'internet' in 2026.] Communication
 and trading off editing a file prevents your work (or your colleagues' work)
 from getting lost when you both try to save the same file on top of each other. One may also use Dropbox as a kind of server for version
-control (though the downside is if your collaborations or project grows large
-(in terms of MBs) you need to switch to Dropbox's paid service). Maarten has been using
-this for most of his collaborations. You can copy files from the Dropbox
-directory into a local working directory so as to avoid clobbering and then
-work on merging changes by hand before copying back to the Dropbox directory
-and replacing existing files. You will however need to agree on an iron file
-and directory naming rule with your collaborators to ensure that the files and directories are always named the same across machines and versions.
-
-Another approach to organizing work on many files, including files for code and
-for writing, is to use a formal [version
-control](https://en.wikipedia.org/wiki/Version_control) strategy. As of 2016,
-the standard for managing large collaborations with many files is the
-[Git](https://en.wikipedia.org/wiki/Git_(software)) system. User friendly free
-front-ends for Git currently make the process of learning and using Git very
-convenient and integrated into the kind of workflow that your future self will
-be proud of: [Open Science Framework](http://osf.io)
-[bitbucket](http://bitbucket.com), [github](http://github.com). Jake uses
-[github](http://github.com) for all of his collaborations with others as well
-as with his future self: his collaborators appreciate some of the nice extra
-features of github, such as the ability to keep a [shared task
-list](https://guides.github.com/features/issues/). Learning about version
-control systems takes a bit of time. We suggest, however, it is well worth the
-time investment as it will save (lots!) of time later on.
+control. See example above. This system is fine untill your collaborations or project grows large
+(in terms of MBs) and you may run out of storage space on your computer or you may need to upgrade your Dropbox's storage space).
+You will however need to agree on an iron rule for file and directory naming with your collaborators to ensure that the files and directories are always named the same across machines and versions (see above).
 
@@
 
+However, best practice we feel is to use formal version control approaches to organizing work [version
+control](https://en.wikipedia.org/wiki/Version_control) strategy. As of 2016,
+the standard for managing large collaborations with many files is the
+[Git](https://en.wikipedia.org/wiki/Git_(software)) system. User friendly free
+front-ends for Git currently make the process of learning and using Git very
+convenient and integrated into the kind of workflow that your future self will
+be proud of: [Open Science Framework](http://osf.io)
+[bitbucket](http://bitbucket.com), [github](http://github.com). Jake uses
+[github](http://github.com) for all of his collaborations with others as well
+as with his future self: his collaborators appreciate some of the nice extra
+features of GitHub, such as the ability to keep a [shared task
+list](https://guides.github.com/features/issues/). Learning about version
+control systems takes a bit of time. We suggest, however, it is well worth the
+time investment as it will save (lots!) of time later on.
+
 **Step 5** Writing is rewriting. Thus, all writing involves versions. When we
@@
 minimize and catch the inevitable mistakes is to include testing in our coding.
-
-If one has a moment to think in between teaching that new class, reading books
-for an awards committee, evaluating application files for the admissions
-committee, feeding popsicles to a sick child, and undertaking the odd bit of
-research, one might say to oneself, "Before I write new code, I should write a
-test of the code. I should write a little bit of code that lets me know that my
-double-bootstrap procedure actually does what it is supposed to do."
 
@@
 
-Lots of people are thinking about 'reproducible research', creating a
-'reproducible workflow' and 'literate programming' these days. Google those
-terms. Of course, the devil is in the details: Here we list a few of our own
+Over the past decades more and more attention is being paid to 'reproducible research', creating a
+'reproducible workflow' and 'literate programming' these days. As always, the devil is in the details: Here we list a few of our own
 attempts at enabling reproducible research. You'll find many other inspiring
@@
 and (Maarten used [atom](https://atom.io/)). We organized our files and
-collaboration on Github. The source code for this document can be accessed,
+collaboration on GitHub. The source code for this document can be accessed,
 downloaded and modified from <https://github.com/jwbowers/workflow>. We also
-used Github Issues to send notes to each other and maintain a task list. The
-nice thing about github is that is enables you to make "releases" of a project
+used GitHub Issues to send notes to each other and maintain a task list. The
+nice thing about GitHub is that is enables you to make "releases" of a project
 which enable you to smooth the reproduction of all of the products of your
@@
 
-There are some noteworthy new developments also. People at UC Berkeley have
-been developing a web app, Jupyter Notebook (http://jupyter.org/), that enables
+There are some noteworthy new developments also. See for example
+a new web app, Jupyter Notebook (http://jupyter.org/), that enables
 you to make and share documents that include text, code, equations and
@@
 Remember, while email is fantastic for fast communication, it is not a tool designed for project management.
-It may work for small projects with few co-authors but when the number of collaborators increase it becomes very difficult to keep track.
-There are some great online systems for task management, most of which have mobile apps also, examples include [Flow](http://www.getflow.com), [Asana](http://www.asana.com), [Wrike](http://www.wrike.com), [Basecamp](http://basecamp.com) and many others.
-There also are great programs for notes.
-For example, Simplenote is very basic but allows you to share and edit notes.
-Others include [Evernote](www.evernote.com) and for Windows users there is OneNote.
+It may work for projects with few co-authors but when the number of collaborators increase it becomes very difficult to keep track.
+Online systems for task management, if which most have mobile apps, include amongst otheres [Flow](http://www.getflow.com), [Asana](http://www.asana.com), [Wrike](http://www.wrike.com), [Basecamp](http://basecamp.com), Simplenote, [Evernote](www.evernote.com) and for Windows users there is OneNote.
 
@@
 
-# Remember that research ought to be credible communication.
+# Research ought to be credible communication.
 
@@
 moment in time as we race against a deadline. All of us collaborate with
-ourselves over time.^[What is a reasonable time-span for which to plan
-for self-collaboration on a single idea?  Ask your advisers how long it took
-them to move from idea to dissertation to publication.]  The time-frames over
+ourselves over time. The time-frames over
 which collaboration are required --- whether among a group of people working
@@
 much longer than any given version of any given software will easily exist.
-Plain text is the exception. Thus, even as we extol version control systems,
-one must have a way to ensure future access to them in a form that will still
-be around when sentient cockroaches finally join political science departments
-(by then dominated by cetaceans after humans are mostly uploads).^[The arrival
-of the six-legged social scientists revives Emacs and finally makes Ctrl-a
-Ctrl-x Esc-x Ctrl-c a [reasonable key
-combination](http://kieran.healy.usesthis.com/). So far, in terms of version
-control systems, Jake has used RCS, CVS, bazaar, subversion, and now mostly
-uses git.]
+Plain text is the exception.
 
@@
 article does not spawn debate? Why then would you spend time to communicate
-with your future self and others? Our own answer to this question is that we
-want our work to be credible and useful to ourselves and other scholars
-regardless.  What we report in our data analyses should have two main
+with your future self and others? Our own answer to this question is that it is efficient and that out of principle we
+want our work to be credible and useful to ourselves and other scholars.  What we report in our data analyses should have two main
 characteristics: (1) the findings of the work should not be a matter of
@@
 in the archive itself. You can even throw in a copy of whatever version of
-linux you used and an open source virtual machine running the whole environment
-using say, docker.] You can say, 'Here is everything you need to reproduce my
+Linux you used and an open source virtual machine running the whole environment
+using say, Docker.] You can say, 'Here is everything you need to reproduce my
 work.' To be extra helpful you can add 'Read the README file for further

### 2016-12-13 workflow.Rmd
Edited up to Meaningful Code..

@@
   affiliation: <maarten.voors@wur.nl>, Wageningen University, Development Economics Group.
-\emph{Acknowledgements:} Many thanks to the EGAP Learning Days 2016 participants in Santiago Chile, to the BITSS team, the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during spring 2016.
+\emph{Acknowledgments:} Many thanks to the EGAP Learning Days 2016 participants in Santiago Chile, to the BITSS team, the Department of Economics and Ted Miguel at UC Berkeley where Maarten was a visiting researcher during spring 2016.
 Maarten gratefully acknowledges financial support from N.W.O.
@@
 This paper extends a previous version by Jake @bowers2011tpm.
-While some text remains the same we have expanded and updated the essay to reflect current developments.
+While most of the text remains the same we have expanded and updated the essay to reflect current developments in technology and thinking about transparency and reproducibility of research.
 bibliography: references.bib
@@
 <!-- An example comment -->
-<!-- Separate the thanks sections from affiliation and integrate? Wasnt sure how to do this here-->
+<!-- Separate the thanks sections from affiliation and integrate? Wasn't sure how to do this here-->
 
@@
 ## To make the html file do
-## render("exploration4.Rmd",output_format=html_document(fig_retina=FALSE))
+## render("exploration4.Rmd", output_format=html_document(fig_retina=FALSE))
 ## To make the pdf file do
-## render("exploration4.Rmd",output_format=pdf_document())
+## render("exploration4.Rmd", output_format=pdf_document())
 ## to debug, try render("thefile.Rmd", clean=FALSE) so that you can see the intermediate files
@@
 opts_chunk$set(
-  tidy=TRUE,     # display code as typed
-  size="footnotesize",    # slightly smaller font for code
-  echo=TRUE,
-  results='markup',
-  strip.white=TRUE,
-  fig.path='figs/fig',
-  cache=FALSE,
-  highlight=TRUE,
-  width.cutoff=100,
-  out.width='.9\\textwidth',
-  message=FALSE,
-  comment=NA)
+  tidy = TRUE,     # display code as typed
+  size = "footnotesize",    # slightly smaller font for code
+  echo = TRUE,
+  results = 'markup',
+  strip.white = TRUE,
+  fig.path = 'figs/fig',
+  cache = FALSE,
+  highlight = TRUE,
+  tidy.opts=list(width.cutoff = 60),
+  out.width = '.9\\textwidth',
+  message = FALSE,
+  comment = NA)
 ```
 
-```{r initialize,echo=FALSE}
+```{r initialize, echo = FALSE}
 ##First, just setup the R environment for today:
@@
 
-options(SweaveHooks=list(fig=function(){
-			   par(mar=c(3.5, 3, 1.1, 0),
-			       pty="s",
-			       mgp=c(1.5,0.5,0),
-			       oma=c(0,0,0,0))},
-			 echo=function(){options(continue=" ") ##Don't show "+" prompts,
-			 options(prompt=" ")
+options(SweaveHooks = list(fig = function(){
+			   par(mar = c(3.5, 3, 1.1, 0),
+			       pty = "s",
+			       mgp = c(1.5,0.5,0),
+			       oma = c(0,0,0,0))},
+			 echo = function(){options(continue = " ") ##Don't show "+" prompts,
+			 options(prompt = " ")
 			 }),
-	digits=4,
-	scipen=8,
-	width=100
+	digits = 4,
+	scipen = 8,
+	width = 80
 	)
 
-#options(error=function(){options(prompt="> ",continue="+ ");NULL})
+#options(error = function(){options(prompt = "> ", continue = "+ ");NULL})
 ```
@@
 Learning requires effort.
-When we do not practice and repeat something that we want to remember, most people forget quickly, are overconfident in their abilities to recall future information [@koriat2005illusions], and may even recall events that never happened.^[See the following site for a nice overview of what we know about memory including the fact that learning requires practice: <http://www.spring.org.uk/2012/10/how-memory-works-10-things-most-people-get-wrong.php>.
+When we do not practice and repeat something that we want to remember, most people forget quickly, are overconfident in their abilities to recall future information [@koriat2005illusions], and may even recall events that never happened.^[See the following site for a nice overview of what we know about memory --- including the fact that learning requires practice: <http://www.spring.org.uk/2012/10/how-memory-works-10-things-most-people-get-wrong.php>.
 On false memory see Wikipedia and linked studies <https://en.wikipedia.org/wiki/False_memory>.] Moreover, most of us live busy lives.
-We write text and code behind our new standing desk at home while the kids run around, cry about an elder sister taking away a favorite new yellow school bus toy, drinks spill on the floor, the laundry needs doing, the kettle boils \ldots Our lives and minds are full and we need to efficiently move from task to task.
+We type text knowing that the laundry needs doing, hearing children play or cry, ignoring the news or the latest journal review.
+Our lives and minds are full and we need to efficiently move from task to task.
 If we cannot count on memory, then how can we do science?
 
-How long does it take from planning a study to publication?
-How much time passes to the first reproduction of a study?
+How long does it take from planning a study to publication and then to the first reproduction of it?
 Is three years too short?
@@
 We also suspect that, if some other scholar decides to re-examine the analyses of a published study, it will occur after publication.
-Moreover, this new scholarly activity of learning from one anothers' data and analyses, can occur at anytime, many years past the initial publication of the article.^[The process of reproducing past findings can occur when one researcher wants to build on the work of another.
+Moreover, this new scholarly activity of learning from one another's data and analyses can occur at anytime, many years past the initial publication of the article.^[The process of reproducing past findings can occur when one researcher wants to build on the work of another.
 It can also occur within the context of classes --- some professors assign reproduction tasks to students to aid learning about data analysis and statistics.
-In addition to those models, recently reproduction of research has been organized to enhance the quality of public policy in the field of economic development by the 3IE Replication Program (@brown2014quality) and to assess the quality of scientific research at the level of sub-disciplines in social psychology (@open2015estimating) and within experimental economics (@camerer2016evaluating).
-In another study 29 research teams recently collaborated on a project focusing on applied statistics to see if the same answers would emerge from re-analyses of the same data set (@silberzahn2015crowdsourced).]
+In addition to those models, reproduction of research has recently been organized to enhance the quality of public policy in the field of economic development by the 3IE Replication Program (@brown2014quality) and to assess the quality of scientific research at within social psychology (@open2015estimating) and within experimental economics (@camerer2016evaluating).
+In another study, 29 research teams recently collaborated on a project focusing on applied statistics to see if the same answers would emerge from re-analyses of the same data set (@silberzahn2015crowdsourced).
+They didn't.]
 
@@
 
-```{r read.data,echo=FALSE,results='hide', cache=TRUE}
+```{r read.data, echo = FALSE, results = 'hide', cache = TRUE}
 ## This data frame will be used throughout the analyses.
@@
 library(readstata13)
-dem.nations.df<-read.dta13("https://www.dropbox.com/s/huhqag5rudwtbno/Democracy%20Cross-National%20Data%20V4.1%2009092015.dta?dl=1&pv=1&raw=1",convert.factors=FALSE)
+dem.nations.df <- read.dta13("https://www.dropbox.com/s/huhqag5rudwtbno/Democracy%20Cross-National%20Data%20V4.1%2009092015.dta?dl=1&pv=1&raw=1", convert.factors = FALSE)
 
-row.names(dem.nations.df)<-dem.nations.df$Nation ##make the row names more helpful
-good.df<-na.omit(dem.nations.df[,c("Gini2004","WVS5_protest","Nation","meanpr")])
-good.df$gini04<-good.df$Gini2004/100
-good.df$protest05<-good.df$WVS5_protest
+row.names(dem.nations.df) <- dem.nations.df$Nation ##make the row names more helpful
+good.df <- na.omit(dem.nations.df[, c("Gini2004","WVS5_protest","Nation","meanpr")])
+good.df$gini04 <- good.df$Gini2004/100
+good.df$protest05 <- good.df$WVS5_protest
 ```
 
-We organize the paper around a series of homilies that lead to certain concrete actions.
+We organize the paper around a series of homilies that lead to certain concrete actions:
+
+ - Data analysis is computer programming.
+ - No data analyst is an island for long.
+ - Meaningful code requires data.
+ - Version control prevents clobbering, reconciles history, and helps organize work.
+ - Testing minimizes error.
+ - Work can be reproducible.
+ - Research ought to be credible communication.
 
@@
 
-All your results (numbers, comparisons, tables, plots, figures) should be produced from
-code, not from a series of mouse clicks or copying and pasting.^[To our future
-selves: both of us are using a computer control device called a track-pad, and
-haven't used an older device called a
-[mouse](https://en.wikipedia.org/wiki/Computer_mouse) in some years. We no
-longer hear a clicking sound when we operate the control device either. We are
-not sure why we talked about mouse clicks just now.] Imagine you wanted to re-create
-a figure and include a new variable, you should be able to do so with just a few edits to the code rather than knowledge
-of how you used a pointing device in your graphical user interface all those
-years ago.
+All your results (numbers, comparisons, tables, plots, figures) should be produced from code, not from a series of mouse clicks or copying and pasting.^[To our future selves: both of us are using a computer control device called a track-pad, and haven't used an older device called a [mouse](https://en.wikipedia.org/wiki/Computer_mouse) in some years.
+Our current track-pads do not click.
+We are not sure why we talked about mouse clicks just now.] Imagine you wanted to re-create a figure and include a new variable, you should be able to do so with just a few edits to the code rather than knowledge of how you used a pointing device in your graphical user interface all those years ago.
 
-Lets look at an example. Using R [@R:2011], a open-source statistical software program, you
+Lets look at an example. Using an open-source statistics programming language called R [@R:2016], you
 might specify that a file, called `fig1.pdf` is produced by the following set
@@
 
-```{r eval=FALSE,tidy=FALSE}
-## This file produces a plot relating the explanatory variable to the outcome.
-thedata <- read.csv("Data/thedata-15-03-2011.csv") ## Read the data
-pdf("fig1.pdf") ## begin writing to the pdf file
+```{r eval = FALSE, tidy = FALSE, highlight = FALSE}
+# This file produces a plot relating the explanatory variable to the outcome.
+## Read the data
+thedata <- read.csv("Data/thedata-15-03-2011.csv")
+## begin writing to the pdf file
+please-open-pdf("fig1.pdf")
 please-plot(outcome by explanatory using thedata. red lines please.)
@@
 ## summary(abs(fitted(model1)-fitted(model2)))
-dev.off() ## stop writing to the pdf file
+## stop writing to the pdf file
+please-close-pdf()
 ```
@@
 will know: (1) 'that plot' is from a file called `fig1.pdf`and (2)
-`fig1.pdf`was created in `makefig1.R`. Any changes to the file in the future (provided R still
-exists), will just require some quick edits of commands already
-written.^[We use data from @norris2015data throughout this paper.]  Even if in the future R ceases to exist, you (or someone else) will at
-least be able to read the plain text commands and use them to write code in a
-new favorite statistical computing language: R scripts are written in [plain
+`fig1.pdf`was created in `makefig1.R`. Any changes to the file in the future
+will just require some quick edits of commands already written  (provided R
+still exists).^[We use data from @norris2015data throughout this paper.]  Even
+if in the future R ceases to exist, you (or someone else) will at least be able
+to read the plain text commands and use them to write code in a new favorite
+statistical computing language: R scripts are written in [plain
 text](http://en.wikipedia.org/wiki/Plain_text), and plain text is a format that
@@
 
-Moreover, coding saves time. Often, a lot of time, especially if projects grow
-in the number of files, partners, or complexity of analysis. Manually importing
-one data file into R, may be effortless. Importing 100 files is another issue,
-and the time costs of each manual action add up quickly (and the probability
-for error increases too).
+Moreover, coding saves time. Often, typing commands into a file saves a lot of
+time, especially if projects grow in the number of files, collaborators, or
+complexity of analysis. Manually importing one data file into R may be
+effortless. Importing 100 files is another issue, and the time costs of each
+manual action add up quickly (and the probability for mistakes increases too).
 
-Also, realize that *file names send messages* to coauthors and to your future self. This means
+Also, realize that **file names send messages** to coauthors and to your future self. This means
 that if you name your files with evocative and descriptive names, your
@@
 can make mistakes. So a better approach is to automatically create the table (in the format you like, with
-horizontal lines, 3 decimals, a title, etc) and output this in whatever file
-type you need (.tex. .pdf, .rtf, etc). Now when we obtain new data a new table can be created
+horizontal lines, 3 decimals, a title, etc) and save this table in whatever file
+type you need (.tex. .pdf, .rtf, etc). Now when we obtain new data, a new table can be created
 quickly, so we make less mistakes and save time!
 
-For example, the following table was created entirely with code using R and the `xtable` package @dahl2016aa.\footnote{@beck2010reg inspired this particular
-presentation of a linear model.}
+For example, the following table was created entirely with code using R and the `xtable` package @dahl2016aa.\footnote{@beck2010reg inspired this particular presentation of a linear model.} We then read this table into the current file using the LaTeX command `\input{protesttable.tex}` which produces a nice looking table in pdf format.
 
-```{r loadlibraries, echo=FALSE}
-makebecktable<-function(lmobj,vars,thecaption,thelabel,filename){
+```{r loadlibraries, echo = FALSE}
+makebecktable <- function(lmobj, vars, thecaption, thelabel, filename){
         require(xtable)
@@
         ##  regression object, following Neal Beck's TPM article 2010.
-	themat<-cbind(summary(lmobj)$coef[,c("Estimate","Std. Error")],confint(lmobj))
-	n<-length(lmobj$fitted.values)
-	r2<-signif(summary(lmobj)$r.squared,2)
-	est.sigma<-signif(summary(lmobj)$sigma,2)
-	colnames(themat)[3:4]<-c("~{}","~{}")
-	row.names(themat)<-vars
-	thextab<-xtable(themat,align=c("l",rep(".",4)),digits=1,
-			caption=thecaption,
-			label=thelabel)
+        themat <- cbind(summary(lmobj)$coef[, c("Estimate","Std. Error")], confint(lmobj))
+        n <- length(lmobj$fitted.values)
+        r2 <- signif(summary(lmobj)$r.squared,2)
+        est.sigma <- signif(summary(lmobj)$sigma,2)
+        colnames(themat)[3:4] <- c("~{}","~{}")
+        row.names(themat) <- vars
+        thextab <- xtable(themat, align = c("l", rep(".",4)), digits = 1,
+                          caption = thecaption,
+                          label = thelabel)
         print(thextab,
-	      file=filename,
-	      include.colnames=FALSE,
-	      hline.after=NULL,
-	      table.placement="H",
-	      comment=FALSE,
-	      sanitize.text.function=function(x){x},
-	      add.to.row=list(
-			      pos=list(0,3),
-			      command=c(
+              file = filename,
+              include.colnames = FALSE,
+              hline.after = NULL,
+              table.placement = "H",
+              comment = FALSE,
+              sanitize.text.function = function(x){x},
+              add.to.row = list(
+                                pos = list(0,3),
+                                command = c(
                                             "\\toprule \n &
@@
                                             \\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ",
-				paste("& \\multicolumn{4}{c}{n: ",n,",",
-				      "   resid.sd: ",signif(est.sigma),
-				      ", R$^2$: ",r2,"} \\\\ \\bottomrule",sep="")
+                                    paste("& \\multicolumn{4}{c}{n: ", n,",",
+                                          "   resid.sd: ", signif(est.sigma),
+                                          ", R$^2$: ", r2,"} \\\\ \\bottomrule", sep = "")
                                     )
@@
 
-```{r maketable1, echo=TRUE, tidy=FALSE}
-lm1<-lm(protest05~gini04+meanpr,data=good.df)  #  Run the regression
-makebecktable(lmobj=lm1, # make the table file
-	      vars=c("Intercept","Income Inequality (lower=more equal)",
-		     "Mean Political Rights (lower=more rights)"),
-	      thecaption="People living in countries with unequal
+```{r maketable1, echo = TRUE, tidy = TRUE, tidy.opts=list(width=50)}
+#  Run the regression
+lm1 <- lm(protest05~gini04+meanpr, data = good.df)
+# make the table file
+makebecktable(lmobj = lm1,
+              vars = c("Intercept","Income Inequality (lower = more equal)",
+                       "Mean Political Rights (lower = more rights)"),
+              thecaption = "People living in countries with unequal
               income distributions report more protest activity to World Values
@@
               rights as measured by Freedom House 1980--2010.",
-	      thelabel="tab:protest",
-	      filename="protesttable.tex")
-```
-
-```
-\input{protesttable.tex} ## read in the table file
+              thelabel = "tab:protest",
+              filename = "protesttable.tex")
 ```
@@
 
-**Step 1** Code everything that can be coded.
+Notice that we might have made one file called `makefig1.R` and another called `protestincomedisttab.R` which in turn produce a pdf file (`fig1.pdf`) and a LaTeX format table (`protesttable.tex`).
+This idea of modularity [@nagler1995coding] in code enables us to quickly find errors, and, if we want to make changes to improve our work, only one file may be changed at a time.
+Further, it enhances collaboration --- Jake can work to make the figure and Maarten can work to make the table without worry about conflicting files.
+
+### Step 1
+Code everything that can be coded.
 If we know the provenance of results, future or current collaborators make less mistakes and can quickly and easily reproduce (and thus change and improve) upon the work.
@@
 you did something with data: comment your code and tightly link your code with
-your writing -- making your code literate.
+your writing---making your code literate.
 
@@
 programming](http://en.wikipedia.org/wiki/Literate_programming) is the practice
-of weaving code into a document --- paragraphs, equations, and diagrams can
+of weaving code into a document such that paragraphs, equations, and diagrams can
 explain the code, and the code can produce numbers, figures, and tables (and
@@
 from a clearly commented plain text file called something like `makefig1.R`.
-An even easier solution would be to directly include a chunk of code to produce
-the figure inside of the paper itself. This paper, for example, was written in
-plain text using markdown markup with R code chunks to make things like Figure
-\ref{fig:giniprot}. This combination of markdown and R is called [R
-Markdown](http://rmarkdown.rstudio.org). Here, is an example of embedding a
-plot inside of a document. In a document sent to a journal one would tend to hide the code chunks and thus one would only see the plot.
+An even *easier* solution would be to directly include a chunk of code to
+produce the figure inside of the paper itself. This paper, for example, was
+written in plain text using markdown markup with R code chunks to make things
+like Figure \ref{fig:giniprot}.^[This combination of markdown and R is called
+[R Markdown](http://rmarkdown.rstudio.org).] Here is an example of embedding a
+plot inside of a document along with the source of this paragraph. In a
+document sent to a journal one would tend to hide the code chunks and thus one
+would only see the plot.
 
-```{r fig1plotcode,eval=FALSE,echo=TRUE}
+```
+For example, in [@sec:data-analys-comp], we suggested that we knew where
+'that plot on page 10' comes from by making sure we had a `fig1.pdf` file
+produced from a clearly commented plain text file called something like
+`makefig1.R`.  An even *easier* solution would be to directly include a chunk
+of code to produce the figure inside of the paper itself. This paper, for
+example, was written in plain text using markdown markup with R code chunks to
+make things like Figure \ref{fig:giniprot}.^[This combination of markdown and R
+is called [R Markdown](http://rmarkdown.rstudio.org).] Here is an example of
+embedding a plot inside of a document along with the source of this
+paragraph. In a document sent to a journal one would tend to hide the code
+chunks and thus one would only see the plot.
+```
+
+```{r fig1plotcode, eval = FALSE, echo = TRUE}
 ##  Make a scatterplot of Protest by Inequality
-par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
-with(good.df,plot(gini04,protest05,
-                  xlab='Gini Coefficient 2004 (UNDP)',
-                  ylab='Proportion Reporting Protest Activities\n(World Values Survey 2005)',
-                  cex=.8))
+par(bty = "n", xpd = TRUE, pty = "s", tcl = -.25)
+with(good.df, plot(gini04, protest05,
+                  xlab = 'Gini Coefficient 2004 (UNDP)',
+                  ylab = 'Proportion Reporting Protest Activities\n(World Values Survey 2005)',
+                  cex = .8))
 ##  Label a few interesting points
 with(good.df[c("Brazil","United Kingdom","United States","Sweden","Chile"),],
-     text(gini04,protest05,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
+     text(gini04, protest05, labels = Nation, srt = 0, cex = .6, pos = 3, offset = .1))
 ```
 
-```{r giniprot,message=FALSE,echo=FALSE,fig=TRUE,include=TRUE,width=4,height=4,fig.cap="Average number of protest activities by income inequality across countries in 2004--2005.",out.width=".5\\textwidth",fig.env="figure",fig.lp="fig:",fig.pos="H",fig.align="center"}
+```{r giniprot, message = FALSE, echo = FALSE, fig = TRUE, include = TRUE, width = 4, height = 4, fig.cap = "Average number of protest activities by income inequality across countries in 2004--2005.", out.width = ".5\\textwidth", fig.env = "figure", fig.lp = "fig:", fig.pos = "H", fig.align = "center"}
 ##  Make a scatterplot of Protest by Inequality
-par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
-with(good.df,plot(gini04,protest05,
-                  xlab='Gini Coefficient 2004 (UNDP)',
-                  ylab='Proportion Reporting Protest Activities\n(World Values Survey 2005)',
-                  cex=.8))
+par(bty = "n", xpd = TRUE, pty = "s", tcl = -.25)
+with(good.df, plot(gini04, protest05,
+                  xlab = 'Gini Coefficient 2004 (UNDP)',
+                  ylab = 'Proportion Reporting Protest Activities\n(World Values Survey 2005)',
+                  cex = .8))
 ##  Label a few interesting points
 with(good.df[c("Brazil","United Kingdom","United States","Sweden","Chile"),],
-     text(gini04,protest05,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
+     text(gini04, protest05, labels = Nation, srt = 0, cex = .6, pos = 3, offset = .1))
 ```
@@
 Markdown (and LaTeX and HTML) all have ways to cross-reference within a
-document. For example, by using `\label{fig:giniprot}` in \LaTeX or
-`{#fig:giniprot}` in the pandoc version of markdown, we do not need to keep
-track of the figure number, nor do extra work when we reorganize the document
-in response to reviewer suggestions. Nor do we need a separate `makefig1.R`file
-or `fig1.pdf` file.  Tables and other numerical results are also possible to
-generate within the source code of a scholarly paper. For example, if we had
-ommitted the filename in `makebecktable` above, we would have seen a LaTeX
-table be generated within this document itself. For Stata users, there now is
+document. For example, by using `\label{fig:giniprot}` in LaTeX or
+`{#fig:giniprot}` in the [pandoc](http://pandoc.org/) version of markdown built
+into [RStudio](https://www.rstudio.com/), we do not need to keep track of the
+figure number, nor do extra work when we reorganize the document in response to
+reviewer suggestions. Nor do we need a separate `makefig1.R`file or `fig1.pdf`
+file.  Tables and other numerical results are also possible to generate within
+the source code of a scholarly paper. For example, if we had omitted the
+filename in `makebecktable` above, the LaTeX formatted table would have appeared
+within this document itself. For Stata users, there now is
 [Markdoc](https://github.com/haghish/MarkDoc) which is very similar to R
@@
 your workflow does not involve R, you can still implement some of the
-principles here. Imagine creating a style in Microsoft Word called 'code 'which
+principles here. Imagine creating a style in Microsoft Word called 'code' which
 hides your code when you print your document, but which allows you to at least
 run each code chunk piece by piece (or perhaps there are ways to extract all
-text of style code from a Microsoft Word document). Or imagine just using some
-other kind of indication linking paragraphs to specific places in code files.
-There are many ways that creative people can program in a literate way.
+text of style 'code' from a Microsoft Word document for use in some other
+program). Or one could just use some other kind of indication linking
+paragraphs to specific places in code files.  There are many ways that creative
+people can program in a literate way.
 
@@
 [@nagler1995coding]. Jake routinely uses several different files that fulfill
-different functions, some of them create \LaTeX code that he can `\input` into
+different functions, some of them create LaTeX code that he can `\input` into
 his `main.tex` file, others setup the data, run simulations, or allow him to
@@
 
-**Step 2** We analyze data in order to explain something about the world to
+### Step 2
+We analyze data in order to explain something about the world to
 other scholars and policy makers. If we focus on explaining how we got our
@@
 did, and we will avoid errors and save time as we ensure that others (including
-our future selves) can retrace our steps. A document than can be 'run' to
-reproduce all of the analyses also instills confidence in readers and can more
+our future selves) can retrace our steps. A paper than can be "run" to
+reproduce all of its analyses also instills confidence in readers and can more
 effectively spur discussion and learning and cumulation of research.
@@
 
-```{bash ,eval=FALSE, tidy=FALSE}
+```{bash , eval = FALSE, tidy = FALSE}
 00_archive/
@@
 
-```{bash, tidy=FALSE, eval=FALSE}
+```{bash, tidy = FALSE, eval = FALSE}
 manuscript\
@@
 
-# Minimize error by testing.
+# Testing minimizes error.
 
@@
 
-```{r unit.test, tidy=FALSE}
+```{r unit.test, tidy = FALSE}
 ##  The test function:
@@
 	##  This function tests times.2.fn
-	if (times.2.fn(thenumber=4) ==  8 &
-	    times.2.fn(thenumber=-4) == -8) {
+	if (times.2.fn(thenumber = 4) == 8 &
+	    times.2.fn(thenumber = -4) == -8) {
 		print("It works!")
@@
 
-```{r results='asis'}
+```{r results = 'asis'}
 test.times.2.fn()
@@
 ```{r}
-smalldat <- subset(good.df,subset=protest05<90)
+smalldat <- subset(good.df, subset = protest05<90)
 stopifnot(all(smalldat$protest05<90))
@@
 
-# Create a reproducible workflow
+# Work can be reproducible.
 
@@
 (unlike perhaps your university drive).  (2) For one paper Jake simply mixed R
-and \LaTeX code in one document (called the "Sweave" format of literate
+and LaTeX code in one document (called the "Sweave" format of literate
 programming) and then added that document and data into into a compressed
@@
 project website (an example is [Maarten's project on conflict and football](
-https://www.dropbox.com/sh/wqosq99019mf04d/AAASTgJLkn6lmZJI_1veUK7ca?dl=0)).
+https://www.dropbox.com/sh/wqosq99019mf04d/AAASTgJLkn6lmZJI_1veUK7ca?dl = 0)).
 
@@
 methods for doing social science
-[@freese2007replication,king1995replication].
+[@freese2007replication, @king1995replication].
 

### 2016-12-14 responsetoreviewers.md
Editing response to reviewers

@@
-Dear Cristian and Diego,
+Hi Cristian and Carsten,
 
@@
 
-In response to your comments and those of the reviewer, we have added XX
-mentions about experiments, although we realize that the issue is really about
-transparency and not experiments in particular.
+In response to your comments and those of the reviewer we revised the paper.
+Because we realize that the issue is really about transparency and not experiments in particular, we did not end up mentioning experiments in our current set of revisions.
+We could add a few mentions here and there if that would help.
 
-About the comment that much of the material appeared in an non-peer reviewed
-journal previously: When I was approached to participate in this journal, wI
-volunteered to send this previously non-peer reviewed article as a submission
-to be peer-reviewed. That idea was accepted. As I worked to update the
-previously published version, I invited Maarten Voors to work on the paper
-with me. So the version that we sent differs a bit more from the 2011 version
-than I had anticipated. I don't currently see us writing a whole new paper on
-this topic. We really appreciate that you allowed our paper to go through the
-peer-review process and very much hope that the version that we attach here is
-satisfactory. It is still more of a version 2.0 than a completely new set of
-thoughts on the topic. We are happy to specifically note the differences from
-the first piece and have done so here.
+About the comment that much of the material appeared in an non-peer reviewed journal previously: When I was approached to participate in this journal, I volunteered to send this previously non-peer reviewed article as a submission to be peer-reviewed.
+That idea was accepted.
+As I worked to update the previously published version, I invited Maarten Voors to work on the paper with me given shared interests and to diversify the paper (i.e.
+to ensure that this was not merely the opinion of one single person based on one set of experiences).
+Maarten has a PhD in economics and works in an economics department, so we both agreed that we would have different perspectives which would enhance the paper.
+So the version that we sent differs a bit more from the 2011 version than I had anticipated.
+That said, I don't currently see us writing a whole new paper on this topic.
+We really appreciate that you allowed our paper to go through the peer-review process and very much hope that the version that we attach here is satisfactory.
+It is still more of a version 2.0 than a completely new set of thoughts on the topic.
 
-We have attached the file here in pdf and LaTeX format. Don't hesitate to let
-us know if you would like other revisions.
+We have attached the file here in pdf format and will happily follow with LaTeX and image sources when needed. Don't hesitate to let us know if you would like other revisions.
 
@@
 
- - We have now pointed out explicitly where we have updated and changed the
-   paper to reflect our own changes thinking and the changes in the technology
-landscape.
+ - We have now pointed out explicitly that we have updated and changed the paper to reflect our own changes thinking and the changes in the technology landscape.
 
- - We have kept the primary lessons the same: we submitted this manuscript to
-   RCP in order to gain more impact from the non-peer reviewed version, and
-this idea was approved by the folks working on the project at that time. We
-also feel like the primary lessons stand the test of time even as the specific
-methods for achieving them might change.
+ - We have kept the primary lessons the same: we submitted this manuscript to RCP in order to gain more impact from the non-peer reviewed version, and this idea was approved by the folks working on the project at that time.
+We also feel like the primary lessons stand the test of time even as the specific methods for achieving them might change.
 
- - The editors highlighted to us that the issue is not primarily about
-   experiments but rather about transparency, so, although we have mentioned
-experiments more in this version, we have not entirely changed the focus on
-the piece.
+ - The editors highlighted to us that the issue is not primarily about experiments but rather about transparency, so we have not mentioned experiments more in this version.
 
- - We hope that we have clarified the role of Dropbox, and about makefiles and
-   the make system.
+ - We hope that we have clarified the role of Dropbox, and about makefiles and the make system.
 
- - We also have edited the paper to try to emphasize that there may be many
-   best practices as long as they all are created to keep in mind the central
-lessons of the paper. (For example, we believe that one can do transparent
-scientific workflow using many different programming languages, computing
-systems, and processes as long as the central lessons are followed.)
+ - We also have edited the paper to try to emphasize that there may be many best practices as long as they all are created to keep in mind the central lessons of the paper.
+(For example, we believe that one can do transparent scientific workflow using many different programming languages, computing systems, and processes as long as the central lessons are followed.)
 
- - We have tried to quash the typos and formatting problems.
+ - We have tried to quash the typos and formatting problems although we expect that your copyeditors will help continue this work.
 

### 2016-12-14 workflow.Rmd
Pre-release version, sent to RCP editors

@@
 On false memory see Wikipedia and linked studies <https://en.wikipedia.org/wiki/False_memory>.] Moreover, most of us live busy lives.
-We type text knowing that the laundry needs doing, hearing children play or cry, ignoring the news or the latest journal review.
+We type text knowing that the laundry needs doing, hearing children play or cry, ignoring the news or the latest journal review, worrying about a friend.
 Our lives and minds are full and we need to efficiently move from task to task.
@@
 Now, in the future if you wonder how 'that plot on page 10' was created, you
-will know: (1) 'that plot' is from a file called `fig1.pdf`and (2)
-`fig1.pdf`was created in `makefig1.R`. Any changes to the file in the future
+will know: (1) 'that plot' is from a file called `fig1.pdf` and (2)
+`fig1.pdf` was created in `makefig1.R`. Any changes to the file in the future
 will just require some quick edits of commands already written  (provided R
@@
 figure number, nor do extra work when we reorganize the document in response to
-reviewer suggestions. Nor do we need a separate `makefig1.R`file or `fig1.pdf`
+reviewer suggestions. Nor do we need a separate `makefig1.R` file or `fig1.pdf`
 file.  Tables and other numerical results are also possible to generate within
@@
 our future selves) can retrace our steps. A paper than can be "run" to
-reproduce all of its analyses also instills confidence in readers and can more
+reproduce all of analyses also instills confidence in readers and can more
 effectively spur discussion and learning and cumulation of research.
 
-# Meaningful code requires data.
+# The territory of a data analysis requires maps.
+
+Data analysis tends to involve multiple people working with multiple files.
+Future collaborators will need a map to understand the flow of inputs (like raw
+data) and outputs (like figures, tables, and individual numbers).
+
+## Meaningful code requires data.
 
 All files containing commands operating on data must refer to a data file. A
-reference to a data file is a line of code the analysis program will use to
-operate on ('load'/ 'open '/ 'get '/ 'use') the data file. One should not have
-to edit this line on different computers or platforms in order to execute this
-command.  Using R, for example, all analysis files should have
-`load("thedata.rda")`or `read.csv("thedata.csv")`or some equivalent line in
-them, and `thedata.csv`should be stored in some easy to find place (like in the
-same directory as the file or perhaps in `"Data/thedata.rda"`). Of course, its
-even better to include a comment pointing to the data file in addition to the
-line loading the file itself.
+reference to a data file is a line of code that the analysis program will use
+to operate on ('load' / 'open' / 'get' / 'use') the data file. One should not
+have to edit this line on different computers or platforms in order to execute
+this command.  Using R, for example, all analysis files should have
+`load("thedata.rda")` or `read.csv("thedata.csv")` or some equivalent line in
+them, and `thedata.csv` should be stored in some easy to find place (like in
+the same directory as the file or perhaps in `"Data/thedata.rda"`). Of course,
+its even better to include a comment pointing to the data file in addition to
+the line loading the file itself. This means one should not see lines like
+`setwd(C:\JakesFiles\theproject)` or `setwd(/Users/maartenvoors/theproject)` at
+the beginning of documents that are meant to be shared with a future self or
+present or future others.
+
+## File organization can be a map itself
 
@@
 processing, recoding, and merging from analysis in different files
-[@nagler1995coding]. So, perhaps your analysis oriented files will
-`load("cleandata.rda")`and a comment in the code will alert the future you
-(among others) that `cleandata.rda`was created from `create-cleandata.R` which
+[@nagler1995coding]. So, perhaps your analysis files will
+`load("cleandata.rda")` and a comment in the code will alert the future you
+(among others) that `cleandata.rda` was created from `create-cleandata.R` which
 in turn begins with
-`read.csv(\url("http://www.greatfreedata.gov/dirtydata.csv"))`. Such a data
+`read.csv(\url("http://data.gov/dirtydata.csv"))`. Such a data
 processing file will typically end with something like
-`save("cleandata.rda")`so that we are doubly certain about the provenance of
+`save("cleandata.rda")` so that we are doubly certain about the provenance of
 the data.^[Of course, if you need math or paragraphs to explain what is
@@
 R+LaTeX files, for which the conventional extension is .Rmd or .Rnw
-respectively. So you'd have `create-cleandata.Rnw` written as a mixture of
-LaTeX and R which might explain and explore the different coding decisions you
+respectively. So you'd have `create-cleandata.Rmd` written as a mixture of
+markdown and R which might explain and explore the different coding decisions you
 made, perhaps involving some diagnostic plots.]
 
-Now, if in the future we wonder where `cleandata.rda`came from, we
-might search for occurrences of `cleandata` in the files on our
-system. However, if such searching among files is a burden, an even
-nicer solution is to maintain a file for each project called
-MANIFEST.txt or INDEX.txt or README.txt which lists the
-data and command files with brief descriptions of their functions and
-relations.
-
-Best practice is to create a file system where you separate folders by function and separate input from output files.
-For example, below we show a folder structure for a paper where we look at inequality.
-The paper is written in .tex, data analysis in Stata (that Maarten prefers, so we see .do and .dta file extensions) in R (the program of choice for Jake, see the .R extensions).
-Maarten likes to make sure that the directory structure in his projects is the same across projects (so, for example, the numbers on the directory names allow for easy default sorting).
+Now, if in the future we wonder where `cleandata.rda` came from, we might
+search for occurrences of `cleandata` in the files on our system. Of course, if
+it is difficult to find the right version of `cleandata` on the system, it
+might help to know where files like `cleandata` tend to be: that is, to have a
+system for project file organization.  Best practice is to create a file system
+where you separate folders by function and separate input from output files.
+For example, below we show a folder structure for a paper where we look at
+inequality. The paper is written in .tex, data analysis in both Stata (that
+Maarten prefers, so we see .do and .dta file extensions) and in R (the program
+of choice for Jake, see the .R extensions). Maarten likes to make sure that the
+directory structure in his projects is the same across projects (so, for
+example, the numbers on the directory names allow for easy default sorting).
 The file naming with full dates also allows for easy sorting:
@@
     02_figures/
-  20160622_analysis.do
-  20160622_analysis.R
-  20160622_prep_data.do
-  20160622_prep_data.R
+      20160622_models.do
+      20160612_simulations.R
+      20160502_prep_data.do
  ```
@@
 Also there is a clear separation between raw data (data that came straight from the field, downloaded data, someone else's replication files, etc) and clean data (data ready to use).
-Note that in the analysis folder we created two files, one for cleaning and merging data called `20160622_prep_data` and another for running the analysis called `20160622_analysis`.
-See also the output folders, with subfolders for graphs and tables.
-These are place holder for R or STATA to output the results from the analysis and keeps things organized.
+Note that in the analysis folder we created two files, one for cleaning and merging data called `20160622_prep_data` and another for running the analysis called `20160622_models`.
+See also the placeholder output folders, with subfolders for figures and tables.
 
-Jake tends to be less organized than Maarten and relies on README.md files to tell his future self what is going on in a given directory, a [version control system](http://github.com) to keep track of versions of files, and a [Makefile](http://kbroman.org/minimal_make/) to keep track of relationships between files.
+## The input-output map can also be in the form of files
+
+Another solution to the problem of finding files and knowing how files relate,
+is to maintain a file for each project called MANIFEST.txt or INDEX.txt or
+README.txt which lists the data and command files with brief descriptions of
+their functions and relations. However, this file may be a burden to maintain.
+
+Jake tends to be less organized than Maarten and relies on README.md files to
+tell his future self what is going on in a given directory, a [version control
+system](http://github.com) to keep track of versions of files (so that he
+doesn't need to add dates to filenames), and a
+[Makefile](http://kbroman.org/minimal_make/) to keep track of relationships
+between files. Notice that the `analysis` subdirectory has its own README.md
+file.
 
@@
 manuscript\
+  paper.pdf
+  paper.Rmd
 figures\
+  boxplots.R
+  boxplots.pdf
+tables\
 analysis\
   README.md
+  models.R
 data\
+  workingdata.R
+  rawdata.csv
+  workingdata.rda
 build\
@@
 
-**Step 3** We should know where the data came from and what operations were performed on which set of data.
+```{r echo=FALSE}
+## from https://stackoverflow.com/questions/20409172/how-to-display-verbatim-inline-r-code-with-backticks-using-rmarkdown
+rinline <- function(code) {
+  sprintf('``` `r %s` ```', code)
+}
+```
+
+Imagine that the `paper.Rmd` shows the `boxplots.pdf` file and also reports on the total size of the data.
+So, the introduction to the paper requires the `workingdata.rda` file (to know the sample size) and `boxplots.pdf`.
+In turn, those files depend on `workingdata.R` which takes `rawdata.csv` and produces the `workingdata.rda` file, and `boxplots.R` which uses `workingdata.rda` to make the plots.
+A Makefile records this web of relationships in a structured way such that one can say `make manuscript/paper.pdf` and all of the files that are required to produce `paper.pdf` will be executed if they have not recently been used or if they have recently been changed.^[See <http://kbroman.org/minimal_make/> for more about Makefiles.] Here is an example that records the relationships just described:
+
+```{bash eval=FALSE}
+manuscript/paper.pdf: manuscript/paper.Rmd figures/boxplot.pdf
+    cd manuscript && Rscript -e "library(rmarkdown); render('paper.Rmd')"
+
+figures/boxplot.pdf: figures/boxplot.R data/workingdata.rda
+    R CMD BATCH figures/boxplot.R
+
+data/workingdata.rda: rawdata.csv workingdata.R
+    R CMD BATCH data/workingdata.R
+```
+
+To produce `paper.pdf` having only downloaded `rawdata.csv` and the command
+files, one would type `make paper.pdf` and the GNU make system would first run
+` R CMD BATCH data/workingdata.R` then, if that is successful, it would run `R
+CMD BATCH figures/boxplot.R` and finally run the line required for creation of
+`paper.pdf`. Later, if only `paper.Rmd` is edited, `make paper.pdf` would only
+run the line for `paper.pdf` because it would know that `boxplot.pdf` and
+`workingdata.rda` are relatively recent and the .csv and .R files required for
+them have not changed.
+
+### Step 3
+We should know where the data came from and what operations were performed on which set of data.
+The authors of this paper have two different versions of this workflow.
+Both fulfill their purpose as a map to the territory of a data analysis.
 
@@
 
-Group work requires version control. Many people are familiar with the 'track
-changes' feature in modern WYSIWYG (what you see is what you get) word
-processors or the fact that Dropbox allows one to recover previous versions of
+Group work requires knowing which versions of files are new, which are old, and
+what changed in between. Many people are familiar with the 'track changes'
+feature in modern WYSIWYG (what you see is what you get) word processors or the
+fact that Dropbox or Google Docs allow one to recover previous versions of
 files. These are both kinds of version control.  More generally, when we
@@
 in favor of previous work while merging the successful parts into the main body
-of the paper; (3) the team can produce 'releases' of the same document
-(one to MPSA, one to APSR, one to their parents) without spawning many possibly
+of the paper; (3) the team can produce 'releases' of the same document (one to
+MPSA, one to APSR, one to their parents) without spawning many possibly
 conflicting copies of the same document; (4) people can work on the same files
@@
 when your future self or your current collaborator saves an old version of a
-file over a new version, erasing good work by accident (or creating a conflicted copy resulting in an arduous back and forth between versions).
+file over a new version, erasing good work by accident (or creating a
+conflicted copy resulting in an arduous back and forth between versions).
 
-Using Dropbox, 'track changes' in Word, or Google Docs is one way to manage your collaborative efforts. They are popular tools, but
-have some downsides as they require you to communicate with other folks
-in your group before you can edit existing files. On Dropbox, you cannot
-edit and save a given file at the same time. Tracking changes with Word has similar limitations (merging the
-changes between two documents is not simple or builtin). Google Docs allows for
-two people to edit the same document at the same time: but you must be online
-with a fairly fast internet connection.^[Note to future selves: change this
-section every five years or so. Probably stop saying 'internet' in 2026.] Communication
-and trading off editing a file prevents your work (or your colleagues' work)
-from getting lost when you both try to save the same file on top of each other. One may also use Dropbox as a kind of server for version
-control. See example above. This system is fine untill your collaborations or project grows large
-(in terms of MBs) and you may run out of storage space on your computer or you may need to upgrade your Dropbox's storage space).
-You will however need to agree on an iron rule for file and directory naming with your collaborators to ensure that the files and directories are always named the same across machines and versions (see above).
+Using Dropbox, 'track changes' in Word, or Google Docs is one way to manage
+your collaborative efforts. They are popular tools, but have some downsides as
+they require you to communicate with other folks in your group before you can
+edit existing files. On Dropbox, you cannot edit and save a given file at the
+same time. Tracking changes with Word has similar limitations (merging the
+changes between two documents is not simple or built-in). Google Docs allows
+for two people to edit the same document at the same time: but you must be
+online with a fairly fast internet connection.^[Note to future selves: change
+this section every five years or so. Probably stop saying 'internet' in 2026.]
+Communication and trading off while editing a file prevents your work (or your
+colleagues' work) from getting lost when you both try to save the same file on
+top of each other. One may also use Dropbox as a kind of server for version
+control. See example above. This system is fine until your collaborations or
+project grows large (in terms of MBs) and you may run out of storage space on
+your computer or you may need to upgrade your Dropbox's storage space. You will
+however need to agree on an iron rule for file and directory naming with your
+collaborators to ensure that the files and directories are always named the
+same across machines and versions plus a clear communication plan for trading off.
 
@@
 within the same day, just before a deadline for example, you may even add a
-time, so it becomes 20160623_4PM_inequalitypaper.docx). This communicates very
-clear what version you are working. Remember the days when you received a file
+time, so it becomes 20160623_4PM_inequalitypaper.docx). This communicates
+version and prevents clobbering unless the old file has the same name or you
+don't update the new file name. Remember the days when you received a file
 called inequalityMV4_APSRsubmit_reallyfinal2.tex? That should quickly become
-something of the past! We find ourselves preaching the gospel to our
+something of the past. We find ourselves preaching this gospel to our
 collaborators frequently and will keep renaming files until a collaborator has
@@
 remember, the life of an idea is measured in years. If you are wise enough to
-have saved your documents as plain text then you can easily compare documents
-using the many utilities available for comparing text files.^[Adobe Acrobat
-allows one to compare differences in pdf files. OpenOffice supports a 'Compare
-Documents' option. Word now does the same. And Google Docs will report on the
-version history of a document.] When you reach certain milestones you can
-rename the file accordingly: thedocAPSA2009.tex --- for the one sent to
-discussants at APSA --- or thedocAPSR2015.tex --- for the version eventually
-sent to the APSR six years after you presented it at APSA. The formal version
-control systems mentioned above all allow this kind of thing and are much more
-elegant and capable, but you can do it by hand too as long as you don't mind
-taking up a lot of disk space and having many 'thedoc... ' files around.
-Indeed, the number of files can add up quickly (especially around submission
-time!) and you may want to create an "0_Archive" folder to store older
-versions. If you do version control by hand, spend a little extra time to
+have saved your documents as plain text then you can easily compare versions of
+the same document using the many utilities available for comparing text
+files.^[Adobe Acrobat allows one to compare differences in pdf files.
+OpenOffice supports a 'Compare Documents' option. Word now does the same. And
+Google Docs will report on the version history of a document. On a Mac the
+FileMerge utility works for plain text files.] When you reach certain
+milestones you can rename the file accordingly: thedocAPSA2009.tex --- for the
+one sent to discussants at APSA --- or thedocAPSR2015.tex --- for the version
+eventually sent to the APSR six years after you presented it at APSA. The
+formal version control systems mentioned above all allow this kind of thing and
+are much more elegant and capable, but you can do it by hand as long as you
+don't mind taking up disk space and having many 'thedoc... ' files
+around.  Indeed, the number of files can add up quickly (especially around
+submission time!) and you may want to create an "0_Archive" folder to store
+older versions. If you do version control by hand, spend a little extra time to
 ensure that you do not clobber files when you make mistakes typing in the
@@
 
-However, best practice we feel is to use formal version control approaches to organizing work [version
-control](https://en.wikipedia.org/wiki/Version_control) strategy. As of 2016,
-the standard for managing large collaborations with many files is the
+We feel the best practice is to use formal version control approaches to
+organizing work [version
+control](https://en.wikipedia.org/wiki/Version_control). As of 2016, the
+standard for managing large collaborations with many files is the
 [Git](https://en.wikipedia.org/wiki/Git_(software)) system. User friendly free
@@
 convenient and integrated into the kind of workflow that your future self will
-be proud of: [Open Science Framework](http://osf.io)
-[bitbucket](http://bitbucket.com), [github](http://github.com). Jake uses
+be proud of:  [github](http://github.com) and [bitbucket](http://bitbucket.com)
+make it easier to use git, and the [Open Science Framework](http://osf.io)
+integrates with github to further ease this task. Jake uses
 [github](http://github.com) for all of his collaborations with others as well
@@
 control systems takes a bit of time. We suggest, however, it is well worth the
-time investment as it will save (lots!) of time later on.
+time investment as it will save lots of time later on.
 
-**Step 5** Writing is rewriting. Thus, all writing involves versions. When we
+### Step 4
+Writing is rewriting. Thus, all writing involves versions. When we
 collaborate with ourselves and others we want to avoid clobbering and we want
@@
 one writes, the more time one has to appreciate problems arising from bugs,
-errors, and typos in data analysis and code. Since mistakes will always be
-made, we should just recognize we need more standardization of practice to
-minimize the mistakes and help others find them. One thing we can do to
+errors, and typos in data analysis and code. One thing we can do to
 minimize and catch the inevitable mistakes is to include testing in our coding.
 
-Of course, this idea, is not new. The desire to avoid error
-looms large when large groups of programmers write code for multi-million
-dollar programs. The idea of [test driven
-development](http://en.wikipedia.org/wiki/Test-driven_development) and the idea
-that one ought to create tests of [small parts of one's
+This idea is not new. The desire to avoid error looms large when large groups
+of programmers write code for multi-million dollar programs. The idea of [test
+driven development](http://en.wikipedia.org/wiki/Test-driven_development) and
+the idea that one ought to create tests of [small parts of one's
 code](http://en.wikipedia.org/wiki/Unit_testing) arose to address such
@@
 ```
+
 Here we use the test function to make sure it works.
 
-```{r results = 'asis'}
+```{r}
 test.times.2.fn()
@@
 
-Ack! we mistyped '+' for '*'. Good thing we wrote the
+Ack! we mistyped $+$ instead of $*$. Good thing we wrote the
 test!^[A more common example of this kind of testing occurs
@@
 
-**Step 6** No one can forsee all of the ways that a
-computer program can fail. One can, however, at least make sure that
-it succeeds in doing the task motivating the writing of the code in
-the first place.
+### Step 5
+No one can forsee all of the ways that a computer program can fail.
+One can, however, at least make sure that it succeeds in doing the task
+motivating the writing of the code in the first place or produces an error in
+the right place and the right time.
 
@@
 
-Over the past decades more and more attention is being paid to 'reproducible research', creating a
-'reproducible workflow' and 'literate programming' these days. As always, the devil is in the details: Here we list a few of our own
-attempts at enabling reproducible research. You'll find many other inspiring
-examples on the web. Luckily, the open source ethos aligns nicely with academic
-incentives, so we are beginning to find more and more people offering their
-files online for copying and improvement. By the way, if you do copy and
-improve, it is polite to alert the person from whom you made the copy about
-your work and to credit them in some way.
+Over the past decades more and more attention is being paid to 'reproducible
+research', and creating a 'reproducible workflow'. As always, the devil is in
+the details: Here we list a few of our own attempts at enabling reproducible
+research. You'll find many other inspiring examples on the web. Luckily, the
+open source ethos aligns nicely with academic incentives, so we are beginning
+to find more and more people offering their files and ideas about workflow
+online for copying and improvement. By the way, if you do copy and improve, it
+is polite to alert the person from whom you made the copy and to credit them in
+some way if not also to offer your ideas for improvement.^[For a formal example
+of copying and improving code, see the GitHub based fork and pull-request
+workflow <https://gist.github.com/Chaser324/ce0505fbed06b947d962>.]
 
-We have experimented with several systems so far:
-
-(1) We wrote this paper in the [R markdown](http://rmarkdown.rstudio.com/)
-literate programming format using two different text editors (Jake used vim)
-and (Maarten used [atom](https://atom.io/)). We organized our files and
-collaboration on GitHub. The source code for this document can be accessed,
-downloaded and modified from <https://github.com/jwbowers/workflow>. We also
-used GitHub Issues to send notes to each other and maintain a task list. The
-nice thing about GitHub is that is enables you to make "releases" of a project
-which enable you to smooth the reproduction of all of the products of your
-research (simulations, data analyses, tables, figures, etc..).  Similar
-features are offered by the [Open Science Framework (OSF)](https://osf.io/).
-Maarten recently participated in a workshop on research transparency organized
-by the [Berkeley Institute for Transparency in the Social
-Sciences](www.bitss.org) in which all [course material](https://osf.io/qh2nr/)
-was accessible through OSF.  OSF integrates with Github nicely and a key
-benefit is that both systems have been created to remain for a long time
-(unlike perhaps your university drive).  (2) For one paper Jake simply mixed R
-and LaTeX code in one document (called the "Sweave" format of literate
-programming) and then added that document and data into into a compressed
-archive [@bowers2005dataverse]; (3) For another more computing intensive paper
-Jake assembled a set of files that enabled reproduction of results using the
-`make` system [@bowers2008dataverse]; (4) Jake also tried the "compendium"
-approach [@gentleman2005reproducible,@gentleman2007statistical] which embeds an
-academic paper within the R package system [@bowers2011dataverse]; The benefit
-of the compendium approach is that one is not required to have access to a
-command line for `make`: the compendium is downloadable from within R using
+We have experimented with several systems so far: (1) We wrote this paper in
+the [R markdown](http://rmarkdown.rstudio.com/) literate programming format
+using two different text editors (Jake used vim) and (Maarten used
+[atom](https://atom.io/)). We organized our files and collaboration on GitHub.
+The source code for this document can be accessed, downloaded and modified from
+<https://github.com/jwbowers/workflow>. We also used GitHub Issues to send
+notes to each other and maintain a task list. The nice thing about GitHub is
+that is enables you to make "releases" of a project which enable you to smooth
+the reproduction of all of the products of your research (simulations, data
+analyses, tables, figures, etc..).  Similar features are offered by the [Open
+Science Framework (OSF)](https://osf.io/).  Maarten recently participated in a
+workshop on research transparency organized by the [Berkeley Institute for
+Transparency in the Social Sciences](www.bitss.org) in which all [course
+material](https://osf.io/qh2nr/) was accessible through OSF.  OSF integrates
+with Github nicely and a key benefit is that both systems have been created to
+remain for a long time (unlike perhaps your university drive).  (2) For one
+paper Jake simply mixed R and LaTeX code in one document (called the "Sweave"
+format of literate programming) and then added that document and data into into
+a compressed archive [@bowers2005dataverse]; (3) For another more computing
+intensive paper Jake's collaborator, Mark Fredrickson, assembled a set of files
+that enabled reproduction of results using the `make` system
+[@bowers2008dataverse]; (4) Jake also tried the "compendium" approach
+[@gentleman2005reproducible,@gentleman2007statistical] which embeds an academic
+paper within the R package system [@bowers2011dataverse]; The benefit of the
+compendium approach is that one is not required to have access to a command
+line for `make`: the compendium is downloadable from within R using
 `install.packages()` and is viewable using the `vignette()` function in any
@@
 It may work for projects with few co-authors but when the number of collaborators increase it becomes very difficult to keep track.
-Online systems for task management, if which most have mobile apps, include amongst otheres [Flow](http://www.getflow.com), [Asana](http://www.asana.com), [Wrike](http://www.wrike.com), [Basecamp](http://basecamp.com), Simplenote, [Evernote](www.evernote.com) and for Windows users there is OneNote.
+Most online systems for task management have mobile apps, including amongst others [Flow](http://www.getflow.com), [Asana](http://www.asana.com), [Wrike](http://www.wrike.com), [Basecamp](http://basecamp.com), Simplenote, [Evernote](www.evernote.com) and for Windows users there is OneNote.
 
-**Step 6** We all learn by doing. When we create a reproducible workflow and share
+### Step 7
+We all learn by doing. When we create a reproducible workflow and share
 reproduction materials we improve both cumulation of knowledge and our
@@
 moment in time as we race against a deadline. All of us collaborate with
-ourselves over time. The time-frames over
-which collaboration are required --- whether among a group of people working
-together or within a single scholar's productive life or probably both --- are
-much longer than any given version of any given software will easily exist.
-Plain text is the exception.
+ourselves over time. The time-frames over which collaboration are required ---
+whether among a group of people working together or within a single scholar's
+productive life or probably both --- are much longer than any given version of
+any given software will easily exist.  Plain text is the exception.
 
@@
 For example, when Jake was invited into the homes and offices of ordinary
-people in Chile in 1991, the "stuff" was recordings of conversations that he
-had about life during the first year of democracy. Now, it might be easy to
-convince others that 'this collection of stuff' is different from 'that
-collection of stuff' if those people were looking over our shoulders the whole
-time that we made decisions about collecting the stuff and broke it up into
-understandable parts and reorganized and summarized it.  Unfortunately, we
-can't assume that people are willing to shadow a researcher throughout her
-career. Rather, we do our work alone or in small groups and want to convince
-other distant and future people to take our analyses and findings seriously.
+people in Chile in 1991, the stuff was recordings of long and semi-structured
+conversations about life during the first year of democracy. Now,
+it might be easy to convince others that 'this collection of stuff' is
+different from 'that collection of stuff' if those others were looking over our
+shoulders the whole time that we made decisions about collecting the stuff and
+broke it up into understandable parts and reorganized and summarized it.
+Unfortunately, we can't assume that people are willing to shadow a researcher
+throughout her career. Rather, we do our work alone or in small groups and want
+to convince other distant and future people to take our analyses and findings
+seriously.
 
@@
 work.' To be extra helpful you can add 'Read the README file for further
-instructions.' And then you can get on with your life: maybe the next great
-idea will occur when your 4-year-old asks a wacky question after stripping and
-painting her overly cooperative 1-year-old brother purple, or teaching a class,
-or in a coffee shop, or on a quiet walk.
+instructions.' And then you can get on with enjoying your life full of friends,
+children, laundry, news, or even journal reviews.
 
